### 2019-10-01 paper/main.tex
adding papaer, incomplete

%% This is a skeleton file demonstrating the use of IEEEtran.cls (requires IEEEtran.cls version 1.8a or later) with an IEEE conference paper.
%%
%% Modified by Khan Reaz( kahn.reaz@ieee.org)
%% Support sites:
%% http://www.ieee.org/

%%***********************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or implied; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk and can modify as s/he wants.

%%***********************************************************

\def\b0{{\bf 0}}
\def\ba{{\bf a}}
\def\bb{{\bf b}}
\def\bc{{\bf c}}
\def\bd{{\bf d}}
\def\be{{\bf e}}
\def\bg{{\bf g}}
\def\bh{{\bf h}}
\def\bi{{\bf i}}
\def\bj{{\bf j}}
\def\bk{{\bf k}}
\def\bl{{\bf l}}
\def\bm{{\bf m}}
\def\bn{{\bf n}}
\def\bo{{\bf o}}
\def\bp{{\bf p}}
\def\bq{{\bf q}}
\def\br{{\bf r}}
\def\bs{{\bf s}}
\def\bt{{\bf t}}
\def\bu{{\bf u}}
\def\bv{{\bf v}}
\def\bw{{\bf w}}
\def\bx{{\bf x}}
\def\by{{\bf y}}
\def\bz{{\bf z}}

\def\bSigma{{\bf \Sigma}}
\def\bA{{\bf A}}
\def\bB{{\bf B}}
\def\bC{{\bf C}}
\def\bD{{\bf D}}
\def\bE{{\bf E}}
\def\bF{{\bf F}}
\def\bG{{\bf G}}
\def\bH{{\bf H}}
\def\bI{{\bf I}}
\def\bJ{{\bf J}}
\def\bK{{\bf K}}
\def\bL{{\bf L}}
\def\bM{{\bf M}}
\def\bN{{\bf N}}
\def\bO{{\bf O}}
\def\bP{{\bf P}}
\def\bQ{{\bf Q}}
\def\bR{{\bf R}}
\def\bS{{\bf S}}
\def\bT{{\bf T}}
\def\bU{{\bf U}}
\def\bV{{\bf V}}
\def\bW{{\bf W}}
\def\bX{{\bf X}}
\def\bY{{\bf Y}}
\def\bZ{{\bf Z}}

%package list
\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\let\labelindent\relax
% \usepackage{breqn}
\usepackage{enumitem}
\usepackage{fleqn}
\usepackage{cite}
\usepackage{graphicx}
\usepackage[varg]{newtxmath}
\graphicspath{ {images/} }
\usepackage{pdfpages}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{lettrine}
\usepackage{amsmath}
% \usepackage{titlesec}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{float}
\usepackage[font={footnotesize}]{caption}
\usepackage[numbers,sort,square,compress]{natbib}
\usepackage[para]{footmisc}
\usepackage{xcolor}
\setlength{\belowcaptionskip}{-12pt}
\newcommand{\highlight}[1]{%
  \colorbox{red!50}{$\displaystyle#1$}}

% \usepackage{parskip}
% \setlength{\parskip}{0.02\baselineskip}

% \fancypagestyle{plain}{
%   \fancyhf{} % sets both header and footer to nothing
% \renewcommand{\headrulewidth}{0pt}
%   \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}% Right header

% }
\pagestyle{plain}% Set page style to plain.

% \pagestyle{fancyplain}
% \fancyhf{}
% \renewcommand{\headrulewidth}{0pt}
% \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}
\DeclareMathOperator*{\argmin}{argmin}
% \usepackage{titlesec}

% \titlespacing*{\section}{0pt}{1pt plus 2p}{1ex}
% \titleformat*{\section}{\fontsize{12}{12}\bfseries}
% \titleformat{\section}
%        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries}{\thesection}{1em}{}

% \titleformat{\subsection}
%        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries\itshape}{\thesubsection}{1em}{}

\begin{document}
%Here goes the title

\title{Heading to be decided}

%Authors List

 \author{\authorblockN{Mohit Madan, Kumar Appaiah, Agrim Gupta}
 \
 \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay\\
 \{mohit,akumar,agrim\}@ee.iitb.ac.in}
% \thanks{Parts of this work was supported by the Bharti Centre for Communication in
% IIT Bombay, and the Visvesvaraya
% PhD Scheme of Ministry of Electronics \& Information Technology,
% Government of India (implemented by the Digital India Corporation).
% }}
}
\maketitle

\thispagestyle{plain}
%Main body starts

% \noindent We consider problem of quantization and interpolation of
% time and frequency varying precoding matrices in wireless MIMO
\begin{abstract}
In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter has the knowledge of the channel state information(CSI). This precoding information is known at the receiver and can be fed back to the transmitter. However, it is not possible to feedback complete information for a limited-feeback OFDM channel since the channel information for multiple subcarriers consume a large amount of data. Therefore to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to minimal number of scalar paramters calculated by decomposing the matrix using Givens Rotation. For an i.i.d flat-fading Rayleigh channel, these parameters are independent which makes its quantization and interpolation easy. We propose an efficient method for quantization of the channel state information at the receiver by using adpative depta modulation for successive time interval. We also propose a method for joint time-frequency interpolation of the channel information at the transmitter.
\end{abstract}

\section{Introduction}
\label{intro}

% no need to write the numbers in your introduction. You start by explaining what is happenning currently and then propose your solution.
MIMO channels can perform significantly better if transmitter has complete knowledge of the Channel State Information (CSIT). In a frequency selective OFDM channel each subcarrier can be modelled as a flat fading MIMO channel. With the CSI information available at the transmitter, beamforming is used to increase the capacity and reduce the BER \cite{love2008overview}. The receiver has the channel knowledge and it could provide feedback to the transmitter after quantizing it. However, it is unrealistic to feeback complete information to the transmitter and therefore information is communicated through a limited feedback channel.

In recent years, codebook design over the Stiefel Manifold \cite{6891198} has been proposed where the manifold structure of the channel matrix is used for its predictive quantization and interpolation across time and frequency. Another such variation of \cite{6891198} has exploited both time and frequency domain together in Stiefel Manifold \cite{Gupt1905:Predictive}. All these methods are good but they do so over Manifolds which has problems of getting stuck in local Maxima. Also there are different challeneged which are faced while interpolating in such a high dimensional space.

Another very different approach in \cite{4114278} have tried to break down the precoder matrix into independent scalar parameters using transformations - Givens rotation and Household Transformation. They are tracking each of these scalar parameters using ADCPM(Adaptive Digital Pulse Code Modulation) for a slowly varying channel in time domain. They have shown capacity achieved with quantized precoding matrix to be very close to the unquantized one. In another paper \cite{4556174}, they have exploited frequency and performed ADPM in frequency domain to achieve similar performance with lower number of feedback bits.

We now propose a method for predictive quantization in time domain over few subcarriers using scalar parameters.We also propose a method for exploiting the joint time-frequency correlation of those scalar parameters on the transmitter side. In \ref{section2}, westart by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix. In \ref{section3} we perform various simulations for our approach and show the benefits from the imporoved BER rates and achievable Rate. 
%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back. In an OFDM channel, this data consume significant amount of bits and therefore, sending the complete data could consume too much bandwidth data. For e.g. an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no. of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame. This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies. Also, the matrices which need to be fed back could be broken down into independent scalar parameters. These scalar quantities could be later quantized and interpolated.
% Broad summary of your work, come to your problem definition.

\section{System Model}
\label{section2}
For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel could be modelled as 
\vspace{-1pt}
\begin{equation}
y_{i,t} = H_{i,t} x_{i,t}+ \eta
\end{equation}

\vspace{-1pt}
where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$. On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$. $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H. The entries of H are assumed to be i.i.d. $\tilde{\mathcal{N}}(0,1)$, complex circularly symmetric Gaussian with mean zero and variance 1. Therefore, the first n column vectors of $V_{i,t}$ have to be quantized and fed back to the transmitter as channel spatial information.

% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).cos(theta) 0 $\leq$ Theta $\leq$ pi/2.
% Householders' transformation and how we are decomposing. -pi$\leq$ phi $\leq$ pi. Uniformly distributed.

% Low-complexity quantization scheme based on a further parameterization of a vector in
% terms of a minimal number of independent scalar parameters. In particular, we consider its application to CSI feedback in slowly time-varying MIMO channels

\subsection{Scalar Paramaterization}
\label{givens}
The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix \cite{4114278}. An orthonormal matrix with $t$ rows and $n$ columns can be represented as

\begin{equation}
\textbf{V} = \Bigg[\prod_{k=1}^{n} D_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} G_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{I} 
\end{equation}

where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k}, \ldots, \phi_{k,t} \big)$ =  $diag\big( 1_{k-1}, e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, $1_{k-1}$ is (k-1) ones; and t x n matrix $\tilde{I}$ = $\big[I_{n}, 0_{n,t-n}\big]^{T}$. The Givens matrix \textbf{G} is given as

\begin{equation}
G_{m-1,m}\big(\theta\big)  = 
\begin{bmatrix}
I_{m-2} & & & \\
& cos\;\theta & - sin\;\theta & \\
& sin\;\theta & cos\;\theta & \\
& & & I_{t-m}
\end{bmatrix}
\end{equation}

% \begin{equation}

For a $4 \times 2$ orthogonal matrix V, 
$$V = D_{1}(\phi_{1,1},\ldots,\phi_{1,4}).G_{3,4}(\theta_{1,1}) .G_{2,3}(\theta_{1,2}) .G_{1,2}(\theta_{1,3})$$
\vspace{-1.4em}
\hspace{1pt}$$.D_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) .G_{3,4}(\theta_{2,1}) .G_{2,3}(\theta_{2,2}).\tilde{I}$$

Here, we denote $\phi_{i}'s$ by phases and $\theta_{i}'s$ by rotation angles. $\phi \in (-\pi, + \pi]$ is uniformaly distributed between the two extremes\cite{4114278}, while $\theta$ has a distribution
\begin{equation}
p(\theta) = 2lsin(\theta)^{2l-1}cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
\end{equation} 
Here all $\theta_i$ and $\phi_i$ are statistically independent which is usefel for tracking them across time as explained in the next subsection.
% may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.

\subsection{Differential Quantization - Channel Tracking}
\label{quantiz}
For a slowly time varying channel, ADPM is an effective method for tracking a paramater over time with using only 1-bit. The 1-bit number $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$ where $x$ is the unquantized vector \& $\hat{x}$ is the quantized vector and $Q(x)$ is the quantization function given as

\begin{equation}
  Q(x)=\begin{cases}
    1, & \text{if $x>0$}.\\
    -1, & \text{otherwise}.
  \end{cases}
\end{equation}

Thus, the quantied vector $\hat{x}_n$ changes with time as 
\begin{equation}
\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n}
\end{equation}
\begin{equation}
\Delta_{n} = \begin{cases}
    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}.\\
    \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
  \end{cases}
\end{equation}

Here we initialize delta using $\Delta_1 = abs(x_{2}-\hat{x}_1)$.

We need to quantize the subcarriers in a way such that we can exploit the joint time-frequency interpolation at the transmitter side. Therefore we select the subcarriers for quantization in an alternating fashion as shown in the figure. Figure shows a 64 subcarrier MIMO channel
% We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using 10 bits.
Explain ADPM here
also the way you are alternatingly
\subsection{Joint Time Frequency Interpolation}
\label{interp}

Here explain about how yo
%Explain prediction and quantization steps here itself ! (succintly)
\section{SIMULATION AND DISCUSSION}
\label{section3}

Initialization is performed by using 10x12 bits for each subcarrier for the first two time frames. Later, the quantization will take place in the time-domain using 1bit for each parameter. Since the bit budget is low and we are considering slowly varying channels, therefore, we will be sending only theta and phi parameters alternatingly in the time domain while keeping the unsent parameters equal to their previous values. This brings down the bit budget to 6 bits average per fed back subcarrier. 
Describing how you performed ADPCM to carry out the 1-bit quantization. Also, the increment rate was 2.4 times the decay rate. Also, need to describe the initialization of the Delta values required in the quantization process. I need to cite that paper. We should also try over different quantization rates to compar \cite{Gupt1905}

While doing interpolation at the transmitter end, future time instances are used to interpolate the rest of the subcarriers indices which are not fed back. To prevent much delay we are using 6 future time instances and using SciPy cubic interpolation over each parameter in 2D. These interpolated values and then used to reconstruct the precoding matrices all together.
Performance Measurements:
Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us. Later we are comparing the ber rates with the ideal completely fed back subcarriers.
Using the idea of joint interpolation given in the other paper.

\begin{figure*}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/system2.pdf}

\caption{Forward Prediction (Green): Both the transmitter and receiver utilize identical reservoirs and quantized precoder estimate at $t-1$ (i.e. $\tilde{\bu}_{i,t-1}$, top left of figure) to obtain $\bP_{i,t}$ via \eqref{fwd_pred}. The transmitter estimates the quantized precoder at $t$ (i.e. $\tilde{\bu}_{i,t}$, top right of figure) using receiver feedback $\text{fb}_{i,t}$ via $q_P$ \eqref{eqn:cn2}. 
\\Feedback Estimation (Cyan): The Receiver estimates the ideal $\text{fb}_{i,t}$ using the exact precoders $\bU_{i,t}$ via \eqref{eqn:cn1}. The transmitter observes only the feedback $\text{fb}_{i,t}$\\ Backward Pass (Blue): $\bW^{\text{out}}_{i,t}$ is updated via \eqref{wout_upd} for optimizing $\mathcal{O}_{i,t}$. This is done at both transmitter and receiver to ensure identical evolution of the reservoirs} 
\label{res_overview}
\vspace{-5pt}
\end{figure*}

\subsection{Reservoir Training Procedure, the backward pass}
\label{train}

\begin{equation}
\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=q_P(f_P(\tilde{\bU}_{i,t-1}),\text{fb}_{i,t})
\label{fpqp}
\end{equation} %its good to label every eqution, section or anything

\subsection{Performance Measurement}
\label{setting}

\noindent Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us. Later we are comparing the ber rates with the ideal completely fed back subcarriers. Or simulations and discussions part 2 criteria could be used. Done similar work as in bhaskar rao paper but along with that we have tried to interpolate the joint interpolation in time and frequency at the transmitter. Also since we want to lower the number of bits we have used the method of interpolation and have achieved significant improvement over predictive quantization in the stiefel manifold paper which came in 2014 and also a variation which came in 2018 by Agrim gupta who tried to use joint interpolation of the stiefel manifold method. In fact we use joint interpolation of the scalar parameters which are easy to use gives better quantization than any of the above methods. 
The advantage they had over the number of bits due to the 6 bit codebook they used is also achieved by us by using smart interpolation techniques, i.e. by dropping the feedback bits in alternating time instances as it is still going to follow the scalar parameters without much difference. 
The problem of values hopping between pi and -pi is also tackled by wrapping the values around which we can show works nicely. The only drawback is that while initializing the parameter, they go out of their actual range and therefore using a uniform quantizer over the prescribed range does not work. But since most of the values lie between the given range we can use most of the initialization bits quantizing the parameters uniformly between the [-$\pi$ , $\pi$] and using smaller amount of bits between the range outside that which could go upto $-3\pi$ and $3\pi$, i.e. we are covering 300% more range outside the given area and that works mostly fine for the rest of the values.

Using the idea of joint interpolation given in the other paper.

%Qtisn error with 1000 chan evols
%Qtisn error with 100 chan evols
%BER
%Achievable rat

\section{Conclusions and Future Work}
\label{section4}

We could reduce more by seeing the effect of sigma values(power distribution). We can save the bits by not sending for those which have very low sigma values. This could result in a very robust method.

\vspace{-4pt}

% \section{Acknowledgment}
% % \label {section6}
% % \input{sections/6_section.tex}
% Parts of this work was supported by the Bharti Centre for Communication in
% IIT Bombay, and the Visvesvaraya
% PhD Scheme of Ministry of Electronics \& Information Technology,
% Government of India, being implemented by Digital India Corporation.

\renewcommand{\bibfont}{\footnotesize}
\bibliography{IEEEabrv,main}
\bibliographystyle{IEEEtran}
\end{document}

% Remove the first para of FSM. State what r1,r2,d,\theta are in the FSM model states themselves. As a concluding statement of STAT-2 state that using the
% inherit geometery of the problem, we get AoA

### 2019-10-01 paper/main.tex
simulation figures added.

@@
 
-\title{Heading to be decided}
+\title{Scalar Parametrization and Joint Time-Frequency interpolation technique for MIMO-OFDM}
 
@@
 Here all $\theta_i$ and $\phi_i$ are statistically independent which is usefel for tracking them across time as explained in the next subsection.
+The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is given by the relation $N_{R}(2*N_{T} - N_{R}) $ \cite{4114278}.
+Number of $\phi = N_{R}(2*N_{T} - N_{R}-1)/2$ while the number of $\theta = N_{T}(2*N_{T} - N_{R}+1)/2$
 % may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
@@
 Therefore we select the subcarriers for quantization in an alternating fashion as shown in the figure.
-Figure shows a 64 subcarrier MIMO channel
+Figure shows a \textbf{S} subcarrier MIMO channel with subcarriers which are going to be quantized at the position sk, k = 0,1,...,S-1/s for even time instance.
+Here s is the gap between two frequencies.
+For odd time instances, subcarriers at the position sk + t , k = 0,1,..., S-1/s - 1 with t = floor(s/2).
+Therefore in ADPCM channel tracking scheme, the arrows in the figure show how previous quantized values along with with the 1-bit number find the new quantized value.
+Now with the above discussion, the bit budget for each subcarrier will be equal to number of scalars * 1-bit.
+On an average, this can be reduced to half if we transmit only $\theta$ values for one time instance and only $\phi$ values for the next time while keeping the untransmitted values same as the previous one.
+This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2*N_{T} - N_{R})/2$.
+One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $sin/cos(\pi) = sin/cos(-\pi)$.
+Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).
+
 % We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using 10 bits.
-Explain ADPM here
-also the way you are alternatingly
 \subsection{Joint Time Frequency Interpolation}
@@
 
-Here explain about how yo
+After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices sk or sk+t.
+To reproduce the rest of the subcarriers will use future time instances to do the interpolation.
+As you can see in the figure, each parameter is interpolated using the subcarriers of the next six time instances.
+We do the interpolation using the spline-cubic interpolation.
+We could also use other standard techniques by calculating the MMSE at the missing subcarrier locations but the spline-cubic interpolation method gives better results over the other method.
+
+Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because 
 %Explain prediction and quantization steps here itself ! (succintly)
@@
 Using the idea of joint interpolation given in the other paper.  
+%talk about this in the simulations, that would be more relevant
+Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found emperically that $m=2.4*M$ i.e. decay
 
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=0.8\textwidth]{images/system2.pdf}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{images/ber_results.pdf}
 
-\caption{Forward Prediction (Green): Both the transmitter and receiver utilize identical reservoirs and quantized precoder estimate at $t-1$ (i.e.
-$\tilde{\bu}_{i,t-1}$, top left of figure) to obtain $\bP_{i,t}$ via \eqref{fwd_pred}.
-The transmitter estimates the quantized precoder at $t$ (i.e.
-$\tilde{\bu}_{i,t}$, top right of figure) using receiver feedback $\text{fb}_{i,t}$ via $q_P$ \eqref{eqn:cn2}.
-\\Feedback Estimation (Cyan): The Receiver estimates the ideal $\text{fb}_{i,t}$ using the exact precoders $\bU_{i,t}$ via \eqref{eqn:cn1}.
-The transmitter observes only the feedback $\text{fb}_{i,t}$\\ Backward Pass (Blue): $\bW^{\text{out}}_{i,t}$ is updated via \eqref{wout_upd} for optimizing $\mathcal{O}_{i,t}$.
-This is done at both transmitter and receiver to ensure identical evolution of the reservoirs}
-\label{res_overview}
+\caption{BER vs SNR for pedestrian channel} 
+\label{ber_overview}
 \vspace{-5pt}
-\end{figure*}
-
-\subsection{Reservoir Training Procedure, the backward pass}
-\label{train}
-
-\begin{equation}
-\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=q_P(f_P(\tilde{\bU}_{i,t-1}),\text{fb}_{i,t})
-\label{fpqp}
-\end{equation} %its good to label every eqution, section or anything
+\end{figure}
 
@@
 Using the idea of joint interpolation given in the other paper.  
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{images/vehicular_ber.png}
+
+\caption{BER vehicular plot comparison between \cite{Gupta1905} and our method} 
+\label{ber_overview}
+\vspace{-5pt}
+\end{figure}
 

### 2019-10-03 paper/main.tex
updated main.tex and added new figures

@@
 
-\title{Scalar Parametrization and Joint Time-Frequency interpolation technique for MIMO-OFDM}
+\title{Scalar Parametrization for Joint Time-Frequency interpolation in a MIMO-OFDM channel}
 
@@
 
- \author{\authorblockN{Mohit Madan, Kumar Appaiah, Agrim Gupta}
+ \author{\authorblockN{Author 1, Author 2, Author 3}
  \
- \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay\\
- \{mohit,akumar,agrim\}@ee.iitb.ac.in}
+ \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay}
 % \thanks{Parts of this work was supported by the Bharti Centre for Communication in
@@
 \begin{abstract}
-In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter has the knowledge of the channel state information(CSI).
-This precoding information is known at the receiver and can be fed back to the transmitter.
-However, it is not possible to feedback complete information for a limited-feeback OFDM channel since the channel information for multiple subcarriers consume a large amount of data.
-Therefore to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to minimal number of scalar paramters calculated by decomposing the matrix using Givens Rotation.
-For an i.i.d flat-fading Rayleigh channel, these parameters are independent which makes its quantization and interpolation easy.
-We propose an efficient method for quantization of the channel state information at the receiver by using adpative depta modulation for successive time interval.
+In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information(CSI).
+The receiver knows the channel information, and it feeds back to the transmitter.
+However, it is not possible to feedback complete information for a limited-feedback OFDM channel since the channel information for multiple subcarriers consumes a large amount of data.
+Therefore to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to a minimal number of scalar parameters.
+The Precoder matrix is decomposed using Givens Rotation.
+For an i.i.d.
+flat-fading Rayleigh channel, these parameters are independent, which makes its quantization and interpolation easy.
+We propose an efficient method for quantization of the channel state information at the receiver by using adaptive delta modulation for successive time interval.
 We also propose a method for joint time-frequency interpolation of the channel information at the transmitter.
+153-words
 \end{abstract}
@@
 % no need to write the numbers in your introduction. You start by explaining what is happenning currently and then propose your solution.
-MIMO channels can perform significantly better if transmitter has complete knowledge of the Channel State Information (CSIT).
-In a frequency selective OFDM channel each subcarrier can be modelled as a flat fading MIMO channel.
+MIMO channels can perform significantly better if the transmitter has complete knowledge of the Channel State Information (CSIT).
+In a frequency-selective OFDM channel, a flat fading MIMO channel approximated each subcarrier.
 With the CSI information available at the transmitter, beamforming is used to increase the capacity and reduce the BER \cite{love2008overview}.
-The receiver has the channel knowledge and it could provide feedback to the transmitter after quantizing it.
-However, it is unrealistic to feeback complete information to the transmitter and therefore information is communicated through a limited feedback channel.
-
+The receiver has the channel knowledge, and it could provide feedback to the transmitter after quantizing it.
+However, it is unrealistic to feedback complete information to the transmitter, and therefore information is communicated through a limited feedback channel.
 In recent years, codebook design over the Stiefel Manifold \cite{6891198} has been proposed where the manifold structure of the channel matrix is used for its predictive quantization and interpolation across time and frequency.
 Another such variation of \cite{6891198} has exploited both time and frequency domain together in Stiefel Manifold \cite{Gupt1905:Predictive}.
-All these methods are good but they do so over Manifolds which has problems of getting stuck in local Maxima.
-Also there are different challeneged which are faced while interpolating in such a high dimensional space.
+All these methods are useful, but they do so over Manifolds which has problems of getting stuck in local Maxima.
+Also, there are different challenged which are faced while interpolating in a high dimensional space.
 
-Another very different approach in \cite{4114278} have tried to break down the precoder matrix into independent scalar parameters using transformations - Givens rotation and Household Transformation.
-They are tracking each of these scalar parameters using ADCPM(Adaptive Digital Pulse Code Modulation) for a slowly varying channel in time domain.
-They have shown capacity achieved with quantized precoding matrix to be very close to the unquantized one.
-In another paper \cite{4556174}, they have exploited frequency and performed ADPM in frequency domain to achieve similar performance with lower number of feedback bits.
+Another very different approach in \cite{4114278} has tried to break down the precoder matrix into independent scalar parameters using transformations, namely Givens rotation and Household Transformation.
+They are tracking each of these scalar parameters using ADCPM(Adaptive Digital Pulse Code Modulation) for a slowly varying channel in the time domain.
+They have shown capacity achieved with the quantized precoding matrix to be very close to the unquantized one.
+In another paper \cite{4556174}, they have exploited frequency and performed ADPM in frequency domain to achieve similar performance with a lower number of feedback bits.
 
-We now propose a method for predictive quantization in time domain over few subcarriers using scalar parameters.We also propose a method for exploiting the joint time-frequency correlation of those scalar parameters on the transmitter side.
-In \ref{section2}, westart by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
-In \ref{section3} we perform various simulations for our approach and show the benefits from the imporoved BER rates and achievable Rate.
+We now propose a method for predictive quantization in time domain over few subcarriers using scalar parameters.
+We also propose a method for exploiting the joint time-frequency correlation of those scalar parameters on the transmitter side.
+In \ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
+In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
 %Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
@@
 
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{images/adpm.pdf}
+
+\caption{Subcarriers which need to be quantized} 
+\label{ber_overvie}
+\vspace{-5pt}
+\end{figure}
+
 For a $4 \times 2$ orthogonal matrix V, 
@@
 Here all $\theta_i$ and $\phi_i$ are statistically independent which is usefel for tracking them across time as explained in the next subsection.
-The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is given by the relation $N_{R}(2*N_{T} - N_{R}) $ \cite{4114278}.
-Number of $\phi = N_{R}(2*N_{T} - N_{R}-1)/2$ while the number of $\theta = N_{T}(2*N_{T} - N_{R}+1)/2$
+The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is given by the relation $N_{R}(2N_{T} - N_{R}) $ \cite{4114278}.
+Number of $\phi = N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta = N_{T}(2N_{T} - N_{R}+1)/2$
 % may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
@@
 On an average, this can be reduced to half if we transmit only $\theta$ values for one time instance and only $\phi$ values for the next time while keeping the untransmitted values same as the previous one.
-This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2*N_{T} - N_{R})/2$.
+This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2N_{T} - N_{R})/2$.
 One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $sin/cos(\pi) = sin/cos(-\pi)$.
@@
 
-Initialization is performed by using 10x12 bits for each subcarrier for the first two time frames.
+Initialization is performed by using $10 \times 12$ bits for each subcarrier for the first two time frames.
 Later, the quantization will take place in the time-domain using 1bit for each parameter.
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber_results.pdf}
+\includegraphics[width=0.5\textwidth]{images/pedestrian.pdf}
 
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/vehicular_ber.png}
+\includegraphics[width=0.5\textwidth]{images/vehicular_ber.pdf}
 
-\caption{BER vehicular plot comparison between \cite{Gupta1905} and our method} 
+\caption{BER vehicular plot comparison between our method and agrims} 
 \label{ber_overview}

### 2019-10-03 ber_results/plot_results.py

@@ -7,10 +7,12 @@ hpang_ber=np.load('eval/hp_ber.npy')
 hpmani_ber=np.load('manifold_interp/hp_ber.npy')
 hpang_ber6=np.load('6bit_eval/hp_ber.npy')
 plt.yscale('log')
-plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
-plt.plot(eb_n0, hpmani_ber, label='hop_manifold')
-plt.plot(eb_n0,hpang_ber6,label='hop_givens_6')
-
+plt.ylabel('BER')
+plt.xlabel(r'$E_{b}/N_{o}(dB)$')
+# plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
+plt.plot(eb_n0, hpang_ber6, label='Pedestrian Manifold', marker='o')
+plt.plot(eb_n0,hpmani_ber,label='Pedestrian Scalar Parameters', marker='+')
+plt.grid()
 plt.legend()
 plt.show()
 

### 2019-10-04 paper/main.tex
Minor addition

@@
 % no need to write the numbers in your introduction. You start by explaining what is happenning currently and then propose your solution.
-MIMO channels can perform significantly better if the transmitter has complete knowledge of the Channel State Information (CSIT).
-In a frequency-selective OFDM channel, a flat fading MIMO channel approximated each subcarrier.
-With the CSI information available at the transmitter, beamforming is used to increase the capacity and reduce the BER \cite{love2008overview}.
-The receiver has the channel knowledge, and it could provide feedback to the transmitter after quantizing it.
-However, it is unrealistic to feedback complete information to the transmitter, and therefore information is communicated through a limited feedback channel.
+The use of multiple antennas at the transmitter and receiver can significantly enhance the performance of wireless systems.
+Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform an better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
+To this end, feeding back the channel state information from the receiver to the transmitter efficiently is necessary.
+In the many wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across the many subcarriers of the system, which places a large feedback demand.
+However, since the channel state information varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency interpolation temporal prediction.
+In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
+
 In recent years, codebook design over the Stiefel Manifold \cite{6891198} has been proposed where the manifold structure of the channel matrix is used for its predictive quantization and interpolation across time and frequency.

### 2019-10-04 paper/main.tex
Fix spaces
Minor intro reword

@@
 However, since the channel state information varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency interpolation temporal prediction.
+In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to enable the transmitter to transmite at capacity as well as to parallelize the channel.
 In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
 
-In recent years, codebook design over the Stiefel Manifold \cite{6891198} has been proposed where the manifold structure of the channel matrix is used for its predictive quantization and interpolation across time and frequency.
+% TODO More coverage of prior work needed.
+The problem of effective quantization and feedback of precoders for MIMO-OFDM attracted significant attention In recent years.
+Many approaches involve codebook design over the Stiefel Manifold~\cite{6891198}, wherein the manifold structure of the precoders is used to obtain tangents for predictive quantization and interpolation across time and frequency.
 Another such variation of \cite{6891198} has exploited both time and frequency domain together in Stiefel Manifold \cite{Gupt1905:Predictive}.
@@
 
-Another very different approach in \cite{4114278} has tried to break down the precoder matrix into independent scalar parameters using transformations, namely Givens rotation and Household Transformation.
-They are tracking each of these scalar parameters using ADCPM(Adaptive Digital Pulse Code Modulation) for a slowly varying channel in the time domain.
-They have shown capacity achieved with the quantized precoding matrix to be very close to the unquantized one.
-In another paper \cite{4556174}, they have exploited frequency and performed ADPM in frequency domain to achieve similar performance with a lower number of feedback bits.
+Other approaches have considered the problem of parametrizing precoding matrices using scalar parameters.
+It has been shown that the space of unitary matrices can be effectively parameterized using scalar parameters that correspond to Givens rotations and Household Transformations of the unitary precoding matrices~\cite{4114278,4556174}.
+By tracking these scalar parameters using adaptive digital pulse code modulation (ADPCM) for slowly varying channels in the time domain and the frequency domain, the link can operate close to capacity while using very few bits for quantization and feedback.
 

### 2019-10-04 paper/main.tex
Remove header

--- 2019-10-04 paper/main.tex
@@
-%% This is a skeleton file demonstrating the use of IEEEtran.cls (requires IEEEtran.cls version 1.8a or later) with an IEEE conference paper.
-%%
-%% Modified by Khan Reaz( kahn.reaz@ieee.org)
-%% Support sites:
-%% http://www.ieee.org/
-
-%%***********************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or implied; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk and can modify as s/he wants.
-
-%%***********************************************************
-
 \def\b0{{\bf 0}}
@@
 \label{intro}
+% no need to write the numbers in your introduction. You start by explaining what is happening currently and then propose your solution.
 
-% no need to write the numbers in your introduction. You start by explaining what is happenning currently and then propose your solution.
 The use of multiple antennas at the transmitter and receiver can significantly enhance the performance of wireless systems.
-Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform an better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
-To this end, feeding back the channel state information from the receiver to the transmitter efficiently is necessary.
-In the many wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across the many subcarriers of the system, which places a large feedback demand.
-However, since the channel state information varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency interpolation temporal prediction.
+Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
+To this end, feeding back the CSI from the receiver to the transmitter efficiently is necessary.
+In wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across all subcarriers, and this places a large feedback demand.
+However, since the CSI varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency domain interpolation temporal prediction.
+In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to achieve link capacity, as well as to parallelize the channel.
 In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
 
-In recent years, codebook design over the Stiefel Manifold \cite{6891198} has been proposed where the manifold structure of the channel matrix is used for its predictive quantization and interpolation across time and frequency.
-Another such variation of \cite{6891198} has exploited both time and frequency domain together in Stiefel Manifold \cite{Gupt1905:Predictive}.
-All these methods are useful, but they do so over Manifolds which has problems of getting stuck in local Maxima.
-Also, there are different challenged which are faced while interpolating in a high dimensional space.
+The problem of effective quantization and feedback of precoders for MIMO-OFDM has attracted significant attention in recent years.
+Many approaches involve codebook design over the Grassmann manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308} or the Stiefel Manifold \cite{6891198,Gupt1905:Predictive}, wherein the manifold structure of the precoders is used to obtain tangents for predictive quantization and interpolation across time and frequency.
+While these methods have been shown to be very effective for precoder feedback,  they involve operations over higher dimensional manifolds.
+In particular, optimization and interpolation is complicated, especially for manifolds where geodesic paths for interpolation may be difficult to obtain, especially for joint time-frequency predictive coding.
+An alternate approach is to parametrize the precoding matrices using scalar parameters.
+The space of unitary matrices parameterized into independent scalar parameters that correspond to Givens rotations and Household Transformations of the unitary precoding matrices \cite{4114278,4556174}.
+These approaches have been combined with adaptive delta modulation effectively track precoding matrices both in the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit separately.
 
-Another very different approach in \cite{4114278} has tried to break down the precoder matrix into independent scalar parameters using transformations, namely Givens rotation and Household Transformation.
-They are tracking each of these scalar parameters using ADCPM(Adaptive Digital Pulse Code Modulation) for a slowly varying channel in the time domain.
-They have shown capacity achieved with the quantized precoding matrix to be very close to the unquantized one.
-In another paper \cite{4556174}, they have exploited frequency and performed ADPM in frequency domain to achieve similar performance with a lower number of feedback bits.
+In this paper, we propose a method for exploiting the joint time-frequency correlation of the precoder's scalar parameters on the transmitter using CSI feedback.
+Typical approaches to predictive quantization and interpolation do not exploit the joint time-frequency correlation to enhance the precoder estimate at the transmitter.
+However, exploiting the correlations jointly not only reduces the feedback requirement, it also is able to enhance the precoder reconstruction accuracy.
+Further, the independence of the scalar parameters reduces the problem to interpolation of separate scalar valued function, as opposed to operations over tangent spaces in manifolds~\cite{Gupt1905:Predictive}.
+Simulations reveal that the proposed method offers matches the performance of manifold based approaches with a much lower implementation complexity and fewer number of bits.
 
-We now propose a method for predictive quantization in time domain over few subcarriers using scalar parameters.
-We also propose a method for exploiting the joint time-frequency correlation of those scalar parameters on the transmitter side.
-In \ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
-In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
+The rest of this paper is organized as follows: Section~\ref{section2} described the system model and the precoder feedback and interpolation techniques.
+Section~\ref{section3} compares the BER and achievable rates for our proposed methods with that of past work.
+Finally, Section~\ref{section4} concludes with discussions on future work.
+
 %Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
-In an OFDM channel, this data consume significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
+In an OFDM channel, this data consumes significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
 For e.g.
@@
 These scalar quantities could be later quantized and interpolated.
+
 % Broad summary of your work, come to your problem definition.
@@
 \section{System Model}
+
 \label{section2}
+
 For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel.
 This flat-fading MIMO channel could be modelled as
+
 \vspace{-1pt}
+
 \begin{equation}
-y_{i,t} = H_{i,t} x_{i,t}+ \eta
+\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
 \end{equation}
@@
 \vspace{-1pt}
-where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
-On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$.
+
+where the channel matrix $\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $\bx_{i,t} \in \mathbb{C}^{N_T}$; the output $\by_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
+On applying thin SVD to H, we get $\bH_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_R}$, $\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_R}$.
 $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H.
@@
 
-% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).cos(theta) 0 $\leq$ Theta $\leq$ pi/2.
+% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).\cos(theta) 0 $\leq$ Theta $\leq$ pi/2.
 % Householders' transformation and how we are decomposing. -pi$\leq$ phi $\leq$ pi. Uniformly distributed.
-
 % Low-complexity quantization scheme based on a further parameterization of a vector in
@@
 
-\subsection{Scalar Paramaterization}
+\subsection{Scalar Parameterization}
 \label{givens}
+
 The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix \cite{4114278}.
@@
 \begin{bmatrix}
+
 I_{m-2} & & & \\
-& cos\;\theta & - sin\;\theta & \\
-& sin\;\theta & cos\;\theta & \\
+& \cos\;\theta & - \sin\;\theta & \\
+& \sin\;\theta & \cos\;\theta & \\
 & & & I_{t-m}
+
 \end{bmatrix}
@@
 
-% \begin{equation}
-
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/adpm.pdf}
-
+\label{adpm-fig}
+%\includegraphics[width=0.45\textwidth]{images/new-adpm-crop.pdf}
 \caption{Subcarriers which need to be quantized}
@@
 For a $4 \times 2$ orthogonal matrix V,
+
 $$V = D_{1}(\phi_{1,1},\ldots,\phi_{1,4}).G_{3,4}(\theta_{1,1}) .G_{2,3}(\theta_{1,2}) .G_{1,2}(\theta_{1,3})$$
+
 \vspace{-1.4em}
+
 \hspace{1pt}$$.D_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) .G_{3,4}(\theta_{2,1}) .G_{2,3}(\theta_{2,2}).\tilde{I}$$
@@
 $\phi \in (-\pi, + \pi]$ is uniformaly distributed between the two extremes\cite{4114278}, while $\theta$ has a distribution
+
 \begin{equation}
-p(\theta) = 2lsin(\theta)^{2l-1}cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
+p(\theta) = 2l\sin(\theta)^{2l-1}\cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
 \end{equation}
+
 Here all $\theta_i$ and $\phi_i$ are statistically independent which is usefel for tracking them across time as explained in the next subsection.
@@
 Number of $\phi = N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta = N_{T}(2N_{T} - N_{R}+1)/2$
+
 % may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
+
 % To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.
@@
 \label{quantiz}
+
 For a slowly time varying channel, ADPM is an effective method for tracking a paramater over time with using only 1-bit.
-The 1-bit number $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$ where $x$ is the unquantized vector \& $\hat{x}$ is the quantized vector and $Q(x)$ is the quantization function given as
+The 1-bit number $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$ where $x$ is the unquantized vector $\&$ $\hat{x}$ is the quantized vector and $Q(x)$ is the quantization function given as
 
@@
 Thus, the quantied vector $\hat{x}_n$ changes with time as
+
 \begin{equation}
@@
 \end{equation}
+
 \begin{equation}
+\label{delta_eqn}
 \Delta_{n} = \begin{cases}
@@
 
-We need to quantize the subcarriers in a way such that we can exploit the joint time-frequency interpolation at the transmitter side.
-Therefore we select the subcarriers for quantization in an alternating fashion as shown in the figure.
-Figure shows a \textbf{S} subcarrier MIMO channel with subcarriers which are going to be quantized at the position sk, k = 0,1,...,S-1/s for even time instance.
-Here s is the gap between two frequencies.
-For odd time instances, subcarriers at the position sk + t , k = 0,1,..., S-1/s - 1 with t = floor(s/2).
-Therefore in ADPCM channel tracking scheme, the arrows in the figure show how previous quantized values along with with the 1-bit number find the new quantized value.
+We need to quantize the subcarriers in a way such that we can exploit
+the joint time-frequency interpolation at the transmitter
+side. Therefore, we select the subcarriers for quantization in an
+alternating fashion as shown in the figure. Figure shows a \textbf{N}
+subcarrier MIMO channel with subcarriers which are going to be
+quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even
+time instance. Here, s is the gap between two frequencies. For odd
+time instances, subcarriers at the position sk + t for k = 0,1,...,
+$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$.
+
+Therefore, in ADPCM channel tracking scheme, the arrows in the figure show how previous quantized values along with the 1-bit number find the new quantized value.
 Now with the above discussion, the bit budget for each subcarrier will be equal to number of scalars * 1-bit.
-On an average, this can be reduced to half if we transmit only $\theta$ values for one time instance and only $\phi$ values for the next time while keeping the untransmitted values same as the previous one.
+On an average, this can be reduced to half if we transmit only $\theta$ values for one-time instance and only $\phi$ values for the next time while keeping the un transmitted values same as the previous one.
 This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2N_{T} - N_{R})/2$.
-One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $sin/cos(\pi) = sin/cos(-\pi)$.
+
+One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $\sin/\cos(\pi) =\sin/\cos(-\pi)$.
 Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).
 
-% We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using 10 bits.
+% We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using $\textbf{B}$ bits.
+
 \subsection{Joint Time Frequency Interpolation}
@@
 Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because
+
 %Explain prediction and quantization steps here itself ! (succintly)
+
 \section{SIMULATION AND DISCUSSION}
@@
 
-Initialization is performed by using $10 \times 12$ bits for each subcarrier for the first two time frames.
-Later, the quantization will take place in the time-domain using 1bit for each parameter.
-Since the bit budget is low and we are considering slowly varying channels, therefore, we will be sending only theta and phi parameters alternatingly in the time domain while keeping the unsent parameters equal to their previous values.
-This brings down the bit budget to 6 bits average per fed back subcarrier.
-Describing how you performed ADPCM to carry out the 1-bit quantization.
-Also, the increment rate was 2.4 times the decay rate.
-Also, need to describe the initialization of the Delta values required in the quantization process.
-I need to cite that paper.
-We should also try over different quantization rates to compar \cite{Gupt1905}
+ We have performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel.
+Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel.
+Performance is measured by averaging over 10 channel realizations with 100 simulations each.
+The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with sampling frequency $T_s = 20MHz$.
+For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$.
+Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s.
+There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel.
+We take s = 33/9 for Vehicular and pedestrian channel respectively I.e.
+32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, ..
+,7 respectively.
+
+\subsection{Initial Quantization}
+For the first and the second simulations/time instances we use of $\textbf{B}$ = 10 bits for quantizing each parameter for effective quantization/initialization.
+Therefore, the number of bits for each subcarrier = $10\times12$ = 120.
+Quantization is performed as given in Section~\ref{section2}.
+
+Simulation: BER is calculated by averaging over all the simulations. For Vehicular Channel, 1024 subcarriers are used with doppler frequency – and sampling rate = $10^{5}$
+
+Initialization is performed by using $10\times12$ bits for each subcarrier for the initial two time frames.
+Later, the quantization will take place in the time-domain using one bit for each parameter.
+Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{adpm-fig}.
+Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s.
+This brings down the bit budget to average of 5(Size of $\Theta$) and 7(Size of $\Phi$) i.e.
+6 bits average per fed back subcarrier.
+
+\bf{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
+This ratio was decided after trying with different ratios.
+This implies that if quantied values loose track of the actual values then after changing delta quickly they start following it again.
+Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare
 
 While doing interpolation at the transmitter end, future time instances are used to interpolate the rest of the subcarriers indices which are not fed back.
-To prevent much delay we are using 6 future time instances and using SciPy cubic interpolation over each parameter in 2D.
+To prevent much delay, we are using 6 future time instances and using SciPy cubic interpolation over each parameter in 2D.
 These interpolated values and then used to reconstruct the precoding matrices all together.
+
 Performance Measurements:
 Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us.
-Later we are comparing the ber rates with the ideal completely fed back subcarriers.
+Later we are comparing the BER rates with the ideal completely fed back subcarriers.
+
 Using the idea of joint interpolation given in the other paper.
+
 %talk about this in the simulations, that would be more relevant
-Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found emperically that $m=2.4*M$ i.e. decay
+Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.4*M$ i.e. decay
 
@@
 \includegraphics[width=0.5\textwidth]{images/pedestrian.pdf}
-
 \caption{BER vs SNR for pedestrian channel}
@@
 \subsection{Performance Measurement}
+
 \label{setting}
 
-\noindent Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us.
-Later we are comparing the ber rates with the ideal completely fed back subcarriers.
+\noindent Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used.
+Later we are comparing the BER rates with the completely fed back subcarriers.
 Or simulations and discussions part 2 criteria could be used.
-Done similar work as in bhaskar rao paper but along with that we have tried to interpolate the joint interpolation in time and frequency at the transmitter.
-Also since we want to lower the number of bits we have used the method of interpolation and have achieved significant improvement over predictive quantization in the stiefel manifold paper which came in 2014 and also a variation which came in 2018 by Agrim gupta who tried to use joint interpolation of the stiefel manifold method.
-In fact we use joint interpolation of the scalar parameters which are easy to use gives better quantization than any of the above methods.
-The advantage they had over the number of bits due to the 6 bit codebook they used is also achieved by us by using smart interpolation techniques, i.e.
+Done similar work as in \cite{4114278} but along with that we have joint interpolated in time and frequency at the transmitter using the fundamental scalar parameters obtained.
+To lower the number of bits we have used the method of interpolation and have achieved significant improvement over predictive quantization method in \cite{6891198} and also in \cite{Gupt1905:Predictive} which tried to use joint interpolation over the tangent space in the Stiefel Manifold.
+In fact, we use joint interpolation of the scalar parameters which are easy to use and give better quantization than any of the above methods.
+
+The advantage they had over the number of bits due to the 6 bit codebook in \cite{6891198,Gupt1905:Predictive} is also achieved by us by using smart interpolation techniques, i.e.
 by dropping the feedback bits in alternating time instances as it is still going to follow the scalar parameters without much difference.
-The problem of values hopping between pi and -pi is also tackled by wrapping the values around which we can show works nicely.
+
+The problem of values hopping between $\pi$ and $-\pi$ is also tackled by wrapping the values around which we can show works nicely.
 The only drawback is that while initializing the parameter, they go out of their actual range and therefore using a uniform quantizer over the prescribed range does not work.
-But since most of the values lie between the given range we can use most of the initialization bits quantizing the parameters uniformly between the [-$\pi$ , $\pi$] and using smaller amount of bits between the range outside that which could go upto $-3\pi$ and $3\pi$, i.e.
+But since most of the values lie between the given range we can use most of the initialization bits quantizing the parameters uniformly between the [-$\pi$ , $\pi$] and using smaller amount of bits between the range outside that which could go up to $-3\pi$ and $3\pi$, i.e.
 we are covering 300% more range outside the given area and that works mostly fine for the rest of the values.
@@
 Using the idea of joint interpolation given in the other paper.
+
 \begin{figure}
 \includegraphics[width=0.5\textwidth]{images/vehicular_ber.pdf}
-
-\caption{BER vehicular plot comparison between our method and agrims} 
+\caption{BER vehicular plot comparison between our method and \cite{Gupt1905:Predictive}}
 \label{ber_overview}
@@
 \end{figure}
-
 %Qtisn error with 1000 chan evols
@@
 %Achievable rat
-
 \section{Conclusions and Future Work}
@@
 
-We could reduce more by seeing the effect of sigma values(power distribution).
+We could reduce more by seeing the effect of sigma values (power distribution).
 We can save the bits by not sending for those which have very low sigma values.
@@
 % \section{Acknowledgment}
+
 % % \label {section6}
+
 % % \input{sections/6_section.tex}
+
 % Parts of this work was supported by the Bharti Centre for Communication in
+
 % IIT Bombay, and the Visvesvaraya
+
 % PhD Scheme of Ministry of Electronics \& Information Technology,
+
 % Government of India, being implemented by Digital India Corporation.
@@
 \renewcommand{\bibfont}{\footnotesize}
+
 \bibliography{IEEEabrv,main}
+
 \bibliographystyle{IEEEtran}
+
 \end{document}
-
-% Remove the first para of FSM. State what r1,r2,d,\theta are in the FSM model states themselves. As a concluding statement of STAT-2 state that using the
-% inherit geometery of the problem, we get AoA

### 2019-10-04 paper/main.tex
Some changes by Kumar

@@
 
+In this paper, we propose a method for predictive quantization in the time domain over the subcarriers of a MIMO-OFDM system using scalar parameters.
+We also exploit the joint time-frequency correlation of these scalar parameters on the transmitter side.
+TODO how we solve it?
+
+The rest of this paper is organized as follows: Section~\ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
+In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
+%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
+In an OFDM channel, this data consume significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
+For e.g.
+an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no.
+of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame.
+This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies.
+Also, the matrices which need to be fed back could be broken down into independent scalar parameters.
+These scalar quantities could be later quantized and interpolated.
 % Broad summary of your work, come to your problem definition.
@@
 \label{section2}
-
-For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel.
-This flat-fading MIMO channel could be modelled as
-
-\vspace{-1pt}
-
 \begin{equation}
@@
 \end{equation}
-
-\vspace{-1pt}
-
-where the channel matrix $\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $\bx_{i,t} \in \mathbb{C}^{N_T}$; the output $\by_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
-On applying thin SVD to H, we get $\bH_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_R}$, $\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_R}$.
+where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
+On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$.
 $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H.
@@
 
-\subsection{Scalar Parameterization}
+\subsection{Scalar Paramatrization}
 \label{givens}
-
-The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix \cite{4114278}.
+The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix~\cite{4114278}.
 An orthonormal matrix with $t$ rows and $n$ columns can be represented as
@@
 
-where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k}, \ldots, \phi_{k,t} \big)$ =  $diag\big( 1_{k-1}, e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, $1_{k-1}$ is (k-1) ones; and t x n matrix $\tilde{I}$ = $\big[I_{n}, 0_{n,t-n}\big]^{T}$.
+where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k}, \ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( 1_{k-1}, e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, $1_{k-1}$ is (k-1) ones; and t x n matrix $\tilde{I}$ = $\big[I_{n}, 0_{n,t-n}\big]^{T}$.
 The Givens matrix \textbf{G} is given as
@@
 I_{m-2} & & & \\
-& \cos\;\theta & - \sin\;\theta & \\
-& \sin\;\theta & \cos\;\theta & \\
+& \cos\theta & -\sin\theta & \\
+& \sin\theta & \cos\theta & \\
 & & & I_{t-m}
@@
 
+%TODO What is t in the equation below?
 \begin{equation}
-p(\theta) = 2l\sin(\theta)^{2l-1}\cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
+p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
 \end{equation}
-
-Here all $\theta_i$ and $\phi_i$ are statistically independent which is usefel for tracking them across time as explained in the next subsection.
-The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is given by the relation $N_{R}(2N_{T} - N_{R}) $ \cite{4114278}.
-Number of $\phi = N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta = N_{T}(2N_{T} - N_{R}+1)/2$
-
+Here all $\theta_i$ and $\phi_i$ are statistically independent, making it useful for tracking them across time as explained in the next subsection.
+The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}.
+The number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
 % may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
@@
 %Explain prediction and quantization steps here itself ! (succintly)
-
-\section{SIMULATION AND DISCUSSION}
+\section{Simulation and discussion}
 \label{section3}

### 2019-10-06 paper/main.tex
linear vs cubic comparison added. report modified

--- 2019-10-04 paper/main.tex
@@
+%% This is a skeleton file demonstrating the use of IEEEtran.cls (requires IEEEtran.cls version 1.8a or later) with an IEEE conference paper.
+
+%%
+
+%% Modified by Khan Reaz( kahn.reaz@ieee.org)
+
+%% Support sites:
+
+%% http://www.ieee.org/
+
+%%***********************************************************
+
+%% Legal Notice:
+
+%% This code is offered as-is without any warranty either expressed or implied; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE!
+
+%% User assumes all risk and can modify as s/he wants.
+
+%%***********************************************************
+
 \def\b0{{\bf 0}}
+
 \def\ba{{\bf a}}
+
 \def\bb{{\bf b}}
+
 \def\bc{{\bf c}}
+
 \def\bd{{\bf d}}
+
 \def\be{{\bf e}}
+
 \def\bg{{\bf g}}
+
 \def\bh{{\bf h}}
+
 \def\bi{{\bf i}}
+
 \def\bj{{\bf j}}
+
 \def\bk{{\bf k}}
+
 \def\bl{{\bf l}}
+
 \def\bm{{\bf m}}
+
 \def\bn{{\bf n}}
+
 \def\bo{{\bf o}}
+
 \def\bp{{\bf p}}
+
 \def\bq{{\bf q}}
+
 \def\br{{\bf r}}
+
 \def\bs{{\bf s}}
+
 \def\bt{{\bf t}}
+
 \def\bu{{\bf u}}
+
 \def\bv{{\bf v}}
+
 \def\bw{{\bf w}}
+
 \def\bx{{\bf x}}
+
 \def\by{{\bf y}}
+
 \def\bz{{\bf z}}
@@
 \def\bSigma{{\bf \Sigma}}
+
 \def\bA{{\bf A}}
+
 \def\bB{{\bf B}}
+
 \def\bC{{\bf C}}
+
 \def\bD{{\bf D}}
+
 \def\bE{{\bf E}}
+
 \def\bF{{\bf F}}
+
 \def\bG{{\bf G}}
+
 \def\bH{{\bf H}}
+
 \def\bI{{\bf I}}
+
 \def\bJ{{\bf J}}
+
 \def\bK{{\bf K}}
+
 \def\bL{{\bf L}}
+
 \def\bM{{\bf M}}
+
 \def\bN{{\bf N}}
+
 \def\bO{{\bf O}}
+
 \def\bP{{\bf P}}
+
 \def\bQ{{\bf Q}}
+
 \def\bR{{\bf R}}
+
 \def\bS{{\bf S}}
+
 \def\bT{{\bf T}}
+
 \def\bU{{\bf U}}
+
 \def\bV{{\bf V}}
+
 \def\bW{{\bf W}}
+
 \def\bX{{\bf X}}
+
 \def\bY{{\bf Y}}
+
 \def\bZ{{\bf Z}}
@@
 %package list
+
 \documentclass[conference]{IEEEtran}
+
 \IEEEoverridecommandlockouts
+
 \let\labelindent\relax
+
 % \usepackage{breqn}
+
 \usepackage{enumitem}
+
 \usepackage{fleqn}
+
 \usepackage{cite}
+
 \usepackage{graphicx}
+
 \usepackage[varg]{newtxmath}
+
 \graphicspath{ {images/} }
+
 \usepackage{pdfpages}
+
 \usepackage{wrapfig}
+
 \usepackage{fancyhdr}
+
 \usepackage{lastpage}
+
 \usepackage{lettrine}
+
 \usepackage{amsmath}
+
 % \usepackage{titlesec}
+
+\usepackage{mathtools}
+
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+
 \usepackage[colorinlistoftodos]{todonotes}
+
 \usepackage{float}
+
 \usepackage[font={footnotesize}]{caption}
+
 \usepackage[numbers,sort,square,compress]{natbib}
+
 \usepackage[para]{footmisc}
+
 \usepackage{xcolor}
+
 \setlength{\belowcaptionskip}{-12pt}
+
 \newcommand{\highlight}[1]{%
+
   \colorbox{red!50}{$\displaystyle#1$}}
@@
 % \usepackage{parskip}
+
 % \setlength{\parskip}{0.02\baselineskip}
@@
 % \fancypagestyle{plain}{
+
 %   \fancyhf{} % sets both header and footer to nothing
+
 % \renewcommand{\headrulewidth}{0pt}
+
 %   \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}% Right header
@@
 % }
+
 \pagestyle{plain}% Set page style to plain.
@@
 % \pagestyle{fancyplain}
+
 % \fancyhf{}
+
 % \renewcommand{\headrulewidth}{0pt}
+
 % \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}
+
 \DeclareMathOperator*{\argmin}{argmin}
+
 % \usepackage{titlesec}
@@
 % \titlespacing*{\section}{0pt}{1pt plus 2p}{1ex}
+
 % \titleformat*{\section}{\fontsize{12}{12}\bfseries}
+
 % \titleformat{\section}
+
 %        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries}{\thesection}{1em}{}
@@
 % \titleformat{\subsection}
+
 %        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries\itshape}{\thesubsection}{1em}{}
@@
 \begin{document}
+
 %Here goes the title
@@
 \author{\authorblockN{Author 1, Author 2, Author 3}
+
 \
+
 \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay}
+
 % \thanks{Parts of this work was supported by the Bharti Centre for Communication in
+
 % IIT Bombay, and the Visvesvaraya
+
 % PhD Scheme of Ministry of Electronics \& Information Technology,
+
 % Government of India (implemented by the Digital India Corporation).
+
 % }}
+
 }
+
 \maketitle
@@
 \thispagestyle{plain}
+
 %Main body starts
@@
 % \noindent We consider problem of quantization and interpolation of
+
 % time and frequency varying precoding matrices in wireless MIMO
+
 \begin{abstract}
-In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information(CSI).
+
+In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information (CSI).
 The receiver knows the channel information, and it feeds back to the transmitter.
 However, it is not possible to feedback complete information for a limited-feedback OFDM channel since the channel information for multiple subcarriers consumes a large amount of data.
-Therefore to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to a minimal number of scalar parameters.
+Therefore, to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to a minimal number of scalar parameters.
 The Precoder matrix is decomposed using Givens Rotation.
@@
 We also propose a method for joint time-frequency interpolation of the channel information at the transmitter.
-153-words
+
 \end{abstract}
@@
 Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
-To this end, feeding back the CSI from the receiver to the transmitter efficiently is necessary.
-In wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across all subcarriers, and this places a large feedback demand.
-However, since the CSI varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency domain interpolation temporal prediction.
-In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to achieve link capacity, as well as to parallelize the channel.
+To this end, feeding back the channel state information from the receiver to the transmitter efficiently is necessary.
+In the many wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across the many subcarriers of the system, which places a large feedback demand.
+However, since the channel state information varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency interpolation temporal prediction.
+In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to enable the transmitter to transmit at capacity as well as to parallelize the channel.
 In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
 
-The problem of effective quantization and feedback of precoders for MIMO-OFDM has attracted significant attention in recent years.
-Many approaches involve codebook design over the Grassmann manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308} or the Stiefel Manifold \cite{6891198,Gupt1905:Predictive}, wherein the manifold structure of the precoders is used to obtain tangents for predictive quantization and interpolation across time and frequency.
-While these methods have been shown to be very effective for precoder feedback,  they involve operations over higher dimensional manifolds.
-In particular, optimization and interpolation is complicated, especially for manifolds where geodesic paths for interpolation may be difficult to obtain, especially for joint time-frequency predictive coding.
-An alternate approach is to parametrize the precoding matrices using scalar parameters.
-The space of unitary matrices parameterized into independent scalar parameters that correspond to Givens rotations and Household Transformations of the unitary precoding matrices \cite{4114278,4556174}.
-These approaches have been combined with adaptive delta modulation effectively track precoding matrices both in the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit separately.
+% TODO More coverage of prior work needed.
 
-In this paper, we propose a method for exploiting the joint time-frequency correlation of the precoder's scalar parameters on the transmitter using CSI feedback.
-Typical approaches to predictive quantization and interpolation do not exploit the joint time-frequency correlation to enhance the precoder estimate at the transmitter.
-However, exploiting the correlations jointly not only reduces the feedback requirement, it also is able to enhance the precoder reconstruction accuracy.
-Further, the independence of the scalar parameters reduces the problem to interpolation of separate scalar valued function, as opposed to operations over tangent spaces in manifolds~\cite{Gupt1905:Predictive}.
-Simulations reveal that the proposed method offers matches the performance of manifold based approaches with a much lower implementation complexity and fewer number of bits.
+The problem of effective quantization and feedback of precoders for MIMO-OFDM attracted significant attention In recent years.
+Many approaches involve codebook design over the Stiefel Manifold \cite{6891198}, wherein the manifold structure of the precoders is used to obtain tangents for predictive quantization and interpolation across time and frequency.
+Another such variation of \cite{6891198} has exploited both time and frequency domain together in Stiefel Manifold \cite{Gupt1905:Predictive}.
+All these methods are useful, but they do so over Manifolds which has problems of getting stuck in local Maxima.
+Also, there are different challenged which are faced while interpolating in a high dimensional space.
 
-The rest of this paper is organized as follows: Section~\ref{section2} described the system model and the precoder feedback and interpolation techniques.
-Section~\ref{section3} compares the BER and achievable rates for our proposed methods with that of past work.
-Finally, Section~\ref{section4} concludes with discussions on future work.
+Other approaches have considered the problem of parametrizing precoding matrices using scalar parameters.
+It has been shown that the space of unitary matrices can be effectively parameterized using scalar parameters that correspond to Givens rotations and Household Transformations of the unitary precoding matrices \cite{4114278,4556174}.
+By tracking these scalar parameters using adaptive digital pulse code modulation (ADPCM) for slowly varying channels in the time domain and the frequency domain, the link can operate close to capacity while using very few bits for quantization and feedback.
+
+We now propose a method for predictive quantization in time domain over few subcarriers using scalar parameters.
+We also propose a method for exploiting the joint time-frequency correlation of those scalar parameters on the transmitter side.
+In \ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
+In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
 
@@
 \begin{equation} 
-\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
+y_{i,t} = H_{i,t} x_{i,t}+ \eta 
 \end{equation}
@@
 
-where the channel matrix $\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $\bx_{i,t} \in \mathbb{C}^{N_T}$; the output $\by_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
-On applying thin SVD to H, we get $\bH_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_R}$, $\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_R}$.
+where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
+On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$.
 $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H.
@@
 
-% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).\cos(theta) 0 $\leq$ Theta $\leq$ pi/2.
+% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).cos(theta) 0 $\leq$ Theta $\leq$ pi/2. 
 % Householders' transformation and how we are decomposing. -pi$\leq$ phi $\leq$ pi. Uniformly distributed. 
@@
 I_{m-2} & & & \\ 
-& \cos\;\theta & - \sin\;\theta & \\
-& \sin\;\theta & \cos\;\theta & \\
+& cos\;\theta & - sin\;\theta & \\ 
+& sin\;\theta & cos\;\theta & \\ 
 & & & I_{t-m}
@@
 \begin{figure} 
-\label{adpm-fig}
-%\includegraphics[width=0.45\textwidth]{images/new-adpm-crop.pdf}
+\includegraphics[width=0.5\textwidth]{images/adpm.pdf} 
 \caption{Subcarriers which need to be quantized} 
@@
 \begin{equation} 
-p(\theta) = 2l\sin(\theta)^{2l-1}\cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
+p(\theta) = 2lsin(\theta)^{2l-1}cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
 \end{equation}
@@
 \begin{equation} 
-\label{delta_eqn}
 \Delta_{n} = \begin{cases} 
@@
 
-We need to quantize the subcarriers in a way such that we can exploit
-the joint time-frequency interpolation at the transmitter
-side. Therefore, we select the subcarriers for quantization in an
-alternating fashion as shown in the figure. Figure shows a \textbf{N}
-subcarrier MIMO channel with subcarriers which are going to be
-quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even
-time instance. Here, s is the gap between two frequencies. For odd
-time instances, subcarriers at the position sk + t for k = 0,1,...,
-$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$.
+We need to quantize the subcarriers in a way such that we can exploit the joint time-frequency interpolation at the transmitter side.
+Therefore, we select the subcarriers for quantization in an alternating fashion as shown in the figure.
+Figure shows a \textbf{N} subcarrier MIMO channel with subcarriers which are going to be quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even time instance.
+Here, s is the gap between two frequencies.
+For odd time instances, subcarriers at the position sk + t for k = 0,1,..., $\frac{N-1}{s}$ with t = $\floor*{\frac{s}{2}}$.
 
@@
 
-One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $\sin/\cos(\pi) =\sin/\cos(-\pi)$.
+One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $sin/cos(\pi) = sin/cos(-\pi)$.
 Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).
@@
 Therefore, the number of bits for each subcarrier = $10\times12$ = 120.
-Quantization is performed as given in Section~\ref{section2}.
+Quantization is performed as given in \ref{section2}.
 
@@
 Later, the quantization will take place in the time-domain using one bit for each parameter.
-Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{adpm-fig}.
-Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s.
-This brings down the bit budget to average of 5(Size of $\Theta$) and 7(Size of $\Phi$) i.e.
-6 bits average per fed back subcarrier.
+Since the bit budget is low and we are considering slowly varying channels, therefore, we will be sending either theta or phi parameters alternatingly in the time domain while keeping the unsent parameters equal to their previous values.
+This brings down the bit budget to 6 bits average per fed back subcarrier.
 
-\bf{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
-This ratio was decided after trying with different ratios.
-This implies that if quantied values loose track of the actual values then after changing delta quickly they start following it again.
-Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare
+Describing how you performed ADPCM to carry out the 1-bit quantization.
+Also, the increment rate was 2.4 times the decay rate.
+Also, need to describe the initialization of the Delta values required in the quantization process.
+I need to cite that paper.
+We should also try over different quantization rates to compare \cite{Gupt1905:Predictive}
 
@@
 \end{document}
+
+% Remove the first para of FSM. State what r1,r2,d,\theta are in the FSM model states themselves. As a concluding statement of STAT-2 state that using the
+
+% inherit geometery of the problem, we get AoA

### 2019-10-06 ber_results/plot_results.py

@@ -3,15 +3,22 @@ import matplotlib.pyplot as plt
 
 eb_n0=np.arange(-6,20,3)
 
-hpang_ber=np.load('eval/hp_ber.npy')
-hpmani_ber=np.load('manifold_interp/hp_ber.npy')
-hpang_ber6=np.load('6bit_eval/hp_ber.npy')
+# cubic=np.load('6bit_eval/hp_ber.npy')
+# linear=np.load('6bit_eval/linear/hp_ber.npy')
+# nearest=np.load('6bit_eval/nearest/hp_ber.npy')
+vehicular_mani=np.load('vehicular_manifold/hp_ber.npy')
+vehicular_scal=np.load('vehicular/hp_ber.npy')
 plt.yscale('log')
 plt.ylabel('BER')
 plt.xlabel(r'$E_{b}/N_{o}(dB)$')
+plt.title('Pedestrian Channel')
+plt.plot(eb_n0,vehicular_scal, label='Vehicular Scalar Paramesters', marker='+')
+plt.plot(eb_n0, vehicular_mani, label='Vehicular Manifold', marker='o')
 # plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
-plt.plot(eb_n0, hpang_ber6, label='Pedestrian Manifold', marker='o')
-plt.plot(eb_n0,hpmani_ber,label='Pedestrian Scalar Parameters', marker='+')
+# plt.plot(eb_n0, cubic, label='Cubic', marker='o', color='b')
+# plt.plot(eb_n0,linear, label='Linear', marker='+', color='g')
+# plt.plot(eb_n0,nearest, label='nearest', marker='1', color='r')
+
 plt.grid()
 plt.legend()
 plt.show()

### 2019-10-06 eval.py

@@ -101,17 +101,17 @@ def main():
     gridz_prev=np.zeros((6,num_subcarriers,2*Nt*Nr-Nr**2))
     for chan_index in range(num_chan):
 
-        tH_allH=np.load('Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU=np.load('Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/th_allU_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU_vec=np.load('Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/th_allU_vec'+str(chan_index+chan_offset)+'.npy')        
-        tHS=np.load('Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/thS_'+str(chan_index+chan_offset)+'.npy')
-        allU=np.load('./Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
-        allU_vec_copy=np.load('./Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
-        onlyt_allU=np.load('./Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/onlyt_allU_'+str(chan_index+chan_offset)+'.npy')
-        allU_vec=np.load('./Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
-        onlyt_allU_vec=np.load('./Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/onlyt_allU_vec'+str(chan_index+chan_offset)+'.npy')
-        # onlyt_allU=np.load('Precoders_generated/Vehicularfast/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
-        interpS=np.load('Precoders_generated/6bit_Vehicularfast/'+str(fdts)+'/interpS_'+str(chan_index+chan_offset)+'.npy')
+        tH_allH=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
+        tH_allU=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/th_allU_'+str(chan_index+chan_offset)+'.npy')
[…]

### 2019-10-09 paper/main.tex
modified adpm figure and added changes to main.tex

@@
 \begin{equation} 
-y_{i,t} = H_{i,t} x_{i,t}+ \eta 
+\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}} 
 \end{equation}
@@
 
-where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
-On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$.
+where the channel matrix $\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $\bx_{i,t} \in \mathbb{C}^{N_T}$; the output $\by_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
+On applying thin SVD to H, we get $\bH_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_R}$, $\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_R}$.
 $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H.
@@
 
-% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).cos(theta) 0 $\leq$ Theta $\leq$ pi/2. 
+% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).\cos(theta) 0 $\leq$ Theta $\leq$ pi/2. 
 % Householders' transformation and how we are decomposing. -pi$\leq$ phi $\leq$ pi. Uniformly distributed. 
@@
 I_{m-2} & & & \\ 
-& cos\;\theta & - sin\;\theta & \\ 
-& sin\;\theta & cos\;\theta & \\ 
+& \cos\;\theta & - \sin\;\theta & \\ 
+& \sin\;\theta & \cos\;\theta & \\ 
 & & & I_{t-m}
@@
 \begin{figure} 
-\includegraphics[width=0.5\textwidth]{images/adpm.pdf} 
+\includegraphics[width=0.45\textwidth]{images/new-adpm-crop.pdf} 
 \caption{Subcarriers which need to be quantized} 
@@
 \begin{equation} 
-p(\theta) = 2lsin(\theta)^{2l-1}cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
+p(\theta) = 2l\sin(\theta)^{2l-1}\cos(\theta), \theta \in [0, \frac{\pi}{2}), for \; l = 1,2,\ldots,t-1
 \end{equation}
@@
 
-One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $sin/cos(\pi) = sin/cos(-\pi)$.
+One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $\sin/\cos(\pi) =\sin/\cos(-\pi)$.
 Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).
@@
 
-Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because
+Now here we can see the advantage of using subcarriers in an alternating fashion as given in the Section~\ref{section1}. With this method we could exploit both time and frequency correlation because
 
@@
 Therefore, the number of bits for each subcarrier = $10\times12$ = 120.
-Quantization is performed as given in \ref{section2}.
+Quantization is performed as given in Section~\ref{section2}.
 

### 2019-10-09 eval.py

@@ -43,6 +43,7 @@ def main():
     # Eb_N0_dB=np.arange(-6,6,2)
     Eb_N0_dB=np.arange(-6,20,3)
     #Store BER here
+    idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
     hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
     otBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
     Ts=constants.Ts # 5e-8
@@ -110,99 +111,104 @@ def main():
         onlyt_allU=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/onlyt_allU_'+str(chan_index+chan_offset)+'.npy')
         allU_vec=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
         onlyt_allU_vec=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/onlyt_allU_vec'+str(chan_index+chan_offset)+'.npy')
-        # onlyt_allU=np.load('Precoders_generated/Pedestrian/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
+        # onlyt_allU=np.load('Precoders_generated/Vehicularfast/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
         interpS=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/interpS_'+str(chan_index+chan_offset)+'.npy')
         for simulation_index in range(number_simulations-6):
         # for simulation_index in range(0,number_simulations-interp_nbr_vals-2):
-            print("---------------------------------------------------------------------------")
-            print ("Starting Eval sim: "+str(simulation_index)+" : of "+ str(number_simulations) + " # of total simulations")
-            print(str(time.strftime("Elapsed Time %H:%M:%S",time.gmtime(time.time()-start_time)))\
-                +str(time.strftime(" Current Time %H:%M:%S",time.localtime(time.time()))))
-            curr_fb_indices=np.arange(feedback_mats)*2 if simulation_index%2==0 else np.arange(feedback_mats-1)*2+1
-            hp_qterr=np.mean([stiefCD(tH_allU[simulation_index][net_fb_indices[i]],allU[simulation_index][net_fb_indices[i]])\
[…]

### 2019-10-09 main.py

@@ -73,7 +73,7 @@ def main():
     # Use Stiefel Chordal Distance as norm
     # norm_fn='stiefCD'
     start_time=time.time()
-    save=True
+    save=False
 
     for chan_index in range(num_chan_realisations):
         print("-----------------------------------------------------------------------")

### 2019-10-09 mimo_tdl_channel.py

@@ -792,7 +792,8 @@ class MIMO_TDL_Channel():
             self.tdl_channels[i].calc_frequency_response(inp_resp,freq_resp , 2*self.fft_size)
             # Store it in chan_freq in appropriate places across 64 matrices
             # print(freq_resp.to_numpy_ndarray().flatten()[0:64]-freq_resp.to_numpy_ndarray().flatten()[64:128])
-            chan_freq[:,row_idx][:,col_idx]=freq_resp.to_numpy_ndarray().flatten()[0:self.fft_size]
+            np_array = np.asarray(freq_resp, dtype=np.ndarray)
+            chan_freq[:,row_idx][:,col_idx]=np.ndarray((freq_resp.rows(), freq_resp.cols()), dtype=np.complex128).flatten()[0:self.fft_size]
         
         chan_freq_list=[]
         for i in range(self.fft_size):

### 2019-10-09 paper/main.tex
modified

@@
 \begin{figure}
+\label{adpm-fig}
 \includegraphics[width=0.45\textwidth]{images/new-adpm-crop.pdf} 
@@
 \begin{equation}
+\label{delta_eqn}
 \Delta_{n} = \begin{cases} 
@@
 
-Now here we can see the advantage of using subcarriers in an alternating fashion as given in the Section~\ref{section1}. With this method we could exploit both time and frequency correlation because
+Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because
 
@@
 Later, the quantization will take place in the time-domain using one bit for each parameter.
-Since the bit budget is low and we are considering slowly varying channels, therefore, we will be sending either theta or phi parameters alternatingly in the time domain while keeping the unsent parameters equal to their previous values.
-This brings down the bit budget to 6 bits average per fed back subcarrier.
+Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{adpm-fig}.
+Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s.
+This brings down the bit budget to average of 5(Size of $\Theta$) and 7(Size of $\Phi$) i.e.
+6 bits average per fed back subcarrier.
 
-Describing how you performed ADPCM to carry out the 1-bit quantization.
-Also, the increment rate was 2.4 times the decay rate.
-Also, need to describe the initialization of the Delta values required in the quantization process.
-I need to cite that paper.
-We should also try over different quantization rates to compare \cite{Gupt1905:Predictive}
+\bf{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
+This ratio was decided after trying with different ratios.
+This implies that if quantied values loose track of the actual values then after changing delta quickly they start following it again.
+Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare
 

### 2019-10-09 paper/main.tex
Fix images etc.

@@
-%% This is a skeleton file demonstrating the use of IEEEtran.cls (requires IEEEtran.cls version 1.8a or later) with an IEEE conference paper.
-
-%%
-
-%% Modified by Khan Reaz( kahn.reaz@ieee.org)
-
-%% Support sites:
-
-%% http://www.ieee.org/
-
-%%***********************************************************
-
-%% Legal Notice:
-
-%% This code is offered as-is without any warranty either expressed or implied; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE!
-
-%% User assumes all risk and can modify as s/he wants.
-
-%%***********************************************************
-
 \def\b0{{\bf 0}}
-
 \def\ba{{\bf a}}
-
 \def\bb{{\bf b}}
-
 \def\bc{{\bf c}}
-
 \def\bd{{\bf d}}
-
 \def\be{{\bf e}}
-
 \def\bg{{\bf g}}
-
 \def\bh{{\bf h}}
-
 \def\bi{{\bf i}}
-
 \def\bj{{\bf j}}
-
 \def\bk{{\bf k}}
-
 \def\bl{{\bf l}}
-
 \def\bm{{\bf m}}
-
 \def\bn{{\bf n}}
-
 \def\bo{{\bf o}}
-
 \def\bp{{\bf p}}
-
 \def\bq{{\bf q}}
-
 \def\br{{\bf r}}
-
 \def\bs{{\bf s}}
-
 \def\bt{{\bf t}}
-
 \def\bu{{\bf u}}
-
 \def\bv{{\bf v}}
-
 \def\bw{{\bf w}}
-
 \def\bx{{\bf x}}
-
 \def\by{{\bf y}}
-
 \def\bz{{\bf z}}
@@
 \def\bSigma{{\bf \Sigma}}
-
 \def\bA{{\bf A}}
-
 \def\bB{{\bf B}}
-
 \def\bC{{\bf C}}
-
 \def\bD{{\bf D}}
-
 \def\bE{{\bf E}}
-
 \def\bF{{\bf F}}
-
 \def\bG{{\bf G}}
-
 \def\bH{{\bf H}}
-
 \def\bI{{\bf I}}
-
 \def\bJ{{\bf J}}
-
 \def\bK{{\bf K}}
-
 \def\bL{{\bf L}}
-
 \def\bM{{\bf M}}
-
 \def\bN{{\bf N}}
-
 \def\bO{{\bf O}}
-
 \def\bP{{\bf P}}
-
 \def\bQ{{\bf Q}}
-
 \def\bR{{\bf R}}
-
 \def\bS{{\bf S}}
-
 \def\bT{{\bf T}}
-
 \def\bU{{\bf U}}
-
 \def\bV{{\bf V}}
-
 \def\bW{{\bf W}}
-
 \def\bX{{\bf X}}
-
 \def\bY{{\bf Y}}
-
 \def\bZ{{\bf Z}}
@@
 %package list
-
 \documentclass[conference]{IEEEtran}
-
 \IEEEoverridecommandlockouts
-
 \let\labelindent\relax
-
 % \usepackage{breqn}
-
 \usepackage{enumitem}
-
 \usepackage{fleqn}
-
 \usepackage{cite}
-
 \usepackage{graphicx}
-
 \usepackage[varg]{newtxmath}
-
 \graphicspath{ {images/} }
-
 \usepackage{pdfpages}
-
 \usepackage{wrapfig}
-
 \usepackage{fancyhdr}
-
 \usepackage{lastpage}
-
 \usepackage{lettrine}
-
 \usepackage{amsmath}
-
 % \usepackage{titlesec}
-
-\usepackage{mathtools}
-
-\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
-
-\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
-
 \usepackage[colorinlistoftodos]{todonotes}
-
 \usepackage{float}
-
 \usepackage[font={footnotesize}]{caption}
-
 \usepackage[numbers,sort,square,compress]{natbib}
-
 \usepackage[para]{footmisc}
-
 \usepackage{xcolor}
-
 \setlength{\belowcaptionskip}{-12pt}
-
 \newcommand{\highlight}[1]{%
-
   \colorbox{red!50}{$\displaystyle#1$}}
@@
 % \usepackage{parskip}
-
 % \setlength{\parskip}{0.02\baselineskip}
@@
 % \fancypagestyle{plain}{
-
 %   \fancyhf{} % sets both header and footer to nothing
-
 % \renewcommand{\headrulewidth}{0pt}
-
 %   \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}% Right header
@@
 % }
-
 \pagestyle{plain}% Set page style to plain.
@@
 % \pagestyle{fancyplain}
-
 % \fancyhf{}
-
 % \renewcommand{\headrulewidth}{0pt}
-
 % \fancyhead[C]{2018 International Conference on Indoor Positioning and Indoor Navigation (IPIN), 24-27 September 2018, Nantes, France}
-
 \DeclareMathOperator*{\argmin}{argmin}
-
 % \usepackage{titlesec}
@@
 % \titlespacing*{\section}{0pt}{1pt plus 2p}{1ex}
-
 % \titleformat*{\section}{\fontsize{12}{12}\bfseries}
-
 % \titleformat{\section}
-
 %        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries}{\thesection}{1em}{}
@@
 % \titleformat{\subsection}
-
 %        {\normalfont\fontfamily{phv}\fontsize{12}{17}\bfseries\itshape}{\thesubsection}{1em}{}
@@
 \begin{document}
-
 %Here goes the title
@@
  \author{\authorblockN{Author 1, Author 2, Author 3}
-
  \
-
  \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay}
-
 % \thanks{Parts of this work was supported by the Bharti Centre for Communication in
-
 % IIT Bombay, and the Visvesvaraya
-
 % PhD Scheme of Ministry of Electronics \& Information Technology,
-
 % Government of India (implemented by the Digital India Corporation).
-
 % }}
-
 }
-
 \maketitle
@@
 \thispagestyle{plain}
-
 %Main body starts
@@
 % \noindent We consider problem of quantization and interpolation of
-
 % time and frequency varying precoding matrices in wireless MIMO
-
 \begin{abstract}
-
-In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information (CSI).
+In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information(CSI).
 The receiver knows the channel information, and it feeds back to the transmitter.
 However, it is not possible to feedback complete information for a limited-feedback OFDM channel since the channel information for multiple subcarriers consumes a large amount of data.
-Therefore, to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to a minimal number of scalar parameters.
+Therefore to resolve this problem, the orthonormality structure of the precoder matrix is exploited to make its one-to-one mapping to a minimal number of scalar parameters.
 The Precoder matrix is decomposed using Givens Rotation.
@@
 We also propose a method for joint time-frequency interpolation of the channel information at the transmitter.
-
+153-words
 \end{abstract}
@@
 
-\subsection{Scalar Parameterization} 
+\subsection{Scalar Paramatrization}
 \label{givens}
-
-The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix \cite{4114278}.
+The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix~\cite{4114278}.
 An orthonormal matrix with $t$ rows and $n$ columns can be represented as
@@
 I_{m-2} & & & \\
-& \cos\;\theta & - \sin\;\theta & \\ 
-& \sin\;\theta & \cos\;\theta & \\ 
+& \cos\theta & -\sin\theta & \\
+& \sin\theta & \cos\theta & \\
 & & & I_{t-m}
@@
 \label{adpm-fig}
-\includegraphics[width=0.45\textwidth]{images/new-adpm-crop.pdf} 
+\includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
 \caption{Subcarriers which need to be quantized}
@@
 
-We need to quantize the subcarriers in a way such that we can exploit the joint time-frequency interpolation at the transmitter side.
-Therefore, we select the subcarriers for quantization in an alternating fashion as shown in the figure.
-Figure shows a \textbf{N} subcarrier MIMO channel with subcarriers which are going to be quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even time instance.
-Here, s is the gap between two frequencies.
-For odd time instances, subcarriers at the position sk + t for k = 0,1,..., $\frac{N-1}{s}$ with t = $\floor*{\frac{s}{2}}$.
+We need to quantize the subcarriers in a way such that we can exploit
+the joint time-frequency interpolation at the transmitter
+side. Therefore, we select the subcarriers for quantization in an
+alternating fashion as shown in the figure. Figure shows a \textbf{N}
+subcarrier MIMO channel with subcarriers which are going to be
+quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even
+time instance. Here, s is the gap between two frequencies. For odd
+time instances, subcarriers at the position sk + t for k = 0,1,...,
+$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$.
 
@@
 %Explain prediction and quantization steps here itself ! (succintly)
-
-\section{SIMULATION AND DISCUSSION} 
+\section{Simulation and discussion}
 \label{section3}

### 2019-10-09 paper/main.tex
Add back lost text
Add more changes

@@
 \section{System Model}
-
+For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver
+antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel can be modelled as
 \label{section2}
@@
 \end{equation}
-where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$.
-On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$.
-$U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H.
-The entries of H are assumed to be i.i.d.
-$\tilde{\mathcal{N}}(0,1)$, complex circularly symmetric Gaussian with mean zero and variance 1.
-Therefore, the first n column vectors of $V_{i,t}$ have to be quantized and fed back to the transmitter as channel spatial information.
+where the channel matrix is
+$\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input is
+$\bx_{i,t} \in \mathbb{C}^{N_T}$; the output is
+$\by_{i,t} \in \mathbb{C}^{N_R}$ and the complex additive white
+Gaussian noise is $\eta$ whose distribution is
+${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R <
+N_T$. The thin SVD of $\bH$ is given by
+$\bH_{i,t} = \bU_{i,t} \Sigma_{i,t} \bV_{i,t}^{H}$, where
+$\bU_{i,t} \in \mathbb{C}^{N_R \times N_R}$,
+$\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$ and
+$\bV_{i,t} \in \mathbb{C}^{N_T \times N_R}$. $\bU_{i,t}$ and
+$\bV_{i,t}$ are matrices whose columns are orthonormal, and
+$\Sigma_{i,t}$ contains the singular values
+$\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of $\bH_{i,t}$. The
+entries of $\bH_{i,t}$ are assumed to be distributed
+i.i.d. $\tilde{\mathcal{N}}(0,1)$. To obtain a precoder at the
+transmitter, we need to quantize and feed back information about
+$\bV_{i,t}$ for each subcarrier.
 
@@
 
-\subsection{Scalar Paramatrization}
+\subsection{Scalar Parametrization}
 \label{givens}
-The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix~\cite{4114278}.
-An orthonormal matrix with $t$ rows and $n$ columns can be represented as
-
+The degrees of freedom in the $\bV_{i,t}$ matrices is smaller than the
+number of real entries in the matrix because of the geometrical
+structure between the columns of the matrix~\cite{4114278}. The
+orthonormal matrix $\bV_{i,t}$ with $N_T$ rows and $N_R$ columns can be
+decomposed as TODO
 \begin{equation}
-\textbf{V} = \Bigg[\prod_{k=1}^{n} D_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} G_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{I}
+\textbf{V} = \Bigg[\prod_{k=1}^{n} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} \bG_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
 \end{equation}
-
-where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k}, \ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( 1_{k-1}, e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, $1_{k-1}$ is (k-1) ones; and t x n matrix $\tilde{I}$ = $\big[I_{n}, 0_{n,t-n}\big]^{T}$.
-The Givens matrix \textbf{G} is given as
-
+where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
+\ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
+e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, ${\bf 1}_{k-1}$ is
+represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
+= $\big[\bI_{n}, {\bf 0}_{n,t-n}\big]^{T}$, and
 \begin{equation}
@@
 \begin{bmatrix}
-
 I_{m-2} & & & \\
@@
 & & & I_{t-m}
-
 \end{bmatrix}
 \end{equation}
-
+Thus, for a $4 \times 2$ orthogonal matrix $\bV_{i,t}$, we have
+\begin{align*}
+  \bV_{i,t} & =
+  \bD_{1}(\phi_{1,1},\ldots,\phi_{1,4})\bG_{3,4}(\theta_{1,1})
+  \bG_{2,3}(\theta_{1,2}) \bG_{1,2}(\theta_{1,3})\times\\
+& \bD_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) \bG_{3,4}(\theta_{2,1}) \bG_{2,3}(\theta_{2,2})\tilde{\bI}
+\end{align*}
+Here, $\phi_{k}$s are referred to as phases and $\theta_{l}$s as
+rotation angles. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
+between the two extremes\cite{4114278}, while $\theta_l$
+is distributed as
+\begin{equation}
+p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
+\end{equation}
+The $\phi_k$ and $\theta_l$ are statistically independent, making this
+representation useful for tracking them separately across time, as
+explained in the following subsection. The total number of parameters
+obtained from the decomposition of a complex orthogonal matrix
+$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The
+number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
+number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
 \begin{figure}
-\label{adpm-fig}
 \includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
-\caption{Subcarriers which need to be quantized}
+\caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
 \label{ber_overvie}
@@
 \end{figure}
-
-For a $4 \times 2$ orthogonal matrix V,
-
-$$V = D_{1}(\phi_{1,1},\ldots,\phi_{1,4}).G_{3,4}(\theta_{1,1}) .G_{2,3}(\theta_{1,2}) .G_{1,2}(\theta_{1,3})$$
-
-\vspace{-1.4em}
-
-\hspace{1pt}$$.D_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) .G_{3,4}(\theta_{2,1}) .G_{2,3}(\theta_{2,2}).\tilde{I}$$
-
-Here, we denote $\phi_{i}'s$ by phases and $\theta_{i}'s$ by rotation angles.
-$\phi \in (-\pi, + \pi]$ is uniformaly distributed between the two extremes\cite{4114278}, while $\theta$ has a distribution
-
-%TODO What is t in the equation below?
-\begin{equation}
-p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
-\end{equation}
-Here all $\theta_i$ and $\phi_i$ are statistically independent, making it useful for tracking them across time as explained in the next subsection.
-The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}.
-The number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
-% may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.
-
 % To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.
@@
 \label{quantiz}
-
-For a slowly time varying channel, ADPM is an effective method for tracking a paramater over time with using only 1-bit.
-The 1-bit number $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$ where $x$ is the unquantized vector $\&$ $\hat{x}$ is the quantized vector and $Q(x)$ is the quantization function given as
-
+For a channel that varies slowly with time, ADPM is an effective
+method for tracking parameters over time using very few bits (often
+just one bit). For example, while tracking a process $x_n$, at time
+$n$, the one-bit jump $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$, where
+$x_n$ is the unquantized value and $\hat{x}_{n-1}$ is the unquantized
+(accurate) value, the quantization function $Q(x)$ is given as
 \begin{equation}
@@
 \end{equation}
-
-Thus, the quantied vector $\hat{x}_n$ changes with time as
-
+Thus, the quantied vector $\hat{x}_n$ evolves with time as
 \begin{equation}
-\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n}
-\end{equation}
-
-\begin{equation}
+\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n} \mbox{ where }
 \label{delta_eqn}
 \Delta_{n} = \begin{cases}
-    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}.\\
+    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
     \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
@@
 \end{equation}
+Here we initialize delta using $\Delta_1 = |x_{2}-\hat{x}_1|$.
 
-Here we initialize delta using $\Delta_1 = abs(x_{2}-\hat{x}_1)$.
-
-We need to quantize the subcarriers in a way such that we can exploit
-the joint time-frequency interpolation at the transmitter
-side. Therefore, we select the subcarriers for quantization in an
-alternating fashion as shown in the figure. Figure shows a \textbf{N}
-subcarrier MIMO channel with subcarriers which are going to be
-quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even
-time instance. Here, s is the gap between two frequencies. For odd
-time instances, subcarriers at the position sk + t for k = 0,1,...,
-$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$.
+To quantize the subcarriers efficiently and exploit the joint
+time-frequency interpolation at the transmitter side, we select the
+subcarriers for quantization in an alternating fashion as shown in
+Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
+$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
+quantized at every $s$-th subcarriers, for even time instances, where,
+$s$ is the gap between two quantized subcarrier indices. For odd time
+instances, subcarriers at the position $sk$ for k = 0,1,...,
+$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$. TODO
 
@@
 
-\bf{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
+{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
 This ratio was decided after trying with different ratios.
-This implies that if quantied values loose track of the actual values then after changing delta quickly they start following it again.
+This implies that if quantied values lose track of the actual values then after changing delta quickly they start following it again.
 Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare

### 2019-10-11 paper/main.tex
todo changes added

@@
 \begin{equation}
-\textbf{V} = \Bigg[\prod_{k=1}^{n} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} \bG_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
+\textbf{V} = \Bigg[\prod_{k=1}^{N_{R}} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,N_{R}} \big) \: \Bigg\{ \prod_{l=1}^{l=N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
 \end{equation}
 where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
-\ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
-e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, ${\bf 1}_{k-1}$ is
+\ldots, \phi_{k,N_T } \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
+e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,N_T }}  \big)$, ${\bf 1}_{k-1}$ is
 represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
-= $\big[\bI_{n}, {\bf 0}_{n,t-n}\big]^{T}$, and
+= $\big[\bI_{N_R }, {\bf 0}_{N_T ,N_T -N_R }\big]^{T}$, and
 \begin{equation}
@@
 \end{align*}
-Here, $\phi_{k}$s are referred to as phases and $\theta_{l}$s as
-rotation angles. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
+Here, $\phi_{k}$s are referred to as phases; their collection is given by $\Phi and $$\theta_{l}$s as rotation angles; their collection is given by $\Theta$.
+$\phi_k \in (-\pi, + \pi]$ is uniformly distributed
 between the two extremes\cite{4114278}, while $\theta_l$
@@
 \begin{equation}
-p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
+p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,N_T -1
 \end{equation}
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.5\textwidth]{images/new-adpm.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
@@
 $N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
-quantized at every $s$-th subcarriers, for even time instances, where,
-$s$ is the gap between two quantized subcarrier indices. For odd time
-instances, subcarriers at the position $sk$ for k = 0,1,...,
-$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$. TODO
+quantized at every $p$-th subcarriers, for even time instances, where,
+$p$ is the gap between two quantized subcarrier indices. For odd time
+instances, subcarriers at the position $pk+q$ for k = 0,1,...,
+$\frac{N-1}{p}$ with q = ${\frac{p}{2}}$. TODO
 
@@
 
-After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices sk or sk+t.
+After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices $pk$ or $pk+q$.
 To reproduce the rest of the subcarriers will use future time instances to do the interpolation.
@@
 Performance is measured by averaging over 10 channel realizations with 100 simulations each.
-The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with sampling frequency $T_s = 20MHz$.
+The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$.
 For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$.
@@
 
-{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} .
-This ratio was decided after trying with different ratios.
-This implies that if quantied values lose track of the actual values then after changing delta quickly they start following it again.
-Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare
+{ADPM Design}: The decrement/decay rate is given by $\bm$ and increment/growth rate is given by $\bM$ in the Equation~\ref{delta_eqn} .
+The ratio of $m/M$ is fixed at 2.5 i.e.
+higher decay than increment of the delta values.
+This way, quantized values are able to follow the curve more precisely.
+Also, need to describe the initialization of the Delta values required in the quantization process.
+Need to cite that paper.
 
-While doing interpolation at the transmitter end, future time instances are used to interpolate the rest of the subcarriers indices which are not fed back.
-To prevent much delay, we are using 6 future time instances and using SciPy cubic interpolation over each parameter in 2D.
+\subsection{Bilinear Interpolation}
+Bilinear interpolation is the equivalent of linear interpolation in two dimensions.
+On the transmitter side, all the scalars - $\Theta$ and $\Phi$ of the missing subcarriers are interpolatted using the values of the neigbours.
 These interpolated values and then used to reconstruct the precoding matrices all together.

### 2019-10-11 ber_results/plot_results.py

@@ -1,25 +1,51 @@
 import numpy as np 
 import matplotlib.pyplot as plt 
+from matplotlib import rc
+from matplotlib.pyplot import figure
 
+# font = {'size'   : 22}
+# rc('font', **font)
+# rc('text', usetex=True)
+# rc('xtick', labelsize=40)
+# rc('ytick', labelsize=40)
+plt.figure(figsize=(8, 6))
 eb_n0=np.arange(-6,20,3)
 
 # cubic=np.load('6bit_eval/hp_ber.npy')
-# linear=np.load('6bit_eval/linear/hp_ber.npy')
+linear17=np.load('6bit_eval/0.00017-linear/hp_ber.npy')
+linear1=np.load('6bit_eval/0.0001-linear/hp_ber.npy')
+linear=np.load('6bit_eval/linear/hp_ber.npy')
+
+linear17i=np.load('6bit_eval/0.00017-linear/ideal_ber.npy')
+linear1i=np.load('6bit_eval/0.0001-linear/ideal_ber.npy')
+lineari=np.load('6bit_eval/ideal_ber.npy')
+
 # nearest=np.load('6bit_eval/nearest/hp_ber.npy')
-vehicular_mani=np.load('vehicular_manifold/hp_ber.npy')
-vehicular_scal=np.load('vehicular/hp_ber.npy')
+# vehicular_mani=np.load('vehicular_manifold/hp_ber.npy')
+# vehicular_scal=np.load('vehicular/hp_ber.npy')
 plt.yscale('log')
-plt.ylabel('BER')
-plt.xlabel(r'$E_{b}/N_{o}(dB)$')
-plt.title('Pedestrian Channel')
-plt.plot(eb_n0,vehicular_scal, label='Vehicular Scalar Paramesters', marker='+')
-plt.plot(eb_n0, vehicular_mani, label='Vehicular Manifold', marker='o')
+plt.ylabel('BER', fontsize=14)
+plt.xlabel(r'$E_{b}/N_{o}(dB)$', fontsize=14)
+plt.title('Pedestrian Channel', fontsize=14)
+# plt.plot(eb_n0,vehicular_scal, label='Vehicular Scalar Paramesters', marker='+')
+# plt.plot(eb_n0, vehicular_mani, label='Vehicular Manifold', marker='o')
 # plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
[…]

### 2019-10-11 eval.py

@@ -52,6 +52,7 @@ def main():
     Nr=constants.Nr #2
     freq_quanta=constants.freq_quanta #4 - 5
     time_quanta=constants.time_quanta #1
+    channel_type="vehicular/"
     #---------------------------------------------------------------------------
     # Run Time variables
     # Stores the previously known estimates of the subcarriers
@@ -102,113 +103,112 @@ def main():
     gridz_prev=np.zeros((6,num_subcarriers,2*Nt*Nr-Nr**2))
     for chan_index in range(num_chan):
 
-        tH_allH=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/th_allU_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU_vec=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/th_allU_vec'+str(chan_index+chan_offset)+'.npy')        
-        tHS=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/thS_'+str(chan_index+chan_offset)+'.npy')
-        allU=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
-        allU_vec_copy=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
-        onlyt_allU=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/onlyt_allU_'+str(chan_index+chan_offset)+'.npy')
-        allU_vec=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
-        onlyt_allU_vec=np.load('./Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/onlyt_allU_vec'+str(chan_index+chan_offset)+'.npy')
-        # onlyt_allU=np.load('Precoders_generated/Vehicularfast/'+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
-        interpS=np.load('Precoders_generated/6bit_Pedestrian/'+str(fdts)+'/interpS_'+str(chan_index+chan_offset)+'.npy')
[…]

### 2019-10-11 main.py

@@ -73,7 +73,7 @@ def main():
     # Use Stiefel Chordal Distance as norm
     # norm_fn='stiefCD'
     start_time=time.time()
-    save=False
+    save=True
 
     for chan_index in range(num_chan_realisations):
         print("-----------------------------------------------------------------------")

### 2019-10-11 mimo_tdl_channel.py

@@ -685,7 +685,7 @@ def stiefCD(A,B):
     CD_2=np.sum([grassCD_2(np.array(A_mat[:,i]),np.array(B_mat[:,i])) for i in range(A_mat.shape[1])])
     if(CD_2<0):
         pdb.set_trace()
-        return -1
+        return 0
     return np.sqrt(CD_2)
 
 
@@ -793,12 +793,12 @@ class MIMO_TDL_Channel():
             # Store it in chan_freq in appropriate places across 64 matrices
             # print(freq_resp.to_numpy_ndarray().flatten()[0:64]-freq_resp.to_numpy_ndarray().flatten()[64:128])
             np_array = np.asarray(freq_resp, dtype=np.ndarray)
-            chan_freq[:,row_idx][:,col_idx]=np.ndarray((freq_resp.rows(), freq_resp.cols()), dtype=np.complex128).flatten()[0:self.fft_size]
-        
+            chan_freq[:,row_idx][:,col_idx]=freq_resp.to_numpy_ndarray().flatten()[0:self.fft_size]
         chan_freq_list=[]
         for i in range(self.fft_size):
             chan_freq_list.append(chan_freq[i])
         #print(chan_freq_list[0])
+        # pdb.set_trace()
         return(chan_freq_list)
 
 

### 2019-10-11 paper/main.tex
Changes by Kumar

@@
 
-\title{Scalar Parametrization for Joint Time-Frequency interpolation in a MIMO-OFDM channel}
+\title{Scalar Feedback based Joint Time-Frequency Precoder
+  interpolation for  MIMO-OFDM Systems}
 
@@
 To quantize the subcarriers efficiently and exploit the joint
-time-frequency interpolation at the transmitter side, we select the
+time-frequency interpolation at the transmitter, we select the
 subcarriers for quantization in an alternating fashion as shown in
 Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
-$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
-quantized at every $p$-th subcarriers, for even time instances, where,
-$p$ is the gap between two quantized subcarrier indices. For odd time
-instances, subcarriers at the position $pk+q$ for k = 0,1,...,
-$\frac{N-1}{p}$ with q = ${\frac{p}{2}}$. TODO
+$N$ subcarrier MIMO-OFDM system with the subcarriers that are
+quantized. Every $p$-th subcarrier starting with $0$ is quantized for
+even time instances, where, $p$ is the gap between two quantized
+subcarrier indices. For odd time instances, subcarriers at the
+position $pk+q$ for $k = 0,1,..., \frac{N-1}{p}$ with $q =
+{\frac{p}{2}}$.
 
-Therefore, in ADPCM channel tracking scheme, the arrows in the figure show how previous quantized values along with the 1-bit number find the new quantized value.
-Now with the above discussion, the bit budget for each subcarrier will be equal to number of scalars * 1-bit.
-On an average, this can be reduced to half if we transmit only $\theta$ values for one-time instance and only $\phi$ values for the next time while keeping the un transmitted values same as the previous one.
-This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2N_{T} - N_{R})/2$.
+For the adaptive channel tracking scheme, the arrows in
+Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
+the 1-bit enhancement are used to find the new quantized value. Since
+we use a single bit adaptation, the amount of feedback required is 1
+bit for each scalar parameter that is being tracked. On an average,
+this can be reduced by half if we transmit only $\Theta$ values for
+one-time instance and only $\Phi$ values for the next time instant in
+an alternating fashion (the untransmitted value is assumed to be the
+same as that at the previous instant). In this case, the average
+number of bits required to quantize a channel matrix will be
+$N_{T}(2N_{T} - N_{R})/2$.
 
-One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $\sin/\cos(\pi) =\sin/\cos(-\pi)$.
-Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).
-
-% We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using $\textbf{B}$ bits.
-
+One issue that is associated with the tracking the $\phi_i$ parameters
+is that they may change abruptly (cite the figure) between $-\pi$ and
+$\pi$ due to jumps of $2\pi$. We have avoided this by unwrapping the
+phases to facilitate continuous tracking.
 \subsection{Joint Time Frequency Interpolation}
 \label{interp}
+After obtaining the $\beta$ values (using the single bit feedback) at
+the transmitter, we can construct the precoder at subcarrier indices
+$pk$ or $pk+q$. To construct the rest of the subcarriers, we will use
+precoder parameters from both the past and future and time instances
+to do the interpolation (using backtracking to ensure causality). In
+Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
+neighbouring subcarriers using bilinear interpolation (equivalent to
+linear interpolation in two dimensions). These interpolated values and
+then used to reconstruct the precoding matrices all together. Thus,
+the approach here is able to exploit the time and frequency
+correlation jointly to enhance performance.
 
-After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices $pk$ or $pk+q$.
-To reproduce the rest of the subcarriers will use future time instances to do the interpolation.
-As you can see in the figure, each parameter is interpolated using the subcarriers of the next six time instances.
-We do the interpolation using the spline-cubic interpolation.
-We could also use other standard techniques by calculating the MMSE at the missing subcarrier locations but the spline-cubic interpolation method gives better results over the other method.
+One reason why a linear predictive quantization is preferred is
+because the underlying parameters can generally be thought to emerge
+from an autoregressive (AR) process. For such processes, it is well known
+that linear prediction based on past samples is optimal. Even though
+the $\Theta$ and $\Phi$ parameters do not manifest as AR processes,
+for small changes, a linear approximation works sufficiently
+well, as in the description in~\cite{4114278}, although that was
+limited to tracking the temporal evolution of the parameters.
 
-Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because
-
-%Explain prediction and quantization steps here itself ! (succintly)
 \section{Simulation and discussion}
 \label{section3}
-
- We have performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel.
+We present some simulations of the scalar feedback based joint
+time-frequency predictive quantization outlined in the previous section.
+performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel.
 Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel.
@@
 Need to cite that paper.
-
-\subsection{Bilinear Interpolation}
-Bilinear interpolation is the equivalent of linear interpolation in two dimensions.
-On the transmitter side, all the scalars - $\Theta$ and $\Phi$ of the missing subcarriers are interpolatted using the values of the neigbours.
-These interpolated values and then used to reconstruct the precoding matrices all together.
 

### 2019-10-11 paper/main.tex
Add changes

@@
 \begin{equation}
-\textbf{V} = \Bigg[\prod_{k=1}^{N_{R}} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,N_{R}} \big) \: \Bigg\{ \prod_{l=1}^{l=N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
+\textbf{V} = \left[\prod_{k=1}^{N_{R}} \bD_{k} \left( \phi_{k,k},\ldots , \phi_{k,N_{R}} \right) \:  \prod_{l=1}^{N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big)  \right] \: \tilde{\bI}
 \end{equation}
@@
 \end{align*}
-Here, $\phi_{k}$s are referred to as phases; their collection is given by $\Phi and $$\theta_{l}$s as rotation angles; their collection is given by $\Theta$.
-$\phi_k \in (-\pi, + \pi]$ is uniformly distributed
-between the two extremes\cite{4114278}, while $\theta_l$
-is distributed as
+Here, $\phi_{k}$s are referred to as phases; their collection is given
+by $\Phi$ and $\theta_{l}$s as rotation angles; their collection is
+given by $\Theta$. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
+between the two extremes~\cite{4114278}, while $\theta_l$ is
+distributed as
 \begin{equation}
-p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,N_T -1
+2l(\sin\theta_l)^{2l-1}\cos\theta_l, \theta_l \in \left[0, \frac{\pi}{2}\right) \mbox{ , } l = 1,2,\ldots,N_T -1
 \end{equation}

### 2019-10-13 ber_results/plot_results.py
changed qpsk performance and linear interp code, added new results

@@ -3,11 +3,12 @@ import matplotlib.pyplot as plt
 from matplotlib import rc
 from matplotlib.pyplot import figure
 
-# font = {'size'   : 22}
-# rc('font', **font)
-# rc('text', usetex=True)
-# rc('xtick', labelsize=40)
-# rc('ytick', labelsize=40)
+font = {'size'   : 18}
+plt.rc('font', **font)
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=20)
+plt.rc('ytick', labelsize=20)
+
 plt.figure(figsize=(8, 6))
 eb_n0=np.arange(-6,20,3)
 
@@ -24,23 +25,25 @@ lineari=np.load('6bit_eval/ideal_ber.npy')
 # vehicular_mani=np.load('vehicular_manifold/hp_ber.npy')
 # vehicular_scal=np.load('vehicular/hp_ber.npy')
 plt.yscale('log')
-plt.ylabel('BER', fontsize=14)
-plt.xlabel(r'$E_{b}/N_{o}(dB)$', fontsize=14)
-plt.title('Pedestrian Channel', fontsize=14)
+plt.ylabel('BER')
+plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
+# plt.title('Pedestrian Channel', fontsize=14)
 # plt.plot(eb_n0,vehicular_scal, label='Vehicular Scalar Paramesters', marker='+')
 # plt.plot(eb_n0, vehicular_mani, label='Vehicular Manifold', marker='o')
 # plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
 # plt.plot(eb_n0, cubic, label='Cubic', marker='o', color='b')
-plt.plot(eb_n0,linear, label='3.5e-4', marker='+')
-plt.plot(eb_n0,linear17, label='1.7e-4', marker='+')
+plt.plot(eb_n0,linear, 'r--o', lw=3, ms=10, label='3.5e-4')
+plt.plot(eb_n0,linear17, label=r'$f_dT_s = 1.7 \times 10^{-4}$', marker='+')
 plt.plot(eb_n0,linear1, label='1e-4', marker='+')
 
 # plt.plot(eb_n0,lineari, label='3.5e-4 ideal', marker='+')
-plt.plot(eb_n0,linear17i, label='1.7e-4 ideal', marker='+')
[…]

### 2019-10-13 eval.py

@@ -52,7 +52,7 @@ def main():
     Nr=constants.Nr #2
     freq_quanta=constants.freq_quanta #4 - 5
     time_quanta=constants.time_quanta #1
-    channel_type="vehicular/"
+    channel_type="Pedestrian/"
     #---------------------------------------------------------------------------
     # Run Time variables
     # Stores the previously known estimates of the subcarriers
@@ -95,12 +95,11 @@ def main():
 
     start_time=time.time()
     sigma_cb=np.load('./Codebooks/Independent_qt/sigma_cb_2bits_10000.npy') #do we need this codebook
-    hp_Qerr=np.zeros(number_simulations-interp_nbr_vals-2)
-    ot_Qerr=np.zeros(number_simulations-interp_nbr_vals-2)
-    hp_Neterr=np.zeros(number_simulations-interp_nbr_vals-2)
-    ot_Neterr=np.zeros(number_simulations-interp_nbr_vals-2)
+    hp_Qerr=np.zeros(number_simulations-1)
+    ot_Qerr=np.zeros(number_simulations-1)
+    hp_Neterr=np.zeros(number_simulations-1)
+    ot_Neterr=np.zeros(number_simulations-1)
     chan_offset=0
-    gridz_prev=np.zeros((6,num_subcarriers,2*Nt*Nr-Nr**2))
     for chan_index in range(num_chan):
 
         tH_allH=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
@@ -113,8 +112,9 @@ def main():
         allU_vec=np.load('./Precoders_generated/6bit_'+channel_type+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
         onlyt_allU_vec=np.load('./Precoders_generated/6bit_'+channel_type+str(fdts)+'/onlyt_allU_vec'+str(chan_index+chan_offset)+'.npy')
         interpS=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/interpS_'+str(chan_index+chan_offset)+'.npy')
+        # pdb.set_trace(    )
         # for simulation_index in range(number_simulations-6):
-        for simulation_index in range(0,number_simulations-interp_nbr_vals-2):
+        for simulation_index in range(1,number_simulations-1):
             print("---------------------------------------------------------------------------")
[…]

### 2019-10-13 sumrate_BER.py

@@ -94,7 +94,7 @@ def calculate_BER_performance_QPSK(H_sequence, precoder_sequence, Eb_N0_dB,flag=
     Eb_N0=10**(0.1*Eb_N0_dB)
     N0=1.0
     #Constellation_pts=np.array([((2*i+1)+1j*(2*j+1))*Eb_N0/16 for i in range(-8,8) for j in range(-8,8)])    
-    Constellation_pts=np.array([-1-1j,-1+1j,1-1j,+1+1j])*(Eb_N0)/np.sqrt(2)
+    Constellation_pts=np.array([-1-1j,-1+1j,1-1j,+1+1j])*np.sqrt(Eb_N0)
     rand_arr=np.random.randint(4,size=(1,number_streams*num_comp_each_stream)).flatten()
     QPSK_stream=np.array([2*Constellation_pts[rand_arr[i]] for i in range(rand_arr.shape[0])]).reshape(number_streams,num_comp_each_stream)
     QPSK_stream_to_bits=np.ravel([bin2conv(rand_arr[i])\
@@ -102,7 +102,8 @@ def calculate_BER_performance_QPSK(H_sequence, precoder_sequence, Eb_N0_dB,flag=
     ])
     
     #print(QAM_stream.shape)
-    Es=2*(np.sum([np.abs(Constellation_pt) for Constellation_pt in Constellation_pts]))*N0
+    # why is that twice as sqrt(Es) = abs(constellation point)
+    Es=np.sum([np.abs(Constellation_pt)**2 for Constellation_pt in Constellation_pts])*N0/len(Constellation_pts)
     #Chan Out
     channel_out=np.matmul(H_F_sequence, np.array(np.hsplit(QPSK_stream, N_subcarrier)))
     noise=(N0/2)*np.random.randn(channel_out.shape[0], channel_out.shape[1], channel_out.shape[2])+(1.0j)*(N0/2)*np.random.randn(channel_out.shape[0], channel_out.shape[1], channel_out.shape[2])
@@ -113,13 +114,14 @@ def calculate_BER_performance_QPSK(H_sequence, precoder_sequence, Eb_N0_dB,flag=
     flattened_op=est_output
     flattened_op=flattened_op.flatten()
     bins=[0]
-    
     real_arr=np.digitize(flattened_op.real,bins)
+    # real_arr=np.digitize(flattened_op.real,bins)channel_out_with_noise
     imag_arr=np.digitize(flattened_op.imag,bins)
     
[…]

### 2019-10-13 archive/codebook_gen.py
removing archive directory:

@@ -1,39 +1,0 @@
-#-----------------------------------------------------------------------------
-# Code to obtain the base codebook: We consider 10,000 independent channel 
-# evolutions. For each one, we use the evolution of the channel over 10 OFDM 
-# frames’ duration to obtain a reliable prediction from both methods. 
-# We collect the vectors representing the skew Hermitian tangents for vector 
-# quantisation (recall that Cayley Exp. map has skew Hermitian matrices as images 
-# (Section III of paper), which form a vector space). 
-# This is done to get 10,000 independent predictions that give the same number 
-# of independent vectors in the collection.  (Done in vec_collec_gen.py code)
-# The code here applies k-means algorithm to get 
-# a 6 bit base codebook for, for both the hopping based and time based schemes
-#-----------------------------------------------------------------------------
-
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-import itpp
-from mimo_tdl_channel import *
-from sklearn.cluster import KMeans
-import sys
-import signal
-import pdb
-
-#---------------------------------------------------------------------------
-#Lambda Functions
-frob_norm= lambda A:np.linalg.norm(A, 'fro')
-diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-#---------------------------------------------------------------------------
-#SIGINT handler
-def sigint_handler(signal, frame):
-    pdb.set_trace()
-    sys.exit(0)
-
-vec_list=np.load('Codebooks/temp/quantcoll.npy')
-vec_list_norm1=[vec/la.norm(vec) for vec in vec_list]
-kmeans64=KMeans(n_clusters=64).fit(vec_list_norm1)
-np.save('Codebooks/temp/base_quant_cb.npy',kmeans64.cluster_centers_)
-pdb.set_trace()

### 2019-10-13 archive/eval.py

@@ -1,327 +1,0 @@
-# Code to generate the generates precoders corresponding to 
-# 100 channel evolutions for 10 independent channel realizations
-# using both the time based and hopping based schemes 
-# (i.e. fill 10 independent 100*`num_subcarriers` 
-# time-frequency bins matrix with generated precoders), and
-# save them as .npy files
-#---------------------------------------------------------------------------
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-import copy
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK,calculate_BER_performance_QAM256,waterfilling
-from munch import munchify
-import yaml
-from config.arguments import parser
-    
-#SIGINT handler
-def sigint_handler(signal, frame):
-    #Do something while breaking
-    pdb.set_trace()
-    sys.exit(0)
-def main():
-    args = parser.parse_args()
-    with open(args.constants, 'r') as f:
-        constants = munchify(yaml.load(f)) 
-
-    #---------------------------------------------------------------------------
-    #Channel Params
-    signal.signal(signal.SIGINT, sigint_handler)
-    #number of db values for which logsum rate will be calculated
-    num_Cappar=7
-    avg_hpcap=np.zeros(num_Cappar)
-    avg_otcap=np.zeros(num_Cappar)
-    avg_maxcap=np.zeros(num_Cappar)
-    # Eb_N0_dB=np.arange(-6,6,2)
-    Eb_N0_dB=np.arange(-6,20,3)
-    #Store BER here
-    hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    otBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    Ts=constants.Ts # 5e-8
-    num_subcarriers=constants.num_subcarriers #64
-    Nt=constants.Nt #4
-    Nr=constants.Nr #2
-    freq_quanta=constants.freq_quanta #4 - 5
-    time_quanta=constants.time_quanta #1
-    #---------------------------------------------------------------------------
-    # Run Time variables
-    # Stores the previously known estimates of the subcarriers
-    past_vals=constants.past_vals #3
-    fdts=constants.fdts
-    count=0
-    num_chan=constants.num_chan_realisations #10
-    feedback_mats=constants.feedback_mats #8
-    freq_jump=(num_subcarriers-1)//(feedback_mats-1)
[…]

### 2019-10-13 archive/interp.py

@@ -1,226 +1,0 @@
-#Import Statements
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK
-from munch import munchify
-import yaml
-from config.arguments import parser
-
-np.random.seed(81)
-#---------------------------------------------------------------------------
-#SIGINT handler
-def sigint_handler(signal, frame):
-    #Do something while breaking
-    pdb.set_trace()
-    sys.exit(0)
-
-def main():
-    args = parser.parse_args()
-    with open(args.constants, 'r') as f:
-        constants = munchify(yaml.load(f)) 
-    #---------------------------------------------------------------------------
-    #Lambda Functions
-    frob_norm= lambda A:np.linalg.norm(A, 'fro')
-    diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-    #---------------------------------------------------------------------------
-    #Channel Params
-    signal.signal(signal.SIGINT, sigint_handler)
-    # itpp.RNG_randomize(81)
-    itpp.RNG_reset(81)
-    #c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
-    c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
-    Ts=constants.Ts
-    num_subcarriers=constants.num_subcarriers
-    Nt=constants.Nt
-    Nr=constants.Nr
-    B=constants.B
-
-    #---------------------------------------------------------------------------
-    # Simulation Parameters
-    number_simulations=constants.number_simulations # equivalent of time
-    feedback_mats=constants.feedback_mats
-    freq_jump=(num_subcarriers-1)//(feedback_mats-1)
-    freq_quanta=constants.freq_quanta
-    time_quanta=constants.time_quanta
-    # for cold start
-    ran_init=False
-    # Number of channels simulated for
-    num_chan_realisations=constants.num_chan_realisations
-    count=0
-    chan_offset=constants.chan_offset
-    fdts=constants.fdts
-    past_vals=constants.past_vals
-    # pdb.set_trace()
-    #---------------------------------------------------------------------------
-    # Run Time variables
[…]

### 2019-10-13 archive/plot_theta.py

@@ -1,165 +1,0 @@
-#Import Statements
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib import pylab
-from mpl_toolkits.mplot3d import Axes3D
-import scipy.linalg as la
-import scipy.fftpack as sci
-from scipy import stats
-from scipy.interpolate import interp1d 
-from scipy.interpolate import InterpolatedUnivariateSpline
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-import copy
-import code
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK
-
-np.random.seed(81)
-
-def dpcm_encode(qtiz_x_0,x,alpha):
-	qtiz_x=np.zeros(np.size(x))
-	qtiz_x[0]=qtiz_x_0
-	delta=abs(x[1]-x[0])
-	beta=np.zeros(np.size(x))
-	beta[0]=0
-	for i in range(1,len(qtiz_x)-1):
-		beta[i]=np.sign(x[i]-qtiz_x[i-1])
-		qtiz_x[i] = qtiz_x[i-1] + beta[i]*delta
-		if(np.sign(x[i+1]-qtiz_x[i])==beta[i]):
-			delta=alpha*delta
-		else:
-			delta=delta/alpha
-	qtiz_x[-1]=qtiz_x[-2]+beta[-1]*delta		
-	# pdb.set_trace()		
-	return beta,abs(x[1]-x[0])
-def dpcm_decode(qtiz_x_0,alpha,beta,delta):
-	delta= delta
-	qtiz_x=np.zeros(np.size(beta))
-	qtiz_x[0]=qtiz_x_0
-	for i in range(1,np.size(beta)-1):
-		qtiz_x[i]=qtiz_x[i-1]+beta[i]*delta
-		if(beta[i+1]==beta[i]):
-			delta=delta*alpha
-		else:
-			delta=delta/alpha
-	qtiz_x[-1]=qtiz_x[-2]+beta[-1]*delta		
-	return qtiz_x
-	
-def dpcm_encode_2d(qtiz_x_0, x, alpha1, alpha2):
-	qtiz_x=np.zeros(np.shape(x))
-	qtiz_x[0][0]=qtiz_x_0
-	delta1=abs(x[0][1])	
-#---------------------------------------------------------------------------
-#SIGINT handler
-def sigint_handler(signal, frame):
[…]

### 2019-10-14 paper/main.tex
more changes

@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
+\by_{i,t} = \bH_{i,t}\bV_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
 \end{equation}
 where the channel matrix is
-$\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input is
-$\bx_{i,t} \in \mathbb{C}^{N_T}$; the output is
-$\by_{i,t} \in \mathbb{C}^{N_R}$ and the complex additive white
+$\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input data vector is
+$\bx_{i,t} \in \mathbb{C}^{N_d}$; the output is
+$\by_{i,t} \in \mathbb{C}^{N_R}$; the precoder matrix is
+$\bV_{i,t} \in \mathbb{C}^{N_T \times N_R}$; and the complex additive white
 Gaussian noise is $\eta$ whose distribution is
 ${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R <
-N_T$. The thin SVD of $\bH$ is given by
+N_T$ and the data vector size $N_d = N_R$.
+The thin SVD of $\bH$ is given by
 $\bH_{i,t} = \bU_{i,t} \Sigma_{i,t} \bV_{i,t}^{H}$, where
@@
 There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel.
-We take s = 33/9 for Vehicular and pedestrian channel respectively I.e.
-32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, ..
+We take s = 33/9 for Vehicular and pedestrian channel respectively i.e.
+32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1,$, ..
+, 31 and 0,1, ..
 ,7 respectively.
@@
 
-Simulation: BER is calculated by averaging over all the simulations. For Vehicular Channel, 1024 subcarriers are used with doppler frequency – and sampling rate = $10^{5}$
+Simulation: BER is calculated by averaging over all the simulations. For Vehicular Channel, 1024 subcarriers are used whereas 64 subcarriers are used in the Pedestrian Channel.
 
 Initialization is performed by using $10\times12$ bits for each subcarrier for the initial two time frames.
-Later, the quantization will take place in the time-domain using one bit for each parameter.
-Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{adpm-fig}.
+Later, the quantization takes place in the time-domain using one bit for each parameter.
+Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{fig:adpm-fig}.
 Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s.
@@
 %talk about this in the simulations, that would be more relevant
-Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.4*M$ i.e. decay
+Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.5*M$ i.e. decay
 

### 2019-10-14 ber_results/plot_results.py

@@ -13,32 +13,25 @@ plt.figure(figsize=(8, 6))
 eb_n0=np.arange(-6,20,3)
 
 # cubic=np.load('6bit_eval/hp_ber.npy')
-linear17=np.load('6bit_eval/0.00017-linear/hp_ber.npy')
-linear1=np.load('6bit_eval/0.0001-linear/hp_ber.npy')
-linear=np.load('6bit_eval/linear/hp_ber.npy')
+linear35=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
+linear35i=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
 
-linear17i=np.load('6bit_eval/0.00017-linear/ideal_ber.npy')
-linear1i=np.load('6bit_eval/0.0001-linear/ideal_ber.npy')
-lineari=np.load('6bit_eval/ideal_ber.npy')
+linear1=np.load('updated_6bit_eval/pedestrian/0.001/hp_ber.npy')
+linear1i=np.load('updated_6bit_eval/pedestrian/0.001/ideal_ber.npy')
 
-# nearest=np.load('6bit_eval/nearest/hp_ber.npy')
-# vehicular_mani=np.load('vehicular_manifold/hp_ber.npy')
-# vehicular_scal=np.load('vehicular/hp_ber.npy')
 plt.yscale('log')
 plt.ylabel('BER')
 plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
-# plt.title('Pedestrian Channel', fontsize=14)
-# plt.plot(eb_n0,vehicular_scal, label='Vehicular Scalar Paramesters', marker='+')
-# plt.plot(eb_n0, vehicular_mani, label='Vehicular Manifold', marker='o')
-# plt.plot(eb_n0,hpang_ber,label='hop_givens_7.5')
-# plt.plot(eb_n0, cubic, label='Cubic', marker='o', color='b')
-plt.plot(eb_n0,linear, 'r--o', lw=3, ms=10, label='3.5e-4')
-plt.plot(eb_n0,linear17, label=r'$f_dT_s = 1.7 \times 10^{-4}$', marker='+')
-plt.plot(eb_n0,linear1, label='1e-4', marker='+')
+
+plt.plot(eb_n0,linear35, 'r--o', lw=3, ms=10, label=r'$3.5 \times 10^{-4}$')
[…]

### 2019-10-14 eval.py

@@ -142,8 +142,8 @@ def main():
                         allU_vec[points[((simulation_index-1)*(2*feedback_mats-1)+1)//2:(simulation_index*(2*feedback_mats-1)+1)//2+(2*feedback_mats-1),0],\
                         points[((simulation_index-1)*(2*feedback_mats-1)+1)//2:(simulation_index*(2*feedback_mats-1)+1)//2+(2*feedback_mats-1),1]][:,i], (ys[0:3],\
                             xs[0:3]), method='linear')
-
             allU_vec_copy[simulation_index]=gridz[1]
+            
             allU[simulation_index]=givens_frame_to_unitary(allU_vec_copy[simulation_index],Nt,Nr)
             # pdb.set_trace()
             for indice_index in range(fb_indices.shape[0]-1):

### 2019-10-14 paper/main.tex
Add changes

@@
 
-\subsection{Differential Quantization - Channel Tracking}
+\subsection{Differential quantization and channel tracking}
 \label{quantiz}
-For a channel that varies slowly with time, ADPM is an effective
-method for tracking parameters over time using very few bits (often
-just one bit). For example, while tracking a process $x_n$, at time
-$n$, the one-bit jump $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$, where
-$x_n$ is the unquantized value and $\hat{x}_{n-1}$ is the unquantized
-(accurate) value, the quantization function $Q(x)$ is given as
-\begin{equation}
-  Q(x)=\begin{cases}
-    1, & \text{if $x>0$}.\\
-    -1, & \text{otherwise}.
-  \end{cases}
-\end{equation}
-Thus, the quantied vector $\hat{x}_n$ evolves with time as
+For channels that vary slowly with time, adaptive differential
+quantization is an effective method for tracking parameters over time
+using very few bits (often just one bit). For example, while tracking
+a slowly varying discrete random process $x_n$ at time $n$, the one
+bit can be used to determine the direction in which to move in order
+to track the parameter. This can be denoted by
+$\beta_{n} = \mbox{sign}(x_{n} - \hat{x}_{n-1})$, where $x_n$ is the
+unquantized value and $\hat{x}_{n-1}$ is the unquantized (accurate)
+value. Further, the step size for the move can be adapted based on the
+speed of channel variation as well the speed for convergence. If
+$\Delta_n$ is the step size, we define our adaptive quantizer as
 \begin{equation}
@@
 \Delta_{n} = \begin{cases}
-    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
-    \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
+    m \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
+    \Delta_{n-1}/M , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
   \end{cases}
 \end{equation}
-Here we initialize delta using $\Delta_1 = |x_{2}-\hat{x}_1|$.
+where $M$ and $m$ are the jumps used during the positive and negative
+moves. We note that, in general, $m = M$, however, we have found that
+using $M > m$ seems to yield better performance. We initialize
+$\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We note that this is
+performed separately for each of the quantized parameters.
 
@@
 We present some simulations of the scalar feedback based joint
-time-frequency predictive quantization outlined in the previous section.
-performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel.
-Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel.
-Performance is measured by averaging over 10 channel realizations with 100 simulations each.
-The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$.
-For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$.
-Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s.
-There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel.
-We take s = 33/9 for Vehicular and pedestrian channel respectively i.e.
-32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1,$, ..
-, 31 and 0,1, ..
-,7 respectively.
+time-frequency predictive quantization outlined in the previous
+section. In particular, we analyze the performance of the proposed
+quantization and interpolation method for time-varying MIMO-OFDM
+systems. The channel is simulated using the typical Jakes Model for
+both the ITU Vehicular-A channel and ITU Pedestrian channel power
+delay profiles CITE. The BER achieved by the proposed predictive
+quantization based precoder is compared against the ideal precoder for
+100 channel evolutions in each run and averaged. The normalised
+Doppler is assumed to be $f_dT_s = 3.5\times10^{-4}$, with symbol time
+period $T_s = 5\times10^{-8} s$. For both the Vehicular and Pedestrian
+channel profiles, we consider $N_T=4$ and $N_R=2$. Therefore, the
+total number of parameters that determine the precoder are
+$N_{R}(2N_{T} - N_{R}) = 12$, with 5$\theta_l$'s and
+7$\phi_i$'s. For the Vehicular channel profile, we assume that the
+number of subcarriers is 1024, while for the Pedestrian channel
+profile it is 64. We take the spacing between fed back subcarrier
+indices $p$ to be $33$ and $9$ for the Vehicular
+and Pedestrian channel profiles respectively (i.e. for the Vehicular
+profile, the subcarriers indices fed back would be $0, 33, 66, \ldots
+1023$ and for the Pedestrian, $0, 9, 18, \ldots 63$).
 
-\subsection{Initial Quantization}
-For the first and the second simulations/time instances we use of $\textbf{B}$ = 10 bits for quantizing each parameter for effective quantization/initialization.
-Therefore, the number of bits for each subcarrier = $10\times12$ = 120.
-Quantization is performed as given in Section~\ref{section2}.
+For the first and the second simulation time instances, since the
+precoder is not available at the transmitter, we use 10 bits for
+quantizing each parameter for effective representation of the
+precoder. Therefore, the number of bits for each subcarrier =
+$10\times12$ = 120. Quantization is performed as given in
+Section~\ref{section2}. We note that this is not significant, since,
+amortized over long durations, this number becomes small. Subsequent
+quantization takes place in the time-domain using one bit for each
+parameter. Since the bit budget is small, if the channel varies
+sufficiently slowly, either the $theta_l$ or the $phi_i$ parameters
+are fed back alternately, while retaining the unsent parameters as
+being to their previous values, as shows in
+Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the the collection of all
+$\theta_l$s and $\Phi$ is the collection of all $\phi_i$s. This brings
+down the bit budget to an average of 5 bits(the length of $\Theta$)
+and 7 bits (the length of $\Phi$) over alternate feedback instances,
+thereby resulting in an average feedback budget of 5 bits.
 
-Simulation: BER is calculated by averaging over all the simulations. For Vehicular Channel, 1024 subcarriers are used whereas 64 subcarriers are used in the Pedestrian Channel.
-
-Initialization is performed by using $10\times12$ bits for each subcarrier for the initial two time frames.
-Later, the quantization takes place in the time-domain using one bit for each parameter.
-Since the bit budget is low and we are considering slowly varying channels, therefore, either theta or phi parameters are sent alternatingly in the time domain while keeping the unsent parameters equal to their previous values as show in the Figure \ref{fig:adpm-fig}.
-Here $\Theta$ is the the collection of all $\theta$'s and $\Phi$ is the collection of all $phi$'s.
-This brings down the bit budget to average of 5(Size of $\Theta$) and 7(Size of $\Phi$) i.e.
-6 bits average per fed back subcarrier.
-
-{ADPM Design}: The decrement/decay rate is given by $\bm$ and increment/growth rate is given by $\bM$ in the Equation~\ref{delta_eqn} .
-The ratio of $m/M$ is fixed at 2.5 i.e.
-higher decay than increment of the delta values.
-This way, quantized values are able to follow the curve more precisely.
-Also, need to describe the initialization of the Delta values required in the quantization process.
-Need to cite that paper.
+To enable adaptive tracking of the channel parameters, we start with
+an initialization of the step size for each parameter, as discussed in
+Equation~\ref{delta_eqn}, with $m = 0.56, M = 1.4$ (the choice is
+motivated by the results in~\ref{5671092}), and track parameters with
+the temporal evolution of the channel. We have found that using
+$M / m = 2.5$ yields optimum performance in terms of channel tracking.
 

### 2019-10-14 eval.py
changed ber code

@@ -41,7 +41,9 @@ def main():
     avg_otcap=np.zeros(num_Cappar)
     avg_maxcap=np.zeros(num_Cappar)
     # Eb_N0_dB=np.arange(-6,6,2)
-    Eb_N0_dB=np.arange(-6,20,3)
+
+    # Eb_N0_dB=np.arange(-3,30,3)
+    Eb_N0_dB=np.arange(-3,36,3)
     #Store BER here
     idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
     hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
@@ -52,6 +54,7 @@ def main():
     Nr=constants.Nr #2
     freq_quanta=constants.freq_quanta #4 - 5
     time_quanta=constants.time_quanta #1
+    # channel_type="Vehicular/"
     channel_type="Pedestrian/"
     #---------------------------------------------------------------------------
     # Run Time variables
@@ -68,9 +71,8 @@ def main():
     # prev_indices=freq_jump*np.arange(feedback_mats-1)
     
     net_fb_indices=np.sort(np.append(fb_indices,prev_indices))
-    
     unsorted_net_fb_indices=np.append(fb_indices,prev_indices)
-     
+
     total_fb_indices=[]
     total_fb_indices=np.array([unsorted_net_fb_indices for i in range(number_simulations//2)]).reshape([-1])
     temp=np.concatenate((np.zeros(np.size(fb_indices)),np.ones(np.size(prev_indices))))
@@ -96,21 +98,17 @@ def main():
     start_time=time.time()
     sigma_cb=np.load('./Codebooks/Independent_qt/sigma_cb_2bits_10000.npy') #do we need this codebook
     hp_Qerr=np.zeros(number_simulations-1)
-    ot_Qerr=np.zeros(number_simulations-1)
     hp_Neterr=np.zeros(number_simulations-1)
-    ot_Neterr=np.zeros(number_simulations-1)
+
     chan_offset=0
     for chan_index in range(num_chan):
-
         tH_allH=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
         tH_allU=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allU_'+str(chan_index+chan_offset)+'.npy')
[…]

### 2019-10-14 main.py

@@ -35,7 +35,8 @@ def main():
     signal.signal(signal.SIGINT, sigint_handler)
     # itpp.RNG_randomize()
     itpp.RNG_reset(81)
-
+    # channel_type="Vehicular"
+    channel_type="Pedestrian"
     # c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
     c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
     Ts=constants.Ts
@@ -54,8 +55,7 @@ def main():
     freq_jump=(num_subcarriers-1)//(feedback_mats-1)
     freq_quanta=constants.freq_quanta
     time_quanta=constants.time_quanta
-    # for cold start
-    # ran_init=False
+    
     # Number of channels simulated for
     num_chan_realisations=constants.num_chan_realisations
     count=0
@@ -74,7 +74,6 @@ def main():
     # norm_fn='stiefCD'
     start_time=time.time()
     save=True
-
     for chan_index in range(num_chan_realisations):
         print("-----------------------------------------------------------------------")
         print ("Starting Chan Realisation: "+str(chan_index)+" : of "+ str(num_chan_realisations) + " # of total channel realisations for "+str(fdts))
@@ -169,12 +168,13 @@ def main():
                     onlyt_Q_error[simulation_index][i] = stiefCD(\
                         onlyt_allU[simulation_index][onlyt_test_index],tH_allU[simulation_index][onlyt_test_index])
 
+                    test_index=simulation_index%2*(freq_jump//2) + freq_jump*i
                     interpS[simulation_index][onlyt_test_index]=sigma_list[onlyt_test_index]
 
                     # hop_pred vectors
                     if(simulation_index%2==1 and i==feedback_mats-1):
                         continue
-                    test_index=simulation_index%2*(freq_jump//2) + freq_jump*i    
+                    
                     if(simulation_index==1):
                         # pdb.set_trace()
                         oU_vec=U_vec[test_index]
@@ -234,11 +234,11 @@ def main():
                     # pdb.set_trace()
[…]

### 2019-10-14 mimo_tdl_channel.py

@@ -749,6 +749,7 @@ class MIMO_TDL_Channel():
         # Delay profile in cspec will be discretized with sampling_time in seconds
         channel = itpp.comm.TDL_Channel(c_spec, sampling_time)
         self.channel_length=channel.taps()
+        # pdb.set_trace()
         # Initialise the Nt*Nr TDL channels
         for i in range(self.Nt*self.Nr):
             self.tdl_channels.append(itpp.comm.TDL_Channel(c_spec,sampling_time))

### 2019-10-14 sumrate_BER.py

@@ -9,19 +9,20 @@ import pdb
 diff_frobenius_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
 
 def waterfilling(alpha,P=1):
-	n=alpha.size
-	ind=np.argsort(alpha)
-	lambda1=0
-	for m in range(n):
-		t=0
-		for j in range(m,n):
-			t=t+(1/alpha[ind[j]])
-		t=(t+P)/(n-m)
-		lambda1=1/t
-		if(lambda1<alpha[ind[m]]):
-			break
-	result=[((1/lambda1)-(1/alpha[j])) if (lambda1<alpha[j]) else 0 for j in range(n)]
-	return result
+    n=alpha.size
+    ind=np.argsort(alpha)
+    lambda1=0
+    for m in range(n):
+        t=0
+        for j in range(m,n):
+            t=t+(1/alpha[ind[j]])
+        t=(t+P)/(n-m)
+        lambda1=1/t
+        if(lambda1<alpha[ind[m]]):
+            break
+    result=[((1/lambda1)-(1/alpha[j])) if (lambda1<alpha[j]) else 0 for j in range(n)]
+    # pdb.set_trace()
+    return result
 
 def leakage_analysis(H_list,U_list,interpolated_U,num_subcarriers,wf_sigma_list,wf_interp_S,Nt,Nr,ret_abs=False):
     fig_merit=np.zeros(num_subcarriers)
@@ -93,17 +94,18 @@ def calculate_BER_performance_QPSK(H_sequence, precoder_sequence, Eb_N0_dB,flag=
     total_number_bits=number_streams*Number_bits_each_stream
     Eb_N0=10**(0.1*Eb_N0_dB)
     N0=1.0
+    Es_N0=2*Eb_N0
     #Constellation_pts=np.array([((2*i+1)+1j*(2*j+1))*Eb_N0/16 for i in range(-8,8) for j in range(-8,8)])    
     Constellation_pts=np.array([-1-1j,-1+1j,1-1j,+1+1j])*np.sqrt(Eb_N0)
     rand_arr=np.random.randint(4,size=(1,number_streams*num_comp_each_stream)).flatten()
-    QPSK_stream=np.array([2*Constellation_pts[rand_arr[i]] for i in range(rand_arr.shape[0])]).reshape(number_streams,num_comp_each_stream)
[…]

### 2019-10-15 paper/main.tex
Further changes by Kumar

@@
 Typical approaches to predictive quantization and interpolation do not exploit the joint time-frequency correlation to enhance the precoder estimate at the transmitter.
-However, exploiting the correlations jointly not only reduces the feedback requirement, it also is able to enhance the precoder reconstruction accuracy.
+However, exploiting the correlations jointly not only reduces the feedback requirement, but also enhances the precoder reconstruction accuracy significantly.
 Further, the independence of the scalar parameters reduces the problem to interpolation of separate scalar valued function, as opposed to operations over tangent spaces in manifolds~\cite{Gupt1905:Predictive}.
@@
 
-The rest of this paper is organized as follows: Section~\ref{section2} described the system model and the precoder feedback and interpolation techniques.
-Section~\ref{section3} compares the BER and achievable rates for our proposed methods with that of past work.
+The rest of this paper is organized as follows: Section~\ref{section2} describes the system model and the precoder feedback and interpolation techniques.
+Section~\ref{section3} compares the BER and achievable rates for our proposed methods with other approaches.
 Finally, Section~\ref{section4} concludes with discussions on future work.
@@
 
-In this paper, we propose a method for predictive quantization in the time domain over the subcarriers of a MIMO-OFDM system using scalar parameters.
-We also exploit the joint time-frequency correlation of these scalar parameters on the transmitter side.
-TODO how we solve it?
-
-The rest of this paper is organized as follows: Section~\ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
+% The rest of this paper is organized as follows: Section~\ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
 In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t}\bV_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
+\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
 \end{equation}
@@
 $\bx_{i,t} \in \mathbb{C}^{N_d}$; the output is
-$\by_{i,t} \in \mathbb{C}^{N_R}$; the precoder matrix is
-$\bV_{i,t} \in \mathbb{C}^{N_T \times N_R}$; and the complex additive white
-Gaussian noise is $\eta$ whose distribution is
-${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R <
-N_T$ and the data vector size $N_d = N_R$.
-The thin SVD of $\bH$ is given by
+$\by_{i,t} \in \mathbb{C}^{N_R}$; the unitary precoder matrix is
+$\tilde{\bV}_{i,t} \in \mathbb{C}^{N_T \times N_R}$; and the complex
+additive white Gaussian noise is $\eta$ whose distribution is
+${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R < N_T$
+and the data vector size $N_d = N_R$.  The ``thin'' SVD (that keeps
+only those right and left singular vectors that correspond to non-zero
+singular values) of $\bH$ is given by
 $\bH_{i,t} = \bU_{i,t} \Sigma_{i,t} \bV_{i,t}^{H}$, where
@@
 
-% Givens rotation definition and how we are decomposing, Theta- pdf = 2lsin(theta)(2l-1).\cos(theta) 0 $\leq$ Theta $\leq$ pi/2.
-% Householders' transformation and how we are decomposing. -pi$\leq$ phi $\leq$ pi. Uniformly distributed.
-% Low-complexity quantization scheme based on a further parameterization of a vector in
-% terms of a minimal number of independent scalar parameters. In particular, we consider its application to CSI feedback in slowly time-varying MIMO channels
-
 \subsection{Scalar Parametrization}
 \label{givens}
-The degrees of freedom in the $\bV_{i,t}$ matrices is smaller than the
-number of real entries in the matrix because of the geometrical
-structure between the columns of the matrix~\cite{4114278}. The
-orthonormal matrix $\bV_{i,t}$ with $N_T$ rows and $N_R$ columns can be
-decomposed as TODO
+The degrees of freedom in the $\tilde{\bV}_{i,t}$ matrices is smaller
+than the number of real entries in the matrix because of the
+geometrical structure between the columns of the
+matrix~\cite{4114278}. The orthonormal matrix $\tilde{\bV}_{i,t}$ with
+$N_T$ rows and $N_R$ columns can be decomposed as follows~\cite{4114278} (we suppress
+the $(i, t)$ subscripts in the decomposition matrices)
 \begin{equation}
-\textbf{V} = \left[\prod_{k=1}^{N_{R}} \bD_{k} \left( \phi_{k,k},\ldots , \phi_{k,N_{R}} \right) \:  \prod_{l=1}^{N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big)  \right] \: \tilde{\bI}
+\tilde{\bV}_{i,t} = \left[\prod_{k=1}^{N_{R}} \bD_{k} \left( \phi_{k,k},\ldots , \phi_{k,N_{R}} \right) \:  \prod_{l=1}^{N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big)  \right] \: \tilde{\bI}
 \end{equation}
-where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
-\ldots, \phi_{k,N_T } \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
-e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,N_T }}  \big)$, ${\bf 1}_{k-1}$ is
-represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
-= $\big[\bI_{N_R }, {\bf 0}_{N_T ,N_T -N_R }\big]^{T}$, and
+where $\bD_{k}$ is a diagonal matrix defined as
+$\bD_{k}\big(\phi_{k,k}, \ldots, \phi_{k,N_T } \big)$ =
+$\mbox{diag}\big( {\bf 1}_{k-1}, e^{j\phi_{k,k}},\ldots,
+e^{j\phi_{k,N_T }} \big)$, ${\bf 1}_{k-1}$ is represents $k-1$ ones,
+and the $N_R \times N_T$ matrix $\tilde{\bI}$ =
+$\big[\bI_{N_R }, {\bf 0}_{N_T ,N_T -N_R}\big]^{T}$. Here,
+${\bf 0}_{N_T ,N_T -N_R }$ is an $N_T\times (N_T - N_R)$ matrix all of
+whose elements are zero. Finally, we also have
 \begin{equation}
-G_{m-1,m}\big(\theta\big)  =
+\bG_{m-1,m}\big(\theta\big)  =
 \begin{bmatrix}
@@
 \end{align*}
-Here, $\phi_{k}$s are referred to as phases; their collection is given
-by $\Phi$ and $\theta_{l}$s as rotation angles; their collection is
-given by $\Theta$. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
-between the two extremes~\cite{4114278}, while $\theta_l$ is
-distributed as
+Here, $\phi_{k}$s are referred to as phases, and their collection for
+the $i$-th subcarrier at time instant $t$ is captured in a vector
+denoted by $\Phi_{i,t}$. $\theta_{l}$s as called rotation angles, and
+their collection is captured in a vector referred to as
+$\Theta_{i,t}$. We note that $\Phi_{i,t}$ and $\Theta_{i,t}$ represent
+the phase and rotation angles for the $i$-th subcarrier at time
+instant $t$, as represented in Fig.~\ref{fig:adpm-fig}. It is known
+that $\phi_k \in (-\pi, + \pi]$ is uniformly distributed between the
+two extremes~\cite{4114278}, while $\theta_l$ is distributed as
 \begin{equation}
@@
 obtained from the decomposition of a complex orthogonal matrix
-$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The
-number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
-number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
+$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$. The
+number of $\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
+number of $\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\Phi_{i,t}$ is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while $\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.
 \begin{figure}
@@
 \caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
-\label{ber_overvie}
-\vspace{-5pt}
 \end{figure}
@@
 quantization is an effective method for tracking parameters over time
-using very few bits (often just one bit). For example, while tracking
-a slowly varying discrete random process $x_n$ at time $n$, the one
-bit can be used to determine the direction in which to move in order
-to track the parameter. This can be denoted by
+using very few bits (often just one bit per parameter). For example,
+while tracking a slowly varying discrete random process $x_n$ at time
+$n$, the one bit can be used to determine the direction in which to
+move in order to track the parameter. This can be denoted by
 $\beta_{n} = \mbox{sign}(x_{n} - \hat{x}_{n-1})$, where $x_n$ is the
@@
 \Delta_{n} = \begin{cases}
-    m \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
+    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
     \Delta_{n-1}/M , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
@@
 \end{equation}
-where $M$ and $m$ are the jumps used during the positive and negative
-moves. We note that, in general, $m = M$, however, we have found that
-using $M > m$ seems to yield better performance. We initialize
-$\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We note that this is
-performed separately for each of the quantized parameters.
+where $M$ is the increment used during the positive and negative
+moves. We initialize $\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We
+note that this is performed separately for each of the quantized
+parameters.
 
@@
 parameter. Since the bit budget is small, if the channel varies
-sufficiently slowly, either the $theta_l$ or the $phi_i$ parameters
+sufficiently slowly, either the $\theta_l$ or the $\phi_i$ parameters
 are fed back alternately, while retaining the unsent parameters as
@@
 an initialization of the step size for each parameter, as discussed in
-Equation~\ref{delta_eqn}, with $m = 0.56, M = 1.4$ (the choice is
-motivated by the results in~\ref{5671092}), and track parameters with
-the temporal evolution of the channel. We have found that using
-$M / m = 2.5$ yields optimum performance in terms of channel tracking.
+Equation~\ref{delta_eqn}, with $M = 1.4$ (the choice is motivated by
+the results in~\cite{5671092}), and track parameters with the temporal
+evolution of the channel, and we find that this choice effectively
+tracks the channel parameters.
 
-Performance Measurements:
-Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us.
+% Performance Measurements:
+% Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us.
 Later we are comparing the BER rates with the ideal completely fed back subcarriers.
 
-Using the idea of joint interpolation given in the other paper.
+% Using the idea of joint interpolation given in the other paper.
 
-%talk about this in the simulations, that would be more relevant
-Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.5*M$ i.e. decay
+% %talk about this in the simulations, that would be more relevant
+% Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.5*M$ i.e. decay
 
@@
 
-\subsection{Performance Measurement}
-
-\label{setting}
-
-\noindent Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used.
-Later we are comparing the BER rates with the completely fed back subcarriers.
-Or simulations and discussions part 2 criteria could be used.
-Done similar work as in \cite{4114278} but along with that we have joint interpolated in time and frequency at the transmitter using the fundamental scalar parameters obtained.
-To lower the number of bits we have used the method of interpolation and have achieved significant improvement over predictive quantization method in \cite{6891198} and also in \cite{Gupt1905:Predictive} which tried to use joint interpolation over the tangent space in the Stiefel Manifold.
-In fact, we use joint interpolation of the scalar parameters which are easy to use and give better quantization than any of the above methods.
+Since the Precoding matrices fundamentally lie on the Stiefel
+manifold, therefore, we are using the chordal distance parameter to
+measure the effectiveness of the quantization method used. Later we
+are comparing the BER rates with the completely fed back
+subcarriers. Or simulations and discussions part 2 criteria could be
+used. Done similar work as in \cite{4114278} but along with that we
+have joint interpolated in time and frequency at the transmitter using
+the fundamental scalar parameters obtained. To lower the number of
+bits we have used the method of interpolation and have achieved
+significant improvement over predictive quantization method in
+\cite{6891198} and also in \cite{Gupt1905:Predictive} which tried to
+use joint interpolation over the tangent space in the Stiefel
+Manifold. In fact, we use joint interpolation of the scalar parameters
+which are easy to use and give better quantization than any of the
+above methods.
 

### 2019-10-15 paper/main.tex
Add more changes

@@
 we use a single bit adaptation, the amount of feedback required is 1
-bit for each scalar parameter that is being tracked. On an average,
-this can be reduced by half if we transmit only $\Theta$ values for
+bit for each scalar parameter that is being tracked. On average, this
+can be reduced by half if we transmit only $\Theta$ values for
 one-time instance and only $\Phi$ values for the next time instant in
@@
 One issue that is associated with the tracking the $\phi_i$ parameters
-is that they may change abruptly (cite the figure) between $-\pi$ and
+is that they may change abruptly between $-\pi$ and
 $\pi$ due to jumps of $2\pi$. We have avoided this by unwrapping the

### 2019-10-17 paper/main.tex
Add text

@@
 \begin{abstract}
-In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information(CSI).
+% In a limited feedback MIMO channel, the performance of the channel can improve significantly if the transmitter knows the channel state information(CSI).
 The receiver knows the channel information, and it feeds back to the transmitter.
@@
 153-words
+
+  In MIMO-OFDM systems, knowledge of the precoding matrix enables
+  effective resource allocation for high rates and low BER. However,
+  feeding back precoders for all subcarriers introduces large
+  overheads. Past work has shown that the orthonormal structure of the
+  precoder enables effective parameterization and quantization using
+  Givens rotations and Householder transformations. In addition, the
+  use of temporal and frequency correlations can further reduce the
+  feedback requirements.  We present a framework that utilizes
+  adaptive feedback with joint time-frequency prediction for precoder
+  construction at the transmitter.  Simulations reveal that the
+  proposed method achieves low BERs for various channel profiles
+  compared to other approaches.
 \end{abstract}
@@
 
-The problem of effective quantization and feedback of precoders for MIMO-OFDM has attracted significant attention in recent years.
-Many approaches involve codebook design over the Grassmann manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308} or the Stiefel Manifold \cite{6891198,Gupt1905:Predictive}, wherein the manifold structure of the precoders is used to obtain tangents for predictive quantization and interpolation across time and frequency.
-While these methods have been shown to be very effective for precoder feedback,  they involve operations over higher dimensional manifolds.
-In particular, optimization and interpolation is complicated, especially for manifolds where geodesic paths for interpolation may be difficult to obtain, especially for joint time-frequency predictive coding.
-An alternate approach is to parametrize the precoding matrices using scalar parameters.
-The space of unitary matrices parameterized into independent scalar parameters that correspond to Givens rotations and Household Transformations of the unitary precoding matrices \cite{4114278,4556174}.
-These approaches have been combined with adaptive delta modulation effectively track precoding matrices both in the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit separately.
+The problem of effective quantization and feedback of precoders for
+MIMO-OFDM has attracted significant attention in recent years. Many
+approaches involve codebook design over the Grassmann
+manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
+the Stiefel manifold \cite{6891198,Gupt1905:Predictive} and the Flag
+manifold~\cite{pitaval2013codebooks}, wherein the manifold structure
+of the precoders is used to obtain tangents for predictive
+quantization and interpolation across time and frequency. While these
+methods have been shown to be very effective for precoder feedback,
+they involve operations over higher dimensional manifolds. In
+particular, optimization and interpolation is complicated, especially
+for manifolds where geodesic paths for interpolation may be difficult
+to obtain, especially for joint time-frequency predictive coding. An
+alternate approach is to parametrize the precoding matrices using
+scalar parameters. The space of unitary matrices parameterized into
+independent scalar parameters that correspond to Givens rotations and
+Household Transformations of the unitary precoding matrices
+\cite{4114278,4556174}. These approaches have been combined with
+adaptive delta modulation effectively track precoding matrices both in
+the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
+separately.
 
@@
 \label{section3}
-We present some simulations of the scalar feedback based joint
+We now present simulations of the scalar feedback based joint
 time-frequency predictive quantization outlined in the previous
@@
 systems. The channel is simulated using the typical Jakes Model for
-both the ITU Vehicular-A channel and ITU Pedestrian channel power
-delay profiles CITE. The BER achieved by the proposed predictive
+both the ITU Vehicular-A channel and ITU Pedestrian-A channel power
+delay profiles~\cite{recommendation1997guidelines}. The BER achieved by the proposed predictive
 quantization based precoder is compared against the ideal precoder for
@@
 total number of parameters that determine the precoder are
-$N_{R}(2N_{T} - N_{R}) = 12$, with 5$\theta_l$'s and
-7$\phi_i$'s. For the Vehicular channel profile, we assume that the
+$N_{R}(2N_{T} - N_{R}) = 12$, with 5$\theta_l$s and
+7$\phi_i$s. For the Vehicular channel profile, we assume that the
 number of subcarriers is 1024, while for the Pedestrian channel
@@
 For the first and the second simulation time instances, since the
-precoder is not available at the transmitter, we use 10 bits for
-quantizing each parameter for effective representation of the
+precoder is not available at the transmitter, we use 2 bits for
+initializing each parameter for effective representation of the
 precoder. Therefore, the number of bits for each subcarrier =
-$10\times12$ = 120. Quantization is performed as given in
+$10\times 2$ = 20 (we note that $\phi_0 = \phi_4 = 0$ due to the
+non-uniqueness of the SVD). Quantization is performed as given in
 Section~\ref{section2}. We note that this is not significant, since,
@@
 are fed back alternately, while retaining the unsent parameters as
-being to their previous values, as shows in
+being to their previous values, as shown in
 Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the the collection of all
@@
 an initialization of the step size for each parameter, as discussed in
-Equation~\ref{delta_eqn}, with $M = 1.4$ (the choice is motivated by
-the results in~\cite{5671092}), and track parameters with the temporal
-evolution of the channel, and we find that this choice effectively
-tracks the channel parameters.
+Equation~\ref{delta_eqn}. The value of $M$ was chosen to be $1.4$, and
+can be adjusted appropriately to track the temporal evolution of
+parameters for different Doppler frequencies.
 
-% Performance Measurements:
-% Since the Precoding matrices fundamentally lie on the Stiefel manifold, therefore, we are using the chordal distance parameter to measure the effectiveness of the quantization method used by us.
-Later we are comparing the BER rates with the ideal completely fed back subcarriers.
+Fig.~\ref{fig:error_decay}, shows the rate at which error diminishes
+with time for the proposed approach and direct quantization on the
+Stiefel manifold using codebooks described
+in~\cite{Gupt1905:Predictive}. We see that, since the adaptation of
+independent scalar parameters causes the error to decay rapidly, the
+error is lower and the convergence is faster.
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{images/stief_veh.pdf}
+\caption{\label{fig:error_decay}A comparison of the $4\times 2$ precoder quantization error vs. time instant for
+  the proposed approach (labelled ``TODO'') and direct quantization on
+the Stiefel manifold for the Vehicular channel profile.}
+\end{figure}
 
-% Using the idea of joint interpolation given in the other paper.
+TODO: Agrim and Mohit discuss and then fill this part.
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{images/ber-pedestrian-35e-4}
+\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
+  channel with the ITU Pedestrian-A channel profile for $f_dT_s = 3.5\times 10^{-4}$.}
+\label{fig:ber_ped}
+\end{figure}
 
-% %talk about this in the simulations, that would be more relevant
-% Since $\Delta_{n} = M\Delta_{n-1} or \Delta_{n-1}/m $, we found empirically that $m=2.5*M$ i.e. decay
+Fig.~\ref{fig:ber_ped} compares the BER obtained using the proposed
+quantization method with that obtained using exact precoders. We can
+observe that the performance of the proposed method is very close to
+the performance with exact precoders. This can be attributed to the
+fact that even the 1 bit update to the scalar parameters is able to
+approximate the precoder very accurately. Since the channel variation
+across subcarriers is gradual, owing to relatively small delay spread
+of the Pedestrian channel profile, linear interpolation and prediction
+yields good performance. Moreover, this method outperforms the Stiefel
+manifold based approach to the fact that the adaptation is much more
+accurate for scalar parameters than the tangent space based adaptation
+used for the Stiefel manifold.
+
+Fig.~\ref{fig:ber_veh} shows a similar comparison for the Vehicular
+channel profile. The larger delay spread of the channel necessitates
+the use of a larger number of subcarriers (1024) than the Pedestrian
+channel (64). The larger number of subcarriers implies that the
+predictive quantization and interpolation needs to be performed across
+a larger frequency range, thereby reducing the accuracy. Even in this
+case, we observe that the scalar parameter based approach outperforms
+manifold based quantization by about 5 dB.
 
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/pedestrian.pdf}
-\caption{BER vs SNR for pedestrian channel}
-\label{ber_overview}
-\vspace{-5pt}
+\includegraphics[width=0.5\textwidth]{images/ber-vehicular-35e-4}
+\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
+  channel with the ITU Vehicular-A channel profile for $f_dT_s = 3.5\times 10^{-4}$.}
+\label{fig:ber_veh}
 \end{figure}
 
-Since the Precoding matrices fundamentally lie on the Stiefel
-manifold, therefore, we are using the chordal distance parameter to
-measure the effectiveness of the quantization method used. Later we
-are comparing the BER rates with the completely fed back
-subcarriers. Or simulations and discussions part 2 criteria could be
-used. Done similar work as in \cite{4114278} but along with that we
-have joint interpolated in time and frequency at the transmitter using
-the fundamental scalar parameters obtained. To lower the number of
-bits we have used the method of interpolation and have achieved
-significant improvement over predictive quantization method in
-\cite{6891198} and also in \cite{Gupt1905:Predictive} which tried to
-use joint interpolation over the tangent space in the Stiefel
-Manifold. In fact, we use joint interpolation of the scalar parameters
-which are easy to use and give better quantization than any of the
-above methods.
+% Since the Precoding matrices fundamentally lie on the Stiefel
+% manifold, therefore, we are using the chordal distance parameter to
+% measure the effectiveness of the quantization method used. Later we
+% are comparing the BER rates with the completely fed back
+% subcarriers. Or simulations and discussions part 2 criteria could be
+% used. Done similar work as in \cite{4114278} but along with that we
+% have joint interpolated in time and frequency at the transmitter using
+% the fundamental scalar parameters obtained. To lower the number of
+% bits we have used the method of interpolation and have achieved
+% significant improvement over predictive quantization method in
+% \cite{6891198} and also in \cite{Gupt1905:Predictive} which tried to
+% use joint interpolation over the tangent space in the Stiefel
+% Manifold. In fact, we use joint interpolation of the scalar parameters
+% which are easy to use and give better quantization than any of the
+% above methods.
 
-The advantage they had over the number of bits due to the 6 bit codebook in \cite{6891198,Gupt1905:Predictive} is also achieved by us by using smart interpolation techniques, i.e.
+% The advantage they had over the number of bits due to the 6 bit codebook in \cite{6891198,Gupt1905:Predictive} is also achieved by us by using smart interpolation techniques, i.e.
 by dropping the feedback bits in alternating time instances as it is still going to follow the scalar parameters without much difference.
 
-The problem of values hopping between $\pi$ and $-\pi$ is also tackled by wrapping the values around which we can show works nicely.
+% The problem of values hopping between $\pi$ and $-\pi$ is also tackled by wrapping the values around which we can show works nicely.
 The only drawback is that while initializing the parameter, they go out of their actual range and therefore using a uniform quantizer over the prescribed range does not work.
@@
 
-Using the idea of joint interpolation given in the other paper.
+% Using the idea of joint interpolation given in the other paper.
 
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{images/vehicular_ber.pdf}
-\caption{BER vehicular plot comparison between our method and \cite{Gupt1905:Predictive}}
-\label{ber_overview}
-\vspace{-5pt}
-\end{figure}
-%Qtisn error with 1000 chan evols
-%Qtisn error with 100 chan evols
-%BER
-%Achievable rat
-\section{Conclusions and Future Work}
+\section{Conclusion}
 \label{section4}
-
-We could reduce more by seeing the effect of sigma values (power distribution).
-We can save the bits by not sending for those which have very low sigma values.
-This could result in a very robust method.
-
-\vspace{-4pt}
+We have presented a predictive quantization based precoder
+reconstruction of a MIMO-OFDM system that employs scalar quantization
+based feedback of precoder parameters. By quantizing and adapting the
+Givens rotation and Householder trasnsformation parameters of the
+precoder, the receiver is able to send minimal feedback to the
+transmitter and yet enable effective precoder adaptation. Exploiting
+the temporal and frequency correlations jointly enhances precoder
+accuracy significantly. Simulations reveal that the proposed approach
+provides faster convergence and smaller errors in the precoder when
+compared to approaches that involve direct quantization on manifold
+tangent spaces. Future work would involve tuning the adaptation rate
+of the scalar parameters and extending this to TODO.
 

### 2019-10-17 ber_results/ber_results.py
adding changing code and ber results

@@ -1,0 +1,52 @@
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import rc
+from matplotlib.pyplot import figure
+
+font = {'size'   : 18}
+plt.rc('font', **font)
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=20)
+plt.rc('ytick', labelsize=20)
+
+plt.figure(figsize=(8, 6))
+eb_n0=np.arange(-3,30,3)
+
+# cubic=np.load('6bit_eval/hp_ber.npy')
+linear0=np.load('updated_6bit_eval/pedestrian/0.001/hp_ber.npy')
+linear1=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
+linear2=np.load('updated_6bit_eval/agrim_manifold_pedestrian/0.00035/hp_ber.npy')
+linear4=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
+linear0=linear0[:11]
+linear1=linear1[:11]
+linear2=linear2[:11]
+linear4=linear4[:11]
+
+plt.yscale('log')
+plt.ylabel('BER')
+plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
+# # pdb.set_trace()
+# plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$\times 10^{-1}$')
+plt.plot(eb_n0,linear4, 'c--o', lw=3, ms=10, label=r'Ideal - Complete Feedback')
+plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$f_{d}T_{s}=10^{-3}$, Scalar param.')
+# plt.plot(eb_n0,linear1,'b--o', lw=3, ms=10, label=r'$f_{d}T_{s}3.5=\times 10^{-4}$-Parameter'))
+plt.plot(eb_n0,linear2,'g--o', lw=3, ms=10, label=r'$f_{d}T_{s}=3.5\times 10^{-4}$, Manifold')
+
+v = plt.axis()
+plt.axis([-3.4, 27.4, v[2], v[3]])
+# plt.plot(eb_n0,lineari, label='3.5e-4 ideal', marker='+')
+# plt.plot(eb_n0,linear17i, 'r-', lw=3, ms=10,label='1.7e-4 ideal', marker='+')
+# plt.plot(eb_n0,linear1i, label='1e-4 ideal', marker='+')
[…]

### 2019-10-17 ber_results/plot_ber.py

@@ -1,0 +1,39 @@
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import rc
+from matplotlib.pyplot import figure
+
+font = {'size'   : 18}
+plt.rc('font', **font)
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=20)
+plt.rc('ytick', labelsize=20)
+
+plt.figure(figsize=(8, 6))
+eb_n0=np.arange(-3,25,3)
+
+ped_scal=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
+ped_mani=np.load('updated_6bit_eval/agrim_manifold_pedestrian/0.00035/hp_ber.npy')
+ped_ideal=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
+
+veh_scal=np.load('updated_6bit_eval/vehicular/0.00035/hp_ber.npy')
+veh_mani=np.load('updated_6bit_eval/agrim_manifold_vehicular/0.00035/hp_ber.npy')
+veh_ideal=np.load('updated_6bit_eval/agrim_manifold_vehicular/0.00035/ideal_ber.npy')
+plt.yscale('log')
+plt.ylabel('BER')
+plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
+
+plt.plot(eb_n0,ped_scal[:10], 'b--o', lw=3, ms=10, label='Scalar')
+plt.plot(eb_n0,ped_mani[:10],'g--o', lw=3, ms=10, label='Manifold')
+plt.plot(eb_n0,ped_ideal[:10],'y--o', lw=3, ms=10, label='64 subcarriers - complete')
+
+# plt.plot(eb_n0,veh_scal[:10], 'g--o', lw=3, ms=10, label='Scalar')
+# plt.plot(eb_n0,veh_mani,'b--o', lw=3, ms=10, label='Manifold')
+# plt.plot(eb_n0,veh_ideal,'y--o', lw=3, ms=10, label='64 subcarriers - complete')
+
+
+plt.grid(which='major', linewidth='0.5', color='black');
+plt.grid(which='minor', linewidth='0.5', color='grey');
+
+plt.legend(fontsize=14.5)
+plt.show()

### 2019-10-17 ber_results/plot_results.py

@@ -1,47 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import rc
-from matplotlib.pyplot import figure
-
-font = {'size'   : 18}
-plt.rc('font', **font)
-plt.rc('text', usetex=True)
-plt.rc('xtick', labelsize=20)
-plt.rc('ytick', labelsize=20)
-
-plt.figure(figsize=(8, 6))
-eb_n0=np.arange(-6,20,3)
-
-# cubic=np.load('6bit_eval/hp_ber.npy')
-linear35=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
-linear35i=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
-
-linear1=np.load('updated_6bit_eval/pedestrian/0.001/hp_ber.npy')
-linear1i=np.load('updated_6bit_eval/pedestrian/0.001/ideal_ber.npy')
-
-plt.yscale('log')
-plt.ylabel('BER')
-plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
-
-plt.plot(eb_n0,linear35, 'r--o', lw=3, ms=10, label=r'$3.5 \times 10^{-4}$')
-plt.plot(eb_n0,linear1, 'g--o', lw=3, ms=10, label=r'$1.0 \times 10^{-3}$')
-plt.plot(eb_n0,linear35i,'b--o', lw=3, ms=10, label=r'Ideal')
-# plt.plot(eb_n0,linear1i,'b--o', lw=3, ms=10, label=r'Ideal $1.0 \times 10^{-3}$'))
-
-
-# plt.plot(eb_n0,lineari, label='3.5e-4 ideal', marker='+')
-# plt.plot(eb_n0,linear17i, 'r-', lw=3, ms=10,label='1.7e-4 ideal', marker='+')
-# plt.plot(eb_n0,linear1i, label='1e-4 ideal', marker='+')
-# plt.plot(eb_n0,nearest, label='nearest', marker='1', color='r')
-
-plt.grid(which='major', linewidth='0.5', color='black');
-plt.grid(which='minor', linewidth='0.5', color='grey');
-
-plt.legend(fontsize=14.5)
-plt.show()
-
-# pylab.rc('text', usetex=True)
-# pylab.rc('font', size=45)
-# pylab.rc('axes', labelsize=40)
-# pylab.rc('legend', fontsize=45)
-# pylab.figure(figsize=(8,6))

### 2019-10-17 eval.py

@@ -42,8 +42,8 @@ def main():
     avg_maxcap=np.zeros(num_Cappar)
     # Eb_N0_dB=np.arange(-6,6,2)
 
-    # Eb_N0_dB=np.arange(-3,30,3)
-    Eb_N0_dB=np.arange(-3,36,3)
+    Eb_N0_dB=np.arange(-3,25,3)
+    # Eb_N0_dB=np.arange(-3,36,3)
     #Store BER here
     idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
     hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
@@ -54,8 +54,8 @@ def main():
     Nr=constants.Nr #2
     freq_quanta=constants.freq_quanta #4 - 5
     time_quanta=constants.time_quanta #1
-    # channel_type="Vehicular/"
-    channel_type="Pedestrian/"
+    channel_type="Vehicular/"
+    # channel_type="Pedestrian/"
     #---------------------------------------------------------------------------
     # Run Time variables
     # Stores the previously known estimates of the subcarriers
@@ -102,14 +102,14 @@ def main():
 
     chan_offset=0
     for chan_index in range(num_chan):
-        tH_allH=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allH_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allU_'+str(chan_index+chan_offset)+'.npy')
-        tH_allU_vec=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/th_allU_vec'+str(chan_index+chan_offset)+'.npy')        
-        tHS=np.load('Precoders_generated/6bit_'+channel_type+str(fdts)+'/thS_'+str(chan_index+chan_offset)+'.npy')
-        allU=np.load('./Precoders_generated/6bit_'+channel_type+str(fdts)+'/allU_'+str(chan_index+chan_offset)+'.npy')
-        allU_vec_copy=np.load('./Precoders_generated/6bit_'+channel_type+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')
-        allU_vec=np.load('./Precoders_generated/6bit_'+channel_type+str(fdts)+'/allU_vec'+str(chan_index+chan_offset)+'.npy')        
[…]

### 2019-10-17 main.py

@@ -23,8 +23,10 @@ def sigint_handler(signal, frame):
     sys.exit(0)
 
 def main():
-    args = parser.parse_args()
-    with open(args.constants, 'r') as f:
+    # channel_type="Pedestrian"
+    channel_type="Vehicular"
+    args = "config/"+ channel_type + ".yml"
+    with open(args, 'r') as f:
         constants = munchify(yaml.load(f)) 
     #---------------------------------------------------------------------------
     #Lambda Functions
@@ -35,10 +37,9 @@ def main():
     signal.signal(signal.SIGINT, sigint_handler)
     # itpp.RNG_randomize()
     itpp.RNG_reset(81)
-    # channel_type="Vehicular"
-    channel_type="Pedestrian"
-    # c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
-    c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
+
+    c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
+    # c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
     Ts=constants.Ts
     num_subcarriers=constants.num_subcarriers
     Nt=constants.Nt
@@ -47,7 +48,6 @@ def main():
     gtheta_size=(2*Nt*Nr-Nr**2-Nr)//2
     vec_size=2*Nt*Nr-Nr**2
     B=constants.B
-
     #---------------------------------------------------------------------------
     # Simulation Parameters
     number_simulations=constants.number_simulations # equivalent of time
@@ -68,12 +68,15 @@ def main():
     qtiz_rate=constants.qtiz_rate
     # qtiz_rate=1.3
 
-    Q_error=np.zeros((number_simulations,feedback_mats),dtype=np.float64) #
+    Q_error_even=np.zeros((number_simulations//2,feedback_mats),dtype=np.float64) #
+    Q_error_odd=np.zeros((number_simulations//2,feedback_mats-1),dtype=np.float64) #
+    
     onlyt_Q_error=np.zeros((number_simulations,feedback_mats),dtype=np.float64) #
+    q_error=np.zeros((10,100))
     # Use Stiefel Chordal Distance as norm
     # norm_fn='stiefCD'
     start_time=time.time()
-    save=True
+    save=False
[…]

### 2019-10-17 mimo_tdl_channel.py

@@ -412,11 +412,11 @@ def qtiz_func(x_vec,B,m,n):
     # cdf = sin^(2l)(theta); icdf = arcsin(y^(1/2l))
     for l in range(m-1):
         theta_level[l]=np.arcsin(y[l]**(1/(2*(l+1))))    
-    phi_level=np.linspace(-np.pi,np.pi,2**B-2**5)
+    phi_level=np.linspace(-3*np.pi,3*np.pi,2**B)
     # pdb.set_trace()
-    phi_level_plus=np.linspace(np.pi,4*np.pi,2**4)
-    phi_level_minus=np.linspace(-4*np.pi,np.pi,2**4)
-    phi_level=np.concatenate((phi_level_minus,phi_level,phi_level_plus))
+    # phi_level_plus=np.linspace(np.pi,4*np.pi,2**4)
+    # phi_level_minus=np.linspace(-4*np.pi,np.pi,2**4)
+    # phi_level=np.concatenate((phi_level_minus,phi_level,phi_level_plus))
     g_theta=x_vec[:(2*m*n-n**2-n)//2]
     d_phi=x_vec[(2*m*n-n**2-n)//2:]
     qtiz_gtheta=np.zeros(np.size(g_theta))
@@ -466,7 +466,7 @@ def dpcm_pred(oU_vec,prev_qU_vec,del_vec):
 def calc_del(u_vec, qtiz_vec,beta,qtiz_rate, prev_del):
     delta=np.zeros(np.size(beta))
     for i in range(np.size(beta)):        
-        if(np.sign(u_vec[i]-qtiz_vec[i])==beta[i]):
+        if((np.sign(u_vec[i]-qtiz_vec[i])==beta[i])):
             delta[i]=qtiz_rate*prev_del[i]
         else:
             delta[i]=prev_del[i]/(qtiz_rate*2.5) # faster decay

### 2019-10-17 new_results/plot_results.py

@@ -1,0 +1,62 @@
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import rc
+from matplotlib.pyplot import figure
+
+font = {'size'   : 18}
+plt.rc('font', **font)
+# plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=20)
+plt.rc('ytick', labelsize=20)
+
+plt.figure(figsize=(8, 6))
+eb_n0=np.arange(-3,25,3)
+
+# cubic=np.load('6bit_eval/hp_ber.npy')
+# linear0=np.load('updated_6bit_eval/pedestrian/0.1/hp_cap.npy')
+# linear00=np.load('updated_6bit_eval/pedestrian/0.1/hp_cap.npy')
+# linear0=linear0/linear00
+
+# linear1=np.load('updated_6bit_eval/pedestrian/0.01/hp_cap.npy')
+# linear11=np.load('updated_6bit_eval/pedestrian/0.01/max_cap.npy')
+# linear1=linear1/linear11
+
+# linear2=np.load('updated_6bit_eval/pedestrian/0.001/hp_cap.npy')
+# linear22=np.load('updated_6bit_eval/pedestrian/0.001/max_cap.npy')
+# linear2=linear2/linear22
+veh_after5=np.array([2.23854902e-02, 5.85075132e-03, 9.43905410e-04, 1.79587744e-04, 7.84925235e-05, 4.36302923e-05, 2.41909232e-05, 1.23645413e-05, 5.51285282e-06, 1.96887601e-06])
+ped_after5=np.load('ped_after5/0.00035/hp_ber.npy')[:10]
+ped_par=np.load('pedestrian/0.00035/hp_ber.npy')[:10]
+ped_man=np.load('agrim_manifold_pedestrian/0.00035/hp_ber.npy')[:10]
+ped_ideal=np.load('pedestrian/0.00035/ideal_ber.npy')[:10]
+
+plt.yscale('log')
+plt.ylabel('BER')
+plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
+# pdb.set_trace()
+# plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$\times 10^{-1}$')
+# plt.plot(eb_n0,linear1, 'g--o', lw=3, ms=10, label=r'$\times 10^{-2}$')
[…]

### 2019-10-17 plot_qerror.py

@@ -1,0 +1,42 @@
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import rc
+from matplotlib.pyplot import figure
+
+font = {'size'   : 18}
+plt.rc('font', **font)
+plt.rc('text', usetex=True)
+plt.rc('xtick', labelsize=20)
+plt.rc('ytick', labelsize=20)
+
+plt.figure(figsize=(8, 6))
+eb_n0=np.arange(0,21,3)
+stief_veh_scal=np.load('qerror_scal/veh_qerror_stief_scalar.npy')
+stief_veh_mani=np.load('qerror_mani/veh_qerror_mani_stief.npy')
+frob_veh_scal=np.load('qerror_scal/veh_frob_scala_qerror.npy')
+frob_veh_mani=np.load('qerror_mani/veh_qerror_mani_frob.npy')
+
+stief_ped_scal=np.load('qerror_scal/ped_qerror_scal_stief.npy')
+stief_ped_mani=np.load('qerror_mani/ped_qerror_mani_stief.npy')
+frob_ped_scal=np.load('qerror_scal/ped_qerror_scal_frob.npy')
+frob_ped_mani=np.load('qerror_mani/ped_qerror_mani_frob.npy')
+
+plt.ylabel('Quantization Error Frobenius avg')
+plt.xlabel('Time')
+
+# plt.plot(stief_veh_mani,label='Vehicular Manifold')
+# plt.plot(stief_veh_scal, label='Vehicular Scalar')
+# plt.plot(stief_ped_mani,label='Pedestrian Manifold')
+# plt.plot(stief_ped_scal, label='Pedestrian Scalar')
+
+# plt.plot(frob_veh_mani,label='Vehicular Manifold')
+# plt.plot(frob_veh_scal, label='Vehicular Scalar')
+plt.plot(frob_ped_mani,label='Pedestrian Manifold')
+plt.plot(frob_ped_scal, label='Pedestrian Scalar')
+
+
+plt.grid(which='major', linewidth='0.5', color='black');
+plt.grid(which='minor', linewidth='0.5', color='grey');
+
+plt.legend(fontsize=14.5)
+plt.show()

### 2019-10-17 qrtz_error.py

@@ -1,0 +1,138 @@
+# Code to generate the generates precoders corresponding to 
+# 100 channel evolutions for 10 independent channel realizations
+# using both the time based and hopping based schemes 
+# (i.e. fill 10 independent 100*`num_subcarriers` 
+# time-frequency bins matrix with generated precoders), and
+# save them as .npy files
+#---------------------------------------------------------------------------
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.linalg as la
+from scipy.interpolate import griddata
+import itpp
+from mimo_tdl_channel import *
+import sys
+import signal
+import pdb
+import time
+import copy
+from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK,calculate_BER_performance_QAM256,waterfilling
+from munch import munchify
+import yaml
+from config.arguments import parser
+    
+#SIGINT handler
+def sigint_handler(signal, frame):
+    #Do something while breaking
+    pdb.set_trace()
+    sys.exit(0)
+def main():
+    diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
+    # channel_type="Vehicular"
+    channel_type="Pedestrian"
+    args = "config/"+ channel_type + ".yml"
+    with open(args, 'r') as f:
+        constants = munchify(yaml.load(f)) 
+
+    #---------------------------------------------------------------------------
+    #Channel Params
+    signal.signal(signal.SIGINT, sigint_handler)
+    #number of db values for which logsum rate will be calculated
+    num_Cappar=7
+    avg_hpcap=np.zeros(num_Cappar)
+    avg_otcap=np.zeros(num_Cappar)
+    avg_maxcap=np.zeros(num_Cappar)
+    # Eb_N0_dB=np.arange(-6,6,2)
+
+    Eb_N0_dB=np.arange(-3,25,3)
+    # Eb_N0_dB=np.arange(-3,36,3)
+    #Store BER here
+    idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
+    hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
+    otBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
+    Ts=constants.Ts # 5e-8
+    num_subcarriers=constants.num_subcarriers #64
+    Nt=constants.Nt #4
+    Nr=constants.Nr #2
[…]

### 2019-10-17 sumrate_BER.py

@@ -110,8 +110,10 @@ def calculate_BER_performance_QPSK(H_sequence, precoder_sequence, Eb_N0_dB,flag=
     channel_out=np.matmul(H_F_sequence, np.array(np.hsplit(QPSK_stream, N_subcarrier)))
     noise=(N0/2)*np.random.randn(channel_out.shape[0], channel_out.shape[1], channel_out.shape[2])+(1.0j)*(N0/2)*np.random.randn(channel_out.shape[0], channel_out.shape[1], channel_out.shape[2])
     channel_out_with_noise=channel_out+noise
+    # pdb.set_trace()
     est_output=apply_MMSE_decoder(H_F_sequence, Es, channel_out_with_noise,N0)
     #MMSE
+
     #print(est_output.shape)
     flattened_op=est_output
     flattened_op=flattened_op.flatten()
@@ -224,7 +226,7 @@ def apply_MMSE_decoder(H_F_sequence, Es, input_sequence, N0=1.0):
     N_subcarriers=H_F_sequence.shape[0]
     H_F_h_sequence=np.conjugate(np.transpose(H_F_sequence, (0,2,1)))
     A_sequence=np.matmul(H_F_h_sequence, H_F_sequence)
-    B_sequence=np.tile(input_sequence.shape[1]*(N0/Es)*np.eye(input_sequence.shape[1]), (N_subcarriers,1,1))
+    B_sequence=np.tile((N0/Es)*np.eye(input_sequence.shape[1]), (N_subcarriers,1,1))
     C_sequence=np.linalg.inv(A_sequence+B_sequence)
     G_sequence=np.matmul(C_sequence, H_F_h_sequence)
     output_sequence=np.matmul(G_sequence, input_sequence)

### 2019-11-01 paper/main.tex
Add minor changes

@@
+\documentclass[journal,10pt,twocolumn]{IEEEtran}
 \def\b0{{\bf 0}}
@@
 %package list
-\documentclass[conference]{IEEEtran}
 \IEEEoverridecommandlockouts
@@
   construction at the transmitter.  Simulations reveal that the
-  proposed method achieves low BERs for various channel profiles
+  proposed method achieves lower BERs for various channel profiles
   compared to other approaches.
@@
 
-The use of multiple antennas at the transmitter and receiver can significantly enhance the performance of wireless systems.
-Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
-To this end, feeding back the CSI from the receiver to the transmitter efficiently is necessary.
-In wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across all subcarriers, and this places a large feedback demand.
-However, since the CSI varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency domain interpolation temporal prediction.
-In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to achieve link capacity, as well as to parallelize the channel.
-In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
+The use of multiple antennas at the transmitter and receiver can
+significantly enhance the performance of wireless systems. Moreover,
+in these multiple-input multiple-output (MIMO) wireless systems, the
+transmitter can allocate resources more efficiently if the channel
+state information (CSI) is made available at the transmitter, both to
+improve data rates and to reduce BER~\cite{love2008overview}. To this
+end, feeding back the CSI from the receiver to the transmitter
+efficiently is necessary. In wireless systems that use orthogonal
+frequency division multiplexing (OFDM), the feedback needs to be
+provided to the transmitter across all subcarriers, and this places a
+large feedback demand. However, since the CSI varies gradually in both
+time and frequency, effective techniques to reduce feedback overheads
+exist, using frequency domain interpolation temporal prediction. In
+particular, knowledge of the precoding matrices (that are typically
+the orthonormal right singular vectors of the channel matrix) is
+essential to achieve link capacity, as well as to parallelize the
+channel. In this paper, we focus on the problem of tracking the
+precoders of a MIMO-OFDM system across time and frequency for
+efficient CSI feedback by tracking a parametrized version of the matrices.
 
 The problem of effective quantization and feedback of precoders for
-MIMO-OFDM has attracted significant attention in recent years. Many
-approaches involve codebook design over the Grassmann
+MIMO-OFDM systems has attracted significant attention in recent
+years. Many approaches involve codebook design over the Grassmann
 manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
@@
 
-In this paper, we propose a method for exploiting the joint time-frequency correlation of the precoder's scalar parameters on the transmitter using CSI feedback.
-Typical approaches to predictive quantization and interpolation do not exploit the joint time-frequency correlation to enhance the precoder estimate at the transmitter.
-However, exploiting the correlations jointly not only reduces the feedback requirement, but also enhances the precoder reconstruction accuracy significantly.
-Further, the independence of the scalar parameters reduces the problem to interpolation of separate scalar valued function, as opposed to operations over tangent spaces in manifolds~\cite{Gupt1905:Predictive}.
-Simulations reveal that the proposed method offers matches the performance of manifold based approaches with a much lower implementation complexity and fewer number of bits.
+In this paper, we propose a method for exploiting the joint
+time-frequency correlation of the precoder's scalar parameters on the
+transmitter using CSI feedback. Typical approaches to predictive
+quantization and interpolation do not exploit the joint time-frequency
+correlation to enhance the precoder estimate at the
+transmitter. However, exploiting the correlations jointly not only
+reduces the feedback requirement, but also enhances the precoder
+reconstruction accuracy significantly. Further, the independence of
+the scalar parameters reduces the problem to one that involves
+interpolation of separate scalar valued functions, as opposed to
+operations over tangent spaces in
+manifolds~\cite{Gupt1905:Predictive}. Simulations reveal that the
+proposed method matches the performance of manifold based approaches
+with a much lower implementation complexity and fewer number of bits
+for quantization.
 
@@
 Finally, Section~\ref{section4} concludes with discussions on future work.
-
-%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
-In an OFDM channel, this data consumes significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
-For e.g.
-an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no.
-of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame.
-This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies.
-Also, the matrices which need to be fed back could be broken down into independent scalar parameters.
-These scalar quantities could be later quantized and interpolated.
-
-% The rest of this paper is organized as follows: Section~\ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
-In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
-%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
-In an OFDM channel, this data consume significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
-For e.g.
-an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no.
-of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame.
-This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies.
-Also, the matrices which need to be fed back could be broken down into independent scalar parameters.
-These scalar quantities could be later quantized and interpolated.
-% Broad summary of your work, come to your problem definition.
 

### 2019-11-01 paper/main.tex
More changes

@@
 For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver
-antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel can be modelled as
+antennas, the channel for the $i$-th subcarrier at time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel can be modelled as
 \label{section2}
@@
 \end{equation}
-where the channel matrix is
-$\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input data vector is
-$\bx_{i,t} \in \mathbb{C}^{N_d}$; the output is
-$\by_{i,t} \in \mathbb{C}^{N_R}$; the unitary precoder matrix is
-$\tilde{\bV}_{i,t} \in \mathbb{C}^{N_T \times N_R}$; and the complex
-additive white Gaussian noise is $\eta$ whose distribution is
-${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R < N_T$
-and the data vector size $N_d = N_R$.  The ``thin'' SVD (that keeps
-only those right and left singular vectors that correspond to non-zero
-singular values) of $\bH$ is given by
+where $\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$ is the channel
+matrix, $\bx_{i,t} \in \mathbb{C}^{N_d}$ is the input data vector,
+$\by_{i,t} \in \mathbb{C}^{N_R}$ is the output vector,
+$\tilde{\bV}_{i,t} \in \mathbb{C}^{N_T \times N_R}$ is the unitary
+precoder, and $\eta$ is the complex additive white Gaussian noise
+distributed as ${\mathcal{CN}}(0,I_{N_R})$. We consider the case where
+$N_R < N_T$, and the data vector size $N_d = N_R$.  The ``thin'' SVD
+(that keeps only those right and left singular vectors that correspond
+to non-zero singular values) of $\bH$ is given by
 $\bH_{i,t} = \bU_{i,t} \Sigma_{i,t} \bV_{i,t}^{H}$, where
@@
 $\Sigma_{i,t}$ contains the singular values
-$\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of $\bH_{i,t}$. The
+$\sigma_1 \geq \ldots \geq \sigma_{N_R} > 0$ of $\bH_{i,t}$. The
 entries of $\bH_{i,t}$ are assumed to be distributed
-i.i.d. $\tilde{\mathcal{N}}(0,1)$. To obtain a precoder at the
+i.i.d. $\\mathcal{CN}(0,1)$. To obtain a precoder at the
 transmitter, we need to quantize and feed back information about

### 2019-11-01 paper/main.tex
Further changes

@@
 $\mbox{diag}\big( {\bf 1}_{k-1}, e^{j\phi_{k,k}},\ldots,
-e^{j\phi_{k,N_T }} \big)$, ${\bf 1}_{k-1}$ is represents $k-1$ ones,
+e^{j\phi_{k,N_T }} \big)$, ${\bf 1}_{k-1}$ represents $k-1$ ones,
 and the $N_R \times N_T$ matrix $\tilde{\bI}$ =
@@
 \begin{bmatrix}
-I_{m-2} & & & \\
+\bI_{m-2} & & & \\
 & \cos\theta & -\sin\theta & \\
 & \sin\theta & \cos\theta & \\
-& & & I_{t-m}
+& & & \bI_{t-m}
 \end{bmatrix}
@@
 instant $t$, as represented in Fig.~\ref{fig:adpm-fig}. It is known
-that $\phi_k \in (-\pi, + \pi]$ is uniformly distributed between the
-two extremes~\cite{4114278}, while $\theta_l$ is distributed as
+that $\phi_k \in (-\pi, \pi]$ is uniformly distributed between the
+two extremes~\cite{4114278}, while $\theta_l$ is distributed as~\cite{4114278}
 \begin{equation}
-2l(\sin\theta_l)^{2l-1}\cos\theta_l, \theta_l \in \left[0, \frac{\pi}{2}\right) \mbox{ , } l = 1,2,\ldots,N_T -1
+2l(\sin\theta_l)^{2l-1}\cos\theta_l, \mbox{  }\theta_l \in \left[0, \frac{\pi}{2}\right) \mbox{, } l = 1,2,\ldots,N_T -1
 \end{equation}
-The $\phi_k$ and $\theta_l$ are statistically independent, making this
-representation useful for tracking them separately across time, as
-explained in the following subsection. The total number of parameters
-obtained from the decomposition of a complex orthogonal matrix
-$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$. The
-number of $\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
-number of $\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\Phi_{i,t}$ is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while $\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.
+In addition, $\phi_k$ and $\theta_l$ are statistically independent,
+making this representation useful for tracking them separately across
+time, as discussed below. The total number of
+parameters obtained from the decomposition of a complex orthogonal
+matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$. The number of
+$\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of
+$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\Phi_{i,t}$
+is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while
+$\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.
 \begin{figure}
@@
 \end{figure}
-% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.
-
 \subsection{Differential quantization and channel tracking}
@@
 value. Further, the step size for the move can be adapted based on the
-speed of channel variation as well the speed for convergence. If
-$\Delta_n$ is the step size, we define our adaptive quantizer as
+speed of channel variation as well the accuracy needed. If $\Delta_n$
+is the step size, we define our adaptive quantizer as
 \begin{equation}
@@
 because the underlying parameters can generally be thought to emerge
-from an autoregressive (AR) process. For such processes, it is well known
-that linear prediction based on past samples is optimal. Even though
-the $\Theta$ and $\Phi$ parameters do not manifest as AR processes,
-for small changes, a linear approximation works sufficiently
-well, as in the description in~\cite{4114278}, although that was
-limited to tracking the temporal evolution of the parameters.
+from an autoregressive (AR) process. For such processes, it is well
+known that linear prediction based on past samples is optimal. Even
+though the $\Theta$ and $\Phi$ parameters do not manifest as AR
+processes, for small changes, a linear approximation works
+sufficiently well, as described in~\cite{4114278}, although that
+consideration was limited to tracking the temporal evolution of the
+parameters.
 

### 2019-11-21 paper/main.tex
adding cost plots

--- 2019-11-01 paper/main.tex
@@
-\documentclass[journal,10pt,twocolumn]{IEEEtran}
 \def\b0{{\bf 0}}
@@
 %package list
+\documentclass[conference]{IEEEtran}
 \IEEEoverridecommandlockouts
@@
   construction at the transmitter.  Simulations reveal that the
-  proposed method achieves lower BERs for various channel profiles
+  proposed method achieves low BERs for various channel profiles
   compared to other approaches.
@@
 
-The use of multiple antennas at the transmitter and receiver can
-significantly enhance the performance of wireless systems. Moreover,
-in these multiple-input multiple-output (MIMO) wireless systems, the
-transmitter can allocate resources more efficiently if the channel
-state information (CSI) is made available at the transmitter, both to
-improve data rates and to reduce BER~\cite{love2008overview}. To this
-end, feeding back the CSI from the receiver to the transmitter
-efficiently is necessary. In wireless systems that use orthogonal
-frequency division multiplexing (OFDM), the feedback needs to be
-provided to the transmitter across all subcarriers, and this places a
-large feedback demand. However, since the CSI varies gradually in both
-time and frequency, effective techniques to reduce feedback overheads
-exist, using frequency domain interpolation temporal prediction. In
-particular, knowledge of the precoding matrices (that are typically
-the orthonormal right singular vectors of the channel matrix) is
-essential to achieve link capacity, as well as to parallelize the
-channel. In this paper, we focus on the problem of tracking the
-precoders of a MIMO-OFDM system across time and frequency for
-efficient CSI feedback by tracking a parametrized version of the matrices.
+The use of multiple antennas at the transmitter and receiver can significantly enhance the performance of wireless systems.
+Moreover, in these multiple-input multiple-output (MIMO) wireless systems, the transmitter can perform better resource allocation if channel state information (CSI) is made available at the transmitter, both to improve data rates and reduce BER~\cite{love2008overview}.
+To this end, feeding back the CSI from the receiver to the transmitter efficiently is necessary.
+In wireless systems that use orthogonal frequency division multiplexing (OFDM), the feedback needs to be provided to the transmitter across all subcarriers, and this places a large feedback demand.
+However, since the CSI varies gradually in both time and frequency, effective techniques to reduce feedback overheads exist, using frequency domain interpolation temporal prediction.
+In particular, knowledge of the precoding matrices (that are typically the orthonormal right singular vectors of the channel matrix) is essential to achieve link capacity, as well as to parallelize the channel.
+In this paper, we focus on the problem of tracking the precoders of a MIMO-OFDM system across time and frequency for efficient CSI feedback.
 
 The problem of effective quantization and feedback of precoders for
-MIMO-OFDM systems has attracted significant attention in recent
-years. Many approaches involve codebook design over the Grassmann
+MIMO-OFDM has attracted significant attention in recent years. Many
+approaches involve codebook design over the Grassmann
 manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
@@
 
-In this paper, we propose a method for exploiting the joint
-time-frequency correlation of the precoder's scalar parameters on the
-transmitter using CSI feedback. Typical approaches to predictive
-quantization and interpolation do not exploit the joint time-frequency
-correlation to enhance the precoder estimate at the
-transmitter. However, exploiting the correlations jointly not only
-reduces the feedback requirement, but also enhances the precoder
-reconstruction accuracy significantly. Further, the independence of
-the scalar parameters reduces the problem to one that involves
-interpolation of separate scalar valued functions, as opposed to
-operations over tangent spaces in
-manifolds~\cite{Gupt1905:Predictive}. Simulations reveal that the
-proposed method matches the performance of manifold based approaches
-with a much lower implementation complexity and fewer number of bits
-for quantization.
+In this paper, we propose a method for exploiting the joint time-frequency correlation of the precoder's scalar parameters on the transmitter using CSI feedback.
+Typical approaches to predictive quantization and interpolation do not exploit the joint time-frequency correlation to enhance the precoder estimate at the transmitter.
+However, exploiting the correlations jointly not only reduces the feedback requirement, but also enhances the precoder reconstruction accuracy significantly.
+Further, the independence of the scalar parameters reduces the problem to interpolation of separate scalar valued function, as opposed to operations over tangent spaces in manifolds~\cite{Gupt1905:Predictive}.
+Simulations reveal that the proposed method offers matches the performance of manifold based approaches with a much lower implementation complexity and fewer number of bits.
 
@@
 Finally, Section~\ref{section4} concludes with discussions on future work.
+
+%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
+In an OFDM channel, this data consumes significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
+For e.g.
+an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no.
+of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame.
+This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies.
+Also, the matrices which need to be fed back could be broken down into independent scalar parameters.
+These scalar quantities could be later quantized and interpolated.
+
+% The rest of this paper is organized as follows: Section~\ref{section2}, we start by explaining the system model and the method for scalar quantization \& interpolation of the precoder matrix.
+In \ref{section3}, we perform various simulations for our approach and show the benefits from the improved BER rates and Achievable Rate.
+%Since $H=U\Sigma V$, it is sufficient to quantize V*(precoder) and send this information back.
+In an OFDM channel, this data consume significant amount of bits and therefore, sending the complete data could consume too much bandwidth data.
+For e.g.
+an OFDM channel with 4 transmitters, 2 receiver antennas, and 64 subcarriers would require 64*(no.
+of bits required to send one precoder ~ 12-16 bits) ~ 896 bits per frame.
+This information could be significantly reduced if we provide better quantization and interpolation techniques/strategies.
+Also, the matrices which need to be fed back could be broken down into independent scalar parameters.
+These scalar quantities could be later quantized and interpolated.
+% Broad summary of your work, come to your problem definition.
 
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber-pedestrian-35e-4}
+\includegraphics[width=0.5\textwidth]{images/cost0035}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
-  channel with the ITU Pedestrian-A channel profile for $f_dT_s = 3.5\times 10^{-4}$.}
+  channel with the ITU Cost207-RA channel profile for $f_dT_s = 3.5\times 10^{-3}$.}
 \label{fig:ber_ped}
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber-vehicular-35e-4}
+\includegraphics[width=0.5\textwidth]{images/cost01}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
-  channel with the ITU Vehicular-A channel profile for $f_dT_s = 3.5\times 10^{-4}$.}
+  channel with the ITU COST207-RA channel profile for $f_dT_s = 1\times 10^{-2}$.}
 \label{fig:ber_veh}

### 2019-12-02 paper/main.tex
Minor changes

@@
 independent scalar parameters that correspond to Givens rotations and
-Household Transformations of the unitary precoding matrices
+Householder transformations of the unitary precoding matrices
 \cite{4114278,4556174}. These approaches have been combined with
@@
 
-The rest of this paper is organized as follows: Section~\ref{section2} describes the system model and the precoder feedback and interpolation techniques.
-Section~\ref{section3} compares the BER and achievable rates for our proposed methods with other approaches.
-Finally, Section~\ref{section4} concludes with discussions on future work.
+The rest of this paper is organized as follows: Section~\ref{section2}
+describes the system model and the precoder feedback and interpolation
+techniques. Section~\ref{section3} compares the BER and achievable
+rates for our proposed methods with other approaches. Finally,
+Section~\ref{section4} concludes with discussions on future work.
 
@@
 For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver
-antennas, the channel for the $i$-th subcarrier at time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel can be modelled as
+antennas, the channel for the $i$-th subcarrier at time instant $t$
+can be modelled as a flat-fading MIMO channel. This flat-fading MIMO
+channel can be modelled as
 \label{section2}
@@
 transmitter, we need to quantize and feed back information about
-$\bV_{i,t}$ for each subcarrier.
+$\bV_{i,t}$ for each subcarrier. However, the frequency coherence and
+temporal correlations would reduce the overall feedback requirement,
+as discussed in subsequent sections.
 
@@
 matrix~\cite{4114278}. The orthonormal matrix $\tilde{\bV}_{i,t}$ with
-$N_T$ rows and $N_R$ columns can be decomposed as follows~\cite{4114278} (we suppress
-the $(i, t)$ subscripts in the decomposition matrices)
+$N_T$ rows and $N_R$ columns can be decomposed as
+follows~\cite{4114278} (we suppress the $(i, t)$ subscripts in the
+decomposition matrices)
 \begin{equation}
@@
 instant $t$, as represented in Fig.~\ref{fig:adpm-fig}. It is known
-that $\phi_k \in (-\pi, \pi]$ is uniformly distributed between the
-two extremes~\cite{4114278}, while $\theta_l$ is distributed as~\cite{4114278}
+that $\phi_k \in (-\pi, \pi]$ is uniformly distributed between the two
+extremes~\cite{4114278}, while $\theta_l$ is distributed
+as~\cite{4114278}
 \begin{equation}
@@
 \includegraphics[width=0.5\textwidth]{images/new-adpm.pdf}
-\caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
+\caption{\label{fig:adpm-fig}Subcarriers which need to be
+  quantized. In the figure, we see that the subcarrier indicated by
+  red arrows has a precoder whose parameters are determined jointly
+  using the neighbouring parameters, both in time and frequency.}
 \end{figure}
@@
 $\pi$ due to jumps of $2\pi$. We have avoided this by unwrapping the
-phases to facilitate continuous tracking.
+phases to facilitate continuous tracking (i.e., we avoid jumps in
+phase of magnitude close to $2\pi$).
 \subsection{Joint Time Frequency Interpolation}
@@
 the transmitter, we can construct the precoder at subcarrier indices
-$pk$ or $pk+q$. To construct the rest of the subcarriers, we will use
+$pk$ or $pk+q$, in keeping with the same notation as the previous
+section. To construct the rest of the subcarriers, we will use
 precoder parameters from both the past and future and time instances
@@
 systems. The channel is simulated using the typical Jakes Model for
-both the ITU Vehicular-A channel and ITU Pedestrian-A channel power
-delay profiles~\cite{recommendation1997guidelines}. The BER achieved by the proposed predictive
-quantization based precoder is compared against the ideal precoder for
-100 channel evolutions in each run and averaged. The normalised
-Doppler is assumed to be $f_dT_s = 3.5\times10^{-4}$, with symbol time
-period $T_s = 5\times10^{-8} s$. For both the Vehicular and Pedestrian
+both the ITU Vehicular-A channel and the COST channel power delay
+profiles~\cite{recommendation1997guidelines,molisch2006cost259,cost1989cost}. The
+BER achieved by the proposed predictive quantization based precoder is
+compared against the ideal precoder for 100 channel evolutions in each
+run and averaged. The normalised Doppler is assumed to be
+$f_dT_s = 3.5\times10^{-4}$, with symbol time period
+$T_s = 5\times10^{-8} s$. For both the Vehicular and Pedestrian
 channel profiles, we consider $N_T=4$ and $N_R=2$. Therefore, the
-total number of parameters that determine the precoder are
-$N_{R}(2N_{T} - N_{R}) = 12$, with 5$\theta_l$s and
-7$\phi_i$s. For the Vehicular channel profile, we assume that the
-number of subcarriers is 1024, while for the Pedestrian channel
-profile it is 64. We take the spacing between fed back subcarrier
-indices $p$ to be $33$ and $9$ for the Vehicular
-and Pedestrian channel profiles respectively (i.e. for the Vehicular
-profile, the subcarriers indices fed back would be $0, 33, 66, \ldots
-1023$ and for the Pedestrian, $0, 9, 18, \ldots 63$).
+total number of parameters that determine the precoder is
+$N_{R}(2N_{T} - N_{R}) = 12$, with five $\theta_l$s and and seven
+$\phi_i$s. For the fast fading channels, we assume that the number of
+subcarriers is 1024, while for the slower varying channel profiles, we
+take 64 subcarriers. We assume that the spacing between fed back
+subcarrier indices $p$ to be $33$ and $9$ for the fast and slow
+channel profiles respectively (i.e. for the slower channels, the
+subcarriers indices fed back would be $0, 33, 66, \ldots 1023$ and for
+the Pedestrian, $0, 9, 18, \ldots 63$).
 

### 2019-12-05 paper/main.tex
Changes by Kumar

@@
 
- \author{\authorblockN{Author 1, Author 2, Author 3}
+ \author{\authorblockN{Author 1, Author 2, Author 3\\}
  \
- \authorblockA{Department of Electrical Engineering, Indian Institute of Technology Bombay}
+ \authorblockA{Department of Electrical Engineering\\Indian Institute of Technology Bombay}
 % \thanks{Parts of this work was supported by the Bharti Centre for Communication in
@@
 \end{figure}
+
+The quantization approach described above is related to the channel
+state information feedback present in several wireless standards,
+including 802.11ac~\cite{lou2013comparison}, although adaptive
+quantization and feedback are not discussed. We show in the further
+sections that the use of adaptive approaches can greatly enhance
+performance with only very minor modifications to the system.
 \subsection{Differential quantization and channel tracking}
@@
 systems. The channel is simulated using the typical Jakes Model for
-both the ITU Vehicular-A channel and the COST channel power delay
-profiles~\cite{recommendation1997guidelines,molisch2006cost259,cost1989cost}. The
-BER achieved by the proposed predictive quantization based precoder is
-compared against the ideal precoder for 100 channel evolutions in each
-run and averaged. The normalised Doppler is assumed to be
-$f_dT_s = 3.5\times10^{-4}$, with symbol time period
-$T_s = 5\times10^{-8} s$. For both the Vehicular and Pedestrian
-channel profiles, we consider $N_T=4$ and $N_R=2$. Therefore, the
-total number of parameters that determine the precoder is
+for the COST207 channel power delay
+profiles~\cite{molisch2006cost259,cost1989cost}. The BER achieved by
+the proposed predictive quantization based precoder is compared
+against the ideal precoder for 100 channel evolutions in each run and
+averaged. Simulations are performed for normalised Doppler values of
+$f_dT_s = 3.5\times 10^{-6}$ (corresponding to a velocity of 36 km/h
+for $f_c = 1$ GHz) with 1024 subcarriers, and with
+$f_dT_s = \times 1.05 \times 10^{-5}$ (velocity of 108 km/h) with 64
+subcarriers. The sampling time period $T_s$ was $5\times10^{-8}$
+s. For both the situations, we considered $N_T=4$ transmit antennas
+and and $N_R=2$ receive antennas. Therefore, the total number of
+parameters that determine the precoder is
 $N_{R}(2N_{T} - N_{R}) = 12$, with five $\theta_l$s and and seven
-$\phi_i$s. For the fast fading channels, we assume that the number of
-subcarriers is 1024, while for the slower varying channel profiles, we
-take 64 subcarriers. We assume that the spacing between fed back
-subcarrier indices $p$ to be $33$ and $9$ for the fast and slow
-channel profiles respectively (i.e. for the slower channels, the
+$\phi_i$s. We assume that the spacing between fed back subcarrier
+indices $p$ to be $33$ and $9$ for the fast and slow channel profiles
+respectively (i.e. for the slower channels with 1024 subcarriers, the
 subcarriers indices fed back would be $0, 33, 66, \ldots 1023$ and for
-the Pedestrian, $0, 9, 18, \ldots 63$).
+the faster channel, with 64 subcarriers, $0, 9, 18, \ldots 63$ would
+be fed back).
 
@@
 
-TODO: Agrim and Mohit discuss and then fill this part.
 \begin{figure}
 \includegraphics[width=0.5\textwidth]{images/cost0035}
-\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
-  channel with the ITU Cost207-RA channel profile for $f_dT_s = 3.5\times 10^{-3}$.}
+\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
+  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
+  for $f_dT_s = 3.5\times 10^{-6}$, corresponding to a velocity of 36 km/h.}
 \label{fig:ber_ped}
@@
 approximate the precoder very accurately. Since the channel variation
-across subcarriers is gradual, owing to relatively small delay spread
-of the Pedestrian channel profile, linear interpolation and prediction
-yields good performance. Moreover, this method outperforms the Stiefel
-manifold based approach to the fact that the adaptation is much more
+across subcarriers is gradual, owing to the limited delay spread of
+the channel profile, linear interpolation and prediction yields good
+performance. Moreover, this method outperforms the Stiefel manifold
+based approach due to the fact that the adaptation is much more
 accurate for scalar parameters than the tangent space based adaptation
-used for the Stiefel manifold.
+used over the Stiefel manifold.
 
-Fig.~\ref{fig:ber_veh} shows a similar comparison for the Vehicular
-channel profile. The larger delay spread of the channel necessitates
-the use of a larger number of subcarriers (1024) than the Pedestrian
-channel (64). The larger number of subcarriers implies that the
+Fig.~\ref{fig:ber_veh} shows a similar comparison for a similar
+situation with a higher speed. The faster variation of the channel
+taps with time necessitates the use of a smaller number of
+subcarriers. The smaller number of subcarriers implies that the
 predictive quantization and interpolation needs to be performed across
-a larger frequency range, thereby reducing the accuracy. Even in this
-case, we observe that the scalar parameter based approach outperforms
-manifold based quantization by about 5 dB.
+a larger frequency range, thereby potentially reducing the
+accuracy. Even in this case, we observe that the scalar parameter
+based approach outperforms manifold based quantization by about 5 dB.
 
@@
 \includegraphics[width=0.5\textwidth]{images/cost01}
-\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ MIMO
-  channel with the ITU COST207-RA channel profile for $f_dT_s = 1\times 10^{-2}$.}
+\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
+  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
+  for $f_dT_s = 1.05\times 10^{-5}$, corresponding to a velocity of 108 km/h.}
 \label{fig:ber_veh}
@@
 tangent spaces. Future work would involve tuning the adaptation rate
-of the scalar parameters and extending this to TODO.
+of the scalar parameters and further reductions of quantization
+redundancy to reduce the feedback rates.
 

### 2019-12-09 paper/main.tex
adding cost207 plots
Add minor fixes

@@
 independent scalar parameters causes the error to decay rapidly, the
-error is lower and the convergence is faster.
+error is lower and the convergence is faster, even though the starting
+error is higher in the case of the scalar quantization approach.
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/stief_veh.pdf}
-\caption{\label{fig:error_decay}A comparison of the $4\times 2$ precoder quantization error vs. time instant for
-  the proposed approach (labelled ``TODO'') and direct quantization on
-the Stiefel manifold for the Vehicular channel profile.}
+\includegraphics[width=0.5\textwidth]{images/qerror.pdf}
+\caption{\label{fig:error_decay}A comparison of the $4\times 2$
+  precoder quantization error vs. time instant for the proposed
+  approach (labelled ``Cost207\_RA Scalar'') and direct quantization
+  on the Stiefel manifold (labelled ``Cost207\_RA Manifold'') for the COST207 channel profile.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/cost0035}
+\includegraphics[width=0.5\textwidth]{images/cost1024_0035}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/cost01}
+\includegraphics[width=0.5\textwidth]{images/cost64_001}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2019-12-12 paper/main.tex
adding plots and small changed in tex

--- 2019-12-05 paper/main.tex
@@
   precoder enables effective parameterization and quantization using
-  Givens rotations and Householder transformations. In addition, the
+  Givens rotations and Householder transformations. Also, the
   use of temporal and frequency correlations can further reduce the
@@
 improve data rates and to reduce BER~\cite{love2008overview}. To this
-end, feeding back the CSI from the receiver to the transmitter
-efficiently is necessary. In wireless systems that use orthogonal
+end, it is necessary to efficiently feedback the CSI from the receiver 
+to the transmitter. In wireless systems that use orthogonal
 frequency division multiplexing (OFDM), the feedback needs to be
@@
 being to their previous values, as shown in
-Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the the collection of all
+Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the collection of all
 $\theta_l$s and $\Phi$ is the collection of all $\phi_i$s. This brings
@@
 independent scalar parameters causes the error to decay rapidly, the
-error is lower and the convergence is faster.
+error is lower, and the convergence is faster.
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/stief_veh.pdf}
+\includegraphics[width=0.5\textwidth]{images/qerror.pdf}
 \caption{\label{fig:error_decay}A comparison of the $4\times 2$ precoder quantization error vs. time instant for
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/cost0035}
+\includegraphics[width=0.5\textwidth]{images/ber1024_035}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
   subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
-  for $f_dT_s = 3.5\times 10^{-6}$, corresponding to a velocity of 36 km/h.}
+  for $f_dT_s = 3.41\times 10^{-6}$, corresponding to a velocity of 37.8 km/h.}
 \label{fig:ber_ped}
@@
 the performance with exact precoders. This can be attributed to the
-fact that even the 1 bit update to the scalar parameters is able to
+fact that even the 1-bit update to the scalar parameters is able to
 approximate the precoder very accurately. Since the channel variation
@@
 a larger frequency range, thereby potentially reducing the
-accuracy. Even in this case, we observe that the scalar parameter
-based approach outperforms manifold based quantization by about 5 dB.
+accuracy. Even in this case, we observe that the scalar parameter-based 
+approach outperforms manifold based quantization by about 5 dB.
 
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/cost01}
+\includegraphics[width=0.5\textwidth]{images/ber64_001}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
   subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
-  for $f_dT_s = 1.05\times 10^{-5}$, corresponding to a velocity of 108 km/h.}
+  for $f_dT_s = 1.56\times 10^{-5}$, corresponding to a velocity of 172.8 km/h.}
 \label{fig:ber_veh}
@@
 based feedback of precoder parameters. By quantizing and adapting the
-Givens rotation and Householder trasnsformation parameters of the
-precoder, the receiver is able to send minimal feedback to the
+Givens rotation and Householder transformation parameters of the
+precoder, the receiver can send minimal feedback to the
 transmitter and yet enable effective precoder adaptation. Exploiting

### 2019-12-12 eval.py

@@ -194,7 +194,7 @@ if __name__ == '__main__':
 
 plt.yscale("log")
 plt.plot(Eb_N0_dB,hpBER_QPSK, marker='o', label="freq_hop")
-plt.plot(Eb_N0_dB,idealBER_QPSK, marker='>', label="ideal")
+# plt.plot(Eb_N0_dB,idealBER_QPSK, marker='>', label="ideal")
 plt.legend()
 plt.ylabel("BER")
 plt.xlabel("Eb/N0(dB)")

### 2019-12-12 paper/main.tex

@@
 independent scalar parameters causes the error to decay rapidly, the
-error is lower, and the convergence is faster.
+error is lower and the convergence is faster, even though the starting
+error is higher in the case of the scalar quantization approach.
+
 \begin{figure}
 \includegraphics[width=0.5\textwidth]{images/qerror.pdf}
-\caption{\label{fig:error_decay}A comparison of the $4\times 2$ precoder quantization error vs. time instant for
-  the proposed approach (labelled ``TODO'') and direct quantization on
-the Stiefel manifold for the Vehicular channel profile.}
+\caption{\label{fig:error_decay}A comparison of the $4\times 2$
+  precoder quantization error vs. time instant for the proposed
+  approach (labelled ``Cost207\_RA Scalar'') and direct quantization
+  on the Stiefel manifold (labelled ``Cost207\_RA Manifold'') for the COST207 channel profile.}
 \end{figure}

### 2019-12-13 paper/main.tex
Add changes

@@
 
-Fig.~\ref{fig:error_decay}, shows the rate at which error diminishes
+Fig.~\ref{fig:error_decay} depicts the rate at which error diminishes
 with time for the proposed approach and direct quantization on the
@@
 error is lower and the convergence is faster, even though the starting
-error is higher in the case of the scalar quantization approach.
+error is higher in the case of the scalar quantization approach. The
+independence of the parameters allows the convergence speed to be very
+fast. The Stiefel manifold based approach uses nested codebooks that
+may be suboptimal when compared to the parametrization using angles.
 
@@
 Fig.~\ref{fig:ber_ped} compares the BER obtained using the proposed
-quantization method with that obtained using exact precoders. We can
-observe that the performance of the proposed method is very close to
-the performance with exact precoders. This can be attributed to the
-fact that even the 1-bit update to the scalar parameters is able to
-approximate the precoder very accurately. Since the channel variation
-across subcarriers is gradual, owing to the limited delay spread of
-the channel profile, linear interpolation and prediction yields good
+quantization method with that obtained using exact precoders for a
+$4\times 2$ 1024 subcarrier system a slow fading
+($f_dT_s = 3.41 \times 10^{-6}$). We can observe that the performance
+of the proposed method tracks the performance with exact precoders for
+a wide range of SNRs. This can be attributed to the fact that even the
+1-bit update to the scalar parameters is able to approximate the
+precoder very accurately. Since the channel variation across
+subcarriers is gradual, owing to the limited delay spread of the
+channel profile, linear interpolation and prediction yields good
 performance. Moreover, this method outperforms the Stiefel manifold
@@
 accurate for scalar parameters than the tangent space based adaptation
-used over the Stiefel manifold.
+used over the Stiefel manifold, and the outperformance is in excess of
+10 dB. The gap from the optimal precoder performance narrows further
+when the speed reduces.
 
@@
 situation with a higher speed. The faster variation of the channel
-taps with time necessitates the use of a smaller number of
-subcarriers. The smaller number of subcarriers implies that the
-predictive quantization and interpolation needs to be performed across
-a larger frequency range, thereby potentially reducing the
-accuracy. Even in this case, we observe that the scalar parameter-based 
-approach outperforms manifold based quantization by about 5 dB.
+taps with time necessitates the use of a smaller number of subcarriers
+(viz. 64 subcarriers). The smaller number of subcarriers implies that
+the predictive quantization and interpolation needs to be performed
+across a larger frequency range, thereby potentially reducing the
+accuracy. Even in this case, we observe that the scalar
+parameter-based approach outperforms manifold based quantization by
+about 1 dB for higher SNRs. Naturally, the rate of channel variation
+is quite fast, which limits the efficacy of the proposed method in
+tracking the channel parameters.
 

### 2019-12-14 plot_qerror.py
adding updated figures

@@ -26,13 +26,13 @@ plt.xlabel('Time')
 
 # plt.plot(stief_veh_mani,label='Vehicular Manifold')
 # plt.plot(stief_veh_scal, label='Vehicular Scalar')
-# plt.plot(stief_ped_mani,label='Pedestrian Manifold')
-# plt.plot(stief_ped_scal, label='Pedestrian Scalar')
+plt.plot(stief_ped_mani,label='Pedestrian Manifold')
+plt.plot(stief_ped_scal, label='Pedestrian Scalar')
 
 # plt.plot(frob_veh_mani,label='Vehicular Manifold')
 # plt.plot(frob_veh_scal, label='Vehicular Scalar')
-plt.plot(frob_ped_mani,label='Pedestrian Manifold')
-plt.plot(frob_ped_scal, label='Pedestrian Scalar')
+# plt.plot(frob_ped_mani,label='Pedestrian Manifold')
+# plt.plot(frob_ped_scal, label='Pedestrian Scalar')
 
 
 plt.grid(which='major', linewidth='0.5', color='black');

### 2019-12-18 main.tex
Adding edits

--- 2019-12-13 paper/main.tex
@@
   precoder enables effective parameterization and quantization using
-  Givens rotations and Householder transformations. Also, the
-  use of temporal and frequency correlations can further reduce the
-  feedback requirements. We present a framework that utilizes
-  adaptive feedback with joint time-frequency prediction for precoder
-  construction at the transmitter.  Simulations reveal that the
-  proposed method achieves lower BERs for various channel profiles
+  Givens rotations and Householder transformations. Also, the use of
+  temporal and frequency correlations can further reduce the feedback
+  requirements. We present a framework that utilizes adaptive feedback
+  with joint time-frequency prediction for precoder construction at
+  the transmitter.  Simulations reveal that the proposed method
+  achieves significantly lower BERs for various channel profiles
   compared to other approaches.
@@
 
-The use of multiple antennas at the transmitter and receiver can
-significantly enhance the performance of wireless systems. Moreover,
-in these multiple-input multiple-output (MIMO) wireless systems, the
-transmitter can allocate resources more efficiently if the channel
-state information (CSI) is made available at the transmitter, both to
-improve data rates and to reduce BER~\cite{love2008overview}. To this
-end, it is necessary to efficiently feedback the CSI from the receiver 
+The use of multiple antennas at the transmitter and receiver (MIMO) can
+significantly enhance the performance of wireless systems. 
+A typical step to get this performance enhancements is estimation of channel
+state information (CSI).
+If transmitter also has the knowledge of CSI, it can allocate 
+resources more efficiently, both to
+improve data rates and to reduce BER~\cite{love2008overview}.
+
+To this end, it is necessary to efficiently feedback the CSI from the receiver 
 to the transmitter. In wireless systems that use orthogonal
 frequency division multiplexing (OFDM), the feedback needs to be
-provided to the transmitter across all subcarriers, and this places a
-large feedback demand. However, since the CSI varies gradually in both
+provided to the transmitter across all subcarriers and in each OFDM frame, 
+which places a large feedback demand. However, since the CSI varies gradually in both
 time and frequency, effective techniques to reduce feedback overheads
-exist, using frequency domain interpolation temporal prediction. In
+exist, using frequency domain interpolation and temporal prediction [CITE?]. In
 particular, knowledge of the precoding matrices (that are typically
-the orthonormal right singular vectors of the channel matrix) is
-essential to achieve link capacity, as well as to parallelize the
+the orthonormal right singular vectors of CSI matrix) is
+essential to achieve the link capacity, as well as to parallelize the
 channel. In this paper, we focus on the problem of tracking the
 precoders of a MIMO-OFDM system across time and frequency for
-efficient CSI feedback by tracking a parametrized version of the matrices.
+efficient CSI feedback by tracking a parametrized version of the precoding matrices.
 
@@
 of the precoders is used to obtain tangents for predictive
-quantization and interpolation across time and frequency. While these
-methods have been shown to be very effective for precoder feedback,
+quantization and interpolation across time and frequency. 
+An alternate approach is to parametrize the precoding matrices using
+scalar parameters. The space of unitary matrices parameterized into
+independent scalar parameters that correspond to Givens rotations and
+Householder transformations of the unitary precoding matrices
+\cite{4114278,4556174} [?].
+
+While the manifold based methods have been shown to be very effective for precoder feedback,
 they involve operations over higher dimensional manifolds. In
@@
 for manifolds where geodesic paths for interpolation may be difficult
-to obtain, especially for joint time-frequency predictive coding. An
-alternate approach is to parametrize the precoding matrices using
-scalar parameters. The space of unitary matrices parameterized into
-independent scalar parameters that correspond to Givens rotations and
-Householder transformations of the unitary precoding matrices
-\cite{4114278,4556174}. These approaches have been combined with
+to obtain, especially (don't repeat especially again) for joint time-frequency predictive coding.
+In addition to being analytically more tractable, the scalar parametrization methods have also been
+implemented in 802.11ac standards \cite{porat2013improved}. Past work in utilizing
+scalar parametrization for precoder feedback have utilized
 adaptive delta modulation effectively track precoding matrices both in
 the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately.
-
-In this paper, we propose a method for exploiting the joint
+separately. In this paper, we propose a method for exploiting the joint
 time-frequency correlation of the precoder's scalar parameters on the
-transmitter using CSI feedback. Typical approaches to predictive
+transmitter using CSI feedback. 
+Typical approaches to predictive
 quantization and interpolation do not exploit the joint time-frequency
@@
 reduces the feedback requirement, but also enhances the precoder
-reconstruction accuracy significantly. Further, the independence of
-the scalar parameters reduces the problem to one that involves
-interpolation of separate scalar valued functions, as opposed to
-operations over tangent spaces in
-manifolds~\cite{Gupt1905:Predictive}. Simulations reveal that the
+reconstruction accuracy significantly.
+
+Most of the papers utilizing manifolds \cite{Gupt1905:Predictive,pitaval2013codebooks,6891198}  and scalar parametrization \cite{4114278,4556174} compare the performance to other manifold/scalar parametrization methods, but don't do the cross comparison to highlight the performance benefits of the chosen method.
+In addition to a novel joint time-frequency feedback scheme, we compare also the results of our
+algorithm with an equivalent manifold based scheme, and illustrate the performance
+benefits entailed by utilizing the scalar parametrization method. 
+%  Further, the independence of
+% the scalar parameters reduces the problem to one that involves
+% interpolation of separate scalar valued functions, as opposed to
+% operations over tangent spaces in
+% manifolds~\cite{Gupt1905:Predictive}. 
+Simulations reveal that the
 proposed method matches the performance of manifold based approaches
@@
 
-The rest of this paper is organized as follows: Section~\ref{section2}
-describes the system model and the precoder feedback and interpolation
-techniques. Section~\ref{section3} compares the BER and achievable
-rates for our proposed methods with other approaches. Finally,
-Section~\ref{section4} concludes with discussions on future work.
+% The rest of this paper is organized as follows: Section~\ref{section2}
+% describes the system model and the precoder feedback and interpolation
+% techniques. Section~\ref{section3} compares the BER and achievable
+% rates for our proposed methods with other approaches. Finally,
+% Section~\ref{section4} concludes with discussions on future work.
 
@@
 can be modelled as a flat-fading MIMO channel. This flat-fading MIMO
-channel can be modelled as
+channel can be written as
 \label{section2}
@@
 entries of $\bH_{i,t}$ are assumed to be distributed
-i.i.d. $\\mathcal{CN}(0,1)$. To obtain a precoder at the
+i.i.d. $\mathcal{CN}(0,1)$. To obtain a precoder at the
 transmitter, we need to quantize and feed back information about
@@
 averaged. Simulations are performed for normalised Doppler values of
-$f_dT_s = 3.5\times 10^{-6}$ (corresponding to a velocity of 36 km/h
+$f_dT_s = 3.41\times 10^{-6}$ (corresponding to a velocity of 37.2 km/h
 for $f_c = 1$ GHz) with 1024 subcarriers, and with
-$f_dT_s = \times 1.05 \times 10^{-5}$ (velocity of 108 km/h) with 64
+$f_dT_s = 1.56 \times 10^{-5}$ (velocity of 172.8 km/h) with 64
 subcarriers. The sampling time period $T_s$ was $5\times10^{-8}$
@@
 $\phi_i$s. We assume that the spacing between fed back subcarrier
-indices $p$ to be $33$ and $9$ for the fast and slow channel profiles
+indices $p$ to be $33$ and $9$ for the slow and fast channel profiles
 respectively (i.e. for the slower channels with 1024 subcarriers, the
@@
 and 7 bits (the length of $\Phi$) over alternate feedback instances,
-thereby resulting in an average feedback budget of 5 bits.
+thereby resulting in an average feedback budget of 6 bits.
 

### 2019-12-18 ber_results/ber_results.py

@@ -1,52 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import rc
-from matplotlib.pyplot import figure
-
-font = {'size'   : 18}
-plt.rc('font', **font)
-plt.rc('text', usetex=True)
-plt.rc('xtick', labelsize=20)
-plt.rc('ytick', labelsize=20)
-
-plt.figure(figsize=(8, 6))
-eb_n0=np.arange(-3,30,3)
-
-# cubic=np.load('6bit_eval/hp_ber.npy')
-linear0=np.load('updated_6bit_eval/pedestrian/0.001/hp_ber.npy')
-linear1=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
-linear2=np.load('updated_6bit_eval/agrim_manifold_pedestrian/0.00035/hp_ber.npy')
-linear4=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
-linear0=linear0[:11]
-linear1=linear1[:11]
-linear2=linear2[:11]
-linear4=linear4[:11]
-
-plt.yscale('log')
-plt.ylabel('BER')
-plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
-# # pdb.set_trace()
-# plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$\times 10^{-1}$')
-plt.plot(eb_n0,linear4, 'c--o', lw=3, ms=10, label=r'Ideal - Complete Feedback')
-plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$f_{d}T_{s}=10^{-3}$, Scalar param.')
-# plt.plot(eb_n0,linear1,'b--o', lw=3, ms=10, label=r'$f_{d}T_{s}3.5=\times 10^{-4}$-Parameter'))
-plt.plot(eb_n0,linear2,'g--o', lw=3, ms=10, label=r'$f_{d}T_{s}=3.5\times 10^{-4}$, Manifold')
-
-v = plt.axis()
-plt.axis([-3.4, 27.4, v[2], v[3]])
-# plt.plot(eb_n0,lineari, label='3.5e-4 ideal', marker='+')
-# plt.plot(eb_n0,linear17i, 'r-', lw=3, ms=10,label='1.7e-4 ideal', marker='+')
-# plt.plot(eb_n0,linear1i, label='1e-4 ideal', marker='+')
[…]

### 2019-12-18 ber_results/plot_ber.py

@@ -1,39 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import rc
-from matplotlib.pyplot import figure
-
-font = {'size'   : 18}
-plt.rc('font', **font)
-plt.rc('text', usetex=True)
-plt.rc('xtick', labelsize=20)
-plt.rc('ytick', labelsize=20)
-
-plt.figure(figsize=(8, 6))
-eb_n0=np.arange(-3,25,3)
-
-ped_scal=np.load('updated_6bit_eval/pedestrian/0.00035/hp_ber.npy')
-ped_mani=np.load('updated_6bit_eval/agrim_manifold_pedestrian/0.00035/hp_ber.npy')
-ped_ideal=np.load('updated_6bit_eval/pedestrian/0.00035/ideal_ber.npy')
-
-veh_scal=np.load('updated_6bit_eval/vehicular/0.00035/hp_ber.npy')
-veh_mani=np.load('updated_6bit_eval/agrim_manifold_vehicular/0.00035/hp_ber.npy')
-veh_ideal=np.load('updated_6bit_eval/agrim_manifold_vehicular/0.00035/ideal_ber.npy')
-plt.yscale('log')
-plt.ylabel('BER')
-plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
-
-plt.plot(eb_n0,ped_scal[:10], 'b--o', lw=3, ms=10, label='Scalar')
-plt.plot(eb_n0,ped_mani[:10],'g--o', lw=3, ms=10, label='Manifold')
-plt.plot(eb_n0,ped_ideal[:10],'y--o', lw=3, ms=10, label='64 subcarriers - complete')
-
-# plt.plot(eb_n0,veh_scal[:10], 'g--o', lw=3, ms=10, label='Scalar')
-# plt.plot(eb_n0,veh_mani,'b--o', lw=3, ms=10, label='Manifold')
-# plt.plot(eb_n0,veh_ideal,'y--o', lw=3, ms=10, label='64 subcarriers - complete')
-
-
-plt.grid(which='major', linewidth='0.5', color='black');
-plt.grid(which='minor', linewidth='0.5', color='grey');
-
-plt.legend(fontsize=14.5)
-plt.show()

### 2019-12-18 code_snippets/adm.py

@@ -1,218 +1,0 @@
-#!/usr/bin/env python3
-
-import array
-
-BYTES = 2   # N bytes arthimetic
-MAX = 2 ** (BYTES * 8 - 1) - 1
-MIN = - (2 ** (BYTES * 8 - 1)) + 1
-
-CHUNK= 1024
-
-MIN_DELTA = 2
-
-def predictive_adm(samples, deltamax = MAX//21, a = 1):
-    """
-    Encodes audio bytearray data with Adaptative Delta Modulation. Return a 
-    BitStream in a list
-
-    Keyword arguments:
-        samples -- Signed 16 bit Audio data in a byte array
-        delta - Delta constant of DM modulation. By default it's 1/21 of Max
-        Sample value
-        a - Sets Integrator decay value. By default it's 1
-
-    """
-    raw = array.array('h')
-    raw.frombytes(samples)
-    stream = []
-    integrator = 0
-    ndeltah = ndeltal = delta = deltamax // 2
-    lastBits = []
-
-    for sample in raw:
-
-        # Adapt Delta value
-        if len(lastBits) >= 1:
-            if lastBits[-1]:
-                ndeltah = min(delta*2, deltamax)
-                ndeltal = max(delta//2, MIN_DELTA)
-            else:
-                ndeltah = max(delta//2, MIN_DELTA)
-                ndeltal = min(delta*2, deltamax)
-
-            lastBits.pop()
-
-        highval = integrator + ndeltah
-        lowval = integrator - ndeltal
-
-        disthigh = abs(highval - sample)
-        distlow = abs(lowval - sample)
-        
-        # Choose integrator with less diference to sample value
-        if disthigh >= distlow:
-            stream.append(0)
-            integrator = lowval
-            delta = ndeltal
-        else:
-            stream.append(1)
-            integrator = highval
-            delta = ndeltah
-        
-        # Clamp to signed 16 bit
-        integrator = max(integrator, MIN)
-        integrator = min(integrator, MAX)
-
-        integrator = round(integrator * a)
-        lastBits.append(stream[-1])
-
-    
-    return stream
-
-
-def decode_adm(stream, deltamax = MAX//21, a = 1):
-    """
-    Decodes a Adaptative Delta Modulation BitStream in Signed 16 bit Audio data
-    in a ByteArray.
-
-    Keywords arguments:
-        stream -- List with the BitStream
[…]

### 2019-12-18 code_snippets/givens_complex.py

@@ -1,126 +1,0 @@
-#extracting the parameters theta and phi out of an orthonormal unitary matrix which can be used to reconstruct it again
-
-import math
-import cmath
-import numpy as np
-import random
-import pdb
-
-def unitary(n):
-    X=(np.random.rand(n,n)+1j*np.random.rand(n,n))/np.sqrt(2)
-    [Q,R]=np.linalg.qr(X)
-    T=np.diag(np.diag(R)/np.abs(np.diag(R)))
-    U=np.matrix(np.matmul(Q,T))
-    # Verify print (np.matmul(U,U.H))
-    return U    
-    
-def rand_stiefel(n,p):
-    H=(np.random.rand(n,p)+1j*np.random.rand(n,p))/np.sqrt(2)
-    U, S, V = np.linalg.svd(H,full_matrices=0)
-    # newU=np.matmul(U,np.diag(np.exp(-1j*np.angle(U[0,:]))))
-
-    return U
-    # return newU
-# extracting givens rotation parameters
-def givens_rot(a, b):
-	if(b == 0):
-		c = 1
-		s = 0
-	elif(abs(b) > abs(a)):
-		r = a/b
-		s = 1/math.sqrt(1 + r**2)
-		c = s*r
-	else:
-		r = b/a
-		c = 1/math.sqrt(1 + r**2)
-		s = c*r
-	return c,s
-
-# extarcting phi and theta out of the matrix	
-def semiunitary_to_givens_vec(A):
-	m, n = np.shape(A)
-	V = A
-	g_theta = []
-	d_phi = []
-	for j in range(n):
-		b = np.angle(V[j:,j])
-		# print b
-		d_phi.append(b)
-		if(j>0):
-			temp = np.ones(j)
-			temp = np.append(temp, np.exp(np.vectorize(complex)(0,b)))
-			a = np.diag(temp)
-			V = np.matmul(np.transpose(np.conj(a)),V)
-		else:
-			a = np.diag(np.exp(np.vectorize(complex)(0,b)))
-			V = np.matmul(np.transpose(np.conj(a)),V)
-
-		for i in range(m-1, j, -1):
-			G = np.identity(m)
-			c,s = givens_rot(np.real(V[i-1,j]), np.real(V[i,j]))
-			g_theta.append(np.arccos(c))
-
[…]

### 2019-12-18 code_snippets/givens_real.py

@@ -1,62 +1,0 @@
-import math
-import numpy as np
-import random
-
-def givens_rot(a, b):
-	if(b == 0):
-		c = 1
-		s = 0
-	# if(abs(b) > abs(a)):
-	else:	
-		r = a/b
-		s = 1/math.sqrt(1 + r**2)
-		c = s*r
-	# else:
-	# 	r = b/a
-	# 	c = 1/math.sqrt(1 + r**2)
-	# 	s = c*r
-
-	return c,s
-	
-def qr_givens(A):
-	m, n = np.shape(A)
-	# Q = np.identity(m, dtype = float)
-	z = np.identity(n)
-	I = np.zeros((m,n))
-	I[:n,:] = z
-	# print I
-	R = A
-	g = []
-
-	for j in range(n):
-		for i in range(m-1, j, -1):
-			G = np.identity(m)
-			c,s = givens_rot(R[i-1,j], R[i,j])
-			G[i-1:i+1, i-1:i+1] = [[c,-s],[s,c]]
-			g.append(G)
-			R = np.matmul(np.transpose(G),R)
-			# Q = np.matmul(Q,G)
-	# return Q,R, g, I
-	return g, I		
-
-def given_recons(g, I):
-	c = np.identity(np.shape(I)[0])
-	
-	for i in range (len(g)):
-		c = np.matmul(c,g[i])
-	
-	c = np.matmul(c,I)
-	return c
-	
-# random.seed()
-x = np.random.rand(4,3)
-x, _ = np.linalg.qr(x)
-
-print x
-g,I = qr_givens(x)
-
-out = given_recons(g,I)
-
-print out
-
-

### 2019-12-18 code_snippets/old.py

@@ -1,95 +1,0 @@
-#extracting the parameters theta and phi out of an orthonormal unitary matrix which can be used to reconstruct it again
-
-import math
-import cmath
-import numpy as np
-import random
-import pdb
-
-def unitary(n):
-    X=(np.random.rand(n,n)+1j*np.random.rand(n,n))/np.sqrt(2)
-    [Q,R]=np.linalg.qr(X)
-    T=np.diag(np.diag(R)/np.abs(np.diag(R)))
-    U=np.matrix(np.matmul(Q,T))
-    # Verify print (np.matmul(U,U.H))
-    return U    
-    
-def rand_stiefel(n,p):
-    H=(np.random.rand(n,p)+1j*np.random.rand(n,p))/np.sqrt(2)
-    U, S, V = np.linalg.svd(H,full_matrices=0)
-    return U
-
-# extracting givens rotation parameters
-def givens_rot(a, b):
-	if(b == 0):
-		c = 1
-		s = 0
-	elif(abs(b) > abs(a)):
-		r = a/b
-		s = 1/math.sqrt(1 + r**2)
-		c = s*r
-	else:
-		r = b/a
-		c = 1/math.sqrt(1 + r**2)
-		s = c*r
-	return c,s
-
-# extarcting phi and theta out of the matrix	
-def qr_givens(A):
-	m, n = np.shape(A)
-	Q = np.identity(m, dtype = float)
-	z = np.identity(n)
-	I = np.zeros((m,n))
-	I[:n,:] = z
-	V = A
-	g = []
-	d = []
-	g_theta = []
-	d_phi = []
-	for j in range(n):
-		b = np.angle(V[j:,j])
-		# print b
-		d_phi.append(b)
-		if(j>0):
-			temp = np.ones(j)
-			temp = np.append(temp, np.exp(np.vectorize(complex)(0,b)))
-			a = np.diag(temp)
-			d.append(a)
-			V = np.matmul(np.transpose(np.conj(a)),V)
-			Q = np.matmul(Q,a)
-		else:
-			a = np.diag(np.exp(np.vectorize(complex)(0,b)))
-			d.append(a)
-			V = np.matmul(np.transpose(np.conj(a)),V)
-			Q = np.matmul(Q,a)
-
-		for i in range(m-1, j, -1):
-			G = np.identity(m)
[…]

### 2019-12-18 config/arguments.py

@@ -1,5 +1,0 @@
-import argparse
-
-parser=argparse.ArgumentParser()
-
-parser.add_argument("-c","--constants",default="config/constants.yml", type=str, help="Predifened constants")

### 2019-12-18 eval.py

@@ -1,208 +1,0 @@
-# Code to generate the generates precoders corresponding to 
-# 100 channel evolutions for 10 independent channel realizations
-# using both the time based and hopping based schemes 
-# (i.e. fill 10 independent 100*`num_subcarriers` 
-# time-frequency bins matrix with generated precoders), and
-# save them as .npy files
-#---------------------------------------------------------------------------
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-from scipy.interpolate import griddata
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-import copy
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK,calculate_BER_performance_QAM256,waterfilling
-from munch import munchify
-import yaml
-from config.arguments import parser
-    
-#SIGINT handler
-def sigint_handler(signal, frame):
-    #Do something while breaking
-    pdb.set_trace()
-    sys.exit(0)
-def main():
-    args = parser.parse_args()
-    with open(args.constants, 'r') as f:
-        constants = munchify(yaml.load(f)) 
-
-    #---------------------------------------------------------------------------
-    #Channel Params
-    signal.signal(signal.SIGINT, sigint_handler)
-    #number of db values for which logsum rate will be calculated
-    num_Cappar=7
-    avg_hpcap=np.zeros(num_Cappar)
-    avg_otcap=np.zeros(num_Cappar)
-    avg_maxcap=np.zeros(num_Cappar)
-    # Eb_N0_dB=np.arange(-6,6,2)
-
-    Eb_N0_dB=np.arange(-3,25,3)
-    # Eb_N0_dB=np.arange(-3,36,3)
-    #Store BER here
-    idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    otBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    Ts=constants.Ts # 5e-8
-    num_subcarriers=constants.num_subcarriers #64
-    Nt=constants.Nt #4
-    Nr=constants.Nr #2
-    freq_quanta=constants.freq_quanta #4 - 5
-    time_quanta=constants.time_quanta #1
-    channel_type="Vehicular/"
-    # channel_type="Pedestrian/"
[…]

### 2019-12-18 interpolation_comparison.py

@@ -1,127 +1,0 @@
-#-----------------------------------------------------------------------------
-# Code to compare interpolation approaches (Caayley Exp. and Geodesic)
-#-----------------------------------------------------------------------------
-
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-import itpp
-from mimo_tdl_channel import *
-from sumrate_BER import waterfilling,leakage_analysis, calculate_BER_performance_QAM256, calculate_BER_performance_QPSK
-import precoder_interpolation
-import sys
-import signal
-import pdb
-import copy
-np.random.seed(81)
-#---------------------------------------------------------------------------
-#Lambda Functions
-frob_norm= lambda A:np.linalg.norm(A, 'fro')
-diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-#---------------------------------------------------------------------------
-#SIGINT handler
-def sigint_handler(signal, frame):
-    pdb.set_trace()
-    sys.exit(0)
-#---------------------------------------------------------------------------
-#Channel Params
-signal.signal(signal.SIGINT, sigint_handler)
-itpp.RNG_randomize()
-c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
-#c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
-Ts=5e-8
-num_subcarriers=1024
-Nt=4
-Nr=2
-B=6
-class_obj=MIMO_TDL_Channel(Nt,Nr,c_spec,Ts,num_subcarriers)
-number_simulations=1000
-feedback_mats=32
-# Why ((num_subcarriers-1)//(feedback_mats-1)) ?
-# We have to cover indices upto (self.num_subcarriers-1) with (self.num_fb_points-1)
-# Since, we have 0 as an implicit input
-fb_indices=np.arange(feedback_mats)*((num_subcarriers-1)//(feedback_mats-1))
-#----------------------------------------------------------------------
-#Leakage arr
[…]

### 2019-12-18 main.py

@@ -1,298 +1,0 @@
- #Import Statements
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK
-from munch import munchify
-import yaml
-from config.arguments import parser
-
-np.random.seed(81)
-#---------------------------------------------------------------------------
-#SIGINT handler
-def sigint_handler(signal, frame):
-    #Do something while breaking
-    pdb.set_trace()
-    sys.exit(0)
-
-def main():
-    # channel_type="Pedestrian"
-    channel_type="Vehicular"
-    args = "config/"+ channel_type + ".yml"
-    with open(args, 'r') as f:
-        constants = munchify(yaml.load(f)) 
-    #---------------------------------------------------------------------------
-    #Lambda Functions
-    frob_norm= lambda A:np.linalg.norm(A, 'fro')
-    diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-    #---------------------------------------------------------------------------
-    #Channel Params
-    signal.signal(signal.SIGINT, sigint_handler)
-    # itpp.RNG_randomize()
-    itpp.RNG_reset(81)
-
-    c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Vehicular_A)
-    # c_spec=itpp.comm.Channel_Specification(itpp.comm.CHANNEL_PROFILE.ITU_Pedestrian_A)
-    Ts=constants.Ts
-    num_subcarriers=constants.num_subcarriers
-    Nt=constants.Nt
-    Nr=constants.Nr
-    dphi_size=(2*Nt*Nr-Nr**2+Nr)//2
-    gtheta_size=(2*Nt*Nr-Nr**2-Nr)//2
-    vec_size=2*Nt*Nr-Nr**2
-    B=constants.B
-    #---------------------------------------------------------------------------
-    # Simulation Parameters
-    number_simulations=constants.number_simulations # equivalent of time
-    feedback_mats=constants.feedback_mats
-    freq_jump=(num_subcarriers-1)//(feedback_mats-1)
-    freq_quanta=constants.freq_quanta
-    time_quanta=constants.time_quanta
-    
-    # Number of channels simulated for
-    num_chan_realisations=constants.num_chan_realisations
-    count=0
[…]

### 2019-12-18 mimo_tdl_channel.py

@@ -1,830 +1,0 @@
-# This python script contains codes for channel generation, lifting 
-# and precoder prediction
-
-from __future__ import division
-import itpp
-import numpy as np
-import scipy.linalg as la
-import pdb
-import math
-
-def unitary(n):
-    X=(np.random.rand(n,n)+1j*np.random.rand(n,n))/np.sqrt(2)
-    [Q,R]=np.linalg.qr(X)
-    T=np.diag(np.diag(R)/np.abs(np.diag(R)))
-    U=np.matrix(np.matmul(Q,T))
-    # Verify print (np.matmul(U,U.H))
-    return U    
-    
-def rand_stiefel(n,p):
-    H=(np.random.rand(n,p)+1j*np.random.rand(n,p))/np.sqrt(2)
-    U, S, V = np.linalg.svd(H,full_matrices=0)
-    return U
-
-# extracting givens rotation parameters
-def givens_rot(a, b):
-    if(b == 0):
-        c = 1
-        s = 0
-    elif(abs(b) > abs(a)):
-        r = a/b
-        s = 1/math.sqrt(1 + r**2)
-        c = s*r
-    else:
-        r = b/a
-        c = 1/math.sqrt(1 + r**2)
-        s = c*r
-    return c,s    
-
-def rand_SH(n):
-    X=2*np.random.rand(n,n)-np.ones((n,n))+1j*(2*np.random.rand(n,n)-np.ones((n,n)))
-    A=np.matrix(X-X.T.conjugate())/2
-    # Verify print(A-A.H)
-    return A
-
-def qtisn(pU,rU,g,num_iter,sH_list,norm_fn,sk=0.0):
-    diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-    trials=0
-    sp=0.0
-    sm=0.0
-    while(trials<num_iter):
-        sp=g**(np.min(sk+1,0))
-        sm=g**(sk-1)
-        # Qp=[np.matmul(la.expm(sp*sH),pU) for sH in sH_list]
-        # Qm=[np.matmul(la.expm(sm*sH),pU) for sH in sH_list]
-        Qp=[sH_retract(pU,sp*sH) for sH in sH_list]
-        Qm=[sH_retract(pU,sm*sH) for sH in sH_list]
-        Q=Qp+Qm
-        if(norm_fn=="diff_frob_norm"):
-            chordal_dists=np.array([diff_frob_norm(Q[i],rU) for i in range(len(Q))])
-        else:
[…]

### 2019-12-18 new_results/plot_results.py

@@ -1,62 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import rc
-from matplotlib.pyplot import figure
-
-font = {'size'   : 18}
-plt.rc('font', **font)
-# plt.rc('text', usetex=True)
-plt.rc('xtick', labelsize=20)
-plt.rc('ytick', labelsize=20)
-
-plt.figure(figsize=(8, 6))
-eb_n0=np.arange(-3,25,3)
-
-# cubic=np.load('6bit_eval/hp_ber.npy')
-# linear0=np.load('updated_6bit_eval/pedestrian/0.1/hp_cap.npy')
-# linear00=np.load('updated_6bit_eval/pedestrian/0.1/hp_cap.npy')
-# linear0=linear0/linear00
-
-# linear1=np.load('updated_6bit_eval/pedestrian/0.01/hp_cap.npy')
-# linear11=np.load('updated_6bit_eval/pedestrian/0.01/max_cap.npy')
-# linear1=linear1/linear11
-
-# linear2=np.load('updated_6bit_eval/pedestrian/0.001/hp_cap.npy')
-# linear22=np.load('updated_6bit_eval/pedestrian/0.001/max_cap.npy')
-# linear2=linear2/linear22
-veh_after5=np.array([2.23854902e-02, 5.85075132e-03, 9.43905410e-04, 1.79587744e-04, 7.84925235e-05, 4.36302923e-05, 2.41909232e-05, 1.23645413e-05, 5.51285282e-06, 1.96887601e-06])
-ped_after5=np.load('ped_after5/0.00035/hp_ber.npy')[:10]
-ped_par=np.load('pedestrian/0.00035/hp_ber.npy')[:10]
-ped_man=np.load('agrim_manifold_pedestrian/0.00035/hp_ber.npy')[:10]
-ped_ideal=np.load('pedestrian/0.00035/ideal_ber.npy')[:10]
-
-plt.yscale('log')
-plt.ylabel('BER')
-plt.xlabel(r'$E_{b}/N_{0}$ (dB)')
-# pdb.set_trace()
-# plt.plot(eb_n0,linear0, 'r--o', lw=3, ms=10, label=r'$\times 10^{-1}$')
-# plt.plot(eb_n0,linear1, 'g--o', lw=3, ms=10, label=r'$\times 10^{-2}$')
[…]

### 2019-12-18 plot_qerror.py

@@ -1,42 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import rc
-from matplotlib.pyplot import figure
-
-font = {'size'   : 18}
-plt.rc('font', **font)
-plt.rc('text', usetex=True)
-plt.rc('xtick', labelsize=20)
-plt.rc('ytick', labelsize=20)
-
-plt.figure(figsize=(8, 6))
-eb_n0=np.arange(0,21,3)
-stief_veh_scal=np.load('qerror_scal/veh_qerror_stief_scalar.npy')
-stief_veh_mani=np.load('qerror_mani/veh_qerror_mani_stief.npy')
-frob_veh_scal=np.load('qerror_scal/veh_frob_scala_qerror.npy')
-frob_veh_mani=np.load('qerror_mani/veh_qerror_mani_frob.npy')
-
-stief_ped_scal=np.load('qerror_scal/ped_qerror_scal_stief.npy')
-stief_ped_mani=np.load('qerror_mani/ped_qerror_mani_stief.npy')
-frob_ped_scal=np.load('qerror_scal/ped_qerror_scal_frob.npy')
-frob_ped_mani=np.load('qerror_mani/ped_qerror_mani_frob.npy')
-
-plt.ylabel('Quantization Error Frobenius avg')
-plt.xlabel('Time')
-
-# plt.plot(stief_veh_mani,label='Vehicular Manifold')
-# plt.plot(stief_veh_scal, label='Vehicular Scalar')
-# plt.plot(stief_ped_mani,label='Pedestrian Manifold')
-# plt.plot(stief_ped_scal, label='Pedestrian Scalar')
-
-# plt.plot(frob_veh_mani,label='Vehicular Manifold')
-# plt.plot(frob_veh_scal, label='Vehicular Scalar')
-plt.plot(frob_ped_mani,label='Pedestrian Manifold')
-plt.plot(frob_ped_scal, label='Pedestrian Scalar')
-
-
-plt.grid(which='major', linewidth='0.5', color='black');
-plt.grid(which='minor', linewidth='0.5', color='grey');
-
-plt.legend(fontsize=14.5)
-plt.show()

### 2019-12-18 qrtz_error.py

@@ -1,138 +1,0 @@
-# Code to generate the generates precoders corresponding to 
-# 100 channel evolutions for 10 independent channel realizations
-# using both the time based and hopping based schemes 
-# (i.e. fill 10 independent 100*`num_subcarriers` 
-# time-frequency bins matrix with generated precoders), and
-# save them as .npy files
-#---------------------------------------------------------------------------
-from __future__ import division
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.linalg as la
-from scipy.interpolate import griddata
-import itpp
-from mimo_tdl_channel import *
-import sys
-import signal
-import pdb
-import time
-import copy
-from sumrate_BER import leakage_analysis, calculate_BER_performance_QPSK,calculate_BER_performance_QAM256,waterfilling
-from munch import munchify
-import yaml
-from config.arguments import parser
-    
-#SIGINT handler
-def sigint_handler(signal, frame):
-    #Do something while breaking
-    pdb.set_trace()
-    sys.exit(0)
-def main():
-    diff_frob_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-    # channel_type="Vehicular"
-    channel_type="Pedestrian"
-    args = "config/"+ channel_type + ".yml"
-    with open(args, 'r') as f:
-        constants = munchify(yaml.load(f)) 
-
-    #---------------------------------------------------------------------------
-    #Channel Params
-    signal.signal(signal.SIGINT, sigint_handler)
-    #number of db values for which logsum rate will be calculated
-    num_Cappar=7
-    avg_hpcap=np.zeros(num_Cappar)
-    avg_otcap=np.zeros(num_Cappar)
-    avg_maxcap=np.zeros(num_Cappar)
-    # Eb_N0_dB=np.arange(-6,6,2)
-
-    Eb_N0_dB=np.arange(-3,25,3)
-    # Eb_N0_dB=np.arange(-3,36,3)
-    #Store BER here
-    idealBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    hpBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    otBER_QPSK=np.zeros(Eb_N0_dB.shape[0])
-    Ts=constants.Ts # 5e-8
-    num_subcarriers=constants.num_subcarriers #64
-    Nt=constants.Nt #4
-    Nr=constants.Nr #2
[…]

### 2019-12-18 sumrate_BER.py

@@ -1,270 +1,0 @@
-#The python script [sumrate\_BER.py](./sumrate_BER.py) 
-#contains the codes used to obtain the BER and achievable rate
-
-from __future__ import division
-import numpy as np
-import scipy.linalg as la
-import pdb
-
-diff_frobenius_norm = lambda A,B:np.linalg.norm(A-B, 'fro')
-
-def waterfilling(alpha,P=1):
-    n=alpha.size
-    ind=np.argsort(alpha)
-    lambda1=0
-    for m in range(n):
-        t=0
-        for j in range(m,n):
-            t=t+(1/alpha[ind[j]])
-        t=(t+P)/(n-m)
-        lambda1=1/t
-        if(lambda1<alpha[ind[m]]):
-            break
-    result=[((1/lambda1)-(1/alpha[j])) if (lambda1<alpha[j]) else 0 for j in range(n)]
-    # pdb.set_trace()
-    return result
-
-def leakage_analysis(H_list,U_list,interpolated_U,num_subcarriers,wf_sigma_list,wf_interp_S,Nt,Nr,ret_abs=False):
-    fig_merit=np.zeros(num_subcarriers)
-    subopt_cap=np.zeros(num_subcarriers)
-    for i in range(num_subcarriers):
-        H=np.matrix(H_list[i])
-        U=np.matrix(U_list[i][:,np.arange(Nr)])
-        U_tilde=np.matrix(interpolated_U[i])
-        opt_vec=wf_sigma_list[Nr*i:Nr*(i+1)]
-        interpopt_vec=wf_interp_S[Nr*i:Nr*(i+1)]
-        #Get Maximal Capacity
-        Q=U*np.diag(opt_vec)*U.H
-        det_mat=np.identity(Nr)+H.H*Q*H
-        max_cap=np.real(np.log(la.det(det_mat)))
-        #Get Suboptimal Capacity
-        Q_tilde=U_tilde*np.diag(interpopt_vec)*U_tilde.H
-        det_mat=np.identity(Nr)+H.H*Q_tilde*H
-        subopt_cap[i]=np.real(np.log(la.det(det_mat)))
-        #print("SubOptCap also is: "+str(subopt_cap))
-        fig_merit[i]=(subopt_cap[i]/max_cap)*100
-    #print(norm_err)
-    if(ret_abs):
-        return subopt_cap
-    else:
-        return fig_merit
-
[…]

### 2019-12-18 paper/main.tex
Minor fix in abstract
Minor fix

@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
+\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ {\bf\eta}_{i,t}
 \end{equation}

### 2019-12-18 paper/main.tex
Enhance changes

@@
 
-The use of multiple antennas at the transmitter and receiver can
-significantly enhance the performance of wireless systems. Moreover,
-in these multiple-input multiple-output (MIMO) wireless systems, the
-transmitter can allocate resources more efficiently if the channel
-state information (CSI) is made available at the transmitter, both to
+The use of multiple antennas at the transmitter and receiver (MIMO)
+can significantly enhance the performance of wireless systems.  A
+typical step to get this performance enhancements is estimation of
+channel state information (CSI).  If transmitter also has the
+knowledge of CSI, it can allocate resources more efficiently, both to
 improve data rates and to reduce BER~\cite{love2008overview}.  To this
-end, it is necessary to efficiently feedback the CSI from the receiver 
-to the transmitter. In wireless systems that use orthogonal
-frequency division multiplexing (OFDM), the feedback needs to be
-provided to the transmitter across all subcarriers, and this places a
-large feedback demand. However, since the CSI varies gradually in both
-time and frequency, effective techniques to reduce feedback overheads
-exist, using frequency domain interpolation temporal prediction. In
-particular, knowledge of the precoding matrices (that are typically
-the orthonormal right singular vectors of the channel matrix) is
-essential to achieve link capacity, as well as to parallelize the
-channel. In this paper, we focus on the problem of tracking the
-precoders of a MIMO-OFDM system across time and frequency for
-efficient CSI feedback by tracking a parametrized version of the matrices.
+end, it is necessary to efficiently feed back CSI from the receiver to
+the transmitter. In wireless systems that use orthogonal frequency
+division multiplexing (OFDM), the feedback needs to be provided to the
+transmitter across all subcarriers and in each OFDM frame, which
+places a large feedback demand. However, since the CSI varies
+gradually in both time and frequency, effective techniques to reduce
+feedback overheads exist, using frequency domain interpolation and
+temporal prediction [CITE?]. In particular, knowledge of the precoding
+matrices (that are typically the orthonormal right singular vectors of
+CSI matrix) is essential to achieve the link capacity, as well as to
+parallelize the channel. In this paper, we focus on the problem of
+tracking the precoders of a MIMO-OFDM system across time and frequency
+for efficient CSI feedback by tracking a parametrized version of the
+precoding matrices.
 
@@
 of the precoders is used to obtain tangents for predictive
-quantization and interpolation across time and frequency. While these
-methods have been shown to be very effective for precoder feedback,
-they involve operations over higher dimensional manifolds. In
-particular, optimization and interpolation is complicated, especially
-for manifolds where geodesic paths for interpolation may be difficult
-to obtain, especially for joint time-frequency predictive coding. An
-alternate approach is to parametrize the precoding matrices using
-scalar parameters. The space of unitary matrices parameterized into
+quantization and interpolation across time and frequency.
+An alternate approach is to parametrize the precoding matrices using
+scalar parameters. The space of unitary matrices can be parameterized into
 independent scalar parameters that correspond to Givens rotations and
 Householder transformations of the unitary precoding matrices
-\cite{4114278,4556174}. These approaches have been combined with
-adaptive delta modulation effectively track precoding matrices both in
-the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately.
+\cite{4114278,4556174}.
 
-In this paper, we propose a method for exploiting the joint
-time-frequency correlation of the precoder's scalar parameters on the
-transmitter using CSI feedback. Typical approaches to predictive
-quantization and interpolation do not exploit the joint time-frequency
-correlation to enhance the precoder estimate at the
+While the manifold based methods have been shown to be very effective
+for precoder feedback, they involve operations over higher dimensional
+manifolds. In particular, optimization and interpolation is
+complicated, especially for manifolds where geodesic paths for
+interpolation may be difficult to obtain, especially (don't repeat
+especially again) for joint time-frequency predictive coding.  In
+addition to being analytically more tractable, the scalar
+parametrization methods have also been implemented in 802.11ac
+standards \cite{porat2013improved}. Past work in utilizing scalar
+parametrization for precoder feedback have utilized adaptive delta
+modulation effectively track precoding matrices both in the
+time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
+separately. In this paper, we propose a method for exploiting the
+joint time-frequency correlation of the precoder's scalar parameters
+on the transmitter using CSI feedback.  Typical approaches to
+predictive quantization and interpolation do not exploit the joint
+time-frequency correlation to enhance the precoder estimate at the
 transmitter. However, exploiting the correlations jointly not only
 reduces the feedback requirement, but also enhances the precoder
-reconstruction accuracy significantly. Further, the independence of
-the scalar parameters reduces the problem to one that involves
-interpolation of separate scalar valued functions, as opposed to
-operations over tangent spaces in
-manifolds~\cite{Gupt1905:Predictive}. Simulations reveal that the
-proposed method matches the performance of manifold based approaches
-with a much lower implementation complexity and fewer number of bits
+reconstruction accuracy significantly.
+
+Prior work that utilizing manifolds based quantization
+\cite{Gupt1905:Predictive,pitaval2013codebooks,6891198} and scalar
+parametrization \cite{4114278,4556174} compare their performance with
+other related quantization methods, but do not present a comparison to
+highlight the performance benefits of the chosen method.  In addition
+to a novel joint time-frequency feedback scheme, we also compare the
+results of our parametrization based adaptive scheme with an
+equivalent manifold based scheme, and illustrate the performance
+benefits entailed by utilizing the scalar parametrization method.
+Simulations reveal that the proposed method outperforms manifold based
+approaches with a much lower implementation complexity and fewer bits
 for quantization.
 
-The rest of this paper is organized as follows: Section~\ref{section2}
-describes the system model and the precoder feedback and interpolation
-techniques. Section~\ref{section3} compares the BER and achievable
-rates for our proposed methods with other approaches. Finally,
-Section~\ref{section4} concludes with discussions on future work.
+% The rest of this paper is organized as follows: Section~\ref{section2}
+% describes the system model and the precoder feedback and interpolation
+% techniques. Section~\ref{section3} compares the BER and achievable
+% rates for our proposed methods with other approaches. Finally,
+% Section~\ref{section4} concludes with discussions on future work.
 

### 2019-12-19 paper/main.tex
adding time based interpolation plots

--- 2019-12-18 paper/main.tex
@@
 
+\usepackage{mathtools}
+  \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
 % \usepackage{parskip}
@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ {\bf\eta}_{i,t}
+\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
 \end{equation}
@@
 is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while
-$\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.
+$\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.~\cite{4114278}
 \begin{figure}
@@
 position $pk+q$ for $k = 0,1,..., \frac{N-1}{p}$ with $q =
-{\frac{p}{2}}$.
+{\floor*{\frac{p}{2}}}$.
 
@@
 bit for each scalar parameter that is being tracked. On average, this
-can be reduced by half if we transmit only $\Theta$ values for
-one-time instance and only $\Phi$ values for the next time instant in
+can be reduced by half if we transmit only $\Theta$(collection of $\theta_{i}s$) values for
+one-time instance and only $\Phi$(collection of $\phi_{i}s$) values for the next time instant in
 an alternating fashion (the untransmitted value is assumed to be the
@@
 against the ideal precoder for 100 channel evolutions in each run and
-averaged. Simulations are performed for normalised Doppler values of
-$f_dT_s = 3.41\times 10^{-6}$ (corresponding to a velocity of 37.2 km/h
-for $f_c = 1$ GHz) with 1024 subcarriers, and with
+averaged over 10 channel realizations. Simulations are performed for 
+normalised Doppler values of $f_dT_s = 3.41\times 10^{-6}$ (corresponding 
+to a velocity of 37.2 km/h for $f_c = 1$ GHz) with 1024 subcarriers, and with
 $f_dT_s = 1.56 \times 10^{-5}$ (velocity of 172.8 km/h) with 64
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber1024_035}
+\includegraphics[width=0.5\textwidth]{images/1024timebased}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber64_001}
+\includegraphics[width=0.5\textwidth]{images/64timebased}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2019-12-25 paper/main.tex
Minor fix
Minor intro rewrite

@@
 
-The use of multiple antennas at the transmitter and receiver (MIMO)
-can significantly enhance the performance of wireless systems.  A
-typical step to get this performance enhancements is estimation of
-channel state information (CSI).  If transmitter also has the
-knowledge of CSI, it can allocate resources more efficiently, both to
-improve data rates and to reduce BER~\cite{love2008overview}.  To this
-end, it is necessary to efficiently feed back CSI from the receiver to
-the transmitter. In wireless systems that use orthogonal frequency
-division multiplexing (OFDM), the feedback needs to be provided to the
-transmitter across all subcarriers and in each OFDM frame, which
-places a large feedback demand. However, since the CSI varies
-gradually in both time and frequency, effective techniques to reduce
-feedback overheads exist, using frequency domain interpolation and
-temporal prediction [CITE?]. In particular, knowledge of the precoding
-matrices (that are typically the orthonormal right singular vectors of
-CSI matrix) is essential to achieve the link capacity, as well as to
-parallelize the channel. In this paper, we focus on the problem of
-tracking the precoders of a MIMO-OFDM system across time and frequency
-for efficient CSI feedback by tracking a parametrized version of the
-precoding matrices.
+Multiple-input multiple-output (MIMO) based wireless communication
+systems that use orthogonal frequency division multiplexing (OFDM)
+have been adapted in many modern wireless systems owing to their
+flexibility and ability to use the wideband channels effectively. One
+key requirement to utilize such channels effectively is for the
+transmitter to possess channel state information (CSI), since that
+permits channel parallelization for low-complexity decoders as well as
+the use of capacity achieving transmission
+strategies~\cite{love2008overview}. This primarily translates to the
+generation of unitary precoding matrices at the transmitter. However,
+feeding back CSI from the receiver to the transmitter requires large
+feedback overhead, since a precoder needs to be fed back for each OFDM
+subcarrier within each coherence interval. Typical past approaches
+have used manifold based approaches that exploit temporal and
+frequency correlation to reduce feedback overheads. Wireless standards
+have largely employed parametric quantization that uses Givens
+rotations based quantization~\cite{lou2013comparison,5671092}, though these do
+not involve any refinement of the estimates using channel
+correlations. In this paper, we propose a joint time-frequency
+interpolation approach that builds upon the parametric quantization
+techniques used by standards that manages to outperform typical
+manifold based quantization techniques.
 
-The problem of effective quantization and feedback of precoders for
-MIMO-OFDM systems has attracted significant attention in recent
-years. Many approaches involve codebook design over the Grassmann
+Most approaches for effective quantization and feedback of precoders
+for MIMO-OFDM systems involve codebook design over the Grassmann
 manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
@@
 of the precoders is used to obtain tangents for predictive
-quantization and interpolation across time and frequency.
-An alternate approach is to parametrize the precoding matrices using
-scalar parameters. The space of unitary matrices can be parameterized into
-independent scalar parameters that correspond to Givens rotations and
-Householder transformations of the unitary precoding matrices
+quantization and interpolation across time and frequency.  An
+alternate approach is to parametrize the precoding matrices using
+scalar parameters. The space of unitary matrices can be parameterized
+into independent scalar parameters that correspond to Givens rotations
+and Householder transformations of the unitary precoding matrices
 \cite{4114278,4556174}.
@@
 parametrization methods have also been implemented in 802.11ac
-standards \cite{porat2013improved}. Past work in utilizing scalar
+standards \cite{lou2013comparison}. Past work in utilizing scalar
 parametrization for precoder feedback have utilized adaptive delta
@@
 joint time-frequency correlation of the precoder's scalar parameters
-on the transmitter using CSI feedback.  Typical approaches to
-predictive quantization and interpolation do not exploit the joint
+on the transmitter using CSI feedback.  Typical approaches that feed
+back scalar precoder parameters problem do not exploit the joint
 time-frequency correlation to enhance the precoder estimate at the
-transmitter. However, exploiting the correlations jointly not only
+transmitter. We show that exploiting correlations jointly not only
 reduces the feedback requirement, but also enhances the precoder

### 2019-12-25 paper/main.tex
minor fixes

@@
 complicated, especially for manifolds where geodesic paths for
-interpolation may be difficult to obtain, especially (don't repeat
-especially again) for joint time-frequency predictive coding.  In
+interpolation may be difficult to obtain, particularly 
+for joint time-frequency predictive coding.  In
 addition to being analytically more tractable, the scalar
@@
 
+\section{Scalar joint time-frequency scheme}
 \subsection{Scalar Parametrization}
@@
 geometrical structure between the columns of the
-matrix~\cite{4114278}. The orthonormal matrix $\tilde{\bV}_{i,t}$ with
+matrix~\cite{4114278}.  Therefore $\tilde{\bV}_{i,t}$ could be paramaterized 
+using fewer parameters.
+The orthonormal matrix $\tilde{\bV}_{i,t}$ with
 $N_T$ rows and $N_R$ columns can be decomposed as
@@
 the $i$-th subcarrier at time instant $t$ is captured in a vector
-denoted by $\Phi_{i,t}$. $\theta_{l}$s as called rotation angles, and
+denoted by $\bf{\Phi_{i,t}}$. $\theta_{l}$s are called rotation angles, and
 their collection is captured in a vector referred to as
-$\Theta_{i,t}$. We note that $\Phi_{i,t}$ and $\Theta_{i,t}$ represent
+$\bf{\Theta_{i,t}}$. We note that $\bf{\Phi_{i,t}}$ and $\bf{\Theta_{i,t}}$ represent
 the phase and rotation angles for the $i$-th subcarrier at time
@@
 \end{equation}
+The given distribution of $\theta_l$ is used for non-uniform quantization
+of $\bf{\Theta}$ during their initialization to reduce the quantization error.  
 In addition, $\phi_k$ and $\theta_l$ are statistically independent,
@@
 $\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of
-$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\Phi_{i,t}$
+$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\bf{\Phi_{i,t}}$
 is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while
-$\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.~\cite{4114278}
+$\bf{\Theta_{i,t}}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.~\cite{4114278}
 \begin{figure}
@@
 bit for each scalar parameter that is being tracked. On average, this
-can be reduced by half if we transmit only $\Theta$(collection of $\theta_{i}s$) values for
-one-time instance and only $\Phi$(collection of $\phi_{i}s$) values for the next time instant in
+can be reduced by half if we transmit only $\bf{\Theta}$ values for
+one-time instance and only $\bf{\Phi}$ values for the next time instant in
 an alternating fashion (the untransmitted value is assumed to be the
@@
 known that linear prediction based on past samples is optimal. Even
-though the $\Theta$ and $\Phi$ parameters do not manifest as AR
+though the $\bf{\Theta}$ and $\bf{\Phi}$ parameters do not manifest as AR
 processes, for small changes, a linear approximation works
@@
 being to their previous values, as shown in
-Fig.~\ref{fig:adpm-fig}. Here, $\Theta$ is the collection of all
-$\theta_l$s and $\Phi$ is the collection of all $\phi_i$s. This brings
-down the bit budget to an average of 5 bits(the length of $\Theta$)
-and 7 bits (the length of $\Phi$) over alternate feedback instances,
+Fig.~\ref{fig:adpm-fig}. 
+% Here, $\bf{\Theta}$ is the collection of all
+% $\theta_l$s and $\bf{\Phi}$ is the collection of all $\phi_i$s. 
+This brings down the bit budget to an average of 5 bits(the length of $\bf{\Theta}$)
+and 7 bits (the length of $\bf{\Phi}$) over alternate feedback instances,
 thereby resulting in an average feedback budget of 6 bits.
@@
 tracking the channel parameters.
-
+In both the situations, the only-time based interpolation using manifold
+has the worst performance as it is not utilizing the joint time $\&$ frequency 
+correlations. 
 \begin{figure}

### 2019-12-27 main.tex
Changes in intro

@@
 
-The use of multiple antennas at the transmitter and receiver (MIMO) can
-significantly enhance the performance of wireless systems. 
-A typical step to get this performance enhancements is estimation of channel
-state information (CSI).
-If transmitter also has the knowledge of CSI, it can allocate 
-resources more efficiently, both to
-improve data rates and to reduce BER~\cite{love2008overview}.
+% Multiple-input multiple-output (MIMO) based wireless communication
+% systems that use orthogonal frequency division multiplexing (OFDM)
+% have been adapted in many modern wireless systems owing to their
+% flexibility and ability to use the wideband channels effectively. One
+% key requirement to utilize such channels effectively is for the
+% transmitter to possess channel state information (CSI), since that
+% permits channel parallelization for low-complexity decoders as well as
+% the use of capacity achieving transmission
+% strategies~\cite{love2008overview}. This primarily translates to the
+% generation of unitary precoding matrices at the transmitter. However,
+% feeding back CSI from the receiver to the transmitter requires large
+% feedback overhead, since a precoder needs to be fed back for each OFDM
+% subcarrier within each coherence interval. Typical past approaches
+% have used manifold based approaches that exploit temporal and
+% frequency correlation to reduce feedback overheads. Wireless standards
+% have largely employed parametric quantization that uses Givens
+% rotations based quantization~\cite{lou2013comparison,5671092}, though these do
+% not involve any refinement of the estimates using channel
+% correlations. In this paper, we propose a joint time-frequency
+% interpolation approach that builds upon the parametric quantization
+% techniques used by standards that manages to outperform typical
+% manifold based quantization techniques.
 
-To this end, it is necessary to efficiently feedback the CSI from the receiver 
-to the transmitter. In wireless systems that use orthogonal
-frequency division multiplexing (OFDM), the feedback needs to be
-provided to the transmitter across all subcarriers and in each OFDM frame, 
-which places a large feedback demand. However, since the CSI varies gradually in both
-time and frequency, effective techniques to reduce feedback overheads
-exist, using frequency domain interpolation and temporal prediction [CITE?]. In
-particular, knowledge of the precoding matrices (that are typically
-the orthonormal right singular vectors of CSI matrix) is
-essential to achieve the link capacity, as well as to parallelize the
-channel. In this paper, we focus on the problem of tracking the
-precoders of a MIMO-OFDM system across time and frequency for
-efficient CSI feedback by tracking a parametrized version of the precoding matrices.
+Multiple-input multiple-output (MIMO) based wireless communication
+systems that use orthogonal frequency division multiplexing (OFDM)
+have been adapted in many modern wireless systems owing to increased 
+data rate and channel capacity.
+One key requirement to achieve the channel capacity is to leverage spatial multiplexing gains obtained by precoding transmissions by Singular Value Decomposition (SVD) of the MIMO channel matrix ~\cite{love2008overview}.
+However, this requires the receiver to feed back the unitary matrix obtained by SVD, to the transmitter. 
+Given that the channel matrix varies with both frequency and time, this feedback has to be ideally done for each subcarrier and time instance, leading to massive feedback overhead.
+To reduce the feedback overhead, past work taps in the inherent temporal and frequency correlations in SVD precoders, in order to perform effective quantization with just a few bits by utilizing predictive quantization \cite{Gupt1905:Predictive,6891198,7370793,sacristan2010differential,5946308,6545375,4114278,4556174}, and reducing number of fed back subcarriers by utilizing interpolation algorithms \cite{Gupt1905:Predictive,5671092,khaled2005quantized}.
 
-The problem of effective quantization and feedback of precoders for
-MIMO-OFDM systems has attracted significant attention in recent
-years. Many approaches involve codebook design over the Grassmann
-manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
-the Stiefel manifold \cite{6891198,Gupt1905:Predictive} and the Flag
-manifold~\cite{pitaval2013codebooks}, wherein the manifold structure
-of the precoders is used to obtain tangents for predictive
-quantization and interpolation across time and frequency. 
-An alternate approach is to parametrize the precoding matrices using
-scalar parameters. The space of unitary matrices parameterized into
-independent scalar parameters that correspond to Givens rotations and
-Householder transformations of the unitary precoding matrices
-\cite{4114278,4556174} [?].
+An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders.
+A large body of works utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
+On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
+While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
+involve operations over higher dimensional manifolds.
+The scalar parametrization algorithms enable usage of the traditional signal processing algorithms over scalars and vectors, and hence are analytically simpler.
+The scalar parametrization methods have also been implemented in 802.11ac standards \cite{lou2013comparison}.
+Surprisingly, there have been no performance comparisons between these two body of works, hence it is not quantifiable from past work if there are performance gains/losses by operating over scalar parametrization instead of manifold.
+In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques.
+One of the key results of our work is that the proposed method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
 
-While the manifold based methods have been shown to be very effective for precoder feedback,
-they involve operations over higher dimensional manifolds. In
-particular, optimization and interpolation is complicated, especially
-for manifolds where geodesic paths for interpolation may be difficult
-to obtain, especially (don't repeat especially again) for joint time-frequency predictive coding.
-In addition to being analytically more tractable, the scalar parametrization methods have also been
-implemented in 802.11ac standards \cite{porat2013improved}. Past work in utilizing
-scalar parametrization for precoder feedback have utilized
-adaptive delta modulation effectively track precoding matrices both in
-the time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately. In this paper, we propose a method for exploiting the joint
-time-frequency correlation of the precoder's scalar parameters on the
-transmitter using CSI feedback. 
-Typical approaches to predictive
-quantization and interpolation do not exploit the joint time-frequency
-correlation to enhance the precoder estimate at the
-transmitter. However, exploiting the correlations jointly not only
-reduces the feedback requirement, but also enhances the precoder
-reconstruction accuracy significantly.
+Past work in utilizing scalar
+parametrization for precoder feedback have utilized adaptive delta
+modulation effectively track precoding matrices both in the
+time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
+separately.
+Typical natural wireless channels don't just have independent time-frequency correlations, but also posses some novel time-frequency cross correlations.
+\cite{Gupt1905:Predictive} framed an equivalent manifold based approach to utilize joint time-frequency correlations for predictive quantization and interpolation, and showed performance benefits compared to utilizing time-frequency correlations independently \cite{6891198,khaled2005quantized}.
+Intuitively, \cite{Gupt1905:Predictive} observes the performance benefits since utilizing joint correlations not only exploits the existing independent correlations, but also the novel cross correlations in time-frequency arising from the underlying channel models.
 
-Most of the papers utilizing manifolds \cite{Gupt1905:Predictive,pitaval2013codebooks,6891198}  and scalar parametrization \cite{4114278,4556174} compare the performance to other manifold/scalar parametrization methods, but don't do the cross comparison to highlight the performance benefits of the chosen method.
-In addition to a novel joint time-frequency feedback scheme, we compare also the results of our
-algorithm with an equivalent manifold based scheme, and illustrate the performance
-benefits entailed by utilizing the scalar parametrization method. 
-%  Further, the independence of
-% the scalar parameters reduces the problem to one that involves
-% interpolation of separate scalar valued functions, as opposed to
-% operations over tangent spaces in
-% manifolds~\cite{Gupt1905:Predictive}. 
-Simulations reveal that the
-proposed method matches the performance of manifold based approaches
-with a much lower implementation complexity and fewer number of bits
-for quantization.
+Hence, our key contributions are two-fold.  We first propose a method for exploiting the
+joint time-frequency correlation of the precoder's scalar parameters
+on the transmitter using the quantized feedback. 
+Secondly, instead of comparing to similar algorithms in the chosen body of work (viz.
+scalar parametrization), as has been the case with past works, we instead compare our proposed algorithm with an equivalent manifold based approach, and show drastic performance benefits over the manifold approach.
 
@@
 $\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of
-$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\Phi_{i,t}$
+$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\bf{\Phi_{i,t}}$
 is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while
-$\Theta_{i,t}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.
+$\bf{\Theta_{i,t}}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.~\cite{4114278}
 \begin{figure}
@@
 bit for each scalar parameter that is being tracked. On average, this
-can be reduced by half if we transmit only $\Theta$ values for
-one-time instance and only $\Phi$ values for the next time instant in
+can be reduced by half if we transmit only $\bf{\Theta}$ values for
+one-time instance and only $\bf{\Phi}$ values for the next time instant in
 an alternating fashion (the untransmitted value is assumed to be the
@@
 against the ideal precoder for 100 channel evolutions in each run and
-averaged. Simulations are performed for normalised Doppler values of
-$f_dT_s = 3.41\times 10^{-6}$ (corresponding to a velocity of 37.2 km/h
-for $f_c = 1$ GHz) with 1024 subcarriers, and with
+averaged over 10 channel realizations. Simulations are performed for
+normalised Doppler values of $f_dT_s = 3.41\times 10^{-6}$ (corresponding
+to a velocity of 37.2 km/h for $f_c = 1$ GHz) with 1024 subcarriers, and with
 $f_dT_s = 1.56 \times 10^{-5}$ (velocity of 172.8 km/h) with 64
@@
 tracking the channel parameters.
-
+In both the situations, the only-time based interpolation using manifold
+has the worst performance as it is not utilizing the joint time $\&$ frequency 
+correlations. 
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/ber64_001}
+\includegraphics[width=0.5\textwidth]{images/64timebased}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-01-01 paper/main.tex
Space hacks

@@
 can be reduced by half if we transmit only $\bf{\Theta}$ values for
-one-time instance and only $\bf{\Phi}$ values for the next time instant in
-an alternating fashion (the untransmitted value is assumed to be the
-same as that at the previous instant). In this case, the average
-number of bits required to quantize a channel matrix will be
-$N_{T}(2N_{T} - N_{R})/2$.
-
-One issue that is associated with the tracking the $\phi_i$ parameters
-is that they may change abruptly between $-\pi$ and
-$\pi$ due to jumps of $2\pi$. We have avoided this by unwrapping the
-phases to facilitate continuous tracking (i.e., we avoid jumps in
-phase of magnitude close to $2\pi$).
+one-time instance and only $\bf{\Phi}$ values for the next time
+instant in an alternating fashion (the untransmitted value is assumed
+to be the same as that at the previous instant). In this case, the
+average number of bits required to quantize a channel matrix will be
+$N_{T}(2N_{T} - N_{R})/2$. While tracking the $\phi_i$ parameters, is
+must be noted they may change abruptly between $-\pi$ and $\pi$ due to
+jumps of $2\pi$. This is avoided by unwrapping the phases to
+facilitate continuous tracking (i.e., jumps in phase of
+magnitude close to $2\pi$ have been eliminated).
 \subsection{Joint Time Frequency Interpolation}
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/qerror.pdf}
+\includegraphics[width=0.4\textwidth]{images/qerror.pdf}
 \caption{\label{fig:error_decay}A comparison of the $4\times 2$
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/1024timebased}
+\includegraphics[width=0.4\textwidth]{images/1024timebased}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 accurate for scalar parameters than the tangent space based adaptation
-used over the Stiefel manifold, and the outperformance is in excess of
-10 dB. The gap from the optimal precoder performance narrows further
-when the speed reduces.
+used over the Stiefel manifold. The proposed approach outperforms the
+manifold based approach significantly, and the gap from the optimal precoder
+performance narrows further when the speed reduces.
 
@@
 is quite fast, which limits the efficacy of the proposed method in
-tracking the channel parameters.
-In both the situations, the only-time based interpolation using manifold
-has the worst performance as it is not utilizing the joint time $\&$ frequency 
-correlations. 
+tracking the channel parameters. In both the situations, the temporal-only
+based interpolation using manifold has the poorest performance since it does
+not utilize the joint time and frequency correlations.
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/64timebased}
+\includegraphics[width=0.4\textwidth]{images/64timebased}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-01-02 eval.py
adding plots with time based of scalar

@@ -73,7 +73,6 @@ def main():
     net_fb_indices=np.sort(np.append(fb_indices,prev_indices))
     unsorted_net_fb_indices=np.append(fb_indices,prev_indices)
 
-    total_fb_indices=[]
     total_fb_indices=np.array([unsorted_net_fb_indices for i in range(number_simulations//2)]).reshape([-1])
     temp=np.concatenate((np.zeros(np.size(fb_indices)),np.ones(np.size(prev_indices))))
     points=np.array([temp+2*i for i in range(number_simulations//2)]).reshape([-1]).astype(int)
@@ -138,7 +137,9 @@ def main():
 
             allU_vec_copy[simulation_index]=gridz[1]
             allU[simulation_index]=givens_frame_to_unitary(allU_vec_copy[simulation_index],Nt,Nr)
-            
+            pdb.set_trace()            
+            onlyt_allU = griddata(points)
+
             for indice_index in range(fb_indices.shape[0]-1):    
                 curr_freq_index=fb_indices[indice_index]
                 next_freq_index=fb_indices[indice_index+1]
@@ -152,6 +153,12 @@ def main():
             
             if(simulation_index>5):
                 # pdb.set_trace()
+                ot_cap=[np.mean(leakage_analysis(tH_allH[simulation_index],tH_allU[simulation_index],\
+                    onlyt_allU[simulation_index],num_subcarriers,\
+                    waterfilling(tHS[simulation_index].flatten(),10**(0.1*p_dB)*num_subcarriers),\
+                    waterfilling(interpS[simulation_index].flatten(),10**(0.1*p_dB)*num_subcarriers),\
+                    Nt,Nr,ret_abs=True)) for p_dB in 5*np.arange(num_Cappar)]
+
                 max_cap=[np.mean(leakage_analysis(tH_allH[simulation_index],tH_allU[simulation_index],\
                     tH_allU[simulation_index],num_subcarriers,\
                     waterfilling(tHS[simulation_index].flatten(),10**(0.1*p_dB)*num_subcarriers),\
@@ -163,27 +170,39 @@ def main():
                     waterfilling(interpS[simulation_index].flatten(),10**(0.1*p_dB)*num_subcarriers),\
                     Nt,Nr,ret_abs=True)) for p_dB in 5*np.arange(num_Cappar)]
 
+                avg_otcap=(count*avg_otcap+np.array(ot_cap))/(count+1)
[…]

### 2020-01-02 paper/main.tex
adding plots with time based of scalar
Minor fixes

@@
 have been adapted in many modern wireless systems owing to their
-flexibility and ability to use the wideband channels effectively. One
-key requirement to utilize such channels effectively is for the
-transmitter to possess channel state information (CSI), since that
-permits channel parallelization for low-complexity decoders as well as
-the use of capacity achieving transmission
+flexibility and ability to use the wideband channels effectively. To
+utilize such channels effectively, the transmitter needs to have the
+channel state information (CSI), since this permits channel
+parallelization for low-complexity decoders as well as the use of
+capacity achieving transmission
 strategies~\cite{love2008overview}. This primarily translates to the
 generation of unitary precoding matrices at the transmitter. However,
-feeding back CSI from the receiver to the transmitter requires large
-feedback overhead, since a precoder needs to be fed back for each OFDM
-subcarrier within each coherence interval. Typical past approaches
-have used manifold based approaches that exploit temporal and
-frequency correlation to reduce feedback overheads. Wireless standards
-have largely employed parametric quantization that uses Givens
-rotations based quantization~\cite{lou2013comparison,5671092}, though these do
-not involve any refinement of the estimates using channel
-correlations. In this paper, we propose a joint time-frequency
-interpolation approach that builds upon the parametric quantization
-techniques used by standards that manages to outperform typical
-manifold based quantization techniques.
+feeding back CSI from the receiver to the transmitter needs large
+feedback overheads, since a precoder needs to be fed back for each
+OFDM subcarrier within each coherence interval. Past approaches have
+typically used manifold based quantization techniques that exploit
+temporal and frequency correlation to reduce feedback
+overheads. Wireless standards have largely employed parametric
+quantization that uses Givens rotations based
+quantization~\cite{lou2013comparison,5671092}, though these do not
+involve any refinement of the estimates using channel correlations. In
+this paper, we propose a joint time-frequency interpolation approach
+that builds upon the parametric quantization techniques used by
+standards that manages to outperform typical manifold based
+quantization techniques.
 
@@
 \begin{figure}
-\includegraphics[width=0.4\textwidth]{images/1024timebased}
+\includegraphics[width=0.4\textwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 \begin{figure}
-\includegraphics[width=0.4\textwidth]{images/64timebased}
+\includegraphics[width=0.4\textwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-01-02 paper/main.tex
Final changes by Kumar

@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{images/new-adpm.pdf}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{images/new-adpm.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers which need to be
@@
   using the neighbouring parameters, both in time and frequency.}
+\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.4\textwidth]{images/qerror.pdf}
-\caption{\label{fig:error_decay}A comparison of the $4\times 2$
+\begin{center}
+\includegraphics[width=0.37\textwidth]{images/qerror.pdf}
+\caption{\label{fig:error_decay}Comparison of the $4\times 2$
   precoder quantization error vs. time instant for the proposed
@@
   on the Stiefel manifold (labelled ``Cost207\_RA Manifold'') for the COST207 channel profile.}
+\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.4\textwidth]{images/1024final_withtime}
+\begin{center}
+\includegraphics[width=0.37\textwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 \label{fig:ber_ped}
+\end{center}
 \end{figure}
@@
 $4\times 2$ 1024 subcarrier system a slow fading
-($f_dT_s = 3.41 \times 10^{-6}$). We can observe that the performance
-of the proposed method tracks the performance with exact precoders for
-a wide range of SNRs. This can be attributed to the fact that even the
-1-bit update to the scalar parameters is able to approximate the
-precoder very accurately. Since the channel variation across
-subcarriers is gradual, owing to the limited delay spread of the
-channel profile, linear interpolation and prediction yields good
-performance. Moreover, this method outperforms the Stiefel manifold
-based approach due to the fact that the adaptation is much more
-accurate for scalar parameters than the tangent space based adaptation
-used over the Stiefel manifold. The proposed approach outperforms the
-manifold based approach significantly, and the gap from the optimal precoder
-performance narrows further when the speed reduces.
+($f_dT_s = 3.41 \times 10^{-6}$). The time based predictive approaches
+correspond to temporal prediction using the manifold as well as scalar
+parameter based approaches without utilizing the joint time-frequency
+correlations. We can observe that the performance of the proposed
+method tracks the performance with exact precoders over a wide range
+of SNRs. This can be attributed to the fact that even the 1-bit update
+to the scalar parameters is able to approximate the precoder very
+accurately. Since the channel variation across subcarriers is gradual,
+owing to the limited delay spread of the channel profile, linear
+interpolation and prediction yields good performance. Moreover, this
+method outperforms the Stiefel manifold based approach due to the fact
+that the adaptation is much more accurate for scalar parameters than
+the tangent space based adaptation used over the Stiefel manifold. The
+proposed approach outperforms the manifold based approach
+significantly, and the gap from the optimal precoder performance
+narrows further when the speed reduces.
 
@@
 \begin{figure}
-\includegraphics[width=0.4\textwidth]{images/64final_withtime}
+\begin{center}
+\includegraphics[width=0.37\textwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
@@
 \label{fig:ber_veh}
+\end{center}
 \end{figure}

### 2020-01-03 paper/main.tex
Minor changes

@@
 
+% Multiple-input multiple-output (MIMO) based wireless communication
+% systems that use orthogonal frequency division multiplexing (OFDM)
+% have been adapted in many modern wireless systems owing to their
+% flexibility and ability to use the wideband channels effectively. One
+% key requirement to utilize such channels effectively is for the
+% transmitter to possess channel state information (CSI), since that
+% permits channel parallelization for low-complexity decoders as well as
+% the use of capacity achieving transmission
+% strategies~\cite{love2008overview}. This primarily translates to the
+% generation of unitary precoding matrices at the transmitter. However,
+% feeding back CSI from the receiver to the transmitter requires large
+% feedback overhead, since a precoder needs to be fed back for each OFDM
+% subcarrier within each coherence interval. Typical past approaches
+% have used manifold based approaches that exploit temporal and
+% frequency correlation to reduce feedback overheads. Wireless standards
+% have largely employed parametric quantization that uses Givens
+% rotations based quantization~\cite{lou2013comparison,5671092}, though these do
+% not involve any refinement of the estimates using channel
+% correlations. In this paper, we propose a joint time-frequency
+% interpolation approach that builds upon the parametric quantization
+% techniques used by standards that manages to outperform typical
+% manifold based quantization techniques.
+
 Multiple-input multiple-output (MIMO) based wireless communication
 systems that use orthogonal frequency division multiplexing (OFDM)
-have been adapted in many modern wireless systems owing to their
-flexibility and ability to use the wideband channels effectively. To
-utilize such channels effectively, the transmitter needs to have the
-channel state information (CSI), since this permits channel
-parallelization for low-complexity decoders as well as the use of
-capacity achieving transmission
-strategies~\cite{love2008overview}. This primarily translates to the
-generation of unitary precoding matrices at the transmitter. However,
-feeding back CSI from the receiver to the transmitter needs large
-feedback overheads, since a precoder needs to be fed back for each
-OFDM subcarrier within each coherence interval. Past approaches have
-typically used manifold based quantization techniques that exploit
-temporal and frequency correlation to reduce feedback
-overheads. Wireless standards have largely employed parametric
-quantization that uses Givens rotations based
-quantization~\cite{lou2013comparison,5671092}, though these do not
-involve any refinement of the estimates using channel correlations. In
-this paper, we propose a joint time-frequency interpolation approach
-that builds upon the parametric quantization techniques used by
-standards that manages to outperform typical manifold based
-quantization techniques.
+have been adapted in many modern wireless systems owing to increased
+data rate and channel capacity.
+One key requirement to achieve the channel capacity is to leverage spatial multiplexing gains obtained by precoding transmissions by Singular Value Decomposition (SVD) of the MIMO channel matrix ~\cite{love2008overview}.
+However, this requires the receiver to feed back the unitary matrix obtained by SVD, to the transmitter.
+Given that the channel matrix varies with both frequency and time, this feedback has to be ideally done for each subcarrier and time instance, leading to massive feedback overhead.
+To reduce the feedback overhead, past work taps in the inherent temporal and frequency correlations in SVD precoders, in order to perform effective quantization with just a few bits by utilizing predictive quantization \cite{Gupt1905:Predictive,6891198,7370793,sacristan2010differential,5946308,6545375,4114278,4556174}, and reducing number of fed back subcarriers by utilizing interpolation algorithms \cite{Gupt1905:Predictive,5671092,khaled2005quantized}.
 
-Most approaches for effective quantization and feedback of precoders
-for MIMO-OFDM systems involve codebook design over the Grassmann
-manifold~\cite{mondal2007quantization,schwarz2013adaptive,5946308},
-the Stiefel manifold \cite{6891198,Gupt1905:Predictive} and the Flag
-manifold~\cite{pitaval2013codebooks}, wherein the manifold structure
-of the precoders is used to obtain tangents for predictive
-quantization and interpolation across time and frequency.  An
-alternate approach is to parametrize the precoding matrices using
-scalar parameters. The space of unitary matrices can be parameterized
-into independent scalar parameters that correspond to Givens rotations
-and Householder transformations of the unitary precoding matrices
-\cite{4114278,4556174}.
+An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders.
+A large body of works utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
+On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
+While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
+involve operations over higher dimensional manifolds.
+The scalar parametrization algorithms enable usage of the traditional signal processing algorithms over scalars and vectors, and hence are analytically simpler.
+Such methods have also been implemented in 802.11ac standards
+\cite{lou2013comparison}. We are not aware of performance comparisons
+between these two bodies of works, thereby making it difficult to
+quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds.
+In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques.
+One of the key results of our work is that the proposed method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
 
-While the manifold based methods have been shown to be very effective
-for precoder feedback, they involve operations over higher dimensional
-manifolds. In particular, optimization and interpolation is
-complicated, especially for manifolds where geodesic paths for
-interpolation may be difficult to obtain, particularly 
-for joint time-frequency predictive coding.  In
-addition to being analytically more tractable, the scalar
-parametrization methods have also been implemented in 802.11ac
-standards \cite{lou2013comparison}. Past work in utilizing scalar
+Past work in utilizing scalar
 parametrization for precoder feedback have utilized adaptive delta
@@
 time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately. In this paper, we propose a method for exploiting the
+separately. Natural wireless channels not only have temporal and
+frequency correlations, but also possess joint time-frequency
+correlations.
+Prior work \cite{Gupt1905:Predictive} has framed an equivalent manifold based approach to utilize joint time-frequency correlations for predictive quantization and interpolation, and showed performance benefits compared to utilizing time-frequency correlations independently \cite{6891198,khaled2005quantized}.
+Intuitively, these performance arise since utilizing joint
+correlations not only exploits the existing independent correlations,
+but also the joint correlations in time-frequency arising from the underlying channels.
+
+Hence, our key contributions are two-fold.  We first propose a method for exploiting the
 joint time-frequency correlation of the precoder's scalar parameters
-on the transmitter using CSI feedback.  Typical approaches that feed
-back scalar precoder parameters problem do not exploit the joint
-time-frequency correlation to enhance the precoder estimate at the
-transmitter. We show that exploiting correlations jointly not only
-reduces the feedback requirement, but also enhances the precoder
-reconstruction accuracy significantly.
-
-Prior work that utilizing manifolds based quantization
-\cite{Gupt1905:Predictive,pitaval2013codebooks,6891198} and scalar
-parametrization \cite{4114278,4556174} compare their performance with
-other related quantization methods, but do not present a comparison to
-highlight the performance benefits of the chosen method.  In addition
-to a novel joint time-frequency feedback scheme, we also compare the
-results of our parametrization based adaptive scheme with an
-equivalent manifold based scheme, and illustrate the performance
-benefits entailed by utilizing the scalar parametrization method.
-Simulations reveal that the proposed method outperforms manifold based
-approaches with a much lower implementation complexity and fewer bits
-for quantization.
+on the transmitter using the quantized feedback.
+Secondly, instead of comparing similar algorithms in the chosen body of work (viz.
+scalar parametrization), as has been the case with past work, we instead compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.
 

### 2020-01-03 paper/main.tex
More minor fixes

@@
 \includegraphics[width=0.4\textwidth]{images/new-adpm.pdf}
-\caption{\label{fig:adpm-fig}Subcarriers which need to be
-  quantized. In the figure, we see that the subcarrier indicated by
+\caption{\label{fig:adpm-fig}Subcarriers that need to be
+  quantized. We see that the subcarrier indicated by
   red arrows has a precoder whose parameters are determined jointly
@@
 including 802.11ac~\cite{lou2013comparison}, although adaptive
-quantization and feedback are not discussed. We show in the further
-sections that the use of adaptive approaches can greatly enhance
+quantization and feedback are not discussed. Further
+sections show that the use of adaptive approaches can greatly enhance
 performance with only very minor modifications to the system.
@@
 proposed approach outperforms the manifold based approach
-significantly, and the gap from the optimal precoder performance
-narrows further when the speed reduces.
+significantly, and performance is close to the optimal precoder when speeds reduce.
 

### 2020-01-03 paper/main.tex
Hacks to bring it to 4 pages

@@
 
-Fig.~\ref{fig:error_decay} depicts the rate at which error diminishes
+Fig.~\ref{fig:error_decay} shows the rate at which error diminishes
 with time for the proposed approach and direct quantization on the
-Stiefel manifold using codebooks described
-in~\cite{Gupt1905:Predictive}. We see that, since the adaptation of
-independent scalar parameters causes the error to decay rapidly, the
-error is lower and the convergence is faster, even though the starting
-error is higher in the case of the scalar quantization approach. The
-independence of the parameters allows the convergence speed to be very
-fast. The Stiefel manifold based approach uses nested codebooks that
-may be suboptimal when compared to the parametrization using angles.
+Stiefel manifold as described in~\cite{Gupt1905:Predictive}. Since the
+adaptation of independent scalar parameters causes the error to decay
+rapidly, the error is lower and the convergence is faster, even though
+the starting error is higher in the case of the scalar quantization
+approach. The independence of the parameters enables fast
+convergence. The Stiefel manifold based approach uses nested codebooks
+that may be suboptimal when compared to the parametrization using
+angles.
 
@@
 \begin{center}
-\includegraphics[width=0.37\textwidth]{images/qerror.pdf}
+\includegraphics[width=0.33\textwidth]{images/qerror.pdf}
 \caption{\label{fig:error_decay}Comparison of the $4\times 2$
-  precoder quantization error vs. time instant for the proposed
-  approach (labelled ``Cost207\_RA Scalar'') and direct quantization
-  on the Stiefel manifold (labelled ``Cost207\_RA Manifold'') for the COST207 channel profile.}
+  precoder quantization error vs. time for the proposed
+  approach (labelled ``Cost207\_RA Scalar'') and direct Stiefel
+  manifold quantization (``Cost207\_RA Manifold'') for the COST207 channel.}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.37\textwidth]{images/1024final_withtime}
+\includegraphics[width=0.35\textwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 the tangent space based adaptation used over the Stiefel manifold. The
-proposed approach outperforms the manifold based approach
-significantly, and performance is close to the optimal precoder when speeds reduce.
+obtained performance is close to the optimal precoder when speeds reduce.
 
@@
 \begin{center}
-\includegraphics[width=0.37\textwidth]{images/64final_withtime}
+\includegraphics[width=0.35\textwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
@@
 We have presented a predictive quantization based precoder
-reconstruction of a MIMO-OFDM system that employs scalar quantization
+reconstruction for MIMO-OFDM systems that employs scalar quantization
 based feedback of precoder parameters. By quantizing and adapting the
 Givens rotation and Householder transformation parameters of the
-precoder, the receiver can send minimal feedback to the
-transmitter and yet enable effective precoder adaptation. Exploiting
-the temporal and frequency correlations jointly enhances precoder
-accuracy significantly. Simulations reveal that the proposed approach
-provides faster convergence and smaller errors in the precoder when
-compared to approaches that involve direct quantization on manifold
-tangent spaces. Future work would involve tuning the adaptation rate
-of the scalar parameters and further reductions of quantization
-redundancy to reduce the feedback rates.
+precoder, the receiver feedback to the transmitter can be minimized
+while retaining effective precoder adaptation. Exploiting the temporal
+and frequency correlations jointly enhances performance
+significantly. Simulations reveal that the proposed approach provides
+faster convergence and smaller errors when compared to
+approaches that involve direct quantization on manifold tangent
+spaces. Future work would involve tuning the adaptation rate of the
+scalar parameters and further reduction of quantization redundancy to
+reduce the feedback rates.
 

### 2020-01-03 paper/cover_letter.tex
Add cover letter

\documentclass{letter}
\usepackage[left=1in,top=0.15in,twoside=false]{geometry}
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.7}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

% \signature{Agrim Gupta}
\address{From \\
Mohit Madan\\
Department of Electrical  Engineering\\
Indian Institute of Technology Bombay, Mumbai\\
India\\
E-mail: mohitmadan71@gmail.com
\date{\today}
}

\begin{document}
\begin{letter}{To\\
Prof. TODO\\
Editor-in-Chief\\
IEEE Wireless Communication Letters}
\vspace{0.5in}

\opening{Dear Prof. TODO,}

Subject: \textbf{Submission of manuscript title ``Scalar Feedback based Joint Time-Frequency Precoder
  interpolation for  MIMO-OFDM Systems'' for potential publication in
  IEEE Wireless Communication Letters.}

Please find enclosed our manuscript titled ``Scalar Feedback based Joint Time-Frequency Precoder
  interpolation for  MIMO-OFDM Systems'' for review and potential
  publication in IEEE Wireless Communication Letters. The co-authors of this manuscripts are
\begin{enumerate}
\item  Agrim Gupta

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: agrim9gupta@gmail.com

\item Prof. Kumar Appaiah

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: akumar@ee.iitb.ac.in

\end{enumerate}

Should you require additional details, please do e-mail me.

Thank you.
\vspace{0.3in}

Sincerely,

Mohit Madan

\end{letter}
\end{document}

### 2020-01-04 paper/main.tex
Add author names
Minor changes in intro

@@
 
- \author{\authorblockN{Author 1, Author 2, Author 3\\}
+ \author{\authorblockN{Mohit Madan, Agrim Gupta, Kumar Appaiah\\}
  \
@@
 involve operations over higher dimensional manifolds.
-The scalar parametrization algorithms enable usage of the traditional signal processing algorithms over scalars and vectors, and hence are analytically simpler.
+The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
 Such methods have also been implemented in 802.11ac standards
@@
 In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques.
-One of the key results of our work is that the proposed method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
+One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
 
@@
 Prior work \cite{Gupt1905:Predictive} has framed an equivalent manifold based approach to utilize joint time-frequency correlations for predictive quantization and interpolation, and showed performance benefits compared to utilizing time-frequency correlations independently \cite{6891198,khaled2005quantized}.
-Intuitively, these performance arise since utilizing joint
+Intuitively, these performance benefits arise because utilization of joint
 correlations not only exploits the existing independent correlations,
@@
 
-Hence, our key contributions are two-fold.  We first propose a method for exploiting the
+Our key contributions are two-fold.  We first propose a method for exploiting the
 joint time-frequency correlation of the precoder's scalar parameters
 on the transmitter using the quantized feedback.
-Secondly, instead of comparing similar algorithms in the chosen body of work (viz.
-scalar parametrization), as has been the case with past work, we instead compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.
+Secondly, instead of just comparing similar algorithms in the chosen body of work (viz.
+scalar parametrization), as has been the case with past work, we also compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.
 

### 2020-01-04 paper/main.tex
Minor changes in result section

@@
 obtained performance is close to the optimal precoder when speeds reduce. 
+Also, similar to results showed in \cite{Gupt1905:Predictive}, the joint time-frequency 
+method outperforms the independent time-frequency method, illustrating the benefits of exploiting 
+joint correlations versus just the independent correlations.
 
@@
 is quite fast, which limits the efficacy of the proposed method in
-tracking the channel parameters. In both the situations, the temporal-only
-based interpolation using manifold has the poorest performance since it does
-not utilize the joint time and frequency correlations.
+tracking the channel parameters.
 \begin{figure}
@@
 
+\vspace{-10pt}
 % \section{Acknowledgment}

### 2020-01-09 paper/main.tex
Add changes

--- 2020-01-04 paper/main.tex
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.8\columnwidth]{images/new-adpm.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers that need to be
@@
 \begin{center}
-\includegraphics[width=0.33\textwidth]{images/qerror.pdf}
+\includegraphics[width=0.66\columnwidth]{images/qerror.pdf}
 \caption{\label{fig:error_decay}Comparison of the $4\times 2$
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{images/1024final_withtime}
+\includegraphics[width=0.78\columnwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 correlations. We can observe that the performance of the proposed
-method tracks the performance with exact precoders over a wide range
-of SNRs. This can be attributed to the fact that even the 1-bit update
-to the scalar parameters is able to approximate the precoder very
-accurately. Since the channel variation across subcarriers is gradual,
-owing to the limited delay spread of the channel profile, linear
-interpolation and prediction yields good performance. Moreover, this
-method outperforms the Stiefel manifold based approach due to the fact
-that the adaptation is much more accurate for scalar parameters than
-the tangent space based adaptation used over the Stiefel manifold. The
-obtained performance is close to the optimal precoder when speeds reduce.
+method is close to the performance with exact precoders over a wide
+range of SNRs. This can be attributed to the fact that even the 1-bit
+update to the scalar parameters is able to approximate the precoder
+very accurately. Since the channel variation across subcarriers is
+gradual, owing to the limited delay spread of the channel profile,
+linear interpolation and prediction yields good performance. Moreover,
+this method outperforms the Stiefel manifold based approach since
+adaptation of scalar parameters is more accurate than tangent space
+based adaptation used over the Stiefel manifold. The obtained
+performance moves closer to the optimal precoder when speeds reduce.
+Also, similar to results shown in \cite{Gupt1905:Predictive}, the
+joint time-frequency method outperforms the independent time-frequency
+method, illustrating the benefits of exploiting joint correlations
+versus just the independent correlations.
 
-Fig.~\ref{fig:ber_veh} shows a similar comparison for a similar
-situation with a higher speed. The faster variation of the channel
-taps with time necessitates the use of a smaller number of subcarriers
-(viz. 64 subcarriers). The smaller number of subcarriers implies that
-the predictive quantization and interpolation needs to be performed
-across a larger frequency range, thereby potentially reducing the
-accuracy. Even in this case, we observe that the scalar
-parameter-based approach outperforms manifold based quantization by
-about 1 dB for higher SNRs. Naturally, the rate of channel variation
-is quite fast, which limits the efficacy of the proposed method in
-tracking the channel parameters. In both the situations, the temporal-only
-based interpolation using manifold has the poorest performance since it does
-not utilize the joint time and frequency correlations.
+Fig.~\ref{fig:ber_veh} shows a similar comparison for a situation with
+a higher speed. The faster variation of the channel taps with time
+necessitates the use of a smaller number of subcarriers (viz. 64
+subcarriers), thereby permitting more accurate tracking of precoders
+on a per subcarrier basis. As in the previous case, we observe that
+the scalar parameter-based approach outperforms manifold based
+quantization by about 1 dB for higher SNRs. Naturally, the rate of
+channel variation is quite fast, which limits the efficacy of the
+proposed method in tracking the channel parameters.
 \begin{figure}
 \begin{center}
-\includegraphics[width=0.35\textwidth]{images/64final_withtime}
+\includegraphics[width=0.78\columnwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-01-09 paper/main.tex
Minor fixes
Add changes

@@
 average number of bits required to quantize a channel matrix will be
-$N_{T}(2N_{T} - N_{R})/2$. While tracking the $\phi_i$ parameters, is
-must be noted they may change abruptly between $-\pi$ and $\pi$ due to
-jumps of $2\pi$. This is avoided by unwrapping the phases to
-facilitate continuous tracking (i.e., jumps in phase of
-magnitude close to $2\pi$ have been eliminated).
+$(N_{R}(2N_{T} - 1) - N_R^2)/2$ (from~\cite{4114278}). While tracking
+the $\phi_i$ parameters, is must be noted they may change abruptly
+between $-\pi$ and $\pi$ due to jumps of $2\pi$. This is avoided by
+unwrapping the phases to facilitate continuous tracking (i.e., jumps
+in phase of magnitude close to $2\pi$ have been eliminated).
 \subsection{Joint Time Frequency Interpolation}
@@
 s. For both the situations, we considered $N_T=4$ transmit antennas
-and and $N_R=2$ receive antennas. Therefore, the total number of
-parameters that determine the precoder is
-$N_{R}(2N_{T} - N_{R}) = 12$, with five $\theta_l$s and and seven
+and and $N_R=2$ receive antennas. Therefore, using the approach
+outlined in~\cite{4114278}, the parameters that determine the precoder is
+$N_{R}(2N_{T} - 1) - N_R^2 = 10$, with five $\theta_l$s and and seven
 $\phi_i$s. We assume that the spacing between fed back subcarrier
@@
 % $\theta_l$s and $\bf{\Phi}$ is the collection of all $\phi_i$s.
-This brings down the bit budget to an average of 5 bits(the length of $\bf{\Theta}$)
-and 7 bits (the length of $\bf{\Phi}$) over alternate feedback instances,
-thereby resulting in an average feedback budget of 6 bits.
+This brings down the bit budget to an average of 5 bits (the length of
+$\bf{\Theta}$) and 5 bits (the number of non-zero elements in
+$\bf{\Phi}$) over alternate feedback instances, thereby resulting in
+an average feedback budget of 5 bits.
 
@@
 \begin{center}
-\includegraphics[width=0.78\columnwidth]{images/1024final_withtime}
+\includegraphics[width=0.72\columnwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 the scalar parameter-based approach outperforms manifold based
-quantization by about 1 dB for higher SNRs. Naturally, the rate of
-channel variation is quite fast, which limits the efficacy of the
-proposed method in tracking the channel parameters.
+quantization by about 1 dB for higher SNRs. Since, on a per subcarrier
+basis, the channel variation is slower, the precoder tracking is
+better than the previous case. Naturally, the rate of channel
+variation is quite fast, which limits the efficacy of the proposed
+method in tracking the channel parameters.
 \begin{figure}
 \begin{center}
-\includegraphics[width=0.78\columnwidth]{images/64final_withtime}
+\includegraphics[width=0.72\columnwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-01-09 paper/main.tex
Final changes

@@
 The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
-Such methods have also been implemented in 802.11ac standards
-\cite{lou2013comparison}. We are not aware of performance comparisons
+Such methods have also been implemented in 802.11 ac and ax standards
+\cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
 between these two bodies of works, thereby making it difficult to
@@
 \begin{center}
-\includegraphics[width=0.8\columnwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.7\columnwidth]{images/new-adpm.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers that need to be
@@
 \begin{center}
-\includegraphics[width=0.72\columnwidth]{images/1024final_withtime}
+\includegraphics[width=0.71\columnwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 basis, the channel variation is slower, the precoder tracking is
-better than the previous case. Naturally, the rate of channel
-variation is quite fast, which limits the efficacy of the proposed
-method in tracking the channel parameters.
+better than the previous case. Faster channel variations would impose
+limits on the efficacy of the proposed method in tracking the channel
+parameters.
 \begin{figure}
 \begin{center}
-\includegraphics[width=0.72\columnwidth]{images/64final_withtime}
+\includegraphics[width=0.71\columnwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64

### 2020-02-24 paper/response_letter.tex
Add response letter

\documentclass[12pt]{letter}
\usepackage[left=1in,twoside=false]{geometry}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\signature{Jinesh C Jacob}
\address{From \\
Mohit Madan\\
Department of Electrical  Engineering\\
Indian Institute of Technology Bombay, Mumbai\\
India\\
E-mail: mohit.madan@iitb.ac.in
\date{\today}
}

\begin{document}
\begin{letter}{To\\
Prof. Xiao Li\\
Editor\\
IEEE Wireless Communications Letters}
\vspace{0.5in}

\opening{Dear Prof. Li,}

Subject: \textbf{Revised submission of manuscript title ``Scalar
  Feedback based Joint Time-Frequency Precoder interpolation for  MIMO-OFDM Systems'' (Paper Number: WCL2020-0046).}

Authors: Mohit Madan, Agrim Gupta, Kumar Appaiah\\
Email: mohit.madan@iitb.ac.in

Thank you for arranging the prompt reviews of our original manuscript
submitted to the IEEE Wireless Communications Letters, and providing
us an opportunity to resubmit this manuscript with modifications
addressing the issues raised during the reviews. I would like to thank
the reviewers for a careful reading of the manuscript, as well as for
the insightful points that they have raised regarding various
technical aspects of the manuscript. Based on these reviews, we have
made extensive additions and revisions to our submission to address
the reviewers concerns. In particular, we have addressed the key issue
raised by reviewer 2 as well as the Editor, regarding the important
feature of our method that provides us better performance using scalar
quantization. We have also addressed all typos to the best of our
ability. We hope that you will reconsider the manuscript for
publication in the IEEE Wireless Communications Letters. Please find
attached our comments and actions taken based on the reviews, along
with a document indicative of all the changes made.

Please do not hesitate to contact me if I may be of any assistance; I
can be reached by e-mail at mohit.madan@iitb.ac.in. Thank you again
for your consideration of this manuscript.

Thank you.
\vspace{0.3in}

Sincerely,

Mohit Madan

\end{letter}

\newpage

\textbf{Editor comments}\\

\textbf{There are several mistakes and typos in the manuscript, as
  also pointed out by both reviewers. Moreover, it is mentioned that
  the temporal correlation based and frequency correlation based
  precoder tracking have been discussed in [8] and [9]. Thus the
  differences between the proposed method and the methods in [8] and
  [9] should be explained and compared, so as to make the
  contributions more clear.}

\textbf{Response:} Response here

\textbf{Reviewer 1}\\

\textbf{This paper presents an interesting case study of quantized
  precoder design with limited feedback channels. The results show
  that quantization base on scalar parameterization is more effective
  than direct quantization on manifolds. There are, however, two
  issues that need to be addressed before publication.}

\textbf{1. It is not clear whether the feedback bandwidth budgets are
equivalent for methods in comparison. The comparison must be fair.}

\textbf{Response:}

\textbf{ The so-called time-frequency interpolation method is just a
  straightforward two-dim filter. So I assume the that the more
  important factor for performance improvement would be the use of an
  efficient parameterization for the precoder matrix, which comes from
  ref. [8]. I would expect this paper shed more light on why this
  particular parameterization is more efficient in characterizing
  precoder matrices than the description based on manifolds.  Would
  the authors make more explanations/comments on this?}

\textbf{Response:}

\textbf{There are some typos/errors to be fixed:}

\textbf{1. Work/works got mixed up for many times.}

\textbf{2. "The quantization approach described above... " Actually,
  parameterization is described without presenting the quantizing
  method.}

\textbf{3. "While tracking the $\phi_i$ parameters, is must be noted... "}

\textbf{4. "we will use precoder parameters from both the past and future and
time instances to do the interpolation... "}

\textbf{5. "the parameters that determine the precoder is ...  = 10, ... "}

\textbf{Response:}
\end{document}

### 2020-02-26 paper/main.tex
corrected few typos

@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
+\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ {\bf{\eta}}_{i,t}
 \end{equation}
@@
 $\tilde{\bV}_{i,t} \in \mathbb{C}^{N_T \times N_R}$ is the unitary
-precoder, and $\eta$ is the complex additive white Gaussian noise
+precoder, and $\bf{\eta}$ is the complex additive white Gaussian noise
 distributed as ${\mathcal{CN}}(0,I_{N_R})$. We consider the case where
@@
 e^{j\phi_{k,N_T }} \big)$, ${\bf 1}_{k-1}$ represents $k-1$ ones,
-and the $N_R \times N_T$ matrix $\tilde{\bI}$ =
-$\big[\bI_{N_R }, {\bf 0}_{N_T ,N_T -N_R}\big]^{T}$. Here,
-${\bf 0}_{N_T ,N_T -N_R }$ is an $N_T\times (N_T - N_R)$ matrix all of
+and the $N_T \times N_R$ matrix $\tilde{\bI}$ =
+$\big[\bI_{N_R }, {\bf 0}_{N_R ,N_T -N_R}\big]^{T}$. Here,
+${\bf 0}_{N_R ,N_T -N_R }$ is an $N_R\times (N_T - N_R)$ matrix all of
 whose elements are zero. Finally, we also have
@@
 \end{align*}
-Here, $\phi_{k}$s are referred to as phases, and their collection for
+Here, $\phi_{k,l}$s are referred to as phases, and their collection for
 the $i$-th subcarrier at time instant $t$ is captured in a vector
-denoted by $\bf{\Phi_{i,t}}$. $\theta_{l}$s are called rotation angles, and
+denoted by $\bf{\Phi}$$_{i,t}$. $\theta_{k,l}$s are called rotation angles, and
 their collection is captured in a vector referred to as
-$\bf{\Theta_{i,t}}$. We note that $\bf{\Phi_{i,t}}$ and $\bf{\Theta_{i,t}}$ represent
+$\bf{\Theta}$$_{i,t}$. We note that $\bf{\Phi}$$_{i,t}$ and $\bf{\Theta}$$_{i,t}$ represent
 the phase and rotation angles for the $i$-th subcarrier at time
 instant $t$, as represented in Fig.~\ref{fig:adpm-fig}. It is known
-that $\phi_k \in (-\pi, \pi]$ is uniformly distributed between the two
-extremes~\cite{4114278}, while $\theta_l$ is distributed
+that $\phi_{k,l} \in (-\pi, \pi]$ for all k and l's is uniformly distributed 
+between the two
+extremes~\cite{4114278}, while $\theta_{k,l}$ is distributed
 as~\cite{4114278}
 \begin{equation}
-2l(\sin\theta_l)^{2l-1}\cos\theta_l, \mbox{  }\theta_l \in \left[0, \frac{\pi}{2}\right) \mbox{, } l = 1,2,\ldots,N_T -1
+2l(\sin\theta_{k,l})^{2l-1}\cos\theta_{k,l}, \mbox{  }\theta_{k,l} \in \left[0, \frac{\pi}{2}\right) \mbox{, }1\leq k \leq N_R, 1\leq l 
+\leq N_T -k
 \end{equation}
-The given distribution of $\theta_l$ is used for non-uniform quantization
+The given distribution of $\theta_{k,l}$ is used for non-uniform quantization
 of $\bf{\Theta}$ during their initialization to reduce the quantization error.
-In addition, $\phi_k$ and $\theta_l$ are statistically independent,
+In addition, $\phi_{k,l}$ and $\theta_{k,l}$ are statistically independent,
 making this representation useful for tracking them separately across
@@
 matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$. The number of
-$\phi_k$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of
-$\theta$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$. Thus, $\bf{\Phi_{i,t}}$
-is a vector of length $N_{R}(2N_{T} - N_{R}-1)/2$, while
-$\bf{\Theta_{i,t}}$ has length $N_{R}(2N_{T} - N_{R}+1)/2$.~\cite{4114278}
+$\phi_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of
+$\theta{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$. Thus, ${\bf{\Phi}}_{i,t}$
+is a vector of length $N_{R}(2N_{T} - N_{R}+1)/2$, while
+${\bf{\Theta}}_{i,t}$ has length $N_{R}(2N_{T} - N_{R}-1)/2$.~\cite{4114278}
 \begin{figure}
@@
 $\beta_{n} = \mbox{sign}(x_{n} - \hat{x}_{n-1})$, where $x_n$ is the
-unquantized value and $\hat{x}_{n-1}$ is the unquantized (accurate)
+unquantized (accurate) value and $\hat{x}_{n-1}$ is the quantized
 value. Further, the step size for the move can be adapted based on the
@@
 quantization and interpolation method for time-varying MIMO-OFDM
-systems. The channel is simulated using the typical Jakes Model for
+systems. The channel is simulated using the typical Jakes Model 
 for the COST207 channel power delay
@@
 outlined in~\cite{4114278}, the parameters that determine the precoder is
-$N_{R}(2N_{T} - 1) - N_R^2 = 10$, with five $\theta_l$s and and seven
-$\phi_i$s. We assume that the spacing between fed back subcarrier
+$N_{R}(2N_{T} - 2N_R )/2 = 12$, with five $\theta_{k,l}$s and and seven
+$\phi_{k,l}$s. We assume that the spacing between fed back subcarrier
 indices $p$ to be $33$ and $9$ for the slow and fast channel profiles

### 2020-02-27 paper/main.tex
corrected few typos

@@
 
-The quantization approach described above is related to the channel
+The parameterization approach described above is related to the channel
 state information feedback present in several wireless standards,
@@
 average number of bits required to quantize a channel matrix will be
-$(N_{R}(2N_{T} - 1) - N_R^2)/2$ (from~\cite{4114278}). While tracking
-the $\phi_i$ parameters, is must be noted they may change abruptly
+$N_{R}(2N_{T} - N_R )/2$ (from~\cite{4114278}). While tracking
+the $\phi_{k,l}$ parameters, it must be noted they may change abruptly
 between $-\pi$ and $\pi$ due to jumps of $2\pi$. This is avoided by

### 2020-03-02 paper/response_letter.tex
adding reviewer 2 questions-resp. letter

@@
 \textbf{Response:}
+
+\textbf{Reviewer 2}\\
+
+\textbf{1. There are several mistakes and typos in the manuscript. Here are some examples.}
+
+\textbf{(a) In the manuscript, boldface letters are used to denote column vectors and matrices and non-bold letters are used to denote scalars.
+Therefore, the subscripts of $\boldsymbol{\eta_{i,t}}$, $\boldsymbol{\Theta_{i,t}}$ and $\boldsymbol{\Phi_{i,t}}$ should be non-bold, and the covariance matrix of $\boldsymbol{\eta}$, should be boldface.}
+
+\textbf{(b) Because the matrix $\tilde{V_{i,t}}$ is $N_T$ by $N_R$ and the vector $x_{i,t}$ is $N_d$ by 1, the product $V_{i,t}x_{i,t}$ in (1) is not well defined}
+
+\textbf{(c) Because the matrix $I_{N_R}$ is $N_R$ by $N_R$ and the matrix $0_{N_T , N_T - N_R }$ is $N_T$ by $N_T - N_R$, the matrix $\tilde{I}$ = [$I_{N_R}, 0_{N_T, N_T - N_R }$] , in (2) is not well-defined}
+
+\textbf{(d) The right-hand side of (3) is a function of t but the left-hand side of (3) is independent of t.}
+
+\textbf{(e) Do $\theta_l$ and $\phi_k$ represent the same things as $\theta_{k,l}$ and $\phi_{k,l}$ , respectively.}
+
+\textbf{(f) It is mentioned in Subsection III.A that the number of $\phi_k$ parameters is $N_R(2N_T - N_R -1 )/2$ while the number of $\theta$ parameters is $N_R(2N_T -N_R + 1)/2$." However, in the provided example with $N_T$ = 4 and $N_R$ = 2, there are 7 $\phi_{k,l}$ 's and 5 $\theta_{k,l}$'s.}
+
+\textbf{(g) It is mentioned in Subsection III.B that where $x_n$ is the unquantized value and $\tilde{x}_{n-1}$ is the unquantized (accurate) value." What does this mean?}
+
+\textbf{(h) It is mentioned in Subsection III.B that where M is the increment used during the positive and negative moves." However, M is not an increment in (5).}
+
+\textbf{(i) (Section IV) "The channel is simulated using the typical Jakes Model for for the COST207 channel power delay profiles." Here is a typo.}
+
+\textbf{(j) (Section IV) the parameters that determine the precoder is $N_R (2N_T - 1) - N^{2}_R$ = 10, with five $\theta_l$s and and seven $\phi_i$s." Here is a typo because 10 $\neq$ 5 + 7.}
+
+\textbf{(k) What does the sentence "The time based predictive approaches correspond to temporal prediction using the manifold as well as scalar parameter based approaches without utilizing the joint time frequency correlations." in Section IV mean?}
+
+\textbf{Response:}
+
+\textbf{2. The quantization of $\phi_{k,l}$'s and $\theta_{k,l}$'s mentioned in Subsection III.A and the interpolation
+process mentioned in Subsection III.C should be explained more clearly.}
+
+\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future precoder parameters are required to conduct the interpolation. How can we obtain the future precoder parameters?}
+
+\textbf{4. Only 10 channel realizations are used to evaluate the BER characteristics. Is it enough?}
+
+\textbf{5.
+In the manuscript, the time-frequency correlation is used for precoder tracking.
+Because it is mentioned in Section I that the precoder tracking have been discussed in [8] (based on temporal correlation) and [9] (based on frequency correlation), the difference among the proposed method and the methods in [8] and [9] should be explained and compared in the manuscript.}
+
+\textbf{6. The reference format is not consistent and there are several typos. The title of the manuscript is not in IEEE format.}
 \end{document}

### 2020-03-02 paper/main.tex
adding reviewer 2 questions-resp. letter

@@
 
+% It has been shown in Agrimcite that jointly exploiting the time and frequency correlation in MIMO-OFDM systems can significantly enhance performance and reduce the feedback burden.
+However, using a manifold based approach complicates quantized interpolation and prediction.
+The use of matrix codebooks and performing arithmetic on manifolds in the time-frequency grid is significantly more complicated than in the case of scalar parameters.
+The approach that we suggest in this paper exploits the fact that the independent parameterization breaks the problem into several smaller scalar quantization and tracking problems.
+It has been shown that such approaches yield good performance for temporal  cite8 and frequency cite9 domain prediction individually.
+In our work, we perform prediction in the time domain, and use this for joint interpolation on the time frequency grid.
+We show that this allows us to exploit the time and frequency coherence jointly.
+We go on to show that this yields much better performance than individual time and frequency domain prediction, without the complexity of manifold based codebooks.
+
 \section{System Model}
@@
 \begin{equation}
-\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ {\bf{\eta}}_{i,t}
+\by_{i,t} = \bH_{i,t}\tilde{\bV}_{i,t} \bx_{i,t}+ {\boldsymbol{\eta}}_{i,t}
 \end{equation}
@@
 outlined in~\cite{4114278}, the parameters that determine the precoder is
-$N_{R}(2N_{T} - 2N_R )/2 = 12$, with five $\theta_{k,l}$s and and seven
+$N_{R}(2N_{T} - N^{2}_R )/2 = 12$, with five $\theta_{k,l}$s and and seven
 $\phi_{k,l}$s. We assume that the spacing between fed back subcarrier

### 2020-03-03 paper/response_letter.tex
adding changes in paper and response letter

@@
 
-\textbf{Response:} Response here
+\textbf{Response:} Thanks for pointing this out. The temporal based precoder tracking discussed in [8] and [9]
+have discussed about precoder tracking in time and requency respectively without any interpolation. 
+We have approached this differently by performing prediction in the time domain and using this for joint interpolation 
+on the time freqeuncy grid. This helps to build a framework for reducing the total bit bandwidth by a huge amount keeping the
+performace similar.
+We have altered the text in the introduction to make this point more clear. The corrections are as follows:
+``\emph{In our work, we perform prediction in time domain, and use this for joint interpolation on the time frequency grid.}''
 
@@
 \textbf{Response:}
+Thanks for pointing this out.
+We have used an average of 5 bits in our method.
+While the number of feedback bits used in agrimcite is 6 bits on an average.
+We have in fact less number of average bits and still we have got a lower BER than agrim's paper.
 
@@
 
+Thanks for pointing this out.
+First of all the interpolation is simple and accurate as compared to manifold based interpolation which uses complex interpolation methods in the manifold space.
+It is also easier to track the scalar parameters with fewer bits as compared to codebook based method which cannot track its parameters with such high accuracy.
+As mentioned in figure 2, the quantization error is much smaller in given's based parametrization which results in higher accuracy as compared to manifolds with similar number of bits.
+
+Additions have been made in the conclusion section as follows `` \emph{With lesser quantization error compared to manifold based method, the 
+errors in feedback are reduced. Moreover, they are further 
+handled easily by using a simple 2d-interpolation over the scalar parameters.}''
+
 \textbf{There are some typos/errors to be fixed:}
@@
 
+\textbf{Response:} Thanks for pointing this out. It has been changed to "The parametrization approach described above..."
+
 \textbf{3. "While tracking the $\phi_i$ parameters, is must be noted... "}
+
+\textbf{Response:} It has been corrected to "While tracking the $\phi_i$ parameters, it must be noted... "
 
@@
 
+\textbf{Response:} The typo have been corrected to "we will use precoder parameters from both the past and future
+time instances to do the interpolation... "
+
 \textbf{5. "the parameters that determine the precoder is ...  = 10, ... "}
 
-\textbf{Response:}
+\textbf{Response:} Thanks for pointing this out. It has been corrected to ``\emph{...the parameters that determine the precoder is
+$N_{R}(2N_{T} - N_R) = 12$, with five $\theta_{k,l}$s and and seven
+$\phi_{k,l}$s.}''
 
@@
 
+\textbf{Response:} Thanks for pointing out this typo. The subscripts of the symbols have been corrected wherever they have incurred. Also the covariance symbol has been made bold.
+
 \textbf{(b) Because the matrix $\tilde{V_{i,t}}$ is $N_T$ by $N_R$ and the vector $x_{i,t}$ is $N_d$ by 1, the product $V_{i,t}x_{i,t}$ in (1) is not well defined}
+
+\textbf{Response:} Thanks for pointing this out. 
+It was \emph{We consider the case where $N_R < N_T$, and the data vector size $N_d = N_R$}
+which makes $x_{i,t}$ dimensions as $N_R$ which makes it well defined.
 
@@
 
+\textbf{Response:} Thanks for pointing this out. 
+It has been corrected to ``\emph{and the $N_T \times N_R$ matrix $\tilde{\textbf{I}}$ = $\big[\textbf{I}_{N_R }, {\boldsymbol{0}}_{N_R ,N_T -N_R}\big]^{T}$}''
+
 \textbf{(d) The right-hand side of (3) is a function of t but the left-hand side of (3) is independent of t.}
+
+\textbf{Response:} ``t'' has been changed to ``$N_T$''. This makes the right-hand side a square matrix of size $N_T \times N_T$
+which is indeed the size of the matrix $G_{m-1,m}\big(\theta\big)$ on the left-hand side.
 
@@
 
+\textbf{Response:} Thanks for pointing this out.
+Yes, $\theta_l$ and $\phi_k$ represent the same things as $\theta_{k,l}$ and $\phi_{k,l}$.
+Therefore those have been corrected wherever they have come.
+Also equation 8 has been modified as follows:
+$$2l(\sin\theta_{k,l})^{2l-1}\cos\theta_{k,l}, \mbox{  }\theta_{k,l} \in \left[0, \frac{\pi}{2}\right) \mbox{, }1\leq k \leq N_R, 1\leq l 
+\leq N_T -k$$
+
 \textbf{(f) It is mentioned in Subsection III.A that the number of $\phi_k$ parameters is $N_R(2N_T - N_R -1 )/2$ while the number of $\theta$ parameters is $N_R(2N_T -N_R + 1)/2$." However, in the provided example with $N_T$ = 4 and $N_R$ = 2, there are 7 $\phi_{k,l}$ 's and 5 $\theta_{k,l}$'s.}
 
+\textbf{Response:} There was an interchange in the expression of number of parameters. Therefore, it has been corrected as follows: ``\emph{The number of
+$\phi_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of
+$\theta_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$.}''
+
 \textbf{(g) It is mentioned in Subsection III.B that where $x_n$ is the unquantized value and $\tilde{x}_{n-1}$ is the unquantized (accurate) value." What does this mean?}
+\textbf{Response:} It has been corrected to ``\emph{where $x_n$ is the
+unquantized (accurate) value and $\tilde{x}_{n-1}$ is the quantized value}''
 
@@
 
+\textbf{Response:} Thanks for pointing this out. M is the scaling factor and it has been corrected as follows: ``\emph{...where M is the scaling factor used during the positive and...}''
+
 \textbf{(i) (Section IV) "The channel is simulated using the typical Jakes Model for for the COST207 channel power delay profiles." Here is a typo.}
+
+\textbf{Response:} It has been corrected to ``\emph{The channel is simulated using the typical Jakes Model for the COST207 channel power delay profiles.}''
 
@@
 
+\textbf{Response:} Thanks for pointing out this typo. the formula for calculating the precoder is corrected to ``\emph{
+$N_{R}(2N_{T} - N_R) = 12$ , with five $\theta_{k,l}$s and and seven
+$\phi_{k,l}$s.}''
+
 \textbf{(k) What does the sentence "The time based predictive approaches correspond to temporal prediction using the manifold as well as scalar parameter based approaches without utilizing the joint time frequency correlations." in Section IV mean?}
 
-\textbf{Response:}
+\textbf{Response:} Thanks for pointing out this error. The time based predictive approach implied the independent time-frequency based approach. The text in the paper has been corrected as follows: 
+``\emph{Independent time-frequency based approaches correspond to the method without joint-time frequency interpolation i.e.
+with prediction in time domain along with interpolation in the frequency domain.}''
 
@@
 
+\textbf{Response:} 
+Thanks for pointing this out.
+The explanation of quantization process is explained more clearly by altering the text in III.A as follows: ``\emph{Say for n bits initialization for each parameter respectively, the interval $(-\pi,\pi]$
+is equally divided into $2^n$ parts for quantizing $\phi_{k,l}$'s and interval $[0,\frac{\pi}{2})$
+is divided into $2^n$ different quantization levels for quantizing $\theta_{k,l}$'s in order to minimize
+the mean square error.}''\\
+III.C is also altered as follows to make it more clear:
+\emph{``By future time instances it means that
+the precoder estimation is improved at a later time instant i.e.
+after receiving the feedback of the next set of subcarriers.''}
+
 \textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future precoder parameters are required to conduct the interpolation. How can we obtain the future precoder parameters?}
 
+Response: Thank you for this comment. By using the phrase future time instances we mean 
+that we are waiting for the feedback of the next time instance and then improving upon our current 
+precoder estimation using that feedback.
+
+We have altered the text in the subsection III.C as follows: 
+``By future time instances it means that the precoder estimation is improved 
+at a later time instant i.e. after receiving the feedback of the next set of
+subcarriers.''
+
 \textbf{4. Only 10 channel realizations are used to evaluate the BER characteristics. Is it enough?}
+
+Thank you for this comment.
+Although we simulate over 10 realizations, the channel evolves beginning with each realization, resulting in sufficient averaging over various channel characteristics.
+Therefore, the results that we present represent simulation over a sufficiently large number of bits transmitted over enough channel variations.
+We have now altered the text in the simulation section to reflect this as follows:
+
+\emph{The BER achieved by
+the proposed predictive quantization based precoder is compared
+against the ideal precoder for 100 channel evolutions in each run and
+averaged over 10 channel realizations (translating to averaging over $250,000$ bits).
+The error properties obtained by simulating transmission over these varying channels is sufficient to obtain good average performance properties.}
 
@@
 
+\textbf{Response:} Thanks for pointing this out.
+First of all the interpolation is simple and accurate as compared to manifold based interpolation which uses complex interpolation methods in the manifold space.
+It is also easier to track the scalar parameters with fewer bits as compared to codebook based method which cannot track its parameters with such high accuracy.
+As mentioned in figure 2, the quantization error is much smaller in given's based parametrization which results in higher accuracy as compared to manifolds with similar number of bits.
+
+Additions have been made in the conclusion section as follows `` \emph{With lesser quantization error compared to manifold based method, the 
+errors in feedback are reduced. Moreover, they are further 
+handled easily by using a simple 2d-interpolation over the scalar parameters.}''
+
 \textbf{6. The reference format is not consistent and there are several typos. The title of the manuscript is not in IEEE format.}

### 2020-03-03 paper/main.tex
adding changes in paper and response letter

@@
 
-Past work in utilizing scalar
-parametrization for precoder feedback have utilized adaptive delta
+Parametrization for precoder feedback have utilized adaptive delta
 modulation effectively track precoding matrices both in the
 time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately. Natural wireless channels not only have temporal and
+separately. In our work, we perform prediction in time domain, and use this for joint interpolation on the time frequency grid. 
+Natural wireless channels not only have temporal and
 frequency correlations, but also possess joint time-frequency
@@
 \begin{equation}
+\label{eq:decompose}
 \tilde{\bV}_{i,t} = \left[\prod_{k=1}^{N_{R}} \bD_{k} \left( \phi_{k,k},\ldots , \phi_{k,N_{R}} \right) \:  \prod_{l=1}^{N_{T}-k} \bG_{N_{T}-l,N_{T} -l+1} \big( \theta_{k,l}\big)  \right] \: \tilde{\bI}
@@
 $\mbox{diag}\big( {\bf 1}_{k-1}, e^{j\phi_{k,k}},\ldots,
-e^{j\phi_{k,N_T }} \big)$, ${\bf 1}_{k-1}$ represents $k-1$ ones,
+e^{j\phi_{k,N_T }} \big)$ where  dimension $N_T \times N_T$. ${\bf 1}_{k-1}$ represents $k-1$ ones,
 and the $N_T \times N_R$ matrix $\tilde{\bI}$ =
@@
 ${\bf 0}_{N_R ,N_T -N_R }$ is an $N_R\times (N_T - N_R)$ matrix all of
-whose elements are zero. Finally, we also have
+whose elements are zero. Finally $\bG_{m-1,m}\big(\theta\big)$ is a matrix with dimension $N_T \times N_T$ and is given as follows:
 \begin{equation}
@@
 & \sin\theta & \cos\theta & \\
-& & & \bI_{t-m}
+& & & \bI_{N_T -m}
 \end{bmatrix}
@@
 of $\bf{\Theta}$ during their initialization to reduce the quantization error.
+Say for n bits initialization for each parameter respectively, the interval $(-\pi,\pi]$ 
+is equally divided into $2^n$ parts for quantizing $\phi_{k,l}$'s and interval $[0,\frac{\pi}{2})$
+is divided into $2^n$ different quantization levels for quantizing $\theta_{k,l}$'s in order to minimize
+the mean square error.
 In addition, $\phi_{k,l}$ and $\theta_{k,l}$ are statistically independent,
@@
 $\phi_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of
-$\theta{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$. Thus, ${\bf{\Phi}}_{i,t}$
+$\theta_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$. Thus, ${\bf{\Phi}}_{i,t}$
 is a vector of length $N_{R}(2N_{T} - N_{R}+1)/2$, while
@@
 \end{equation}
-where $M$ is the increment used during the positive and negative
+where $M$ is the scaling factor used during the positive and negative
 moves. We initialize $\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We
@@
 precoder parameters from both the past and future and time instances
-to do the interpolation (using backtracking to ensure causality). In
+to do the interpolation (using backtracking to ensure causality). 
+By future time instances it means that the precoder estimation is 
+improved at a later time instant i.e. after receiving the feedback of the next set of
+subcarriers. In
 Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
 neighbouring subcarriers using bilinear interpolation (equivalent to
-linear interpolation in two dimensions). These interpolated values and
-then used to reconstruct the precoding matrices all together. Thus,
+linear interpolation in two dimensions). These interpolated values are
+then used to reconstruct the precoding matrices all together by applying 
+the reverse transformation of breaking the precoder into independent parameters as 
+given in Equation~\ref{eq:decompose}. Thus,
 the approach here is able to exploit the time and frequency
@@
 against the ideal precoder for 100 channel evolutions in each run and
-averaged over 10 channel realizations. Simulations are performed for
+averaged over 10 channel realizations (translating to averaging over $250,000$ bits).
+The error properties obtained by simulating transmission over these varying channels is sufficient to obtain good average performance properties.
+Simulations are performed for
 normalised Doppler values of $f_dT_s = 3.41\times 10^{-6}$ (corresponding
@@
 outlined in~\cite{4114278}, the parameters that determine the precoder is
-$N_{R}(2N_{T} - N^{2}_R )/2 = 12$, with five $\theta_{k,l}$s and and seven
+$N_{R}(2N_{T} - N_R) = 12$, with five $\theta_{k,l}$s and and seven
 $\phi_{k,l}$s. We assume that the spacing between fed back subcarrier
@@
 $4\times 2$ 1024 subcarrier system a slow fading
-($f_dT_s = 3.41 \times 10^{-6}$). The time based predictive approaches
-correspond to temporal prediction using the manifold as well as scalar
-parameter based approaches without utilizing the joint time-frequency
-correlations. We can observe that the performance of the proposed
+($f_dT_s = 3.41 \times 10^{-6}$).
+Independent time-frequency based approaches correspond to the method without joint-time frequency interpolation i.e.
+with prediction in time domain along with interpolation in the frequency domain.
+ % The time based predictive approaches
+% correspond to temporal prediction using the manifold as well as scalar
+% parameter based approaches without utilizing the joint time-frequency
+% correlations. 
+We can observe that the performance of the proposed
 method is close to the performance with exact precoders over a wide
@@
 precoder, the receiver feedback to the transmitter can be minimized
-while retaining effective precoder adaptation. Exploiting the temporal
+while retaining effective precoder adaptation. 
+Exploiting the temporal
 and frequency correlations jointly enhances performance
@@
 approaches that involve direct quantization on manifold tangent
-spaces. Future work would involve tuning the adaptation rate of the
+spaces. With lesser quantization error compared to manifold based method, the 
+errors in feedback are reduced. Moreover, they are further 
+handled easily by using a simple 2d-interpolation over the scalar parameters.
+Future work would involve tuning the adaptation rate of the
 scalar parameters and further reduction of quantization redundancy to

### 2020-03-03 paper/response_letter.tex
Changes by Kumar

--- 2020-03-02 paper/response_letter.tex
@@
 
-\textbf{Response:} Response here
+\textbf{Response:} Thank you for pointing this out. The temporal
+correlation based precoder tracking discussed in [8] and [9] have
+considered about precoder tracking in time and frequency respectively,
+albeit separately, without any joint interpolation.  We have
+approached this differently by performing prediction in the time
+domain and using this for joint interpolation on the time-freqeuncy
+grid. This helps build a framework for reducing the total bit
+bandwidth by a signficant amount while maintaining similar performance
+in terms of BER.  We have altered the text in the introduction to make
+this point more clear,  as follows:
+
+``\emph{In our work, we perform prediction in time domain, and use this for joint
+  interpolation on the time frequency grid.}''
 
@@
 
-\textbf{Response:}
+\textbf{Response:} Thank you for pointing this out. We have used an
+average of 5 bits in our method. While the number of feedback bits
+used in [2] is 6 bits per OFDM frame, the approach that we suggest
+uses a smaller number bits on average while still achieving similar or
+better BER performance. We have now added a brief indication of this
+in the text as follows
+
+\emph{Since the bit budget is small, if the channel varies
+sufficiently slowly, the $\theta_{k,l}$ and the $\phi_{k,l}$ parameters
+are fed back alternately, while retaining the unsent parameters as
+their previous values, as shown in
+Fig. 1. This brings down the bit budget to an average of 5 bits for
+$\bf{\Theta}$ and 5 bits for $\bf{\Phi}$ over alternate feedback
+instances, thereby resulting in an average feedback budget of 5 bits,
+as opposed to the 6 bits of feedback used
+in [2].}
 
@@
 
+Thank you for the feedback. The interpolation approach that we have
+proposed is simple and ends up becoming more accurate over time when
+compared to the manifold based interpolation that uses complex
+interpolation methods in the manifold space. It is also easier to
+track the adaptation of the scalar parameters using few bits. As
+observed in Fig. 2, the quantization error obtained when using the
+parameter based approach taken in this manuscript is much smaller in
+that obtained using the conventional codebooks on the Stiefel
+manifold, as discussed in [2]. This results in higher accuracy as
+compared to manifolds with similar number of bits.
+
+To emphasize these points, we have made the following addition:
+
+\emph{With smaller
+quantization error as compared to manifold based approaches, the
+feedback results in more accurate precoder tracking, along with simple
+2D-interpolation over the scalar parameters to optimize the precoders.}
+
 \textbf{There are some typos/errors to be fixed:}
@@
 
-\textbf{2. "The quantization approach described above... " Actually,
+Thank you for this comment. We have fixed all instances of this error.
+
+\textbf{2. ``The quantization approach described above... '' Actually,
   parameterization is described without presenting the quantizing
@@
 
+\textbf{Response:} We thank the reviewer for pointing out this
+error. We have altered the text to include this:
+
+\emph{The distribution of $\theta_{k,l}$ is used for
+quantizing $\bf{\Theta}$ during initialization to reduce the
+quantization error.  If $n$ bits are allocated to quantize a
+parameter for the first time, the range of values that each parameter
+takes is divided into $2^n$ optimal intervals using Lloyd's
+quantization [17].}
+
 \textbf{3. "While tracking the $\phi_i$ parameters, is must be noted... "}
+
+\textbf{Response:} It has been corrected to \emph{``While tracking the $\phi_i$ parameters, it must be noted... ''}.
 
@@
 
+\textbf{Response:} Thank you. The typo has now been corrected to read
+``we will use precoder parameters from both the past and future time instances to do the interpolation... ''
+
 \textbf{5. "the parameters that determine the precoder is ...  = 10, ... "}
 
-\textbf{Response:}
+\textbf{Response:} Thanks for pointing this out. It has been corrected to ``\emph{...the parameters that determine the precoder is
+$N_{R}(2N_{T} - N_R) = 12$, with five $\theta_{k,l}$s and and seven
+$\phi_{k,l}$s.}''
 
@@
 
-\textbf{(b) Because the matrix $\tilde{V_{i,t}}$ is $N_T$ by $N_R$ and the vector $x_{i,t}$ is $N_d$ by 1, the product $V_{i,t}x_{i,t}$ in (1) is not well defined}
+\textbf{Response:} Thanks for pointing out this typo. The subscripts
+of the symbols have been corrected wherever they have incurred. In addition,
+the covariance symbol have now been made bold.
 
-\textbf{(c) Because the matrix $I_{N_R}$ is $N_R$ by $N_R$ and the matrix $0_{N_T , N_T - N_R }$ is $N_T$ by $N_T - N_R$, the matrix $\tilde{I}$ = [$I_{N_R}, 0_{N_T, N_T - N_R }$] , in (2) is not well-defined}
+\textbf{(b) Because the matrix $\tilde{V_{i,t}}$ is $N_T$ by $N_R$ and
+  the vector $x_{i,t}$ is $N_d$ by 1, the product $V_{i,t}x_{i,t}$ in
+  (1) is not well defined}.
+
+\textbf{Response:} Thanks for pointing this out.  It now reads
+\emph{We consider the case where $N_R < N_T$, and the data vector size
+  $N_d = N_R$} which makes $x_{i,t}$ dimensions as $N_R$ which now
+makes it well defined.
+
+\textbf{(c) Because the matrix $I_{N_R}$ is $N_R$ by $N_R$ and the
+  matrix $0_{N_T , N_T - N_R }$ is $N_T$ by $N_T - N_R$, the matrix
+  $\tilde{I}$ = [$I_{N_R}, 0_{N_T, N_T - N_R }$] , in (2) is not
+  well-defined}
+
+\textbf{Response:} Thanks for pointing this out.  We have now fixed
+this to read:
+
+``\emph{and the $N_T \times N_R$ matrix
+  $\tilde{\textbf{I}}$ =
+  $\big[\textbf{I}_{N_R }, {\boldsymbol{0}}_{N_R ,N_T
+    -N_R}\big]^{T}$}''
 
@@
 
+\textbf{Response:} ``t'' has been changed to ``$N_T$''. This makes the
+right-hand side a square matrix of size $N_T \times N_T$ which is
+indeed the size of the matrix $G_{m-1,m}\big(\theta\big)$ on the
+left-hand side.
+
 \textbf{(e) Do $\theta_l$ and $\phi_k$ represent the same things as $\theta_{k,l}$ and $\phi_{k,l}$ , respectively.}
 
-\textbf{(f) It is mentioned in Subsection III.A that the number of $\phi_k$ parameters is $N_R(2N_T - N_R -1 )/2$ while the number of $\theta$ parameters is $N_R(2N_T -N_R + 1)/2$." However, in the provided example with $N_T$ = 4 and $N_R$ = 2, there are 7 $\phi_{k,l}$ 's and 5 $\theta_{k,l}$'s.}
+\textbf{Response:} Thanks for pointing this out. Yes, $\theta_l$ and
+$\phi_k$ represent the same things as $\theta_{k,l}$ and
+$\phi_{k,l}$. Therefore, those references have now been corrected
+wherever they appeared. In addition, we have also specified the pdf of
+the $\theta_{k,l}$ clearly.
+
+\textbf{(f) It is mentioned in Subsection III.A that the number of
+  $\phi_k$ parameters is $N_R(2N_T - N_R -1 )/2$ while the number of
+  $\theta$ parameters is $N_R(2N_T -N_R + 1)/2$." However, in the
+  provided example with $N_T$ = 4 and $N_R$ = 2, there are 7
+  $\phi_{k,l}$ 's and 5 $\theta_{k,l}$'s.}
+
+\textbf{Response:} There was an error and we had interchanged the
+expression for the number of parameters. Therefore, we have now
+corrected it as follows:
+
+``\emph{The number of $\phi_{k,l}$ parameters is
+  $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of $\theta_{k,l}$
+  parameters is $N_{R}(2N_{T} - N_{R}-1)/2$.}''
 
@@
 
+\textbf{Response:} We have now corrected this to ``\emph{where $x_n$ is
+  the unquantized (accurate) value and $\tilde{x}_{n-1}$ is the
+  quantized value}''
+
 \textbf{(h) It is mentioned in Subsection III.B that where M is the increment used during the positive and negative moves." However, M is not an increment in (5).}
+
+\textbf{Response:} Thanks for pointing this out. $M$ is the scaling
+factor, and we have now corrected the reference to $M$ as follows:
+
+``\emph{...where M is the scaling factor used during the positive and...}''
 
@@
 
+\textbf{Response:} It has now been corrected to ``\emph{The channel is simulated using the typical Jakes Model for the COST207 channel power delay profiles.}''
+
 \textbf{(j) (Section IV) the parameters that determine the precoder is $N_R (2N_T - 1) - N^{2}_R$ = 10, with five $\theta_l$s and and seven $\phi_i$s." Here is a typo because 10 $\neq$ 5 + 7.}
+
+\textbf{Response:} Thanks for pointing out this typo. The formula for
+calculating the precoder has now been corrected to ``\emph{
+  $N_{R}(2N_{T} - N_R) = 12$ , with five $\theta_{k,l}$s and and seven
+  $\phi_{k,l}$s.}''
 
@@
 
-\textbf{Response:}
+\textbf{Response:} Thanks for pointing out this error. The time based
+predictive approach was meant to imply the independent time-frequency based
+approach. The text in the paper has been corrected as follows:
+
+``\emph{Independent time-frequency based approaches correspond to the
+  method without joint-time frequency interpolation i.e. with
+  prediction in time domain along with separate interpolation in the frequency
+  domain.}''
 
@@
 
-\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future precoder parameters are required to conduct the interpolation. How can we obtain the future precoder parameters?}
+\textbf{Response:}
+Thanks for pointing this out. The explanation of the quantization
+process has now been made more  clear by altering the text in III.A as
+follows:
+
+``\emph{If $n$ bits are allocated to quantize a parameter
+for the first time, the range of values that each parameter takes is
+divided into $2^n$ optimal intervals using Lloyd's
+quantization [17].}''\\
+
+In addition, Section III.C is also altered as follows to make it more
+clear:
+
+\emph{``By
+  future time instances it means that the precoder estimation is
+  improved at a later time instant i.e.  after receiving the feedback
+  of the next set of subcarriers.''}
+
+\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future
+  precoder parameters are required to conduct the interpolation. How
+  can we obtain the future precoder parameters?}
+
+\textbf{Response:} AGRIM TODO
+
+We have altered the text in the subsection III.C as follows:
+``By future time instances it means that the precoder estimation is improved
+at a later time instant i.e. after receiving the feedback of the next set of
+subcarriers.''
 
 \textbf{4. Only 10 channel realizations are used to evaluate the BER characteristics. Is it enough?}
+
+Thank you for this comment. Although we simulate over 10 realizations,
+the channel evolves beginning with each realization, resulting in
+sufficient averaging over various channel characteristics. Therefore,
+the results that we present represent simulation over a sufficiently
+large number of bits transmitted over enough channel variations. We
+have now altered the text in the simulation section to reflect this as
+follows:
+
+\emph{The BER achieved by the proposed predictive quantization based
+  precoder is compared against the ideal precoder for 100 channel
+  evolutions in each run and averaged over 10 channel realizations
+  (translating to averaging over $125,000$ bits for 64-point DFT and
+  $2\times 10^6$ bits for the 1024-point DFT case). The error
+  properties obtained by simulating transmission over these varying
+  channels is sufficient to obtain good average performance
+  properties.}
 
@@
 
+\textbf{Response:} Thanks for pointing this out.  First of all the
+interpolation is simple and accurate as compared to manifold based
+interpolation which uses complex interpolation methods in the manifold
+space. It is also easier to track the scalar parameters with fewer
+bits as compared to codebook based method which cannot track its
+parameters with such high accuracy. As mentioned in figure 2, the
+quantization error is much smaller in given's based parametrization
+which results in higher accuracy as compared to manifolds with similar
+number of bits.
+
+Additions have been made in the conclusion section as follows `` \emph{With lesser quantization error compared to manifold based method, the
+errors in feedback are reduced. Moreover, they are further
+handled easily by using a simple 2d-interpolation over the scalar parameters.}''
+
 \textbf{6. The reference format is not consistent and there are several typos. The title of the manuscript is not in IEEE format.}

### 2020-03-03 paper/main.tex
Changes by Kumar

@@
 
+% It has been shown in Agrimcite that jointly exploiting the time and frequency correlation in MIMO-OFDM systems can significantly enhance performance and reduce the feedback burden.
+However, using a manifold based approach complicates quantized interpolation and prediction.
+The use of matrix codebooks and performing arithmetic on manifolds in the time-frequency grid is significantly more complicated than in the case of scalar parameters.
+The approach that we suggest in this paper exploits the fact that the independent parameterization breaks the problem into several smaller scalar quantization and tracking problems.
+It has been shown that such approaches yield good performance for temporal  cite8 and frequency cite9 domain prediction individually.
+In our work, we perform prediction in the time domain, and use this for joint interpolation on the time frequency grid.
+We show that this allows us to exploit the time and frequency coherence jointly.
+We go on to show that this yields much better performance than individual time and frequency domain prediction, without the complexity of manifold based codebooks.
+
 An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders.
-A large body of works utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
+A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
 On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
@@
 \cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
-between these two bodies of works, thereby making it difficult to
+between these two bodies of work, thereby making it difficult to
 quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds.
@@
 time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
-separately. In our work, we perform prediction in time domain, and use this for joint interpolation on the time frequency grid. 
-Natural wireless channels not only have temporal and
-frequency correlations, but also possess joint time-frequency
-correlations. 
-Prior work \cite{Gupt1905:Predictive} has framed an equivalent manifold based approach to utilize joint time-frequency correlations for predictive quantization and interpolation, and showed performance benefits compared to utilizing time-frequency correlations independently \cite{6891198,khaled2005quantized}.
-Intuitively, these performance benefits arise because utilization of joint
-correlations not only exploits the existing independent correlations,
-but also the joint correlations in time-frequency arising from the underlying channels.
+separately. In our work, we perform prediction in time domain, and use
+this for joint interpolation on the time frequency grid. Natural
+wireless channels not only have temporal and frequency correlations,
+but also possess joint time-frequency correlations.  Prior work
+\cite{Gupt1905:Predictive} has framed an equivalent manifold based
+approach to utilize joint time-frequency correlations for predictive
+quantization and interpolation, and showed performance benefits
+compared to utilizing time-frequency correlations independently
+\cite{6891198,khaled2005quantized}.  Intuitively, these performance
+benefits arise because utilization of joint correlations not only
+exploits the existing independent correlations, but also the joint
+correlations in time-frequency arising from the underlying channels.
 
@@
 % Section~\ref{section4} concludes with discussions on future work.
-
-% It has been shown in Agrimcite that jointly exploiting the time and frequency correlation in MIMO-OFDM systems can significantly enhance performance and reduce the feedback burden.
-However, using a manifold based approach complicates quantized interpolation and prediction.
-The use of matrix codebooks and performing arithmetic on manifolds in the time-frequency grid is significantly more complicated than in the case of scalar parameters.
-The approach that we suggest in this paper exploits the fact that the independent parameterization breaks the problem into several smaller scalar quantization and tracking problems.
-It has been shown that such approaches yield good performance for temporal  cite8 and frequency cite9 domain prediction individually.
-In our work, we perform prediction in the time domain, and use this for joint interpolation on the time frequency grid.
-We show that this allows us to exploit the time and frequency coherence jointly.
-We go on to show that this yields much better performance than individual time and frequency domain prediction, without the complexity of manifold based codebooks.
 
@@
 \end{align*}
-Here, $\phi_{k,l}$s are referred to as phases, and their collection for
-the $i$-th subcarrier at time instant $t$ is captured in a vector
-denoted by $\bf{\Phi}$$_{i,t}$. $\theta_{k,l}$s are called rotation angles, and
-their collection is captured in a vector referred to as
-$\bf{\Theta}$$_{i,t}$. We note that $\bf{\Phi}$$_{i,t}$ and $\bf{\Theta}$$_{i,t}$ represent
-the phase and rotation angles for the $i$-th subcarrier at time
-instant $t$, as represented in Fig.~\ref{fig:adpm-fig}. It is known
-that $\phi_{k,l} \in (-\pi, \pi]$ for all k and l's is uniformly distributed 
-between the two
-extremes~\cite{4114278}, while $\theta_{k,l}$ is distributed
-as~\cite{4114278}
-\begin{equation}
-2l(\sin\theta_{k,l})^{2l-1}\cos\theta_{k,l}, \mbox{  }\theta_{k,l} \in \left[0, \frac{\pi}{2}\right) \mbox{, }1\leq k \leq N_R, 1\leq l 
-\leq N_T -k
-\end{equation}
-The given distribution of $\theta_{k,l}$ is used for non-uniform quantization
-of $\bf{\Theta}$ during their initialization to reduce the quantization error.
-Say for n bits initialization for each parameter respectively, the interval $(-\pi,\pi]$ 
-is equally divided into $2^n$ parts for quantizing $\phi_{k,l}$'s and interval $[0,\frac{\pi}{2})$
-is divided into $2^n$ different quantization levels for quantizing $\theta_{k,l}$'s in order to minimize
-the mean square error.
-In addition, $\phi_{k,l}$ and $\theta_{k,l}$ are statistically independent,
-making this representation useful for tracking them separately across
-time, as discussed below. The total number of
-parameters obtained from the decomposition of a complex orthogonal
-matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$. The number of
-$\phi_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of
-$\theta_{k,l}$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$. Thus, ${\bf{\Phi}}_{i,t}$
+Here, $\phi_{k,l}$s are referred to as phases, and their collection
+for the $i$-th subcarrier at time instant $t$ is captured in a vector
+denoted by $\bf{\Phi}$$_{i,t}$.
+$\theta_{k,l}$s are called rotation angles, and their collection is
+captured in a vector referred to as
+$\bf{\Theta}$$_{i,t}$. We note that
+$\bf{\Phi}$$_{i,t}$ and $\bf{\Theta}$$_{i,t}$ represent the phase and
+rotation angles for the $i$-th subcarrier at time instant $t$, as
+represented in Fig.~\ref{fig:adpm-fig}. It is known~\cite{4114278}
+that $\phi_{k,l} \in (-\pi, \pi]$ for all k and l's is uniformly
+distributed between the two extremes~\cite{4114278}, while
+$\theta_{k,l}$ is distributed as
+$2l(\sin\theta_{k,l})^{2l-1}\cos\theta_{k,l}$, for
+$\theta_{k,l} \in \left[0, \frac{\pi}{2}\right)$, $1\leq k \leq N_R$,
+$1\leq l \leq N_T -k$.  The distribution of $\theta_{k,l}$ is used for
+quantizing $\bf{\Theta}$ during initialization to reduce the
+quantization error.  If $n$ bits are allocated to quantize a parameter
+for the first time, the range of values that each parameter takes is
+divided into $2^n$ optimal intervals using Lloyd's
+quantization~\cite{lloyd1982least}. In addition, $\phi_{k,l}$ and
+$\theta_{k,l}$ are statistically independent, making this
+representation useful for tracking them separately across time, as
+discussed below. The total number of parameters obtained from the
+decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is
+$N_{R}(2N_{T} - N_{R})$. The number of $\phi_{k,l}$ parameters is
+$N_{R}(2N_{T} - N_{R}+1)/2$ while the number of $\theta_{k,l}$
+parameters is $N_{R}(2N_{T} - N_{R}-1)/2$. Thus, ${\bf{\Phi}}_{i,t}$
 is a vector of length $N_{R}(2N_{T} - N_{R}+1)/2$, while
-${\bf{\Theta}}_{i,t}$ has length $N_{R}(2N_{T} - N_{R}-1)/2$.~\cite{4114278}
+${\bf{\Theta}}_{i,t}$ has length
+$N_{R}(2N_{T} - N_{R}-1)/2$.~\cite{4114278}
 \begin{figure}
@@
 parameter. Since the bit budget is small, if the channel varies
-sufficiently slowly, either the $\theta_l$ or the $\phi_i$ parameters
+sufficiently slowly, the $\theta_{k,l}$ and the $\phi_{k,l}$ parameters
 are fed back alternately, while retaining the unsent parameters as
-being to their previous values, as shown in
+their previous values, as shown in
 Fig.~\ref{fig:adpm-fig}.
@@
 % $\theta_l$s and $\bf{\Phi}$ is the collection of all $\phi_i$s.
-This brings down the bit budget to an average of 5 bits (the length of
-$\bf{\Theta}$) and 5 bits (the number of non-zero elements in
-$\bf{\Phi}$) over alternate feedback instances, thereby resulting in
-an average feedback budget of 5 bits.
+This brings down the bit budget to an average of 5 bits for
+$\bf{\Theta}$ and 5 bits for $\bf{\Phi}$ over alternate feedback
+instances, thereby resulting in an average feedback budget of 5 bits,
+as opposed to the 6 bits of feedback used
+in~\cite{Gupt1905:Predictive}.
 
@@
 $4\times 2$ 1024 subcarrier system a slow fading
-($f_dT_s = 3.41 \times 10^{-6}$).
-Independent time-frequency based approaches correspond to the method without joint-time frequency interpolation i.e.
-with prediction in time domain along with interpolation in the frequency domain.
+($f_dT_s = 3.41 \times 10^{-6}$). Independent time-frequency based
+approaches correspond to the method without joint-time frequency
+interpolation i.e. with prediction in time domain along with separate interpolation in the frequency domain.
  % The time based predictive approaches
@@
 precoder, the receiver feedback to the transmitter can be minimized
-while retaining effective precoder adaptation. 
-Exploiting the temporal
-and frequency correlations jointly enhances performance
+while retaining effective precoder adaptation.  Exploiting the
+temporal and frequency correlations jointly enhances performance
 significantly. Simulations reveal that the proposed approach provides
-faster convergence and smaller errors when compared to
-approaches that involve direct quantization on manifold tangent
-spaces. With lesser quantization error compared to manifold based method, the 
-errors in feedback are reduced. Moreover, they are further 
-handled easily by using a simple 2d-interpolation over the scalar parameters.
-Future work would involve tuning the adaptation rate of the
-scalar parameters and further reduction of quantization redundancy to
-reduce the feedback rates.
+faster convergence and smaller errors when compared to approaches that
+involve direct quantization on manifold tangent spaces. With smaller
+quantization error as compared to manifold based approaches, the
+feedback results in more accurate precoder tracking, along with simple
+2D-interpolation over the scalar parameters to optimize the precoders.
+Future work would involve tuning the adaptation rate of the scalar
+parameters and further reduction of quantization redundancy to reduce
+the feedback rates.
 

### 2020-03-03 paper/response_letter.tex
Add more changes

@@
 grid. This helps build a framework for reducing the total bit
-bandwidth by a signficant amount while maintaining similar performance
+bandwidth by a significant amount while maintaining similar performance
 in terms of BER.  We have altered the text in the introduction to make
@@
 
-\textbf{Response:} Thanks for pointing this out.  First of all the
-interpolation is simple and accurate as compared to manifold based
-interpolation which uses complex interpolation methods in the manifold
-space. It is also easier to track the scalar parameters with fewer
-bits as compared to codebook based method which cannot track its
-parameters with such high accuracy. As mentioned in figure 2, the
-quantization error is much smaller in given's based parametrization
-which results in higher accuracy as compared to manifolds with similar
-number of bits.
+Thank you for the feedback. The interpolation approach that we have
+proposed is simple and ends up becoming more accurate over time when
+compared to the manifold based interpolation that uses complex
+interpolation methods in the manifold space. It is also easier to
+track the adaptation of the scalar parameters using few bits. As
+observed in Fig. 2, the quantization error obtained when using the
+parameter based approach taken in this manuscript is much smaller in
+that obtained using the conventional codebooks on the Stiefel
+manifold, as discussed in [2]. This results in higher accuracy as
+compared to manifolds with similar number of bits.
 
-Additions have been made in the conclusion section as follows `` \emph{With lesser quantization error compared to manifold based method, the
-errors in feedback are reduced. Moreover, they are further
-handled easily by using a simple 2d-interpolation over the scalar parameters.}''
+To emphasize these points, we have made the following addition:
+
+\emph{With smaller
+quantization error as compared to manifold based approaches, the
+feedback results in more accurate precoder tracking, along with simple
+2D-interpolation over the scalar parameters to optimize the precoders.}
+
+In addition, the temporal correlation based precoder tracking
+discussed in [8] and [9] have considered about precoder tracking in
+time and frequency respectively, albeit separately, without any joint
+interpolation.  We have approached this differently by performing
+prediction in the time domain and using this for joint interpolation
+on the time-freqeuncy grid. This helps build a framework for reducing
+the total bit bandwidth by a significant amount while maintaining
+similar performance in terms of BER.  We have altered the text in the
+introduction to make this point more clear, as follows:
+
+``\emph{In our work, we perform prediction in time domain, and use this for joint
+  interpolation on the time frequency grid.}''
+
+This is in addition to various changes in other sections of the paper
+that further emphasize the advantages of our approach.
 

### 2020-03-04 paper/main.tex
Add changes

@@
 
+\def\imwidth{0.65}
 %package list
@@
 where $M$ is the scaling factor used during the positive and negative
-moves. We initialize $\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. We
-note that this is performed separately for each of the quantized
+moves. We initialize $\Delta_1$ as $\Delta_1 = |x_{2}-\hat{x}_1|$. This is performed separately for each of the quantized
 parameters.
@@
 for the COST207 channel power delay
-profiles~\cite{molisch2006cost259,cost1989cost}. The BER achieved by
+profiles~\cite{molisch2006cost259}. The BER achieved by
 the proposed predictive quantization based precoder is compared
@@
 instances, thereby resulting in an average feedback budget of 5 bits,
-as opposed to the 6 bits of feedback used
-in~\cite{Gupt1905:Predictive}.
+as opposed to the 6 bits in~\cite{Gupt1905:Predictive}.
 
@@
 \begin{center}
-\includegraphics[width=0.66\columnwidth]{images/qerror.pdf}
+\includegraphics[width=\imwidth\columnwidth]{images/qerror.pdf}
 \caption{\label{fig:error_decay}Comparison of the $4\times 2$
@@
 \begin{center}
-\includegraphics[width=0.71\columnwidth]{images/1024final_withtime}
+\includegraphics[width=\imwidth\columnwidth]{images/1024final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
@@
 \begin{center}
-\includegraphics[width=0.71\columnwidth]{images/64final_withtime}
+\includegraphics[width=\imwidth\columnwidth]{images/64final_withtime}
 \caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
@@
 reconstruction for MIMO-OFDM systems that employs scalar quantization
-based feedback of precoder parameters. By quantizing and adapting the
-Givens rotation and Householder transformation parameters of the
-precoder, the receiver feedback to the transmitter can be minimized
-while retaining effective precoder adaptation.  Exploiting the
-temporal and frequency correlations jointly enhances performance
-significantly. Simulations reveal that the proposed approach provides
-faster convergence and smaller errors when compared to approaches that
-involve direct quantization on manifold tangent spaces. With smaller
-quantization error as compared to manifold based approaches, the
-feedback results in more accurate precoder tracking, along with simple
-2D-interpolation over the scalar parameters to optimize the precoders.
-Future work would involve tuning the adaptation rate of the scalar
-parameters and further reduction of quantization redundancy to reduce
-the feedback rates.
+based feedback of precoder parameters. Quantizing and adapting the
+Givens rotation and Householder transformation parameters of precoders
+minimizes receiver feedback to the transmitter while retaining
+effective precoder adaptation. Exploiting the temporal and frequency
+correlations jointly enhances performance significantly. Simulations
+reveal that the proposed approach provides faster convergence and
+smaller errors when compared to manifold based approaches. Therefore,
+our feedback approach with simple 2D-interpolation over the
+scalar parameters to optimize the precoders tracks precoders
+effectively.  Future work would involve tuning the adaptation rate of
+the scalar parameters and further reduction of quantization redundancy
+to reduce the feedback rates.
 

### 2020-03-04 paper/main.tex
Add changes

@@
 
+\setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
 \def\imwidth{0.65}
@@
 %Authors List
-
  \author{\authorblockN{Mohit Madan, Agrim Gupta, Kumar Appaiah\\}
@@
 % }}
+\vspace{-2.3\baselineskip}
 }
@@
 section. To construct the rest of the subcarriers, we will use
-precoder parameters from both the past and future and time instances
-to do the interpolation (using backtracking to ensure causality).
-By future time instances it means that the precoder estimation is
-improved at a later time instant i.e. after receiving the feedback of the next set of
-subcarriers. In
-Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
-neighbouring subcarriers using bilinear interpolation (equivalent to
-linear interpolation in two dimensions). These interpolated values are
-then used to reconstruct the precoding matrices all together by applying
-the reverse transformation of breaking the precoder into independent parameters as
-given in Equation~\ref{eq:decompose}. Thus,
-the approach here is able to exploit the time and frequency
-correlation jointly to enhance performance.
+precoder parameters from both the ``past'' and ``future'' and time
+instances to do the interpolation (using backtracking to ensure
+causality). By future time instances it means that the precoder
+estimation is improved at a later time instant i.e. after receiving
+the feedback of the next set of subcarriers. Note that the precoder
+used for transmission is causally obtained, while the ``future''
+correction of past precoders merely improves the prediction
+accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
+using the neighbouring subcarriers using bilinear interpolation
+(equivalent to linear interpolation in two dimensions). These
+interpolated values are then used to reconstruct the precoding
+matrices all together by applying the reverse transformation of
+breaking the precoder into independent parameters as given in
+Equation~\ref{eq:decompose}. Thus, the approach here is able to
+exploit the time and frequency correlation jointly to enhance
+performance.
 
@@
 % Using the idea of joint interpolation given in the other paper.
-
+\vspace{-1\baselineskip}
 \section{Conclusion}
@@
 Givens rotation and Householder transformation parameters of precoders
-minimizes receiver feedback to the transmitter while retaining
-effective precoder adaptation. Exploiting the temporal and frequency
+minimizes receiver feedback to the transmitter while permitting
+effective precoder adaptation. Exploiting temporal and frequency
 correlations jointly enhances performance significantly. Simulations
-reveal that the proposed approach provides faster convergence and
-smaller errors when compared to manifold based approaches. Therefore,
-our feedback approach with simple 2D-interpolation over the
-scalar parameters to optimize the precoders tracks precoders
-effectively.  Future work would involve tuning the adaptation rate of
-the scalar parameters and further reduction of quantization redundancy
-to reduce the feedback rates.
+reveal that the proposed approach converges faster and has smaller
+errors when compared to manifold based approaches. Our feedback
+approach, with simple 2D-interpolation over the scalar parameters,
+tracks precoders effectively. Future work would involve tuning the
+adaptation rates of scalar parameters and further reduction of
+redundancies to improve quantization.
 

### 2020-03-04 paper/response_letter.tex
Add changes

@@
 divided into $2^n$ optimal intervals using Lloyd's
-quantization [17].}''\\
-
-In addition, Section III.C is also altered as follows to make it more
-clear:
-
-\emph{``By
-  future time instances it means that the precoder estimation is
-  improved at a later time instant i.e.  after receiving the feedback
-  of the next set of subcarriers.''}
+quantization [17].}''
 
@@
 
-\textbf{Response:} AGRIM TODO
+\textbf{Response:} We thank the reviewer for pointing this out this
+missing piece of information. We
+wish to clarify that the precoder used for transmission is used for
+transmission. However, the past precoder estimates are improved using
+the new information, and this provides slight improvement when
+performing prediction in the further steps. This was not stated
+clearly in the previous manuscript. We have now altered the text in the subsection III.C as follows:
 
-We have altered the text in the subsection III.C as follows:
-``By future time instances it means that the precoder estimation is improved
-at a later time instant i.e. after receiving the feedback of the next set of
-subcarriers.''
+\emph{To construct the rest of the subcarriers, we will use
+precoder parameters from both the ``past'' and ``future'' and time
+instances to do the interpolation (using backtracking to ensure
+causality). By future time instances it means that the precoder
+estimation is improved at a later time instant i.e. after receiving
+the feedback of the next set of subcarriers. Note that the precoder
+used for transmission is causally obtained, while the ``future''
+correction of past precoders merely improves the prediction
+accuracy.}
 
@@
 
-\textbf{6. The reference format is not consistent and there are several typos. The title of the manuscript is not in IEEE format.}
+\textbf{6. The reference format is not consistent and there are
+  several typos. The title of the manuscript is not in IEEE format.}
+
+Thank you for the comment. We have now improved the references and
+fixed the errors.
 \end{document}

### 2020-03-13 paper/response_letter.tex
make changes

@@
 \textbf{Response:} We thank the reviewer for pointing out this
-error. We have altered the text to include this:
+error. We have reworded Section III to clarify the parameterization and quantization
+in a suitable order. In addition, we have altered the text to include this:
 
@@
 
+In addition we have clarified the interpolation process by precisely stating
+how we are combining temporaly separated precoders as follows:
+
+\emph{To construct precoders for rest of the subcarriers, we will use
+precoder parameters from both the ``past'' and ``future'' time
+instances to do the interpolation (using backtracking to ensure
+causality). When using future time instances, the implication is that the precoder
+estimation is improved at a later time instant i.e. after receiving
+the feedback of the next set of subcarriers. Note that the precoder
+used for transmission is causally obtained, while the ``future''
+correction of past precoders merely improves the prediction
+accuracy.}
+
 \textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future

### 2020-03-13 paper/main.tex
make changes

@@
 \title{Scalar Feedback based Joint Time-Frequency Precoder
-  interpolation for  MIMO-OFDM Systems}
+  Interpolation for  MIMO-OFDM Systems}
 
@@
 An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders.
-A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{mondal2007quantization,schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
+A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
 On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
@@
 One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
+With smaller quantization error as compared to manifold based approaches, the feedback
+results in more accurate precoder tracking, along with simple 2D-interpolation over the
+scalar parameters to optimize the precoders
 
@@
 rotation angles for the $i$-th subcarrier at time instant $t$, as
-represented in Fig.~\ref{fig:adpm-fig}. It is known~\cite{4114278}
+represented in Fig.~\ref{fig:adpm-fig}. It is known 
 that $\phi_{k,l} \in (-\pi, \pi]$ for all k and l's is uniformly
@@
 the transmitter, we can construct the precoder at subcarrier indices
-$pk$ or $pk+q$, in keeping with the same notation as the previous
-section. To construct the rest of the subcarriers, we will use
-precoder parameters from both the ``past'' and ``future'' and time
+$pk$ or $pk+q$, keeping with the same notation as the previous
+section. To construct precoders for rest of the subcarriers, we will use
+precoder parameters from both the ``past'' and ``future'' time
 instances to do the interpolation (using backtracking to ensure
-causality). By future time instances it means that the precoder
+causality). When using future time instances, the implication is that the precoder
 estimation is improved at a later time instant i.e. after receiving

### 2020-03-19 paper/response_letter.tex
Add changes

@@
 
-\textbf{Response:} Thank you for pointing this out. The temporal
-correlation based precoder tracking discussed in [8] and [9] have
-considered about precoder tracking in time and frequency respectively,
-albeit separately, without any joint interpolation.  We have
-approached this differently by performing prediction in the time
-domain and using this for joint interpolation on the time-freqeuncy
-grid. This helps build a framework for reducing the total bit
-bandwidth by a significant amount while maintaining similar performance
-in terms of BER.  We have altered the text in the introduction to make
-this point more clear,  as follows:
+Thank you for the feedback. The interpolation approach that we have
+proposed is simple and ends up becoming more accurate over time when
+compared to the manifold based interpolation that uses complex
+interpolation methods in the manifold space. It is also easier to
+track the adaptation of the scalar parameters using few bits. As
+observed in Fig. 2, the quantization error obtained when using the
+parameter based approach taken in this manuscript is much smaller in
+that obtained using the conventional codebooks on the Stiefel
+manifold, as discussed in [2]. This results in higher accuracy as
+compared to manifolds with similar number of bits.
+
+To emphasize these points, we have made the following addition:
+
+\emph{With smaller
+quantization error as compared to manifold based approaches, the
+feedback results in more accurate precoder tracking, along with simple
+2D-interpolation over the scalar parameters to optimize the precoders.}
+
+In addition, the temporal correlation based precoder tracking
+discussed in [8] and [9] have considered about precoder tracking in
+time and frequency respectively, albeit separately, without any joint
+interpolation.  We have approached this differently by performing
+prediction in the time domain and using this for joint interpolation
+on the time-freqeuncy grid. This helps build a framework for reducing
+the total bit bandwidth by a significant amount while maintaining
+similar performance in terms of BER.  We have altered the text in the
+introduction to make this point more clear, as follows:
 
@@
   interpolation on the time frequency grid.}''
+
+This is in addition to various changes in other sections of the paper
+that further emphasize the advantages of our approach.
 
@@
 
-\textbf{2. ``The quantization approach described above... '' Actually,
+\textbf{2. `` The quantization approach described above... '' Actually,
   parameterization is described without presenting the quantizing
@@
 takes is divided into $2^n$ optimal intervals using Lloyd's
-quantization [17].}
+quantization [16].}
 
-\textbf{3. "While tracking the $\phi_i$ parameters, is must be noted... "}
+\textbf{3. ``While tracking the $\phi_i$ parameters, is must be noted...''}
 
-\textbf{Response:} It has been corrected to \emph{``While tracking the $\phi_i$ parameters, it must be noted... ''}.
+\textbf{Response:} It has been corrected to \emph{``While tracking the $\phi_{k,l}$ parameters, it must be noted...''}.
 
-\textbf{4. "we will use precoder parameters from both the past and future and
-time instances to do the interpolation... "}
+\textbf{4. ``we will use precoder parameters from both the past and future and
+time instances to do the interpolation...''}
 
 \textbf{Response:} Thank you. The typo has now been corrected to read
-``we will use precoder parameters from both the past and future time instances to do the interpolation... ''
+``we will use precoder parameters from both the past and future time instances to do the interpolation...''
 
-\textbf{5. "the parameters that determine the precoder is ...  = 10, ... "}
+\textbf{5. ``the parameters that determine the precoder is ...  = 10, ...''}
 
@@
 divided into $2^n$ optimal intervals using Lloyd's
-quantization [17].}''
+quantization [16].}''
 

### 2020-03-19 paper/main.tex
Add changes

@@
 $\mbox{diag}\big( {\bf 1}_{k-1}, e^{j\phi_{k,k}},\ldots,
-e^{j\phi_{k,N_T }} \big)$ where  dimension $N_T \times N_T$. ${\bf 1}_{k-1}$ represents $k-1$ ones,
+e^{j\phi_{k,N_T }} \big)$ with  dimension $N_T \times N_T$. ${\bf 1}_{k-1}$ represents $k-1$ ones,
 and the $N_T \times N_R$ matrix $\tilde{\bI}$ =
@@
 performance with only very minor modifications to the system.
+
+\vspace{-0.5\baselineskip}
 \subsection{Differential quantization and channel tracking}
@@
 in phase of magnitude close to $2\pi$ have been eliminated).
+
+\vspace{-1\baselineskip}
 \subsection{Joint Time Frequency Interpolation}
@@
 against the ideal precoder for 100 channel evolutions in each run and
-averaged over 10 channel realizations (translating to averaging over $250,000$ bits).
-The error properties obtained by simulating transmission over these varying channels is sufficient to obtain good average performance properties.
+averaged over 10 channel realizations (translating to averaging over $125,000$ bits for 64-point DFT and
+$2\times 10^6$ bits for the 1024-point DFT case). 
+The error properties obtained by simulating transmission over these 
+varying channels is sufficient to obtain good average performance properties. 
 Simulations are performed for

### 2020-03-19 paper/main.tex
add changes, spelling correction

@@
 A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}.
-On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
+On the other hand, there are numerous works which utilize scalar parameterization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}.
 While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
 involve operations over higher dimensional manifolds.
-The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
+The scalar parameterization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
 Such methods have also been implemented in 802.11 ac and ax standards
@@
 between these two bodies of work, thereby making it difficult to
-quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds.
+quantify from past work whether performance gains/losses can accrue by operating over scalar parameterization instead of manifolds.
 In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques.
-One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
+One of the key results of our work is that the proposed scalar parameterization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
 With smaller quantization error as compared to manifold based approaches, the feedback
@@
 
-Parametrization for precoder feedback have utilized adaptive delta
+Past work in utilizing scalar parameterization for precoder feedback have utilized adaptive delta
 modulation effectively track precoding matrices both in the
@@
 Secondly, instead of just comparing similar algorithms in the chosen body of work (viz.
-scalar parametrization), as has been the case with past work, we also compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.
+scalar parameterization), as has been the case with past work, we also compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.
 
@@
 \section{Scalar joint time-frequency scheme}
-\subsection{Scalar Parametrization}
+\subsection{Scalar Parameterization}
 \label{givens}
@@
 convergence. The Stiefel manifold based approach uses nested codebooks
-that may be suboptimal when compared to the parametrization using
+that may be suboptimal when compared to the parameterization using
 angles.

### 2020-03-26 paper/response_letter.tex
add changes, spelling correction
Adding edits

@@
 
-Thank you for the feedback. The interpolation approach that we have
+Thank you for the feedback.
+
+It has been recently shown that for asymmetric MIMO channels ($N_t\neq N_R$), precoders that achieve channel capacity via waterfilling reside on non-trivial manifolds like Stiefel and Flag manifolds (Cite Pitavals papers here), which are harder to analyze and characterize than the traditionally considered Grassmannian and Unitary manifolds.
+A careful study of the manifold methods presented to interpolate and predictively quantize reveals that they also can be interpreted as straightforward 2-D filters using manifold operations like lifting/retraction maps.
+However, most of the times these operations themselves are suboptimal, like [2] uses quasi-geodesics instead of true geodesics.
+This leads to a performance degradation.
+However, the interpolation/prediction methods presented in our manuscript doesn't suffer from these mathematical artefacts and hence outperforms the manifold based methods.
+To stress explicitly, although both the manifold and the Girvans' rotation parametrization method may be equivalent mathematically, the latter offers more tractability in applying algorithms over them, and gives an hint as to why the latter was also chosen in standards to perform quantization instead of the former.
+
+The interpolation approach that we have
 proposed is simple and ends up becoming more accurate over time when
@@
 
-\textbf{2. `` The quantization approach described above... '' Actually,
+\textbf{2. ``The quantization approach described above...'' Actually,
   parameterization is described without presenting the quantizing
@@
 In addition we have clarified the interpolation process by precisely stating
-how we are combining temporaly separated precoders as follows:
+how we are combining temporally separated precoders as follows:
 
@@
 
-\textbf{Response:} We thank the reviewer for pointing this out this
-missing piece of information. We
-wish to clarify that the precoder used for transmission is used for
-transmission. However, the past precoder estimates are improved using
-the new information, and this provides slight improvement when
-performing prediction in the further steps. This was not stated
-clearly in the previous manuscript. We have now altered the text in the subsection III.C as follows:
+\textbf{Response:} 
+We thank the reviewer for raising this subtle point.
+We divide the problem we consider into two slightly uncoupled aspects.
+The receiver estimates the channel matrix and obtains the accurate unquantized precoders.
+First aspect of the problem considers effective quantization of the precoder by quantizing just the difference between the past and the current precoder, as is typically done in delta quantization approaches.
+For this aspect we don not use future parameters.
+
+Now, the receiver feeds back these quantized estimates back to the transmitter.
+The transmitter has to estimate the precoder at the non-fed back time-frequency points, for which it utilizes wait and decode principle to get future precoder data to efficiently interpolate over the previous time precoders.
+The window size used to get future parameters is low enough to be within the channel coherence time, to make this method physically feasible, however, we would like to highlight that this does add an extra delay to the system, at the expense of improved interpolation error.
 
@@
 used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the prediction
+correction of past precoders merely improves the prediction (Do you mean interpolation here?)
 accuracy.}

### 2020-03-27 paper/response_letter.tex
Minor fixes

@@
 
-It has been recently shown that for asymmetric MIMO channels ($N_t\neq N_R$), precoders that achieve channel capacity via waterfilling reside on non-trivial manifolds like Stiefel and Flag manifolds (Cite Pitavals papers here), which are harder to analyze and characterize than the traditionally considered Grassmannian and Unitary manifolds.
-A careful study of the manifold methods presented to interpolate and predictively quantize reveals that they also can be interpreted as straightforward 2-D filters using manifold operations like lifting/retraction maps.
-However, most of the times these operations themselves are suboptimal, like [2] uses quasi-geodesics instead of true geodesics.
-This leads to a performance degradation.
-However, the interpolation/prediction methods presented in our manuscript doesn't suffer from these mathematical artefacts and hence outperforms the manifold based methods.
-To stress explicitly, although both the manifold and the Girvans' rotation parametrization method may be equivalent mathematically, the latter offers more tractability in applying algorithms over them, and gives an hint as to why the latter was also chosen in standards to perform quantization instead of the former.
+It has recently been shown that, for asymmetric MIMO channels
+($N_t\neq N_R$), precoders that achieve channel capacity via
+waterfilling reside on non-trivial manifolds such as the Stiefel and
+Flag manifolds (see references [12-13]) that are harder to analyze and
+characterize than the traditionally considered Grassmannian and
+Unitary manifolds. A careful study of the manifold methods presented
+here to interpolate and predictively quantize reveals that they also
+can be interpreted as 2-D filters using manifold operations like
+lifting/retraction maps. However, most of these operations themselves
+are suboptimal. For instance, the approach in [2] uses quasi-geodesics
+instead of true geodesics. This leads to a performance
+degradation. However, the interpolation/prediction methods presented
+in our manuscript does not suffer from these mathematical artifacts,
+and hence outperforms the manifold based methods. To stress this
+explicitly, although both the manifold and the Givens' rotation
+parametrization method may be close (or equivalent) mathematically,
+the latter offers more tractability in applying interpolation
+algorithms, and gives hints as to why the latter was also chosen in
+standards to perform quantization instead of the former.
 
-The interpolation approach that we have
-proposed is simple and ends up becoming more accurate over time when
-compared to the manifold based interpolation that uses complex
-interpolation methods in the manifold space. It is also easier to
-track the adaptation of the scalar parameters using few bits. As
-observed in Fig. 2, the quantization error obtained when using the
-parameter based approach taken in this manuscript is much smaller in
-that obtained using the conventional codebooks on the Stiefel
-manifold, as discussed in [2]. This results in higher accuracy as
-compared to manifolds with similar number of bits.
+The interpolation approach that we have proposed is simple and ends up
+becoming more accurate over time when compared to the manifold based
+interpolation that uses complex interpolation methods in the manifold
+space. It is also easier to track the adaptation of the scalar
+parameters using few bits. As observed in Fig. 2, the quantization
+error obtained when using the parameter based approach taken in this
+manuscript is much smaller in that obtained using the conventional
+codebooks on the Stiefel manifold, as discussed in [2]. This results
+in higher accuracy as compared to manifolds with similar number of
+bits.
 
@@
 
-\textbf{Response:} 
-We thank the reviewer for raising this subtle point.
-We divide the problem we consider into two slightly uncoupled aspects.
-The receiver estimates the channel matrix and obtains the accurate unquantized precoders.
-First aspect of the problem considers effective quantization of the precoder by quantizing just the difference between the past and the current precoder, as is typically done in delta quantization approaches.
-For this aspect we don not use future parameters.
+\textbf{Response:} We thank the reviewer for raising this subtle
+point. We divide the problem we consider into two slightly uncoupled
+aspects. The receiver estimates the channel matrix and obtains the
+accurate unquantized precoders. The first aspect of the problem
+considers effective quantization of the precoder by quantizing just
+the difference between the past and the current precoder, as is
+typically done in delta quantization approaches. For this aspect we
+don not use future parameters.
 
-Now, the receiver feeds back these quantized estimates back to the transmitter.
-The transmitter has to estimate the precoder at the non-fed back time-frequency points, for which it utilizes wait and decode principle to get future precoder data to efficiently interpolate over the previous time precoders.
-The window size used to get future parameters is low enough to be within the channel coherence time, to make this method physically feasible, however, we would like to highlight that this does add an extra delay to the system, at the expense of improved interpolation error.
+Now, the receiver feeds back these quantized estimates back to the
+transmitter. The transmitter has to estimate the precoder at the
+non-fed back time-frequency points, for which it utilizes wait and
+decode principle to get future precoder data to efficiently
+interpolate over the previous time precoders. The window size used to
+get future parameters is low enough to be within the channel coherence
+time, to make this method physically feasible, however, we would like
+to highlight that this does add an extra delay to the system, at the
+expense of improved interpolation error.
 
@@
 used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the prediction (Do you mean interpolation here?)
+correction of past precoders merely improves the interpolation
 accuracy.}

### 2020-03-27 paper/main.tex
Minor fixes

@@
 \setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
-\def\imwidth{0.65}
+\def\imwidth{0.625}
 %package list
@@
 used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the prediction
+correction of past precoders merely improves the interpolation
 accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
@@
 
-Fig.~\ref{fig:error_decay} shows the rate at which error diminishes
-with time for the proposed approach and direct quantization on the
-Stiefel manifold as described in~\cite{Gupt1905:Predictive}. Since the
-adaptation of independent scalar parameters causes the error to decay
-rapidly, the error is lower and the convergence is faster, even though
-the starting error is higher in the case of the scalar quantization
-approach. The independence of the parameters enables fast
-convergence. The Stiefel manifold based approach uses nested codebooks
-that may be suboptimal when compared to the parameterization using
-angles.
+Fig.~\ref{fig:error_decay} shows the error decay rate with time for
+the proposed approach and direct quantization on the Stiefel manifold
+as described in~\cite{Gupt1905:Predictive}. Since the adaptation of
+independent scalar parameters causes the error to decay rapidly, the
+error is lower and the convergence is faster, even for larger starting
+errors in the scalar quantization approach. The independence of the
+parameters enables fast convergence. The Stiefel manifold based
+approach with nested codebooks may thus be suboptimal when compared to
+the parameterization using angles.
 

### 2020-03-27 paper/main.tex
minor fix

@@
 antennas, the channel for the $i$-th subcarrier at time instant $t$
-can be modelled as a flat-fading MIMO channel. This flat-fading MIMO
+can be modeled as a flat-fading MIMO channel. This flat-fading MIMO
 channel can be written as
@@
 $pk$ or $pk+q$, keeping with the same notation as the previous
-section. To construct precoders for rest of the subcarriers, we will use
+section. To construct precoders for the rest of the subcarriers, we will use
 precoder parameters from both the ``past'' and ``future'' time
@@
 precoder is not available at the transmitter, we use 2 bits for
-initializing each parameter for effective representation of the
+initializing each parameter for the effective representation of the
 precoder. Therefore, the number of bits for each subcarrier =

### 2020-03-28 paper/cover_letter.tex
Add editor name
Update cover letter

@@
 \begin{letter}{To\\
-Prof. TODO\\
+Prof. Wei Zhang\\
 Editor-in-Chief\\
@@
 
-\opening{Dear Prof. TODO,}
+\opening{Dear Prof. Wei Zhang,}
 
@@
   interpolation for  MIMO-OFDM Systems'' for potential publication in
-  IEEE Wireless Communication Letters.}
+  IEEE Wireless Communication Letters (old reference number  WCL2020-0046).}
 
-Please find enclosed our manuscript titled ``Scalar Feedback based Joint Time-Frequency Precoder
-  interpolation for  MIMO-OFDM Systems'' for review and potential
-  publication in IEEE Wireless Communication Letters. The co-authors of this manuscripts are
+Thank you for arranging the prompt reviews of our original manuscript
+titled ``Scalar Feedback based Joint Time-Frequency Precoder
+interpolation for MIMO-OFDM Systems'' (WCL2020-0046) submitted to the
+IEEE Wireless Communications Letters, and providing us an opportunity
+to resubmit this manuscript with modifications addressing the issues
+raised during the reviews. I would like to thank the reviewers for a
+careful reading of the manuscript, as well as for the insightful
+points that they have raised regarding various technical aspects of
+the manuscript. Based on these reviews, we have made extensive
+additions and revisions to our submission to address the reviewers
+concerns. In particular, we have addressed the key issues pertaining
+to the novelty of the method and the use of causal information for
+interpolation of the precoders. Please find attached our comments and
+actions taken based on the reviews, along with a document indicative
+of all the changes made.
+
+The co-authors of this manuscript are
 \begin{enumerate}

### 2020-04-26 paper/main.tex
adding changes

@@
 \setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
-\def\imwidth{0.625}
+\def\imwidth{0.7}
 %package list
@@
 
-\vspace{-0.5\baselineskip}
+% \vspace{-0.5\baselineskip}
 \subsection{Differential quantization and channel tracking}
@@
 
-\vspace{-1\baselineskip}
+% \vspace{-1\baselineskip}
 \subsection{Joint Time Frequency Interpolation}
@@
 $pk$ or $pk+q$, keeping with the same notation as the previous
-section. To construct precoders for the rest of the subcarriers, we will use
-precoder parameters from both the ``past'' and ``future'' time
-instances to do the interpolation (using backtracking to ensure
-causality). When using future time instances, the implication is that the precoder
-estimation is improved at a later time instant i.e. after receiving
-the feedback of the next set of subcarriers. Note that the precoder
-used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the interpolation
-accuracy. In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
+section. 
+% To construct precoders for the rest of the subcarriers, we will use
+% precoder parameters from both the ``past'' and ``future'' time
+% instances to do the interpolation (using backtracking to ensure
+% causality). When using future time instances, the implication is that the precoder
+% estimation is improved at a later time instant i.e. after receiving
+% the feedback of the next set of subcarriers. Note that the precoder
+% used for transmission is causally obtained, while the ``future''
+% correction of past precoders merely improves the interpolation
+% accuracy. 
+In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
 using the neighbouring subcarriers using bilinear interpolation
@@
 breaking the precoder into independent parameters as given in
-Equation~\ref{eq:decompose}. Thus, the approach here is able to
-exploit the time and frequency correlation jointly to enhance
-performance.
+Equation~\ref{eq:decompose}. Inorder to improve the precoder estimates 
+and interpolation accuracy we walkback to improve the past estimates by interpolating
+upon the current estimates. Thus, the approach here is able to exploit the time and frequency correlation jointly to enhance performance.
+% To construct the precoders for the rest of the subcarriers, we will
+% use the  parameters from the previously determined subcarriers.
+%
 
@@
 % Using the idea of joint interpolation given in the other paper.
-\vspace{-1\baselineskip}
+% \vspace{-1\baselineskip}
 \section{Conclusion}
@@
 
-\vspace{-10pt}
+% \vspace{-10pt}
 % \section{Acknowledgment}

### 2020-04-26 paper/main.tex
Minor fix

@@
 % accuracy. 
-In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
-using the neighbouring subcarriers using bilinear interpolation
-(equivalent to linear interpolation in two dimensions). These
-interpolated values are then used to reconstruct the precoding
-matrices all together by applying the reverse transformation of
-breaking the precoder into independent parameters as given in
-Equation~\ref{eq:decompose}. Inorder to improve the precoder estimates 
-and interpolation accuracy we walkback to improve the past estimates by interpolating
-upon the current estimates. Thus, the approach here is able to exploit the time and frequency correlation jointly to enhance performance.
+In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
+neighbouring subcarriers using bilinear interpolation (equivalent to
+linear interpolation in two dimensions). These interpolated values are
+then used to reconstruct the precoding matrices all together by
+applying the reverse transformation of breaking the precoder into
+independent parameters as given in Equation~\ref{eq:decompose}. To
+improve the precoder estimates and interpolation accuracy, we ``walk
+back'' to improve the past estimates by interpolating using current
+estimates (although the current estimates only use data
+causally). Thus, the approach here is able to exploit the time and
+frequency correlation jointly to enhance performance.
 % To construct the precoders for the rest of the subcarriers, we will

### 2020-04-26 paper/main.tex
modifying figure 1

--- 2020-04-26 paper/main.tex
@@
 \begin{center}
-\includegraphics[width=0.7\columnwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.7\columnwidth]{images/new-adpm1.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers that need to be
@@
 In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
-using the neighbouring subcarriers using bilinear interpolation
+using the neighbouring subcarriers (red arrows) using bilinear interpolation
 (equivalent to linear interpolation in two dimensions). These
@@
 breaking the precoder into independent parameters as given in
-Equation~\ref{eq:decompose}. Inorder to improve the precoder estimates 
-and interpolation accuracy we walkback to improve the past estimates by interpolating
+Equation~\ref{eq:decompose}. Inorder to improve the precoder estimates we walkback to improve the past estimates (black arrows) by interpolating
 upon the current estimates. Thus, the approach here is able to exploit the time and frequency correlation jointly to enhance performance.

### 2020-04-30 paper/response_letter_2.tex
Add cover letter, changes

--- 2020-03-19 paper/response_letter.tex
@@
 Subject: \textbf{Revised submission of manuscript title ``Scalar
-  Feedback based Joint Time-Frequency Precoder interpolation for  MIMO-OFDM Systems'' (Paper Number: WCL2020-0046).}
+  Feedback based Joint Time-Frequency Precoder interpolation for  MIMO-OFDM Systems'' (Paper Number: WCL2020-0511).}
 
@@
 technical aspects of the manuscript. Based on these reviews, we have
-made extensive additions and revisions to our submission to address
-the reviewers concerns. In particular, we have addressed the key issue
-raised by reviewer 2 as well as the Editor, regarding the important
-feature of our method that provides us better performance using scalar
-quantization. We have also addressed all typos to the best of our
-ability. We hope that you will reconsider the manuscript for
-publication in the IEEE Wireless Communications Letters. Please find
-attached our comments and actions taken based on the reviews, along
-with a document indicative of all the changes made.
+made revisions to our submission to address the reviewers concerns. In
+particular, we have addressed the key issue raised by reviewer 2 on
+the aspect of improving past precoder estimates, and have clarified
+the persisting confusion. We hope that you will reconsider the
+manuscript for publication in the IEEE Wireless Communications
+Letters. Please find attached our comments and actions taken based on
+the reviews, along with a document indicative of all the changes made.
 
@@
 
-\textbf{Editor comments}\\
-
-\textbf{There are several mistakes and typos in the manuscript, as
-  also pointed out by both reviewers. Moreover, it is mentioned that
-  the temporal correlation based and frequency correlation based
-  precoder tracking have been discussed in [8] and [9]. Thus the
-  differences between the proposed method and the methods in [8] and
-  [9] should be explained and compared, so as to make the
-  contributions more clear.}
-
-Thank you for the feedback. The interpolation approach that we have
-proposed is simple and ends up becoming more accurate over time when
-compared to the manifold based interpolation that uses complex
-interpolation methods in the manifold space. It is also easier to
-track the adaptation of the scalar parameters using few bits. As
-observed in Fig. 2, the quantization error obtained when using the
-parameter based approach taken in this manuscript is much smaller in
-that obtained using the conventional codebooks on the Stiefel
-manifold, as discussed in [2]. This results in higher accuracy as
-compared to manifolds with similar number of bits.
-
-To emphasize these points, we have made the following addition:
-
-\emph{With smaller
-quantization error as compared to manifold based approaches, the
-feedback results in more accurate precoder tracking, along with simple
-2D-interpolation over the scalar parameters to optimize the precoders.}
-
-In addition, the temporal correlation based precoder tracking
-discussed in [8] and [9] have considered about precoder tracking in
-time and frequency respectively, albeit separately, without any joint
-interpolation.  We have approached this differently by performing
-prediction in the time domain and using this for joint interpolation
-on the time-freqeuncy grid. This helps build a framework for reducing
-the total bit bandwidth by a significant amount while maintaining
-similar performance in terms of BER.  We have altered the text in the
-introduction to make this point more clear, as follows:
-
-``\emph{In our work, we perform prediction in time domain, and use this for joint
-  interpolation on the time frequency grid.}''
-
-This is in addition to various changes in other sections of the paper
-that further emphasize the advantages of our approach.
-
 \textbf{Reviewer 1}\\
 
-\textbf{This paper presents an interesting case study of quantized
-  precoder design with limited feedback channels. The results show
-  that quantization base on scalar parameterization is more effective
-  than direct quantization on manifolds. There are, however, two
-  issues that need to be addressed before publication.}
+\textbf{My concerns have been addressed not very thoroughly, but
+  acceptably. The paper is not good at theorizing the issues at hands,
+  but nevertheless reports an important case
++of assembling existing techniques to achieve better performance than
+theory-driven approaches do.
+}
 
-\textbf{1. It is not clear whether the feedback bandwidth budgets are
-equivalent for methods in comparison. The comparison must be fair.}
-
-\textbf{Response:} Thank you for pointing this out. We have used an
-average of 5 bits in our method. While the number of feedback bits
-used in [2] is 6 bits per OFDM frame, the approach that we suggest
-uses a smaller number bits on average while still achieving similar or
-better BER performance. We have now added a brief indication of this
-in the text as follows
-
-\emph{Since the bit budget is small, if the channel varies
-sufficiently slowly, the $\theta_{k,l}$ and the $\phi_{k,l}$ parameters
-are fed back alternately, while retaining the unsent parameters as
-their previous values, as shown in
-Fig. 1. This brings down the bit budget to an average of 5 bits for
-$\bf{\Theta}$ and 5 bits for $\bf{\Phi}$ over alternate feedback
-instances, thereby resulting in an average feedback budget of 5 bits,
-as opposed to the 6 bits of feedback used
-in [2].}
-
-\textbf{ The so-called time-frequency interpolation method is just a
-  straightforward two-dim filter. So I assume the that the more
-  important factor for performance improvement would be the use of an
-  efficient parameterization for the precoder matrix, which comes from
-  ref. [8]. I would expect this paper shed more light on why this
-  particular parameterization is more efficient in characterizing
-  precoder matrices than the description based on manifolds.  Would
-  the authors make more explanations/comments on this?}
-
-\textbf{Response:}
-
-Thank you for the feedback. The interpolation approach that we have
-proposed is simple and ends up becoming more accurate over time when
-compared to the manifold based interpolation that uses complex
-interpolation methods in the manifold space. It is also easier to
-track the adaptation of the scalar parameters using few bits. As
-observed in Fig. 2, the quantization error obtained when using the
-parameter based approach taken in this manuscript is much smaller in
-that obtained using the conventional codebooks on the Stiefel
-manifold, as discussed in [2]. This results in higher accuracy as
-compared to manifolds with similar number of bits.
-
-To emphasize these points, we have made the following addition:
-
-\emph{With smaller
-quantization error as compared to manifold based approaches, the
-feedback results in more accurate precoder tracking, along with simple
-2D-interpolation over the scalar parameters to optimize the precoders.}
-
-\textbf{There are some typos/errors to be fixed:}
-
-\textbf{1. Work/works got mixed up for many times.}
-
-Thank you for this comment. We have fixed all instances of this error.
-
-\textbf{2. `` The quantization approach described above... '' Actually,
-  parameterization is described without presenting the quantizing
-  method.}
-
-\textbf{Response:} We thank the reviewer for pointing out this
-error. We have reworded Section III to clarify the parameterization and quantization
-in a suitable order. In addition, we have altered the text to include this:
-
-\emph{The distribution of $\theta_{k,l}$ is used for
-quantizing $\bf{\Theta}$ during initialization to reduce the
-quantization error.  If $n$ bits are allocated to quantize a
-parameter for the first time, the range of values that each parameter
-takes is divided into $2^n$ optimal intervals using Lloyd's
-quantization [16].}
-
-\textbf{3. ``While tracking the $\phi_i$ parameters, is must be noted...''}
-
-\textbf{Response:} It has been corrected to \emph{``While tracking the $\phi_{k,l}$ parameters, it must be noted...''}.
-
-\textbf{4. ``we will use precoder parameters from both the past and future and
-time instances to do the interpolation...''}
-
-\textbf{Response:} Thank you. The typo has now been corrected to read
-``we will use precoder parameters from both the past and future time instances to do the interpolation...''
-
-\textbf{5. ``the parameters that determine the precoder is ...  = 10, ...''}
-
-\textbf{Response:} Thanks for pointing this out. It has been corrected to ``\emph{...the parameters that determine the precoder is
-$N_{R}(2N_{T} - N_R) = 12$, with five $\theta_{k,l}$s and and seven
-$\phi_{k,l}$s.}''
+We appreciate the reviewer's comments. We have strived to make things
+more clear with minor changes in the updated manuscript, while
+ensuring that we adhere to the page limit. We hope that these changes
+would present our idea better.
 
@@
 
-\textbf{1. There are several mistakes and typos in the manuscript. Here are some examples.}
+\textbf{It is still not clear how the transmitter can obtain the
+  future precoder parameters.
+  Could you use an example to explain it?}
 
-\textbf{(a) In the manuscript, boldface letters are used to denote column vectors and matrices and non-bold letters are used to denote scalars.
-Therefore, the subscripts of $\boldsymbol{\eta_{i,t}}$, $\boldsymbol{\Theta_{i,t}}$ and $\boldsymbol{\Phi_{i,t}}$ should be non-bold, and the covariance matrix of $\boldsymbol{\eta}$, should be boldface.}
+We appreciate the reviewer's concern. We feel that the word ``future''
+is incorrect, and have clarified in the manuscript that past estimates
+are improved as parameters come in. Here is an example.
 
-\textbf{Response:} Thanks for pointing out this typo. The subscripts
-of the symbols have been corrected wherever they have incurred. In addition,
-the covariance symbol have now been made bold.
-
-\textbf{(b) Because the matrix $\tilde{V_{i,t}}$ is $N_T$ by $N_R$ and
-  the vector $x_{i,t}$ is $N_d$ by 1, the product $V_{i,t}x_{i,t}$ in
-  (1) is not well defined}.
-
-\textbf{Response:} Thanks for pointing this out.  It now reads
-\emph{We consider the case where $N_R < N_T$, and the data vector size
-  $N_d = N_R$} which makes $x_{i,t}$ dimensions as $N_R$ which now
-makes it well defined.
-
-\textbf{(c) Because the matrix $I_{N_R}$ is $N_R$ by $N_R$ and the
-  matrix $0_{N_T , N_T - N_R }$ is $N_T$ by $N_T - N_R$, the matrix
-  $\tilde{I}$ = [$I_{N_R}, 0_{N_T, N_T - N_R }$] , in (2) is not
-  well-defined}
-
-\textbf{Response:} Thanks for pointing this out.  We have now fixed
-this to read:
-
-``\emph{and the $N_T \times N_R$ matrix
-  $\tilde{\textbf{I}}$ =
-  $\big[\textbf{I}_{N_R }, {\boldsymbol{0}}_{N_R ,N_T
-    -N_R}\big]^{T}$}''
-
-\textbf{(d) The right-hand side of (3) is a function of t but the left-hand side of (3) is independent of t.}
-
-\textbf{Response:} ``t'' has been changed to ``$N_T$''. This makes the
-right-hand side a square matrix of size $N_T \times N_T$ which is
-indeed the size of the matrix $G_{m-1,m}\big(\theta\big)$ on the
-left-hand side.
-
-\textbf{(e) Do $\theta_l$ and $\phi_k$ represent the same things as $\theta_{k,l}$ and $\phi_{k,l}$ , respectively.}
-
-\textbf{Response:} Thanks for pointing this out. Yes, $\theta_l$ and
-$\phi_k$ represent the same things as $\theta_{k,l}$ and
-$\phi_{k,l}$. Therefore, those references have now been corrected
-wherever they appeared. In addition, we have also specified the pdf of
-the $\theta_{k,l}$ clearly.
-
-\textbf{(f) It is mentioned in Subsection III.A that the number of
-  $\phi_k$ parameters is $N_R(2N_T - N_R -1 )/2$ while the number of
-  $\theta$ parameters is $N_R(2N_T -N_R + 1)/2$." However, in the
-  provided example with $N_T$ = 4 and $N_R$ = 2, there are 7
-  $\phi_{k,l}$ 's and 5 $\theta_{k,l}$'s.}
-
-\textbf{Response:} There was an error and we had interchanged the
-expression for the number of parameters. Therefore, we have now
-corrected it as follows:
-
-``\emph{The number of $\phi_{k,l}$ parameters is
-  $N_{R}(2N_{T} - N_{R}+1)/2$ while the number of $\theta_{k,l}$
-  parameters is $N_{R}(2N_{T} - N_{R}-1)/2$.}''
-
-\textbf{(g) It is mentioned in Subsection III.B that where $x_n$ is the unquantized value and $\tilde{x}_{n-1}$ is the unquantized (accurate) value." What does this mean?}
-
-\textbf{Response:} We have now corrected this to ``\emph{where $x_n$ is
-  the unquantized (accurate) value and $\tilde{x}_{n-1}$ is the
-  quantized value}''
-
-\textbf{(h) It is mentioned in Subsection III.B that where M is the increment used during the positive and negative moves." However, M is not an increment in (5).}
-
-\textbf{Response:} Thanks for pointing this out. $M$ is the scaling
-factor, and we have now corrected the reference to $M$ as follows:
-
-``\emph{...where M is the scaling factor used during the positive and...}''
-
-\textbf{(i) (Section IV) "The channel is simulated using the typical Jakes Model for for the COST207 channel power delay profiles." Here is a typo.}
-
-\textbf{Response:} It has now been corrected to ``\emph{The channel is simulated using the typical Jakes Model for the COST207 channel power delay profiles.}''
-
-\textbf{(j) (Section IV) the parameters that determine the precoder is $N_R (2N_T - 1) - N^{2}_R$ = 10, with five $\theta_l$s and and seven $\phi_i$s." Here is a typo because 10 $\neq$ 5 + 7.}
-
-\textbf{Response:} Thanks for pointing out this typo. The formula for
-calculating the precoder has now been corrected to ``\emph{
-  $N_{R}(2N_{T} - N_R) = 12$ , with five $\theta_{k,l}$s and and seven
-  $\phi_{k,l}$s.}''
-
-\textbf{(k) What does the sentence "The time based predictive approaches correspond to temporal prediction using the manifold as well as scalar parameter based approaches without utilizing the joint time frequency correlations." in Section IV mean?}
-
-\textbf{Response:} Thanks for pointing out this error. The time based
-predictive approach was meant to imply the independent time-frequency based
-approach. The text in the paper has been corrected as follows:
-
-``\emph{Independent time-frequency based approaches correspond to the
-  method without joint-time frequency interpolation i.e. with
-  prediction in time domain along with separate interpolation in the frequency
-  domain.}''
-
-\textbf{2. The quantization of $\phi_{k,l}$'s and $\theta_{k,l}$'s mentioned in Subsection III.A and the interpolation
-process mentioned in Subsection III.C should be explained more clearly.}
-
-\textbf{Response:}
-Thanks for pointing this out. The explanation of the quantization
-process has now been made more  clear by altering the text in III.A as
-follows:
-
-``\emph{If $n$ bits are allocated to quantize a parameter
-for the first time, the range of values that each parameter takes is
-divided into $2^n$ optimal intervals using Lloyd's
-quantization [16].}''
-
-In addition we have clarified the interpolation process by precisely stating
-how we are combining temporaly separated precoders as follows:
-
-\emph{To construct precoders for rest of the subcarriers, we will use
-precoder parameters from both the ``past'' and ``future'' time
-instances to do the interpolation (using backtracking to ensure
-causality). When using future time instances, the implication is that the precoder
-estimation is improved at a later time instant i.e. after receiving
-the feedback of the next set of subcarriers. Note that the precoder
-used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the prediction
-accuracy.}
-
-\textbf{3. As mentioned in Fig. 1 and Subsection III.C, the future
-  precoder parameters are required to conduct the interpolation. How
-  can we obtain the future precoder parameters?}
-
-\textbf{Response:} We thank the reviewer for pointing this out this
-missing piece of information. We
-wish to clarify that the precoder used for transmission is used for
-transmission. However, the past precoder estimates are improved using
-the new information, and this provides slight improvement when
-performing prediction in the further steps. This was not stated
-clearly in the previous manuscript. We have now altered the text in the subsection III.C as follows:
-
-\emph{To construct the rest of the subcarriers, we will use
-precoder parameters from both the ``past'' and ``future'' and time
-instances to do the interpolation (using backtracking to ensure
-causality). By future time instances it means that the precoder
-estimation is improved at a later time instant i.e. after receiving
-the feedback of the next set of subcarriers. Note that the precoder
-used for transmission is causally obtained, while the ``future''
-correction of past precoders merely improves the prediction
-accuracy.}
-
-\textbf{4. Only 10 channel realizations are used to evaluate the BER characteristics. Is it enough?}
-
-Thank you for this comment. Although we simulate over 10 realizations,
-the channel evolves beginning with each realization, resulting in
-sufficient averaging over various channel characteristics. Therefore,
-the results that we present represent simulation over a sufficiently
-large number of bits transmitted over enough channel variations. We
-have now altered the text in the simulation section to reflect this as
-follows:
-
-\emph{The BER achieved by the proposed predictive quantization based
-  precoder is compared against the ideal precoder for 100 channel
-  evolutions in each run and averaged over 10 channel realizations
-  (translating to averaging over $125,000$ bits for 64-point DFT and
-  $2\times 10^6$ bits for the 1024-point DFT case). The error
-  properties obtained by simulating transmission over these varying
-  channels is sufficient to obtain good average performance
-  properties.}
-
-\textbf{5.
-In the manuscript, the time-frequency correlation is used for precoder tracking.
-Because it is mentioned in Section I that the precoder tracking have been discussed in [8] (based on temporal correlation) and [9] (based on frequency correlation), the difference among the proposed method and the methods in [8] and [9] should be explained and compared in the manuscript.}
-
-Thank you for the feedback. The interpolation approach that we have
-proposed is simple and ends up becoming more accurate over time when
-compared to the manifold based interpolation that uses complex
-interpolation methods in the manifold space. It is also easier to
-track the adaptation of the scalar parameters using few bits. As
-observed in Fig. 2, the quantization error obtained when using the
-parameter based approach taken in this manuscript is much smaller in
-that obtained using the conventional codebooks on the Stiefel
-manifold, as discussed in [2]. This results in higher accuracy as
-compared to manifolds with similar number of bits.
-
-To emphasize these points, we have made the following addition:
-
-\emph{With smaller
-quantization error as compared to manifold based approaches, the
-feedback results in more accurate precoder tracking, along with simple
-2D-interpolation over the scalar parameters to optimize the precoders.}
-
-In addition, the temporal correlation based precoder tracking
-discussed in [8] and [9] have considered about precoder tracking in
-time and frequency respectively, albeit separately, without any joint
-interpolation.  We have approached this differently by performing
-prediction in the time domain and using this for joint interpolation
-on the time-freqeuncy grid. This helps build a framework for reducing
-the total bit bandwidth by a significant amount while maintaining
-similar performance in terms of BER.  We have altered the text in the
-introduction to make this point more clear, as follows:
-
-``\emph{In our work, we perform prediction in time domain, and use this for joint
-  interpolation on the time frequency grid.}''
-
-This is in addition to various changes in other sections of the paper
-that further emphasize the advantages of our approach.
-
-\textbf{6. The reference format is not consistent and there are
-  several typos. The title of the manuscript is not in IEEE format.}
-
-Thank you for the comment. We have now improved the references and
-fixed the errors.
+Consider the 10th OFDM frame. All precoder parameters are calculated
+using the parameters obtained at the 10th, 9th, 8th ... frames to fill
+in the missing precoders. While at this process, the information
+obtained during the 10th precoder can improve the precoder estimates
+for the 9th, 8th etc. precoders. Naturally, these ``past'' precoder
+parameters cannot be used directly for precoding, but the reduced
+error that they have can improve the prediction and interpolation for
+the precoders in the 11th frame. It is in this spirit that we ``walk
+back'' to improve past estimates so that their contribution to future
+predictions would possess less error. We have now clarified this in
+the manuscript as well.
 \end{document}

### 2020-04-30 paper/cover_letter.tex
Add cover letter, changes

@@
 
-Subject: \textbf{Submission of manuscript title ``Scalar Feedback based Joint Time-Frequency Precoder
+Subject: \textbf{Submission of updated manuscript titled ``Scalar Feedback based Joint Time-Frequency Precoder
   interpolation for  MIMO-OFDM Systems'' for potential publication in
-  IEEE Wireless Communication Letters (old reference number  WCL2020-0046).}
+  IEEE Wireless Communication Letters (old reference number  WCL2020-0511).}
 
@@
 titled ``Scalar Feedback based Joint Time-Frequency Precoder
-interpolation for MIMO-OFDM Systems'' (WCL2020-0046) submitted to the
+interpolation for MIMO-OFDM Systems'' (WCL2020-0511) submitted to the
 IEEE Wireless Communications Letters, and providing us an opportunity
@@
 points that they have raised regarding various technical aspects of
-the manuscript. Based on these reviews, we have made extensive
-additions and revisions to our submission to address the reviewers
-concerns. In particular, we have addressed the key issues pertaining
-to the novelty of the method and the use of causal information for
-interpolation of the precoders. Please find attached our comments and
-actions taken based on the reviews, along with a document indicative
-of all the changes made.
+the manuscript. Based on these reviews, we have madeadditions and
+revisions to our submission to address the reviewers concerns. In
+particular, we have clarified on the issue of improving past precoder
+estimate, as mentioned by Reviwer 2. Please find attached our comments
+and actions taken based on the reviews, along with a document
+indicative of all the changes made.
 

### 2020-04-30 paper/main.tex
minor image size fix
Minor changes

--- 2020-04-26 paper/main.tex
@@
 \setlength{\textfloatsep}{0.7\baselineskip plus 0.2\baselineskip minus 0.5\baselineskip}
-\def\imwidth{0.7}
+\def\imwidth{0.62}
 %package list
@@
 \begin{center}
-\includegraphics[width=0.7\columnwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.68\columnwidth]{images/new-adpm1.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers that need to be
-  quantized. We see that the subcarrier indicated by
-  red arrows has a precoder whose parameters are determined jointly
-  using the neighbouring parameters, both in time and frequency.}
+  quantized. Subcarriers indicated by red arrows have precoders whose
+  parameters are determined using the neighbouring parameters, both in
+  time and frequency. The black arrow indicates refinement of past
+  estimates.}
 \end{center}
@@
 % accuracy.
-In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated
-using the neighbouring subcarriers using bilinear interpolation
-(equivalent to linear interpolation in two dimensions). These
-interpolated values are then used to reconstruct the precoding
-matrices all together by applying the reverse transformation of
-breaking the precoder into independent parameters as given in
-Equation~\ref{eq:decompose}. Inorder to improve the precoder estimates 
-and interpolation accuracy we walkback to improve the past estimates by interpolating
-upon the current estimates. Thus, the approach here is able to exploit the time and frequency correlation jointly to enhance performance.
+
+In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
+neighbouring subcarriers using bilinear interpolation (equivalent to
+linear interpolation in two dimensions). These interpolated values are
+then used to reconstruct the precoding matrices all together by
+applying the reverse transformation of breaking the precoder into
+independent parameters as given in
+Equation~\ref{eq:decompose}. Precoder estimates are calculated using
+the current and past data, though the earlier precoder estimates
+arexsa improved by refining the past past estimates through
+interpolation with current estimates (indicated by the black arrow in
+Fig.~\ref{fig:adpm-fig}). Thus, the approach here is able to exploit
+the time and frequency correlation jointly to enhance performance.
+
 % To construct the precoders for the rest of the subcarriers, we will

### 2020-04-30 paper/response_letter_2.tex
Add changes

@@
 predictions would possess less error. We have now clarified this in
-the manuscript as well.
+the manuscript as well, as follows:
+
+\emph{Precoder estimates are calculated using
+the current and past data, though the earlier precoder estimates
+arexsa improved by refining the past past estimates through
+interpolation with current estimates (indicated by the black arrow in
+Fig. 1).}
+
+In addition, we have now indicated the reverse arrow in Fig. 1 with a
+different colour and have clarified the refinement aspect in the
+caption. We believe that this should clarify this aspect suitably.
 \end{document}

### 2020-05-03 paper/main.tex
updating figure and description

--- 2020-04-26 paper/main.tex
@@
 \begin{center}
-\includegraphics[width=0.7\columnwidth]{images/new-adpm.pdf}
+\includegraphics[width=0.68\columnwidth]{images/new-adpm2.pdf}
 \caption{\label{fig:adpm-fig}Subcarriers that need to be
-  quantized. We see that the subcarrier indicated by
-  red arrows has a precoder whose parameters are determined jointly
-  using the neighbouring parameters, both in time and frequency.}
+  quantized. Subcarriers indicated by red shaded cells have precoders 
+  whose parameters are determined using the neighbouring parameters
+  (shown via red arrows), both in time and frequency. The black arrow 
+  indicates refinement of past precoders as fresher estimates are made
+  available.}
 \end{center}
@@
 % accuracy.
+
 In Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
@@
 applying the reverse transformation of breaking the precoder into
-independent parameters as given in Equation~\ref{eq:decompose}. To
-improve the precoder estimates and interpolation accuracy, we ``walk
-back'' to improve the past estimates by interpolating using current
-estimates (although the current estimates only use data
-causally). Thus, the approach here is able to exploit the time and
-frequency correlation jointly to enhance performance.
+independent parameters as given in
+Equation~\ref{eq:decompose}. Precoder estimates are calculated using
+the current and past data, though the earlier precoder estimates
+are improved by refining the past past estimates through
+interpolation with current estimates (indicated by the black arrow in
+Fig.~\ref{fig:adpm-fig}). Thus, the approach here is able to exploit
+the time and frequency correlation jointly to enhance performance.
+
 % To construct the precoders for the rest of the subcarriers, we will
